### 2012-07-23 2dconvolution.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

% Please do NOT edit this file.
% It has been automatically generated
% by a perl script from the original cxx sources
% in the Insight/Examples directory

% Any changes should be made in the file
% src/2dconvolution.c

The source code for this section can be found in the file \verb|2dconvolution.c|.\\  This file contains the necessary functions to compute the convolution between an 
  input array and a kernel (usually the input array is larger than the kernel). \\

	Includes:
\small
\begin{lstlisting}
	#include <stdlib.h> // malloc, calloc, free
	#include <stdio.h> // fprintf
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{free\_gaussian\_kernel}} \\

	\normalsize
	This function frees the memory allocated in the function \texttt{get\_neighborhood}. It receives as parameter the pointer to the array to be freed. \\
\small
\begin{lstlisting}
void free_neighborhood(double* neighborhood){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Free memory: \\
\small
\begin{lstlisting}
	free(neighborhood);
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{free\_gaussian\_kernel}} \\

	\normalsize
	This function frees the memory allocated in the function \texttt{get\_neighbors\_offset}. It receives as parameter the pointer to the array to be freed. \\
\small
\begin{lstlisting}
void free_neighbors_offsets(int* offsets){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Free memory: \\
\small
\begin{lstlisting}
	free(offsets);
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{get\_neighbors\_offset}} \\

	\normalsize
	This function computes the linear array index offsets required to access the neighbor pixels in 
	a neighborhood of size $n\times n$. This computation requires to know the width $w$ of the image. \\
\small
\begin{lstlisting}
int *get_neighbors_offset(int w, int n){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Memory allocation for the index offsets array: \\
\small
\begin{lstlisting}
	int *aux = malloc(n*n*sizeof(int));
\end{lstlisting}
\vspace{1ex}
\normalsize
	Computation of the index offsets array: \\
\small
\begin{lstlisting}
	int imax = n*n;
	int dif_fila_col = (n-1)/2;
	for (i=0;i<imax;i++){
		delta_fila = (int)(i/n)-dif_fila_col;
		delta_col = i-n*((int)(i/n))-dif_fila_col;
		aux[i] = delta_fila*w+delta_col;
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Return: \\
\small
\begin{lstlisting}
	return aux;
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{get\_neighborhood}} \\

	\normalsize
	This function returns an array of size $n\times n$ containing the neighbors of the 
  pixel at the position \texttt{pos} of the image \texttt{im}. For this calculation, 
	the index offsets array calculated with the function \texttt{get\_neighbors\_offset} 
	is needed.
\small
\begin{lstlisting}
double *get_neighborhood(double *im, int pos, int n, int* offsets){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Memory allocation for the neighborhood array: \\
\small
\begin{lstlisting}
	double *aux = malloc(n*n*sizeof(double));
\end{lstlisting}
\vspace{1ex}
\normalsize
	Assigning values ​​of the neighborhood: \\
\small
\begin{lstlisting}
	int i;
	int imax = n*n;
	for (i=0;i<imax;i++){
		aux[i] = im[pos+offsets[i]];
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Return: \\
\small
\begin{lstlisting}
	return aux;
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{conv2d}} \\

	\normalsize
	This function calculates the convolution of image \texttt{input} (size $w\times h$) with the kernel \texttt{kernel} (size $n\times n$).
	Returns an image of size $(w+n-1)\times(h+n-1)$ (due to padding). The integer \texttt{padding\_method} manages the padding method: $0$
  means zero-padding, $1$ means image boundary reflection. \\
\small
\begin{lstlisting}
double *conv2d(double *input, int w, int h, double *kernel, int n, int padding_method){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Zero padding: reserve memory for an array of size $(w+2(n-1))\times(h+2(n-1))$. \\
\small
\begin{lstlisting}
	int wx = (w+2*(n-1));
	int hx = (h+2*(n-1));
	double *aux = calloc(wx*hx,sizeof(double));
\end{lstlisting}
\vspace{1ex}
\normalsize
	Fill in the values ​​of the image \texttt{aux}, centering the original image \texttt{input} on it (zero-padding). \\
\small
\begin{lstlisting}
	int i,j,fila,col;
	int imax = wx*hx;
	if (padding_method == 0) {
		for(i=0;i<imax;i++){
			fila = (int)(i/wx);
			col = i-(wx*fila);
			if ( (fila>=n-1)&&(col>=n-1)&&(fila<h+n-1)&&(col<w+n-1) ) {
				aux[i] = input[(col-n+1)+(w*(fila-n+1))];
			}
		}
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
 Other padding method: reflection of the image (if \texttt{padding\_method} equal to $1$). \\
\small
\begin{lstlisting}
	if (padding_method == 1) {
		int fila_refl, col_refl;
		for(i=0;i<imax;i++){
			fila = (int)(i/wx);
			col = i-(wx*fila);
			if (fila<n-1) {
				fila_refl = 2*n - 3 - fila;
				if (col<n-1) { //zone1
					col_refl = 2*n - 3 - col;
				} else if (col<w+n-1) {	//zone2
					col_refl = col;
				} else { //zone3
					col_refl = 2*w + 2*n - 3 - col;
				}
			} else if (fila<h+n-1) {
				fila_refl = fila;
				if (col<n-1) { //zone4
					col_refl = 2*n - 3 - col;
				} else if (col<w+n-1) { //image
					col_refl = col;
				} else { //zone5
					col_refl =  2*w + 2*n - 3 - col;
				}
			} else {
				fila_refl = 2*h + 2*n - 3 - fila;
				if (col<n-1) { //zone6
					col_refl =	2*n - 3 - col;
				} else if (col<w+n-1) {	//zone7
					col_refl = col;
				} else { //zone8
					col_refl =  2*w + 2*n - 3 - col;
				}
			}
			aux[i] = input[(col_refl-n+1)+(w*(fila_refl-n+1))];
		} //for
	} //if
\end{lstlisting}
\vspace{1ex}
\normalsize
	Reserve memory for the output array of size $(w+n-1)\times(h+n-1)$. \\
\small
\begin{lstlisting}
	double *out = malloc((w+n-1)*(h+n-1)*sizeof(double));
\end{lstlisting}
\vspace{1ex}
\normalsize
	Compute the convolution. Most of the operations are intended to calculate the relative 
	positions between the images \texttt{aux} and \texttt{out}. \\
\small
\begin{lstlisting}
	double acum = 0;
	int pos;
	// Convolution
	imax = (w+n-1)*(h+n-1);
	int jmax = n*n;
	int *offsets = get_neighbors_offset(wx, n);
	int dif_fila_col = (n-1)/2;
	for(i=0;i<imax;i++){
		fila = (int)(i/(w+n-1));
		col = i-((w+n-1)*fila);
		fila += dif_fila_col;
		col += dif_fila_col;
		pos = wx*fila + col;
		// compute convolution:
		acum = 0;
		for (j=0;j<jmax;j++){
			acum += aux[pos+offsets[j]]*kernel[j];
		}
		out[i] = acum;
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Free and return: \\
\small
\begin{lstlisting}
	free(aux);
	free_neighbors_offsets(offsets);
	return out;
	//return aux; //TEST
\end{lstlisting}
\vspace{1ex}
\normalsize

### 2012-07-23 edge-detection-review-v10.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v10
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

\documentclass{ipol}
\ipolSetYear{2012}
\ipolSetMonth{06}
\ipolSetDay{06}
%\ipolSetID{arced}
\ipolSetTitle{A Review of Classic Edge Detectors}
\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
\ipolSetAffiliations{%
$^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,array}
\usepackage[hang,center]{subfigure}
\usepackage[table]{xcolor}
\usepackage{multicol}
\usepackage{listings}
\usepackage{color}
%\usepackage{algorithm}
\usepackage{algorithmic}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
%\numberwithin{figure}{section}
\begin{document}
\bibliographystyle{ieeetr}
\lstset{
	backgroundcolor=\color[rgb]{0.94,0.94,0.94},
    tabsize=4,
    inputencoding=utf8x,
         extendedchars=\true,
         language=C,
         basicstyle=\scriptsize,
                 showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

%-------------------------------------------------------------------------------

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second reviewed family it is used second derivative, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time) are compared, considering different ways to convolve an 
image with a kernel (step required in some of the algorithms).
\end{abstract}

%-------------------------------------------------------------------------------

\begin{ipolCode}
For all the algorithms reviewed, an open source C implementation is provided and can be downloaded from 
\href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}. You can also follow 
this code project in this \href{git://github.com/haldos/edges.git}{GIT repository}. An 
\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration} is also available, where 
you can test and reproduce our results.
\end{ipolCode}

%-------------------------------------------------------------------------------

%\begin{ipolSupp}
%\end{ipolSupp}

%-------------------------------------------------------------------------------

\section{Introduction}
\label{sec:intro}

Edge detection is one of the first approaches published in the literature for 
segmenting images. The basic idea is to detect abrupt changes in intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives. \\

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} the 
behavior of such algorithms is described, as an introduction to a more sophisticated 
analysis of the edge detection process.\\

In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to approximate locally the image around a point. Then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters. \\

Along with this paper, a detailed and well commented source code is presented, which 
implements the described algorithms. In section \ref{sec:appendix1} some common mathematical developments are presented. 
A commented explanation of the source code can be found in section \ref{sec:appendix2}. This work also provides tools for code documentation 
(see section \ref{sec:appendix2}). \\

Results of the implemented algorithms are presented in section \ref{sec:results}, along with examples 
to compare their performance. Conclusions are detailed in section \ref{sec:conclusions}.

\nocite{IPOL}

%-------------------------------------------------------------------------------

\section{Algorithms based on first derivative}
\label{sec:first}

Algorithms based on first derivative studied in this paper have a common scheme, with 
the only difference in the type of filtering. Figure \ref{fig:blockdiagram1} presents
a block diagram of these algorithms. \\

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\textwidth]{blockdiagram1.pdf}
	\caption{Block diagram of first derivative edge detection algorithms.}
	\label{fig:blockdiagram1}
\end{figure}

The usual tool to find the amplitude and direction of changes in 
intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
as the vector
\begin{equation}
	\nabla f = 
				\begin{bmatrix} 
					g_x \\ g_y
				\end{bmatrix}
	=				
				\begin{bmatrix} 
					\cfrac{\partial f}{\partial x} \\ \cfrac{\partial f}{\partial y} \\
				\end{bmatrix}.
\end{equation}

The magnitude ($M$) and direction ($\alpha$) of the gradient vector $\nabla f$ at location $(x,y)$
are calculated as
\begin{equation}
\label{eq:mag_alpha}
	\begin{cases}
		M(x,y) = \sqrt{g_x^2 + g_y^2} \\
		\alpha(x,y) = \tan^{-1} \begin{bmatrix} g_x \\ g_y \end{bmatrix} .\\
	\end{cases}
\end{equation}

The direction of an edge at an arbitrary location $(x,y)$ of the image is 
orthogonal to the direction $\alpha(x,y)$ of the gradient vector.\\

To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image is required. When dealing with digital images, numerical approximations 
of the partial derivatives are required, calculated in a neighborhood of each point. In the following,
the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.

%-------------------------------------------------------------------------------

\subsection{The \textit{Roberts} operators}

The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with
a small $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
The image is composed by discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. But when finding 
diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to normalize the masks, in order to have gain equal to $1$. \\

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.13\textwidth]{1dmasks_n.pdf}
	\caption{One-dimensional normalized masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{SCfigure}

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.15\textwidth]{roberts_n.pdf}
	\caption{Normalized \textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{SCfigure}

%-------------------------------------------------------------------------------

\subsection{The \textit{Prewitt} operators}

Masks of size $2\times2$, although conceptually simple, are not symmetrical with respect to the central points. 
Having symmetrical edges is a desirable property and can only be achieved with oddly sized masks, the smallest 
of them being the 3x3. These masks provide more information to find the direction of the edges, because they take 
into account information on opposite sides of the central point. \\

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators}. \\

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{prewitt_n.pdf}
	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
smoothing properties in the direction opposite to the gradient.

%-------------------------------------------------------------------------------

\subsection{The \textit{Sobel} operators}

A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
difference. It can be shown that using a $2$ in the center location provides better image smoothing. This
variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}. \\

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{sobel_n.pdf}
	\caption{Normalized \textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\

The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, 
because of the weighted shape of $[1\ 2\ 1]$.

%-------------------------------------------------------------------------------

\subsection{Computation of the edges}

As mentioned before, edges can be seen as abrupt changes in intensity, i.e.\ higher values of gradient.
An example of this behavior is shown in Figures \ref{fig:profiles3d-b} and \ref{fig:profiles3d-c}. \\

\begin{figure}[t!]
	\centering
	\subfigure[Grayscale image.]{\label{fig:profiles3d-a}\includegraphics[width=0.3\textwidth]{molino_crop_bw.jpg}}

	\subfigure[3D image representation.]{\label{fig:profiles3d-b}\includegraphics[width=0.48\textwidth]{profile3d_image.pdf}}
	\subfigure[3D gradient representation.]{\label{fig:profiles3d-c}\includegraphics[width=0.48\textwidth]{profile3d_gradient.pdf}}
	\caption{Example figure, 3D image and gradient representations.}
	\label{fig:profiles3d}
\end{figure}

\begin{figure}[b!]
	\centering
	\subfigure[$th = 0.1*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_1.png}}
	\quad
	\subfigure[$th = 0.2*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_2.png}}
	\quad
	\subfigure[$th = 0.4*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
	\caption{Thresholded images. Different values of the threshold applied to the magnitude of the gradient.}
	\label{fig:thresholding}
\end{figure}

Then, once the respective operators are applied, an approximation of the gradient (stored in two images) 
is obtained, containing the approximation of the partial derivatives $g_x$ and $g_y$. Gradient magnitude 
image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}). \\

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
\begin{equation*}
	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{th} \\
									1,& \mbox{if}\ M[i]<\mbox{th}
					  \end{cases}
\end{equation*}
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.

%-------------------------------------------------------------------------------

\subsection{Pseudo-code}

The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
\clearpage

\begin{algorithm}[t]
\caption{First derivative edge detection algorithms.}
\label{algo:fded}
\begin{algorithmic}[1]
\REQUIRE input image, threshold $th$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define $operator_x$ and $operator_y$ \COMMENT{Roberts, Prewitt or Sobel.}
\STATE $g_x \leftarrow$ convolution($im$,$operator_x$)
\STATE $g_y \leftarrow$ convolution($im$,$operator_y$)
\STATE $max_M \leftarrow 0$
\FORALL{pixel $i$ in image}
	\STATE $M[i] \leftarrow \sqrt{g_x^2+g_y^2}$ \COMMENT{Gradient magnitude.}
	\IF{$M[i]>max_M$}
		\STATE $max_M \leftarrow M[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image}
	\IF{$M[i] \geq th \times max_M$}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

%-------------------------------------------------------------------------------

\section{Algorithms based on second derivative}
\label{sec:second}

The edge detection methods discussed in the previous section are simply based on filtering the 
image with different masks, without taking into account the characteristics of the edges or 
noise in the image. \\

The two algorithms presented in this section (Marr-Hildreth \cite{AIM-518} [1980] and Haralick \cite{bb20239} [1987]) 
are based on the second derivative of the image, and both take steps to reduce noise before 
detecting edges in the image.

%-------------------------------------------------------------------------------

\subsection{The \textit{Marr} and \textit{Hildreth} algorithm}

The Marr-Hildreth algorithm is a method of detecting edges in digital 
images. It is based on finding zero crossing points of the second derivative
of the image. This can be done in several ways. Two different ways of doing 
this are implemented implemented in this work (see block diagram in Figure 
\ref{fig:blockdiagram2}); convolving the image with a Gaussian kernel and then 
approximating the second derivative (Laplacian) with a 3x3 kernel, or 
convolving the image with a kernel calculated as the Laplacian of a 
Gaussian function. There are more ways to do so, for example, using 
recursive Gaussian filters \cite{Deriche1993Recursively}. \\
\clearpage

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.9\textwidth]{blockdiagram2.pdf}
	\caption{Block diagram of Marr-Hildreth algorithm.}
	\label{fig:blockdiagram2}
\end{figure}

The algorithm is divided in three steps, each one described later:
\begin{enumerate}
	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
	\item Convolution of the image with:
	\begin{itemize}
		\item a Laplacian of Gaussian (LoG) kernel. (or)
		\item a Gaussian kernel and then a Laplacian operator.
	\end{itemize}
	\item Search of zero crossing points in the filtered image.
\end{enumerate}

Some auxiliary functions were implemented for computing the operations
needed in the algorithm, such as Gaussian kernel and Laplacian of a Gaussian 
kernel generation, and 2-D convolution of an image with a given kernel, 
with different boundary conditions (these operations will be discussed 
in detail in section \ref{sec:appendix1}).

%-------------------------------------------------------------------------------

\subsubsection{Gaussian and LoG kernels}

The Marr-Hildreth algorithm consists on convolving the input image $f(x,y)$ with a LoG kernel;
\begin{equation}\label{eq:log}
  g(x,y) = [\nabla^2G(x,y)]\star f(x,y), 
\end{equation}
and then finding the zero crossings of $g(x,y)$ to determine the location of edges in $f(x,y)$. 
Because these are linear processes, equation \ref{eq:log} can be written also as
\begin{equation}
  g(x,y) = \nabla^2[G(x,y)\star f(x,y)]
\end{equation}
indicating that image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result.\\

Then, generation of both Gaussian and LoG kernels is required (see section \ref{sec:appendix1}).\\

The Marr-Hildreth edge-detection algorithm may be summarized as follows:
\begin{enumerate}
	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling equation \ref{eq:gaussian_function}.
	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask:
			$\begin{bmatrix}
			1 &  1 & 1 \\
			1 & -8 & 1 \\
			1 &  1 & 1 \\
			\end{bmatrix}$\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
	\item Find the zero crossings of the image from step 2.
\end{enumerate}

%-------------------------------------------------------------------------------

\subsubsection{Zero crossing}

A zero crossing at pixel $p$ implies that the signs of at least two opposite neighboring pixels are 
different. There are four cases to test: left/right, up/down, and the two diagonals. In this case 
a threshold is used, so that not only the signs of the opposite pixels must differ, also their 
difference in absolute value must be greater than a certain threshold. \\

The zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
image (both Gaussian and LoG kernels). Each pixel $p$ has eight neighbors, named according to their position 
as follows.

\begin{center}
\begin{tabular}{ c c c c c }
	$p_{up,left}$		& 					& $p_{up,middle}$	&					& $p_{up,right}$ 		\\
						& $\nwarrow$		& $\uparrow$		& $\nearrow$		&						\\
	$p_{middle,left}$	& $\leftarrow$		& $p$				& $\rightarrow$		& $p_{middle,right}$	\\
						& $\swarrow$		& $\downarrow$		& $\searrow$		&						\\
	$p_{down,left}$		&					& $p_{down,middle}$	&					& $p_{down,right}$.		\\  
\end{tabular}
\end{center}

Then a pixel $p$ is considered as edge pixel if any of the following conditions is true 
(for simplicity the Laplacian image is denoted as $\mathcal{L}$):
\begin{itemize}
	\item $(\mbox{sign}(\mathcal{L}[p_{up,left}])\neq\mbox{sign}(\mathcal{L}[p_{down,right}])$ \& $|\mathcal{L}[p_{up,left}]-\mathcal{L}[p_{down,right}]|>th_{ZC}*max_L$
	\item $(\mbox{sign}(\mathcal{L}[p_{up,middle}])\neq\mbox{sign}(\mathcal{L}[p_{down,middle}])$ \& $|\mathcal{L}[p_{up,middle}]-\mathcal{L}[p_{down,middle}]|>th_{ZC}*max_L$
	\item $(\mbox{sign}(\mathcal{L}[p_{down,left}])\neq\mbox{sign}(\mathcal{L}[p_{up,right}])$ \& $|\mathcal{L}[p_{down,left}]-\mathcal{L}[p_{up,right}]|>th_{ZC}*max_L$
	\item $(\mbox{sign}(\mathcal{L}[p_{middle,left}])\neq\mbox{sign}(\mathcal{L}[p_{middle,right}])$ \& $|\mathcal{L}[p_{middle,left}]-\mathcal{L}[p_{middle,right}]|>th_{ZC}*max_L$. \\
\end{itemize}

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
accuracy could be used (e.g.\ marching squares \cite{marching_cubes}).

%-------------------------------------------------------------------------------

\subsubsection{Pseudo-code}

The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}
\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
\IF{Gaussian kernel}
	\STATE Define Laplacian approx. operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

%-------------------------------------------------------------------------------

\subsection{The \textit{Haralick} algorithm}

In this section the original work of Haralick \cite{bb20239} on edge detection is presented in detail.
The main idea behind this algorithm is identical to that of the previous method; find zeros in 
the second derivative of the image. In this method, the input image is smoothly approximated through local bi-cubic
polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
its parameters.%\footnote{This implementation is slightly different from the traditional implementation 
%of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
%in order to implement two edge detection methods using only the second derivative (along the 
%Marr-Hildreth algorithm).}

%-------------------------------------------------------------------------------

\subsubsection{Bi-cubic polynomial fitting}
\label{sec:bicubic}

Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}. \\

The surrounding neighborhood of a point $(x,y)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
\end{equation}

To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

\begin{equation*}
	\begin{array}{l}
		f_1 = f(x_1,y_1) = k_1 + k_2x_1 + k_3y_1 + k_4x_1^2 + k_5x_1y_1 + k_6y_1^2 + k_7x_1^3 + k_8x_1^2y_1 + k_9x_1y_1^2 + k_{10}y_1^3 \\
		f_2 = f(x_2,y_2) = k_1 + k_2x_2 + k_3y_2 + k_4x_2^2 + k_5x_2y_2 + k_6y_2^2 + k_7x_2^3 + k_8x_2^2y_2 + k_9x_2y_2^2 + k_{10}y_2^3 \\
		\vdots \\
		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3 . \\
	\end{array}
\end{equation*}

Using matrix notation, the system can be rewritten as

\begin{equation*}
	\begin{bmatrix} 
		f_1		\\ 
		f_2		\\ 
		\vdots	\\
		f_{25}
	\end{bmatrix} 
	= 
	\begin{bmatrix} 
		1 		& x_1 		& y_1 		& x_1^2 	& x_1y_1 		& y_1^2 	& \hdots 	& y_1^3 	\\
		1 		& x_2 		& y_2 		& x_2^2 	& x_2y_2 		& y_2^2 	& \hdots 	& y_2^3 	\\
		\vdots	& \vdots	& \vdots	& \vdots	& \vdots		& \vdots	& \ddots	& \vdots	\\
		1 		& x_{25}	& y_{25}	& x_{25}^2 	& x_{25}y_{25} 	& y_{25}^2 	& \hdots 	& y_{25}^3
	\end{bmatrix}
	\times
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix}
	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k} .\\
\end{equation*}

Then, the least squares problem can be solved by using the normal equations,

\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
\end{equation*}

$\mathbf{B}$ is a $10\times25$ matrix,

\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix} .\\
\end{equation*}

For each coefficient ($i$ from $1$ to $10$),

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
\end{equation}

where

\begin{equation}
	\label{eq:b_i}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	& b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}

\vspace{1ex}
Using equation \ref{eq:coefficients} and the masks given in Table \ref{table:b_i}, it is possible to compute 
the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.\\

\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
\begin{table}
\centering
\subfigure[$\mathbf{b_1}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
275 & 125 &  75 & 125 & 275 \\ 
\hline
225 &  75 &  25 &  75 & 225 \\
\hline
275 & 125 &  75 & 125 & 275 \\
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_2}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-2260 & -620 & 0 & 620 & 2260 \\ 
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-1460 & -220 & 0 & 220 & 1460 \\
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-2260 & -620 & 0 & 620 & 2260 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_3}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
2260  &  1660 &  1460 &  1660 &  2260 \\ 
\hline
620   &   320 &   220 &   320 &   620 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-620  &  -320 &  -220 &  -320 &  -620 \\ 
\hline
-2260 & -1660 & -1460 & -1660 & -2260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_4}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 620 & 450 & 620 & 1130 \\ 
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
730  & 220 &  50 & 220 &  730 \\
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
1130 & 620 & 450 & 620 & 1130 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_5}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-400 & -200 & 0 &  200 &  400 \\ 
\hline
-200 & -100 & 0 &  100 &  200 \\ 
\hline
0    &    0 & 0 &    0 &    0 \\
\hline
200  &  100 & 0 & -100 & -200 \\ 
\hline
400  &  200 & 0 & -200 & -400 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_6}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 830 & 730 & 830 & 1130 \\ 
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
450  & 150 &  50 & 150 &  450 \\
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
1130 & 830 & 730 & 830 & 1130 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_7}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\ 
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-5540 &  -820 & 0 &  820 & 5540 \\
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_8}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
5640  &  3600 &  2920 &  3600 &  5640 \\ 
\hline
1800  &   780 &   440 &   780 &  1800 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-1800 &  -780 &  -440 &  -780 & -1800 \\ 
\hline
-5640 & -3600 & -2920 & -3600 & -5640 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_9}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\ 
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-2920 &  -440 & 0 &  440 & 2920 \\
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_{10}}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
8260  &  6220 &  5540 &  6220 &  8260 \\ 
\hline
2180  &  1160 &   820 &  1160 &  2180 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-2180 & -1160 &  -820 & -1160 & -2180 \\ 
\hline
-8260 & -6220 & -5540 & -6220 & -8260 \\
\hline
\end{tabular}} \\
\caption{Masks to compute the coefficients of the bicubic fit.}
\label{table:b_i}
\end{table}

%-------------------------------------------------------------------------------

\subsubsection{Analytical calculation of the second derivative}
\label{sec:secderivative}

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as

\begin{align}
\label{eq:sincos}
	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
\end{align}

Now substituting the variables $x$ and $y$ in polar form as

\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
\end{equation*}

in the bi-cubic polynomial, we obtain

\begin{equation}
	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3 ,
\end{equation}

where

\begin{align}
\label{eq:c}
	C_0 & = k_1 \nonumber \nonumber \\
	C_1 & = k_2\sin(\theta) + k_3\cos(\theta) \nonumber \\
	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). \nonumber \\
\end{align}

The derivatives are obtained as follows.

\begin{align}
	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
	f'''_{\theta}(\rho) = 6C_3 .\nonumber \\
\end{align}

Then, the condition that the second derivative is equal to zero becomes

\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}

where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
value is between $0.4$ and $0.6$.\\

The other condition required is that the third derivative to be negative, i.e.\

\begin{equation}
	f'''_{\theta}(\rho) = 6C_3 < 0 \ \ \Rightarrow \ \ C_3 < 0.
\end{equation}

Given the direction defined by $\theta$, with $\rho>0$, one edge will always be an ascending step. This 
indicates that first derivative in this direction has a maximum and that second derivative has a zero 
crossing with negative slope. That is, third derivative less than zero.

%-------------------------------------------------------------------------------

\subsubsection{Algorithm}

The Haralick edge detection algorithm is summarized in the following 4 steps:

\begin{enumerate}
	\item Find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
	\item Compute $\sin(\theta)$ and $\cos(\theta)$ (equations \ref{eq:sincos}).
	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$ and $C_3 < 0$, then that point is an edge point.
\end{enumerate}

%-------------------------------------------------------------------------------

\subsubsection{Pseudo-code}

The pseudo-code of Haralick's algorithm is shown in Algorithm \ref{algo:haralick}.

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

%-------------------------------------------------------------------------------

\section{Common Mathematical Operations}
\label{sec:appendix1}

All algorithms implemented use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

%-------------------------------------------------------------------------------

\subsection{Grayscale conversion}
\label{sec:grayscale}

The first step in every algorithm presented above is to convert color images to gray intensity images. 
For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float} to avoid truncated results. See code in section \ref{app:marr-hildreth}.}.

%-------------------------------------------------------------------------------

\subsection{Kernel generation}

Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
kernels are generated by sampling the corresponding analytical function, which in each case depends 
on the standard deviation $\sigma$ of the Gaussian function. The result is an array of 
size $n$ by $n$.

%-------------------------------------------------------------------------------

\subsubsection{Gaussian kernel}

Gaussian kernel is generated by sampling the 2-D Gaussian function
%\footnote{Note that for simplicity the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = \frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A value of $n$ large enough is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\

Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
(first odd integer greater than $6\sigma=24$).

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
large enough to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

%-------------------------------------------------------------------------------

\subsubsection{LoG kernel}

The Laplacian of Gaussian $\nabla^2G(x,y)$ can be obtained analyically first and then a discrete mask 
can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
have to be performed, one with the kernel and another with the Laplacian operator).\\

\begin{equation}
	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
\end{equation}

We first compute

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}

where the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted. Therefore we obtain

\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}\\

Now the LoG kernel is generated by sampling the function defined in equation \ref{eq:log_function}. 
Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

%-------------------------------------------------------------------------------

\subsubsection{Gaussian and LoG functions comparison}

As mentioned before, the size $n$ of the kernel and the standard deviation $\sigma$ of the exponential function 
are not independent of each one. This is because the function LoG has wider support than the
Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to 
generate a correctly sampled LoG kernel than in the case of the Gaussian kernel, with the same $\sigma$.\\

Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
be $29$.

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernels.pdf}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{SCfigure}

%-------------------------------------------------------------------------------

\subsection{Convolution}

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.

%-------------------------------------------------------------------------------

\section{Implementation}
\label{sec:appendix2}

In this section a detailed explanation of the source code of the different implemented algorithms
is presented. \\

To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}. There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}. All these links were last checked in July 2012. \\

This code documentation was generated using a perl script to extract some parts of the 
code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\

Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\

Required external libraries: 
\begin{itemize}
	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
\end{itemize}

\subsection{Roberts, Prewitt and Sobel}
\input{test_fded.tex}

\subsection{Marr-Hildreth}
\label{app:marr-hildreth}
\input{test_mh.tex}

\subsection{Haralick}
\input{test_haralick.tex}

\subsection{Gaussian kernel generation}
\input{gaussian_kernel.tex}

\subsection{2D convolution}
\input{2dconvolution.tex}

%-------------------------------------------------------------------------------

\section{Results}
\label{sec:results}

The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used. \\

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.2\textwidth]{results/square127.png}
	\caption{Test image 1: white square on black background (127$\times$127 pixels).}
	\label{fig:original1}
\end{figure}

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm. \\

Now the performance of each algorithm is analyzed, using a more realistic test image shown in Figure \ref{fig:original2}. \\

\begin{figure}[t!]
	\centering
	\subfigure[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
	\quad
	\subfigure[Image converted to greyscale.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
	\caption{Test image 2: windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

First the results of the first derivative edge detection algorithms is shown Figures \ref{fig:result2-a} to \ref{fig:result2-c}. Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm. Table \ref{exectime2} shows the execution time for each one of the algorithms. \\

\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}

	\subfigure[Sobel. $th=0.1$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\quad
	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}

	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using Figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative. It can be seen in Figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither the Haralick algorithm is able to detect it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel. \\

Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes. \\

\begin{figure}[h!]
	\centering
	\subfigure[Original.]{\label{fig:result3-a}\includegraphics[width=0.48\textwidth]{results/molino_bw_zoom.png}}

	\subfigure[Roberts.]{\label{fig:result3-b}\includegraphics[width=0.48\textwidth]{results/molino_roberts_zoom.png}}
	\quad
	\subfigure[Prewitt.]{\label{fig:result3-c}\includegraphics[width=0.48\textwidth]{results/molino_prewitt_zoom.png}}

	\subfigure[Sobel.]{\label{fig:result3-d}\includegraphics[width=0.48\textwidth]{results/molino_sobel_zoom.png}}
	\quad
	\subfigure[Marr-Hildreth (Gaussian).]{\label{fig:result3-e}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth_zoom.png}}

	\subfigure[Marr-Hildreth (LoG).]{\label{fig:result3-f}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth-log_zoom.png}}
	\quad
	\subfigure[Haralick.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
	\caption{Results of the algorithms using Figure \ref{fig:original2} as input image, enlarging an interesting area.}
	\label{fig:result3}
\end{figure}

\clearpage

%-------------------------------------------------------------------------------

\subsection{Further examples}
\label{sec:examples}

Further examples obtained with the implemented algorithms are shown in Figures \ref{fig:example1} and \ref{fig:example2}. More examples can be found in the online \href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{demo}. \\

\begin{figure}[h!]
	\centering
	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
	\quad
	\subfigure[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}

	\subfigure[Roberts. $th=0.1$.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=2$, $n=21$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=2$, $n=25$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.35$.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
	\caption{Example: Lena (512$\times$512 pixels).}
	\label{fig:example1}
\end{figure}

\clearpage

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity of the edges. They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g. Canny edge detector \cite{Canny1986}). \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail. This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in Figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

\begin{figure}[t!]
	\centering
	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
	\quad
	\subfigure[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}

	\subfigure[Roberts. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=2$, $n=21$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=2$, $n=25$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
	\caption{Example: Oranges (536$\times$480 pixels).}
	\label{fig:example2}
\end{figure}

%-------------------------------------------------------------------------------

\subsection{Video}

This is an example of applying the edge detection algorithms described here, frame by frame, to a video: 
\begin{itemize}
	\centering
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (43 Mb).
\end{itemize}
\begin{multicols}{2}
\begin{itemize}
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts-wide_0.1.mov}{roberts version} (7.9 Mb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt-wide_0.1.mov}{prewitt version} (9.1 Mb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel-wide_0.1.mov}{sobel version} (9.1 Mb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian-wide_3_19_0.04.mov}{marr-hildreth-gaussian version} (12 Mb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log-wide_3_25_0.04.mov}{marr-hildreth-log version} (14 Mb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick-wide_0.5.mov}{haralick version} (12 Mb).
\end{itemize}
\end{multicols}

\clearpage

%-------------------------------------------------------------------------------

\section{Conclusions}
\label{sec:conclusions}

Some of the most traditional methods of edge detection in digital images were discussed and carefully implemented in this work. The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods. \\

The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation. Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels). The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges. Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make more fair the comparison with the second derivative methods. \\

The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed. In real images, they have better behavior than first derivative algorithms. \\

Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one. The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel). It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels. \\

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

All algorithms have their advantages and disadvantages. Choosing one or the other may depend on requirements of the application. Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.\\

%To conclude, this paper is a summary of some classical edge detection algorithms. In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods. \\

As an addendum to this work a detailed implementation of the described algorithms is presented, 
available in C language, along with an online demo where users can run the algorithms 
with arbitrary images uploaded by themselves. Also some code documentation tools are supplied, 
which were used to generate the detailed code description presented in section \ref{sec:appendix2}. We think this tool could be useful for future IPOL articles.

\clearpage

%-------------------------------------------------------------------------------

\bibliography{bibliography}

%-------------------------------------------------------------------------------

\end{document}

### 2012-07-23 edge-detection-review-v3.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

%-------------------------------------------------------------------------------
% Edge detectors review
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------
\documentclass{ipol}
\ipolSetYear{2012}
\ipolSetDOI{10.XXXX/ipol.2012.sc-red}

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,subfig,array}
\usepackage[table]{xcolor}
\usepackage{multicol}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\begin{document}

%-------------------------------------------------------------------------------
\title{Review of Edge Detectors}
\author{Haldo Spont\'on,\\
        Juan Cardelino}
\date{}
\ipolMaketitle

%-------------------------------------------------------------------------------
\begin{ipolAbstract}
In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
like Marr-Hildreth and Haralick.\\
We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).\\

%The Marr-Hildreth algorithm is a method of detecting edges in digital 
%images. It's based on convolving the image with the Laplacian of the 
%Gaussian function. Then, zero crossings are detected in the filtered 
%result to obtain the edges points.
\end{ipolAbstract}

%-------------------------------------------------------------------------------
\begin{ipolCode}
The C implementation of the algorithms implemented in this work (beta version) can be
downloaded from:
\begin{itemize}
	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
\end{itemize}
\end{ipolCode}

%-------------------------------------------------------------------------------
%\begin{ipolSupp}
%\end{ipolSupp}

%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

Edge detection is one of the first approaches published in the literature for 
segmenting images. The basic idea is to detect abrupt changes in intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives.

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image. In section ~\ref{sec:first} we study the 
performance of such algorithms, as an introduction to a more sophisticated 
analysis into the edge finding process.\\

In 1980 Marr and Hildreth argued that intensity changes are not independent 
of image scale, so the edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. The bulk of this work focuses on the 
study of this algorithm, presented in Section ~\ref{sec:second}.\\

In section ~\ref{sec:examples} we present some examples to compare the 
performance of the algorithms studied in this work. We also present a 
video created using the Marr-Hildreth algorithm, applied frame by frame.\\

%-------------------------------------------------------------------------------
\section{Algorithms based on first derivative}
\label{sec:first}

The selected tool to find the amplitude and direction of changes in 
intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
as the vector:
\begin{equation}
	\nabla f = 
				\begin{bmatrix} 
					g_x \\ g_y
				\end{bmatrix}
	=				
				\begin{bmatrix} 
					\cfrac{\partial f}{\partial x} \\ \cfrac{\partial f}{\partial y} \\
				\end{bmatrix}
\end{equation}

The magnitude ($M$) and direction ($\alpha$) of the gradient vector $\nabla f$ at location $(x,y)$
are calculated as:
\begin{equation}
	\begin{cases}
		M(x,y) = \sqrt{g_x^2 + g_y^2} \\
		\alpha(x,y) = \tan^{-1} \begin{bmatrix} g_x \\ g_y \end{bmatrix} \\
	\end{cases}
\end{equation}

The direction of an edge at an arbitrary location $(x,y)$ of the image is 
orthogonal to the direction $\alpha(x,y)$ of the gradient vector.\\

We need to compute partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image. Since we are dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. Next we will study 
the methods of Roberts, Prewitt and Sobel, whose main difference is how they approach this calculation.\\

\subsection{The \textit{Roberts} operators}

The most usual approach of the first derivative is to take the Taylor expansion of order zero with
a small $h$. Approximating the first derivative:
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}\\
\end{equation}
The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, we approximate the 
first derivative of the image based on the intensity values ​​at points of the image as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(x+1,y) - f(x,y)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(x,y+1) - f(x,y)\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in figure \ref{fig:1dmasks}. But when we want 
to find diagonal edges, we need 2-D masks. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks with this purpose. These operators are based on implementing the diagonal 
diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}.\\

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.14\textwidth]{1dmasks.pdf}
	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{SCfigure}

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.15\textwidth]{roberts.pdf}
	\caption{\textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{SCfigure}

\subsection{The \textit{Prewitt} operators}

Masks of size $2\times2$ are conceptually simple, but not being symmetrical to a central point implies 
that they are not as useful for detecting edges as symmetrical ones. The smaller mask with these characteristics is 
$3\times3$. These masks provide more information to find the direction of the edges, because they take 
into account information on opposite sides of the central point.\\

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are given 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators}.\\

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{prewitt.pdf}
	\caption{\textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

\subsection{The \textit{Sobel} operators}

A slight variation of the Prewitt operators use more weight on the central coefficients of the 
difference. It can be shown that using a $2$ in the center location provides image smoothing. This
variation is implemented using masks in figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}.\\

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{sobel.pdf}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

The Prewitt masks are simpler to implement than the Sobel masks, but the computational difference
between them is not an issue. It is preferable to use Sobel operators, because they have better
noise-suppression characteristics (smoothing). Noise suppression is an important issue when dealing
with derivatives.\\

%-------------------------------------------------------------------------------
\section{Algorithms based on second derivative}
\label{sec:second}

The edge detection methods discussed in the previous section are simply based on filtering the 
image with different masks, without taking into account the characteristics of the edges or 
noise in the image.\\

The two algorithms that we will see in this section (Marr-Hildreth [1980] and Haralick [1987]) 
are based on the second derivative of the image, and both take steps to reduce noise before 
detecting edges in the image.\\

\subsection{The \textit{Marr} and \textit{Hildreth} algorithm}

The Marr-Hildreth algorithm is a method of detecting edges in digital 
images. It's based on find zero crossing points of the second derivative
of the image. This can be done in several ways. In this work we implemented 
two different ways; convolving the image with a Gaussian kernel and then 
approximating the second derivative (Laplacian) with a 3x3 kernel, or 
convolving the image with a kernel calculated as the Laplacian of a 
Gaussian function. There are more ways to do so, for example, using 
recursive Gaussian filters.\\ % poner referencia!

The algorithm is divided in three steps, each one described later:
\begin{enumerate}
	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
	\item Convolution of the image with:
	\begin{itemize}
		\item a Laplacian of Gaussian (LoG) kernel. (or)
		\item a Gaussian kernel and then a Laplacian operator.
	\end{itemize}
	\item Search of zero crossing points in the filtered image.\\
\end{enumerate}

Some auxiliary functions were implemented for computing the operations
needed in the algorithm, such as Gaussian kernel and Laplacian of a Gaussian 
kernel generation, and 2-D convolution of an image with a given kernel, 
with zero padding.\\

%-------------------------------------------------------------------------------
\subsubsection{Gaussian and LoG kernels}

The Marr-Hildreth algorithm consists of convolving the LoG filter with an input image, $f(x,y)$.
\begin{equation}\label{eq:log}
  g(x,y) = [\nabla^2G(x,y)]\star f(x,y)
\end{equation}
and then finding the zero crossings of $g(x,y)$ to determine the location of edges in $f(x,y)$. 
Because these are linear processes, equation \ref{eq:log} can be written also as
\begin{equation}
  g(x,y) = \nabla^2[G(x,y)\star f(x,y)]
\end{equation}
indicating that we can smooth the image first with a Gaussian filter and then compute the Laplacian of the result.\\

Then we need to generate both Gaussian kernel and LoG kernel (see section \ref{sec:appendix2}).\\

The Marr-Hildreth edge-detection algorithm may be summarized as follows:
\begin{enumerate}
	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling equation \ref{eq:gaussian_function}.
	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask:
			$\begin{bmatrix}
			1 &  1 & 1 \\
			1 & -8 & 1 \\
			1 &  1 & 1 \\
			\end{bmatrix}$\footnote{Steps 1 and 2 can be made into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
	\item Find the zero crossings of the image from step 2.
\end{enumerate}

%-------------------------------------------------------------------------------
\subsubsection{Zero crossing}

A zero crossing at pixel $p$ implies that the signs of at least two opposite neighboring pixels are 
different. There are four cases to test: left/right, up/down, and the two diagonals. In this case 
we work with a threshold, so that not only the signs of the opposite pixels must differ, also their 
difference in absolute value must be greater than a certain threshold.\\

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, methods for finding zero crossings with subpixel accuracy 
should be used.\\

%-------------------------------------------------------------------------------
\subsection{The \textit{Haralick} algorithm}

The idea behind the Haralick edge detector is identical to that of the previous method; find zeros in 
the second derivative of the image. In this method, we smooth the input image through local bi-cubic
polynomial fitting. Then, calculating the second derivative analytically, we can find an expression 
equivalent to cancel the second derivative of the polynomial as a function of the parameters of it.\footnote{This implementation is slightly 
different to the traditional implementation of the Haralick algorithm. We do not use the condition concerning the third derivative.}\\

\subsubsection{Bi-cubic polynomial fitting}
\label{sec:bicubic}

We adjust the surrounding neighborhood of a point $(x,y)$ in the image $f$ using the following bi-cubic polynomial:
\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3\\
\end{equation}

To solve this problem, we need to take more neighbors than coefficients to be adjusted. The smallest 
neighborhood of odd dimensions has size $5\times5$. Now it is possible to convolve the input image
with some precomputed masks, instead of finding the least square solution to find the coefficients 
$k_1\dots k_{10}$ at each point $(x,y)$.\\

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. We use a least square
formulation to obtain the coefficients (25 points give equal number of equations):

\begin{equation*}
	\begin{array}{l}
		f_1 = f(x_1,y_1) = k_1 + k_2x_1 + k_3y_1 + k_4x_1^2 + k_5x_1y_1 + k_6y_1^2 + k_7x_1^3 + k_8x_1^2y_1 + k_9x_1y_1^2 + k_{10}y_1^3\\
		f_2 = f(x_2,y_2) = k_1 + k_2x_2 + k_3y_2 + k_4x_2^2 + k_5x_2y_2 + k_6y_2^2 + k_7x_2^3 + k_8x_2^2y_2 + k_9x_2y_2^2 + k_{10}y_2^3\\
		\vdots \\
		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3\\
	\end{array}
\end{equation*}

Using matrix notation:

\begin{equation*}
	\begin{bmatrix} 
		f_1		\\ 
		f_2		\\ 
		\vdots	\\
		f_{25}
	\end{bmatrix} 
	= 
	\begin{bmatrix} 
		1 		& x_1 		& y_1 		& x_1^2 	& x_1y_1 		& y_1^2 	& \hdots 	& y_1^3 	\\
		1 		& x_2 		& y_2 		& x_2^2 	& x_2y_2 		& y_2^2 	& \hdots 	& y_2^3 	\\
		\vdots	& \vdots	& \vdots	& \vdots	& \vdots		& \vdots	& \ddots	& \vdots	\\
		1 		& x_{25}	& y_{25}	& x_{25}^2 	& x_{25}y_{25} 	& y_{25}^2 	& \hdots 	& y_{25}^3
	\end{bmatrix}
	\times
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix}
	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k}\\
\end{equation*}

Then:

\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f}\\
\end{equation*}

$\mathbf{B}$ is a $10\times25$ matrix:

\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix}\\
\end{equation*}

For each coefficient ($i$ from $1$ to $10$):

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i}\\
\end{equation}

where:

\begin{equation}
	\label{eq:b_i}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	& b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}\\
\end{equation}

Using equation \ref{eq:coefficients} and the masks given in table \ref{table:b_i}, we can compute 
the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.\\

\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
\begin{table}
\centering
\subfloat[$\mathbf{b_1}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
275 & 125 &  75 & 125 & 275 \\ 
\hline
225 &  75 &  25 &  75 & 225 \\
\hline
275 & 125 &  75 & 125 & 275 \\
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
\end{tabular}}
\qquad\qquad
\subfloat[$\mathbf{b_2}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-2260 & -620 & 0 & 620 & 2260 \\ 
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-1460 & -220 & 0 & 220 & 1460 \\
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-2260 & -620 & 0 & 620 & 2260 \\
\hline
\end{tabular}} \\
\subfloat[$\mathbf{b_3}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
2260  &  1660 &  1460 &  1660 &  2260 \\ 
\hline
620   &   320 &   220 &   320 &   620 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-620  &  -320 &  -220 &  -320 &  -620 \\ 
\hline
-2260 & -1660 & -1460 & -1660 & -2260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfloat[$\mathbf{b_4}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 620 & 450 & 620 & 1130 \\ 
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
730  & 220 &  50 & 220 &  730 \\
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
1130 & 620 & 450 & 620 & 1130 \\
\hline
\end{tabular}} \\
\subfloat[$\mathbf{b_5}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-400 & -200 & 0 &  200 &  400 \\ 
\hline
-200 & -100 & 0 &  100 &  200 \\ 
\hline
0    &    0 & 0 &    0 &    0 \\
\hline
200  &  100 & 0 & -100 & -200 \\ 
\hline
400  &  200 & 0 & -200 & -400 \\
\hline
\end{tabular}}
\qquad\qquad
\subfloat[$\mathbf{b_6}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 830 & 730 & 830 & 1130 \\ 
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
450  & 150 &  50 & 150 &  450 \\
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
1130 & 830 & 730 & 830 & 1130 \\
\hline
\end{tabular}} \\
\subfloat[$\mathbf{b_7}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\ 
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-5540 &  -820 & 0 &  820 & 5540 \\
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfloat[$\mathbf{b_8}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
5640  &  3600 &  2920 &  3600 &  5640 \\ 
\hline
1800  &   780 &   440 &   780 &  1800 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-1800 &  -780 &  -440 &  -780 & -1800 \\ 
\hline
-5640 & -3600 & -2920 & -3600 & -5640 \\
\hline
\end{tabular}} \\
\subfloat[$\mathbf{b_9}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\ 
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-2920 &  -440 & 0 &  440 & 2920 \\
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\
\hline
\end{tabular}}
\qquad\qquad
\subfloat[$\mathbf{b_{10}}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
8260  &  6220 &  5540 &  6220 &  8260 \\ 
\hline
2180  &  1160 &   820 &  1160 &  2180 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-2180 & -1160 &  -820 & -1160 & -2180 \\ 
\hline
-8260 & -6220 & -5540 & -6220 & -8260 \\
\hline
\end{tabular}} \\
\caption{Masks to compute the coefficients of the bicubic fit.}
\label{table:b_i}
\end{table}
\clearpage

\subsubsection{Analytical calculation of the second derivative}
\label{sec:secderivative}

We approximate the neighborhood of each point of the image using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If we just take the first order terms of this 
polynomial, the Gradient angle, defined with positive y-axis is defined as:

\begin{align}
\label{eq:sincos}
	\sin(\theta) & = \frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = \frac{k_3}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
\end{align}

Now substituting the variables $x$ and $y$ in polar form as:

\begin{equation*}
	x = \rho\cos{\theta} \: , \: y = \rho\sin(\theta) \\
\end{equation*}

in the bi-cubic polynomial, we obtain:

\begin{equation}
	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3
\end{equation}

where:

\begin{align}
\label{eq:c}
	C_0 & = k_1 \nonumber \nonumber \\
	C_1 & = k_2\sin(\theta) + k_3\cos(\theta) \nonumber \\
	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). \nonumber \\
\end{align}

The derivatives are obtained as follows:

\begin{align}
	f'{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
\end{align}

Then, the condition of the second derivative equal to zero becomes:

\begin{equation}
	f''{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0.
\end{equation}

\subsubsection{Algorithm}

The Haralick edge detection algorithm is summarized in the following 4 steps:

\begin{enumerate}
	\item Find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
	\item Compute $\sin(\theta)$ and $\cos(\theta)$ (equations \ref{eq:sincos}).
	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$, then that point is an edge point.
\end{enumerate}

%-------------------------------------------------------------------------------
\section{Results}
\label{sec:results}

First we see the results of each of the algorithms on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.2\textwidth]{results/square.png}
	\caption{Test image 1: white square on black background.}
	\label{fig:original1}
\end{figure}

The figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm. Execution time\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result1-a}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_roberts.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result1-b}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_prewitt.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Sobel.]{\label{fig:result1-c}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_sobel.png}\hspace*{0.5cm}}

	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result1-d}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result1-e}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth-log.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Haralick.]{\label{fig:result1-f}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_haralick.png}\hspace*{0.5cm}}
	\caption{Results of the algorithms using the figure \ref{fig:original1} as input image.}
	\label{fig:result1}
\end{figure}

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using the figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. Let us now analyze the performance of each of the algorithmos using a more realistic test image, shown in figure \ref{fig:original2}. \\

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.5\textwidth]{results/molino.jpg}
	\caption{Test image 2: windmill.}
	\label{fig:original2}
\end{figure}

First we show in figure \ref{fig:result2} the results of the first derivative edge detection algorithms.

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}

	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\caption{Results of the first derivative algorithms using the figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

Now we show in figures \ref{fig:result3-a} and \ref{fig:result3-b} the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and in figure \ref{fig:result3-c} the results obtained using the Haralick algorithm. In table \ref{exectime2} we show the execution time for each one of the algorithms. \\

\begin{figure}[h!]
	\centering
	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-a}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-b}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}

	\subfloat[Haralick.]{\label{fig:result3-c}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the Marr-Hildreth and the Haralick algorithms, using the figure \ref{fig:original2} as input image.}
	\label{fig:result3}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using the figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}
\vspace{0.5cm}

It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. \\

\clearpage
%-------------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusions}

\clearpage
%-------------------------------------------------------------------------------
\section{Appendix: Common operations}
\label{sec:appendix1}

All algorithms implemented use some common operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care.\\

\subsection{Grayscale conversion}

The first step in every algorithm presented above is to convert color images to gray intensity images. 
We use the libpng coefficients to do this:
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively.\\

%-------------------------------------------------------------------------------
\subsection{Kernel generation}
\label{sec:appendix2}

Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
kernels are generated by sampling the corresponding analytical function, which in each case depends 
on the parameter $\sigma$ (standard deviation of the Gaussian function). The result is an array of 
size $n$ by $n$.\\

\subsubsection{Gaussian kernel}

We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{space constant}).\\

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. We need a value 
of $n$ large enough to ensure that no information is lost when creating the kernel $G$. To ensure this, we take 
$n$ equal to the first odd integer greater than $6\sigma$. Larger values of $n$ does not add more 
significant samples ​​of $G$, and increases the number of operations in the convolution.\\

Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
(first odd integer greater than $6\sigma=24$).\\

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
sufficient to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

\subsubsection{LoG kernel}

We can obtain analytically the Laplacian of Gaussian $\nabla^2G(x,y)$ first, and generate a kernel 
with the result function. Use this kernel for edge detection involves only one convolution with 
the input image (unlike the case of Gaussian kernel, where we have to make two convolutions, one 
with the kernel and another with the Laplacian operator).\\

\begin{equation}
	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
\end{equation}

We first obtain:

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2}
\end{equation}

And then:

\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
\end{equation}

Similarly:

\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
\end{equation}

Therefore we obtain:

\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}
\end{equation}\\

Now we generate the LoG kernel by sampling this function. Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the 
values $\sigma=4$ and $n=31$.\\

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

\subsubsection{Gaussian and LoG functions comparison}

As we said before, the size of the kernel $n$ and the standard deviation of the exponential function 
$\sigma$ are not independent of each one. This is because the function LoG is "wider" that the 
Gaussian function (i.e. LoG function has a slower decay), so we need a greater value of $n$ to 
generate a correctly sampled LoG kernel than the needed in the case of the Gaussian kernel, with the same $\sigma$.\\

Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
be $29$.\\

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernels.pdf}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{SCfigure}

\clearpage
%-------------------------------------------------------------------------------
\section{Examples}
\label{sec:examples}

\subsection{Images}

Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} to \ref{fig:example3}. More examples can be found in the online demo. \\

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
	\caption{Example: Lena.}
	\label{fig:example1}
\end{figure}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
	\caption{Example: Oranges.}
	\label{fig:example2}
\end{figure}

\vspace{2cm}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
	\caption{Example: Shapes.}
	\label{fig:example3}
\end{figure}

%-------------------------------------------------------------------------------
\subsection{Video}

%TODO: videos de ejemplo con todos los métodos.

This is an example of applying the edge detection algorithms described here, frame by frame, to a video: 
\begin{itemize}
	\centering
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (xxxMb).
\end{itemize}
\begin{multicols}{2}
\begin{itemize}
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts.mov}{roberts version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt.mov}{prewitt version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel.mov}{sobel version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian.mov}{marr-hildreth-gaussian version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log.mov}{marr-hildreth-log version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick.mov}{haralick version} (xxxMb).
\end{itemize}
\end{multicols}

%-------------------------------------------------------------------------------
\begin{thebibliography}{9}

\end{thebibliography}

\end{document}
%-------------------------------------------------------------------------------

### 2012-07-23 edge-detection-review-v4.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

--- 2012-07-23 edge-detection-review-v3.tex
@@
 \usepackage{multicol}
+\usepackage{listings}
+\usepackage{color}
 \newtheorem{theorem}{Theorem}
@@
 \begin{document}
+\bibliographystyle{ieeetr}
+\lstset{
+	backgroundcolor=\color[rgb]{0.9,0.9,0.9},
+    tabsize=4,
+    inputencoding=utf8x,
+         extendedchars=\true,
+         language=C++,
+         basicstyle=\scriptsize,
+                 showtabs=false,
+        showspaces=false,
+        showstringspaces=false,
+        identifierstyle=\ttfamily,
+        keywordstyle=\color[rgb]{0,0,1},
+        commentstyle=\color[rgb]{0.133,0.545,0.133},
+        stringstyle=\color[rgb]{0.627,0.126,0.941},
+}
 
@@
 As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
-such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
-like Marr-Hildreth and Haralick.\\
+such as Sobel, Prewitt and Roberts \cite{Gonzalez2007Digital}. The second derivative is also used, for example in algorithms 
+like Marr-Hildreth \cite{AIM-518} and Haralick \cite{bb20239}.\\
 We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
@@
 The C implementation of the algorithms implemented in this work (beta version) can be
-downloaded from:
-\begin{itemize}
-	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
-	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
-	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
-\end{itemize}
+downloaded from \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}.
+%\begin{itemize}
+%	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
+%	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
+%	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
+%\end{itemize}
+% TODO referencia a github
 \end{ipolCode}
@@
 Detecting those changes in intensity for the purpose of finding edges in images 
-can be accomplished using first or second order derivatives.
+can be accomplished using first or second order derivatives. \\
 
@@
 crossing in the second derivative. The bulk of this work focuses on the 
-study of this algorithm, presented in Section ~\ref{sec:second}.\\
-
-In section ~\ref{sec:examples} we present some examples to compare the 
+study of this algorithm, presented in Section ~\ref{sec:second}. In this section 
+we also present the Haralick algorithm, which is also based on the second derivative.
+This algorithm has the particularity of proposing a model to locally approximate 
+the image around a point. Then, with this model, second derivative can be calculated 
+analytically (rather approximate) and finding edges is done by imposing a 
+condition over the model parameters. \\
+
+Results and conclusions are shown in sections \ref{sec:results} and 
+\ref{sec:conclusions} respectively. In section ~\ref{sec:examples} we present some examples to compare the 
 performance of the algorithms studied in this work. We also present a 
-video created using the Marr-Hildreth algorithm, applied frame by frame.\\
+video created using the implemented algorithms, applied frame by frame.\\
 
@@
 
-The selected tool to find the amplitude and direction of changes in 
+The usual tool to find the amplitude and direction of changes in 
 intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
@@
 at every pixel in the image. Since we are dealing with digital images, numerical approximations 
-of the partial derivatives over a neighborhood about a point are required. Next we will study 
-the methods of Roberts, Prewitt and Sobel, whose main difference is how they approach this calculation.\\
+of the partial derivatives over a neighborhood about a point are required. In the following we will study 
+the methods of Roberts, Prewitt and Sobel, whose main difference is how they perform this calculation.\\
 
@@
 
-The most usual approach of the first derivative is to take the Taylor expansion of order zero with
+The most usual approach of the first derivative is to take the Taylor expansion of first order with
 a small $h$. Approximating the first derivative:
@@
 \label{eq:roberts1}
-	g_x = \frac{\partial f(x,y)}{\partial x} = f(x+1,y) - f(x,y)
+	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
 \end{equation}
@@
 \label{eq:roberts2}
-	g_y = \frac{\partial f(x,y)}{\partial y} = f(x,y+1) - f(x,y)\\
+	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j)\\
 \end{equation}
@@
 to find diagonal edges, we need 2-D masks. The \textit{Roberts operators} are one of the earliest 
-attempts to use 2-D masks with this purpose. These operators are based on implementing the diagonal 
+attempts to use 2-D masks for this purpose. These operators are based on implementing the diagonal 
 diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}.\\
@@
 	\includegraphics[width=0.2\textwidth]{results/square.png}
-	\caption{Test image 1: white square on black background.}
+	\caption{Test image 1: white square on black background (128$\times$128 pixels).}
 	\label{fig:original1}
@@
 
-The figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
+Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
 The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm.
-Execution time\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
+Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
 \\
@@
 It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster.
-Let us now analyze the performance of each of the algorithmos using a more realistic test image, shown in figure \ref{fig:original2}.
-\\
+All algorithms show consistent results, but the Haralick algorithm shows some rare artifacts out of the edges.
+This may occur because this is the only algorithm that assumes a certain regularity (third degree) in the neighborhood of a pixel.
+Synthetic edges are far from this regularity.
+This may cause the adjustment of the bicubic polynomial function around these points to be quite different from reality.
+\\
+
+Let us now analyze the performance of each of the algorithms using a more realistic test image, shown in figure \ref{fig:original2}. \\
 
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{results/molino.jpg}
-	\caption{Test image 2: windmill.}
+	\subfloat[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
+	\quad
+	\subfloat[Grayscale image.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
+	\caption{Test image 2: windmill (1000$\times$563 pixels).}
 	\label{fig:original2}
@@
 
-First we show in figure \ref{fig:result2} the results of the first derivative edge detection algorithms.
-
+First we show in figures \ref{fig:result2-a} to \ref{fig:result2-c} the results of the first derivative edge detection algorithms.
+Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm.
+In table \ref{exectime2} we show the execution time for each one of the algorithms.
+\\
 \begin{figure}[h!]
@@
 	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
-	\caption{Results of the first derivative algorithms using the figure \ref{fig:original2} as input image.}
+	\quad
+	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
+
+	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
+	\quad
+	\subfloat[Haralick.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
+	\caption{Results of the first derivative, Marr-Hildreth and Haralick algorithms using the figure \ref{fig:original2} as input image.}
 	\label{fig:result2}
-\end{figure}
-
-Now we show in figures \ref{fig:result3-a} and \ref{fig:result3-b} the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and in figure \ref{fig:result3-c} the results obtained using the Haralick algorithm.
-In table \ref{exectime2} we show the execution time for each one of the algorithms.
-\\
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-a}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-b}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
-
-	\subfloat[Haralick.]{\label{fig:result3-c}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
-	\caption{Results of the Marr-Hildreth and the Haralick algorithms, using the figure \ref{fig:original2} as input image.}
-	\label{fig:result3}
 \end{figure}
@@
 
-It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. \\
+It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones.
+We can see in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to better view of the detail.
+Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded).
+Neither Haralick algorithm detects it.
+However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.
+\\
+
+Another observation is that the edges detected by the Haralick algorithm appear to be thicker than those detected with the other algorithms.
+This may be due the regularity that Haralick assumes, previously mentioned.
+\\
+
+\begin{figure}[h!]
+	\centering
+	\subfloat[Original.]{\label{fig:result3-a}\includegraphics[width=0.48\textwidth]{results/molino_bw_zoom.png}}
+
+	\subfloat[Roberts.]{\label{fig:result3-b}\includegraphics[width=0.48\textwidth]{results/molino_roberts_zoom.png}}
+	\quad
+	\subfloat[Prewitt.]{\label{fig:result3-c}\includegraphics[width=0.48\textwidth]{results/molino_prewitt_zoom.png}}
+
+	\subfloat[Sobel.]{\label{fig:result3-d}\includegraphics[width=0.48\textwidth]{results/molino_sobel_zoom.png}}
+	\quad
+	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-e}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth_zoom.png}}
+
+	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-f}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth-log_zoom.png}}
+	\quad
+	\subfloat[Haralick.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
+	\caption{Results of the algorithms using the figure \ref{fig:original2} as input image, enlarging an interesting area.}
+	\label{fig:result3}
+\end{figure}
 
@@
 
-\clearpage
-%-------------------------------------------------------------------------------
-\section{Appendix: Common operations}
-\label{sec:appendix1}
-
-All algorithms implemented use some common operations that are independent of the algorithms themselves. 
-These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
-implemented with care.\\
-
-\subsection{Grayscale conversion}
-
-The first step in every algorithm presented above is to convert color images to gray intensity images. 
-We use the libpng coefficients to do this:
-\begin{equation}
-    Y = (6968 R + 23434 G + 2366 B) / 32768
-\end{equation}
-where $R$, $G$ and $B$ are the red, green and blue components respectively.\\
-
-%-------------------------------------------------------------------------------
-\subsection{Kernel generation}
-\label{sec:appendix2}
-
-Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
-kernels are generated by sampling the corresponding analytical function, which in each case depends 
-on the parameter $\sigma$ (standard deviation of the Gaussian function). The result is an array of 
-size $n$ by $n$.\\
-
-\subsubsection{Gaussian kernel}
-
-We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
-we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
-\begin{equation}
-	\label{eq:gaussian_function}
-	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
-\end{equation}
-where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{space constant}).\\
-
-The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
-input parameters, but these are not strictly independent of each other. We need a value 
-of $n$ large enough to ensure that no information is lost when creating the kernel $G$. To ensure this, we take 
-$n$ equal to the first odd integer greater than $6\sigma$. Larger values of $n$ does not add more 
-significant samples ​​of $G$, and increases the number of operations in the convolution.\\
-
-Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
-(first odd integer greater than $6\sigma=24$).\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
-	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
-sufficient to have a good approximation of the Gaussian function in the kernel.}
-	\label{fig:gaussian_kernel}
-\end{SCfigure}
-
-\subsubsection{LoG kernel}
-
-We can obtain analytically the Laplacian of Gaussian $\nabla^2G(x,y)$ first, and generate a kernel 
-with the result function. Use this kernel for edge detection involves only one convolution with 
-the input image (unlike the case of Gaussian kernel, where we have to make two convolutions, one 
-with the kernel and another with the Laplacian operator).\\
-
-\begin{equation}
-	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
-\end{equation}
-
-We first obtain:
-
-\begin{equation} 
-	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2}
-\end{equation}
-
-And then:
-
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
-\end{equation}
-
-Similarly:
-
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
-\end{equation}
-
-Therefore we obtain:
-
-\begin{equation}
-	\label{eq:log_function}
-	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}
-\end{equation}\\
-
-Now we generate the LoG kernel by sampling this function. Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the 
-values $\sigma=4$ and $n=31$.\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
-	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
-to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
-Gaussian kernel.}
-	\label{fig:log_kernel}
-\end{SCfigure}
-
-\subsubsection{Gaussian and LoG functions comparison}
-
-As we said before, the size of the kernel $n$ and the standard deviation of the exponential function 
-$\sigma$ are not independent of each one. This is because the function LoG is "wider" that the 
-Gaussian function (i.e. LoG function has a slower decay), so we need a greater value of $n$ to 
-generate a correctly sampled LoG kernel than the needed in the case of the Gaussian kernel, with the same $\sigma$.\\
-
-Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
-required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
-example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
-first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
-be $29$.\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernels.pdf}
-	\caption{Comparison of the Gaussian and LoG functions.}
-	\label{fig:kernels}
-\end{SCfigure}
+We have discussed and carefully implemented some of the more traditional methods of edge detection in digital images.
+We have tested the implemented algorithms with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
+\\
+
+The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation.
+Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
+The results obtained with these methods are quite good, considering its simplicity, but they have problems such as noise and discontinuity at the edges.
+\\
+
+The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
+In real images, they have better behavior towards first derivative algorithms.
+\\
+
+Comparing the two versions of the Marr-Hildreth algorithm, is significantly faster the version with Gaussian kernel and numerical approximation of the Laplacian.
+The LoG kernel version, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
+It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
+\\
+
+Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
+It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function.
+This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
+\\
+
+All algorithms have their advantages and disadvantages. Choosing one or other may depend on requirements after edge detection. \\
+
+% At the end:
+Finally, this paper is a summary of some classical edge detection algorithms.
+In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods.
+\\
 
@@
 
-Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} to \ref{fig:example3}. More examples can be found in the online demo. \\
+Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} and \ref{fig:example2}. More examples can be found in the online \href{http://link_to_demo}{demo}. \\
 
@@
 	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
+	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
+	\quad
+	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}
 
@@
 	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
-	\caption{Example: Lena.}
+	\caption{Example: Lena (512$\times$512 pixels).}
 	\label{fig:example1}
@@
 	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
+	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
+	\quad
+	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}
 
@@
 	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
-	\caption{Example: Oranges.}
+	\caption{Example: Oranges (536$\times$480 pixels).}
 	\label{fig:example2}
@@
 
-\vspace{2cm}
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
-	\caption{Example: Shapes.}
-	\label{fig:example3}
-\end{figure}
+Some observations: 
+\begin{itemize}
+	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity in the edges.
+They are also very affected by noise in images, appearing unwanted isolated edges.
+\\
+	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
+This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel.
+This behavior can be seen in Lena's hair (figure \ref{fig:example1}), or in the inside of the oranges (example \ref{fig:example2}).
+\\
+	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order).
+This causes thicker edges, as shown in both examples.
+But, note that some edges are only detected with this algorithm, e.g.
+the lower edges of the oranges, which are in the shadow, in figure \ref{fig:example2}.
+These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods.
+\\
+\end{itemize}
+
+%\vspace{2cm}
+
+%\begin{figure}[t!]
+%	\centering
+%	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}
+
+%	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
+%	\quad
+%	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
+%	\quad
+%	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}
+
+%	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
+%	\quad
+%	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
+%	\quad
+%	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
+%	\caption{Example: Shapes.}
+%	\label{fig:example3}
+%\end{figure}
 
@@
 %-------------------------------------------------------------------------------
-\begin{thebibliography}{9}
-
-\end{thebibliography}
-
+%\begin{thebibliography}{9}
+%\end{thebibliography}
+
+\bibliography{bibliography}
+
+\clearpage
+%-------------------------------------------------------------------------------
+\section{Appendix 1: Common operations}
+\label{sec:appendix1}
+
+All algorithms implemented use some common operations that are independent of the algorithms themselves. 
+These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
+implemented with care.\\
+
+\subsection{Grayscale conversion}
+
+The first step in every algorithm presented above is to convert color images to gray intensity images. 
+We use the libpng coefficients to do this:
+\begin{equation}
+    Y = (6968 R + 23434 G + 2366 B) / 32768
+\end{equation}
+where $R$, $G$ and $B$ are the red, green and blue components respectively.\\
+
+%-------------------------------------------------------------------------------
+\subsection{Kernel generation}
+\label{sec:appendix2}
+
+Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
+kernels are generated by sampling the corresponding analytical function, which in each case depends 
+on the parameter $\sigma$ (standard deviation of the Gaussian function). The result is an array of 
+size $n$ by $n$.\\
+
+\subsubsection{Gaussian kernel}
+
+We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
+we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
+\begin{equation}
+	\label{eq:gaussian_function}
+	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
+\end{equation}
+where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\
+
+The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
+input parameters, but these are not strictly independent of each other. We need a value 
+of $n$ large enough to ensure that no information is lost when creating the kernel $G$. To ensure this, we take 
+$n$ equal to the first odd integer greater than $6\sigma$. Larger values of $n$ does not add more 
+significant samples ​​of $G$, and increases the number of operations in the convolution.\\
+
+Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
+(first odd integer greater than $6\sigma=24$).\\
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
+	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
+sufficient to have a good approximation of the Gaussian function in the kernel.}
+	\label{fig:gaussian_kernel}
+\end{SCfigure}
+
+\subsubsection{LoG kernel}
+
+We can obtain analytically the Laplacian of Gaussian $\nabla^2G(x,y)$ first, and generate a kernel 
+with the result function. Use this kernel for edge detection involves only one convolution with 
+the input image (unlike the case of Gaussian kernel, where we have to make two convolutions, one 
+with the kernel and another with the Laplacian operator).\\
+
+\begin{equation}
+	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
+\end{equation}
+
+We first obtain:
+
+\begin{equation} 
+	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2}
+\end{equation}
+
+And then:
+
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
+\end{equation}
+
+Similarly:
+
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
+\end{equation}
+
+Therefore we obtain:
+
+\begin{equation}
+	\label{eq:log_function}
+	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}
+\end{equation}\\
+
+Now we generate the LoG kernel by sampling this function. Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the 
+values $\sigma=4$ and $n=31$.\\
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
+	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
+to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
+Gaussian kernel.}
+	\label{fig:log_kernel}
+\end{SCfigure}
+
+\subsubsection{Gaussian and LoG functions comparison}
+
+As we said before, the size of the kernel $n$ and the standard deviation of the exponential function 
+$\sigma$ are not independent of each one. This is because the function LoG is "wider" that the 
+Gaussian function (i.e. LoG function has a slower decay), so we need a greater value of $n$ to 
+generate a correctly sampled LoG kernel than the needed in the case of the Gaussian kernel, with the same $\sigma$.\\
+
+Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
+required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
+example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
+first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
+be $29$.\\
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernels.pdf}
+	\caption{Comparison of the Gaussian and LoG functions.}
+	\label{fig:kernels}
+\end{SCfigure}
+
+\clearpage
+%-------------------------------------------------------------------------------
+\section{Appendix 2: Implementation}
+\label{sec:appendix2}
+
+\subsection{Roberts, Prewitt and Sobel}
+\input{test_fded.tex}
+
+\subsection{Marr-Hildreth}
+\label{app:marr-hildreth}
+\input{test_mh.tex}
+
+\subsection{Haralick}
+\input{test_haralick.tex}
+
+\subsection{Gaussian kernel generation}
+\input{gaussian_kernel.tex}
+
+\subsection{2D convolution}
+\input{2dconvolution.tex}
+
+%-------------------------------------------------------------------------------
 \end{document}
-%-------------------------------------------------------------------------------

### 2012-07-23 edge-detection-review-v8.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

--- 2012-07-23 edge-detection-review-v10.tex
@@
 %-------------------------------------------------------------------------------
-% A Review of Classic Edge Detectors - v10
+% A Review of Classic Edge Detectors - v8
 % by Haldo Spontón & Juan Cardelino
@@
 \numberwithin{table}{section}
-%\numberwithin{figure}{section}
+\numberwithin{figure}{section}
 \begin{document}
@@
 The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
-such as Sobel, Prewitt and Roberts. In the second reviewed family it is used second derivative, for example in algorithms 
+such as Sobel, Prewitt and Roberts. In the second family reviewed, second derivative is used, for example in algorithms 
 like Marr-Hildreth and Haralick. \\
@@
 \begin{ipolCode}
-For all the algorithms reviewed, an open source C implementation is provided and can be downloaded from 
+For all the algorithsm reviewed, an open source C implementation is provided and can be downloaded from 
 \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}. You can also follow 
@@
 behavior of such algorithms is described, as an introduction to a more sophisticated 
-analysis of the edge detection process.\\
+analysis into the edge detection process.\\
 
@@
 which is also reviewed in section \ref{sec:second}. This algorithm has the particularity of 
-proposing a model to approximate locally the image around a point. Then, using this model, 
+proposing a model to locally approximate the image around a point. Then, using this model, 
 an approximation to the second derivative of the image can be calculated analytically and 
-finding edges is achieved by imposing a condition over the model parameters. \\
+finding edges is done by imposing a condition over the model parameters. \\
 
@@
 at every pixel in the image is required. When dealing with digital images, numerical approximations 
-of the partial derivatives are required, calculated in a neighborhood of each point. In the following,
+of the partial derivatives over a neighborhood about a point are required. In the following,
 the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.
@@
 
-Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
+Finally, the edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
 \begin{equation*}
-	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{th} \\
-									1,& \mbox{if}\ M[i]<\mbox{th}
+	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{threshold} \\
+									1,& \mbox{if}\ M[i]<\mbox{threshold}.
 					  \end{cases}
 \end{equation*}
-where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
+A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
+It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
 for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
-of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
+of the image to be processed. The advantage of this is to adapt the threshold to the dynamic range of the image. 
 Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
@@
 			1 &  1 & 1 \\
-			\end{bmatrix}$\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
+			\end{bmatrix}$\footnote{Steps 1 and 2 can be made into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
 	\item Find the zero crossings of the image from step 2.
@@
 
-The zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
+Zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
 image (both Gaussian and LoG kernels). Each pixel $p$ has eight neighbors, named according to their position 
@@
 
-Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}. \\
+Here, the interpolation method used by Haralick is presented, although there are other options to do this \cite{getreuer}. \\
 
@@
 points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
-In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
-computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
+For speeding up the computation, it is possible to convolve the input image with some precomputed masks, instead 
+of finding the least square solution to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
 
@@
 expression in equation \ref{eq:bicubic}. If just the first order terms of this 
-polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as
+polynomial are taken, the gradient angle $\theta$, defined with negative y-axis, can be approxmated as
 
@@
 
-where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
-of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
-value is between $0.4$ and $0.6$.\\
+where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. \\
 
@@
 \end{equation}
-
-Given the direction defined by $\theta$, with $\rho>0$, one edge will always be an ascending step. This 
-indicates that first derivative in this direction has a maximum and that second derivative has a zero 
-crossing with negative slope. That is, third derivative less than zero.
 
@@
 	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
-	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
+	\IF{$|C_2/3C_3|\leq\rho_0$}
 		\STATE $im_{OUT}[i] \leftarrow 255$
@@
 \end{equation}
-where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float} to avoid truncated results.
-See code in section \ref{app:marr-hildreth}.}.
+where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float}. See code in section \ref{app:marr-hildreth}.}.
 
@@
 
-Gaussian kernel is generated by sampling the 2-D Gaussian function
-%\footnote{Note that for simplicity the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
+Gaussian kernel is generated by sampling the 2-D Gaussian function\footnote{Note that for simplicity
+the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
 \begin{equation}
 	\label{eq:gaussian_function}
-	G(x,y) = \frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{x^2+y^2}{2\sigma^2}}
+	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
 \end{equation}
@@
 input parameters, but these are not strictly independent of each other. 
-A value of $n$ large enough is needed to ensure that no information is lost when creating the kernel $G$.
+A value of $n$ big enough is needed to ensure that no information is lost when creating the kernel $G$.
 To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$.
@@
 \begin{equation} 
-	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
-\end{equation}
-
-where the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted. Therefore we obtain
+	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
+\end{equation}
+
+Therefore we obtain
 
@@
 	\label{eq:log_function}
-	LoG(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
+	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
 \end{equation}\\
@@
 kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
-relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
+relative to the edge of the image. Convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
 or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.
@@
 
-To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
-There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
-All these links were last checked in July 2012.
-\\
-
 This code documentation was generated using a perl script to extract some parts of the 
-code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
-
-Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
-
+code and comments into latex files. This code documentation tool is available as a GIT
+project: 
+\begin{center} 
+\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
+\end{center}
+Also a Doxygen generated documentation of the functions used is available from:
+\begin{center}
+\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1} \\
+\end{center}
 Required external libraries: 
-\begin{itemize}
-	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
-	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
-\end{itemize}
+\begin{center}
+\texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}) \\
+\texttt{libpng} (\url{http://www.libpng.org/})
+\end{center}
 
@@
 \input{2dconvolution.tex}
+
+\clearpage
 
@@
 	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.
-The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm.
-Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
+The first three (\ref{fig:result1-a}, \ref{fig:result1-b} and \ref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\ref{fig:result1-d} and \ref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\ref{fig:result1-f}) corresponds to the Haralick algorithm.}
 	\label{fig:result1}
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 	\begin{center}
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
+	\caption{Execution time of the algorithms (including i/o) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
 	\label{exectime1}
@@
 All algorithms show consistent results.
-Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.
+Thicker edges appear in the results of Haralick, due to smooth edges model that is imposed in this algorithm.
 \\
@@
 	\quad
-	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
-
-	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
+	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3.5$, $n=25$, $th_{ZC}=0.05$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
+
+	\subfigure[Marr-Hildreth (LoG). $\sigma=3.5$, $n=29$, $th_{ZC}=0.1$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
 	\quad
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including I/O) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
+	\caption{Execution time of the algorithms (including i/o) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
 	\label{exectime2}
@@
 	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
-This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel.
+This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel.
 This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}).
@@
 The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges.
-Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make more fair the comparison with the second derivative methods.
 \\
@@
 Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
-It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function.
+It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function.
 This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
@@
 
-All algorithms have their advantages and disadvantages.
-Choosing one or the other may depend on requirements of the application.
-Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.\\
+All algorithms have their advantages and disadvantages. Choosing one or the other may depend on requirements of the application. \\
 
@@
 with arbitrary images uploaded by themselves. Also some code documentation tools are supplied, 
-which were used to generate the detailed code description presented in section \ref{sec:appendix2}. We think this tool could be useful for future IPOL articles.
+which were used to generate the detailed code description presented in section \ref{sec:appendix2} (Appendix 2). We think this tool could be useful for future IPOL articles.
 

### 2012-07-23 gaussian_kernel.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

% Please do NOT edit this file.
% It has been automatically generated
% by a perl script from the original cxx sources
% in the Insight/Examples directory

% Any changes should be made in the file
% src/gaussian_kernel.c

The source code for this section can be found in the file \verb|gaussian_kernel.c|.\\	This file implements the necessary functions for generating 
	Gaussian and LoG (Laplacian of a Gaussian) kernels. It is also 
	done here the alloc and free of the memory needed. \\

	Includes:
\small
\begin{lstlisting}
	#include <math.h> // exp
	#include <stdlib.h> // malloc, calloc, free
	#include <stdio.h> // fprintf
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{gaussian\_kernel}} \\

	\normalsize
	This function generates a Gaussian kernel of size $n\times n$ and standard deviation $\sigma$. \\
\small
\begin{lstlisting}
double *gaussian_kernel(int n, float sigma){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Memory allocation for the kernel: \\
\small
\begin{lstlisting}
	double *kernel = malloc(n*n*sizeof(double));
\end{lstlisting}
\vspace{1ex}
\normalsize
	A normalized Gaussian kernel is generated using the expression $e^{-\frac{x^2+y^2}{2\sigma^2}}$. \\
\small
\begin{lstlisting}
	int i;
	int fila, col, x, y;
	double suma = 0;
	int imax = n*n;
	for(i=0;i<imax;i++){
		fila = (int)(i/n);
		col = i-(n*fila);
		y = ((int)(n/2))-fila;
		x = col-((int)(n/2));
		kernel[i] = exp(-(x*x + y*y)/(2*sigma*sigma));
		suma += kernel[i];
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Kernel normalization, using the sum of its components. \\
\small
\begin{lstlisting}
	for(i=0;i<n*n;i++){
		kernel[i] = kernel[i]/suma;
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Return: \\
\small
\begin{lstlisting}
	return kernel;
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{free\_gaussian\_kernel}} \\

	\normalsize
	This function frees the memory allocated in the function \texttt{gaussian\_kernel}. It receives as parameter the pointer to the array to be freed. \\
\small
\begin{lstlisting}
void free_gaussian_kernel(double* kernel){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Free memory: \\
\small
\begin{lstlisting}
	free(kernel);
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Function \texttt{LoG\_kernel}} \\

	\normalsize
	This function generates a Laplacian of a Gaussian kernel (LoG kernel) of size $n\times n$ and standard deviation $\sigma$. \\
\small
\begin{lstlisting}
double *LoG_kernel(int n, float sigma){
\end{lstlisting}
\vspace{1ex}
\normalsize
	Memory allocation for the kernel: \\
\small
\begin{lstlisting}
	double *kernel = malloc(n*n*sizeof(double));
\end{lstlisting}
\vspace{1ex}
\normalsize
	We generate a Laplacian of a Gaussian kernel using the expression $\frac{x^2 + y^2 - 2\sigma^2}{\sigma^4}e^{-\frac{x^2+y^2}{2\sigma^2}}$. \\
\small
\begin{lstlisting}
	int i;
	int fila, col, x, y;
	double suma = 0;
	int imax = n*n;
	for(i=0;i<imax;i++){
		fila = (int)(i/n);
		col = i-(n*fila);
		y = ((int)(n/2))-fila;
		x = col-((int)(n/2));
		kernel[i] = ( (x*x + y*y - 2*sigma*sigma)/(sigma*sigma*sigma*sigma) )
					*exp(-(x*x + y*y)/(2*sigma*sigma));
	}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Return: \\
\small
\begin{lstlisting}
	return kernel;
\end{lstlisting}
\vspace{1ex}
\normalsize

### 2012-07-23 test_fded.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

% Please do NOT edit this file.
% It has been automatically generated
% by a perl script from the original cxx sources
% in the Insight/Examples directory

% Any changes should be made in the file
% src/test_fded.c

The source code for this section can be found in the file \verb|test_fded.c|.\\  First derivative edge detectors (Roberts, Prewitt and Sobel), main C file.\\

  Parameters:
	\begin{itemize}
		\item \texttt{input\_image} - Input image.
		\item \texttt{threshold} - Threshold for gradient image ($0\leq th \leq 1$).
		\item \texttt{padding\_method} - Padding method flag (in convolution): 0 means zero-padding, 1 means image boundary reflection.
	\end{itemize}

	\textit{Note: Output images are saved with the filenames ``\texttt{roberts.png}'', ``\texttt{prewitt.png}'' and ``\texttt{sobel.png}''.} \\

	Includes:
\small
\begin{lstlisting}
	#include "iio.c"
	#include "2dconvolution.c"
	#include <time.h>
\end{lstlisting}
\vspace{1ex}
\normalsize
  Macros:
\small
\begin{lstlisting}
#define MAX(x, y) (((x) > (y)) ? (x) : (y))
#define MIN(x, y) (((x) < (y)) ? (x) : (y))
#define THRESHOLD(x, th) (((x) > (th)) ? (255) : (0))
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Main function} \\
\small
\begin{lstlisting}
int main(int argc, char *argv[]) {
\end{lstlisting}
\vspace{1ex}
\normalsize
	Load input image (using \textit{iio}): \\
\small
\begin{lstlisting}
		int w, h, pixeldim;
		float *im_orig = iio_read_image_float_vec(argv[1], &w, &h, &pixeldim);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Grayscale conversion (if necessary): explained in \ref{sec:grayscale}. \\ \\
	Define the normalized Roberts, Prewitt and Sobel operators: \\
	(We use $3\times 3$ operators)
	\begin{itemize}
		\item	Roberts:
				$$
				R_1 = \begin{bmatrix} -1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0 \end{bmatrix}
				$$
				$$
				R_2 = \begin{bmatrix} 0 & -1 & 0 \\ 1 & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}
				$$
		\item	Prewitt:
				$$
				P_1 = \begin{bmatrix} \frac{-1}{6} & \frac{-1}{6} & \frac{-1}{6} \\ 0 & 0 & 0 \\ \frac{1}{6} & \frac{1}{6} & \frac{1}{6} \end{bmatrix}
				$$
				$$
				P_2 = \begin{bmatrix} \frac{-1}{6} & 0 & \frac{1}{6} \\ \frac{-1}{6} & 0 & \frac{1}{6} \\ \frac{-1}{6} & 0 & \frac{1}{6} \end{bmatrix}
				$$
		\item	Sobel:
				$$
				S_1 = \begin{bmatrix} \frac{-1}{8} & \frac{-1}{4} & \frac{-1}{8} \\ 0 & 0 & 0 \\ \frac{1}{8} & \frac{1}{4} & \frac{1}{8} \end{bmatrix}
				$$
				$$
				S_2 = \begin{bmatrix} \frac{-1}{8} & 0 & \frac{1}{8} \\ \frac{-1}{4} & 0 & \frac{1}{4} \\ \frac{-1}{8} & 0 & \frac{1}{8} \end{bmatrix}
				$$
	\end{itemize}
\small
\begin{lstlisting}
		double roberts_1[9] = {-1, 0, 0, 0, 1, 0, 0, 0, 0};		// ROBERTS
		double roberts_2[9] = { 0,-1, 0, 1, 0, 0, 0, 0, 0};		// OPERATORS
		//---------------------------------------------------------------------------------
		double prewitt_1[9] = {-1,-1,-1, 0, 0, 0, 1, 1, 1};		// PREWITT
		double prewitt_2[9] = {-1, 0, 1,-1, 0, 1,-1, 0, 1};		// OPERATORS
		//---------------------------------------------------------------------------------
		double sobel_1[9] = {-1,-2,-1, 0, 0, 0, 1, 2, 1};		// SOBEL
		double sobel_2[9] = {-1, 0, 1,-2, 0, 2,-1, 0, 1};		// OPERATORS
		//---------------------------------------------------------------------------------
		for (z=0;z<9;z++) {										// NORMALIZATION
			roberts_1[z] /= sqrt(2);
			roberts_2[z] /= sqrt(2);
			prewitt_1[z] /= sqrt(6);
			prewitt_2[z] /= sqrt(6);
			sobel_1[z] /= sqrt(12);
			sobel_2[z] /= sqrt(12);
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	The input image is convolved with the defined operatos, using the \texttt{conv2d} function in \texttt{2dconvolution.c}:
\small
\begin{lstlisting}
		int padding_method = atoi(argv[3]);
		double *im_r1 = conv2d(im, w, h, roberts_1, 3, padding_method);
		double *im_r2 = conv2d(im, w, h, roberts_2, 3, padding_method);
		double *im_p1 = conv2d(im, w, h, prewitt_1, 3, padding_method);
		double *im_p2 = conv2d(im, w, h, prewitt_2, 3, padding_method);
		double *im_s1 = conv2d(im, w, h, sobel_1, 3, padding_method);
		double *im_s2 = conv2d(im, w, h, sobel_2, 3, padding_method);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Allocate memory for final images:
\small
\begin{lstlisting}
		float *im_roberts = malloc(w*h*sizeof(float));
		float *im_prewitt = malloc(w*h*sizeof(float));
		float *im_sobel = malloc(w*h*sizeof(float));
\end{lstlisting}
\vspace{1ex}
\normalsize
	For each method, two images are obtained (one for each operator). Then the gradient magnitude image is constructed using $M=\sqrt{g_x^2+g_y^2}$. \\
	Also the absolute maximum value of the constructed images is computed, for each method. \\
\small
\begin{lstlisting}
		int i,j, fila, col;
		double max_r = 0;
		double max_p = 0;
		double max_s = 0;
		int imax = w*h;
		for (i=0;i<imax;i++){
			fila = (int)(i/w);
			col = i - w*fila + 1;
			fila += 1;
			j = col + (w+2)*fila;
			// Max in each case
			im_roberts[i] = sqrt(im_r1[j]*im_r1[j] + im_r2[j]*im_r2[j]);
			im_prewitt[i] = sqrt(im_p1[j]*im_p1[j] + im_p2[j]*im_p2[j]);
			im_sobel[i] = sqrt(im_s1[j]*im_s1[j] + im_s2[j]*im_s2[j]);
			// Absolute max
			max_r = MAX(max_r,im_roberts[i]);
			max_p = MAX(max_p,im_prewitt[i]);
			max_s = MAX(max_s,im_sobel[i]);
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Thresholded images of each method are created, using the THRESHOLD macro: \\
\small
\begin{lstlisting}
		float th = atof(argv[2]);
		for (i=0;i<imax;i++){
			im_roberts[i] = THRESHOLD(im_roberts[i],th*max_r);
			im_prewitt[i] = THRESHOLD(im_prewitt[i],th*max_p);
			im_sobel[i] = THRESHOLD(im_sobel[i],th*max_s);
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Save output image (using \textit{iio}): \\
\small
\begin{lstlisting}
		iio_save_image_float_vec("roberts.png", im_roberts, w, h, 1);
		iio_save_image_float_vec("prewitt.png", im_prewitt, w, h, 1);
		iio_save_image_float_vec("sobel.png", im_sobel, w, h, 1);
\end{lstlisting}
\vspace{1ex}
\normalsize

### 2012-07-23 test_haralick.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

% Please do NOT edit this file.
% It has been automatically generated
% by a perl script from the original cxx sources
% in the Insight/Examples directory

% Any changes should be made in the file
% src/test_haralick.c

The source code for this section can be found in the file \verb|test_haralick.c|.\\  Haralick edge detectors, main C file.\\

  Parameters:
	\begin{itemize}
		\item \texttt{input\_image} - Input image.
		\item \texttt{rhozero} - Threshold for the Haralick condition $|\frac{C_2}{2C_3}|\leq\rho_0$.
		\item \texttt{padding\_method} - Padding method flag (in convolution): 0 means zero-padding, 1 means image boundary reflection.
		\item \texttt{output\_image} - Output image (edges).
	\end{itemize}

	Includes:
\small
\begin{lstlisting}
	#include "iio.c"
	#include "2dconvolution.c"
	#include <time.h>
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Main function} \\
\small
\begin{lstlisting}
int main(int argc, char *argv[]) {
\end{lstlisting}
\vspace{1ex}
\normalsize
	Load input image (using \textit{iio}): \\
\small
\begin{lstlisting}
		int w, h, pixeldim;
		float *im_orig = iio_read_image_float_vec(argv[1], &w, &h, &pixeldim);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Grayscale conversion (if necessary): explained in \ref{app:marr-hildreth}. \\ \\
	Masks calculated by 2-d fitting (using LS) with the function: 
	$$
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6*y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3 \\
	$$
\small
\begin{lstlisting}
		double masks[10][25] = { {   425,   275,  225,  275,  425,   
									 275,   125,   75,  125,  275,   
                                     225,    75,   25,   75,  225,   
									 275,   125,   75,  125,  275,   
									 425,   275,  225,  275,  425},
								 { -2260,  -620,    0,  620,  2260, 
								   -1660,  -320,    0,  320,  1660, 
								   -1460,  -220,    0,  220,  1460,
                                   -1660,  -320,    0,  320,  1660, 
                                   -2260,  -620,    0,  620, 2260},
								// matrix continues, too large to display in documentation.
\end{lstlisting}
\vspace{1ex}
\normalsize
	Initialise edge image using \texttt{calloc}:
\small
\begin{lstlisting}
		float *edges = calloc(w*h,sizeof(float));
\end{lstlisting}
\vspace{1ex}
\normalsize
	Padding: a larger auxiliar image \texttt{aux} is required to compute the coefficients $k_1$ to $k_{10}$ in every pixel of the original image. \\
	Two different methods are implemented: zero-padding (\texttt{padding\_method}$=0$) and reflection of original image (\texttt{padding\_method}$=1$).
\small
\begin{lstlisting}
		int wx = (w+8);
		int hx = (h+8);
		double *aux = calloc(wx*hx,sizeof(double));
		int fila,col;
		int imax = wx*hx;
		if (padding_method == 0) {
			for(i=0;i<imax;i++){
				fila = (int)(i/wx);
				col = i-(wx*fila);	
				if ( (fila>=4)&&(col>=4)&&(fila<h+4)&&(col<w+4) ) {
					aux[i] = im[(col-4)+(w*(fila-4))];
				}
			}
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
\small
\begin{lstlisting}
		if (padding_method == 1) {
			int fila_refl, col_refl;
			for(i=0;i<imax;i++){
				fila = (int)(i/wx);
				col = i-(wx*fila);
				if (fila<4) {
					fila_refl = 7 - fila;
					if (col<4) { //zone1
						col_refl = 7 - col;
					} else if (col<w+4) {	//zone2
						col_refl = col;
					} else { //zone3
						col_refl = 2*w + 7 - col;
					}
				} else if (fila<h+4) {
					fila_refl = fila;
					if (col<4) { //zone4
						col_refl = 7 - col;
					} else if (col<w+4) { //image
						col_refl = col;
					} else { //zone5
						col_refl =  2*w + 7 - col;
					}
				} else {
					fila_refl = 2*h + 7 - fila;
					if (col<4) { //zone6
						col_refl =	7 - col;
					} else if (col<w+4) {	//zone7
						col_refl = col;
					} else { //zone8
						col_refl =  2*w + 7 - col;
					}
				}
				aux[i] = im[(col_refl-4)+(w*(fila_refl-4))];
			} //for
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Haralick algorithm: coefficients $k_1$ to $k_{10}$ are computed in every pixel of the original image 
	(using the function \texttt{get\_neighbors\_offset} to get the index offsets of the neighbor pixels and 
	the function \texttt{get\_neighborhood} to get the neighborhood of a pixel using those index offsets). 
	Once the coefficients are calculated, are computed
	$$
	C_2 = \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}
	$$
	and
	$$
	C_3 = \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3},
	$$
	and then the edge condition is evaluated in every pixel; $|\frac{C_2}{2C_3}|\leq\rho_0$. \\
\small
\begin{lstlisting}
		int i_zp, u, v, num_edges;
		num_edges = 0;
		double k[10];
		int *offsets = get_neighbors_offset(wx, 5);
		double acum;
		double C2, C3, denom, sintheta, costheta;
		for(fila=0;fila<h;fila++){
			for(col=0;col<w;col++){
				i = col + w*fila;				// original image & edges image index
				i_zp = (col+4) + wx*(fila+4);	// padded image index
				double *neighborhood = get_neighborhood(aux, i_zp, 5, offsets);
				// k1 to k10 (note: k1 (u=0) is not necessary)
				for(u=0;u<10;u++){
					acum = 0;
					for(v=0;v<25;v++){
						acum += neighborhood[v]*masks[u][v];
					}
					k[u] = acum;
				}
				// compute C2 and C3
				denom = sqrt( k[1]*k[1] + k[2]*k[2] );
				sintheta = - k[1] / denom;
				costheta = - k[2] / denom;
				C2 = k[3]*sintheta*sintheta + k[4]*sintheta*costheta + k[5]*costheta*costheta;
				C3 = k[6]*sintheta*sintheta*sintheta + k[7]*sintheta*sintheta*costheta +
					 k[8]*sintheta*costheta*costheta + k[9]*costheta*costheta*costheta;
				//if ((fabs(C2 / (3*C3))<=rhozero)&&(C3<=0)) {
				if ((fabs(C2 / (3*C3))<=rhozero)) {
					edges[i] = 255;
					num_edges += 1;
				}
				// free neighborhood
				free_neighborhood(neighborhood);
			}
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Save output image (using \textit{iio}): \\
\small
\begin{lstlisting}
		iio_save_image_float_vec(argv[4], edges, w, h, 1);
\end{lstlisting}
\vspace{1ex}
\normalsize

### 2012-07-23 test_mh.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

% Please do NOT edit this file.
% It has been automatically generated
% by a perl script from the original cxx sources
% in the Insight/Examples directory

% Any changes should be made in the file
% src/test_mh.c

The source code for this section can be found in the file \verb|test_mh.c|.\\  Marr-Hildreth edge detector, main C file.\\

  Parameters:
	\begin{itemize}
		\item \texttt{input\_image} - Input image.
		\item \texttt{sigma} - Standard deviation $\sigma$ of the Gaussian kernel.
		\item \texttt{n} - Size $n$ of the Gaussian kernel ($n\times n$).
		\item \texttt{tzc} - Threshold in the Zero-Crossing calculation ($0\leq t_{zc}\leq 1$).
		\item \texttt{padding\_method} - Padding method flag (in convolution): 0 means zero-padding, 1 means image boundary reflection.
		\item \texttt{output\_image} - Output image (edges).
	\end{itemize}

	Includes:
\small
\begin{lstlisting}
	#include "iio.c"
	#include "gaussian_kernel.c"
	#include "2dconvolution.c"
	#include <time.h>
\end{lstlisting}
\vspace{1ex}
\normalsize
	\vspace{0.5cm}
	\Large{Main function} \\
\small
\begin{lstlisting}
int main(int argc, char *argv[]) {
\end{lstlisting}
\vspace{1ex}
\normalsize
	Load input image (using \textit{iio}): \\
\small
\begin{lstlisting}
		int w, h, pixeldim;
		float *im_orig = iio_read_image_float_vec(argv[1], &w, &h, &pixeldim);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Grayscale conversion (if necessary): \\
	First is allocated the memory for the grayscale image \texttt{im}, with
	the corresponding correct allocation check. Then the number of channels of the image is checked: if
	\texttt{pixeldim}=3, RGB image is assumed and conversion is needed, else, single channel image (grayscale) is assumed and
	no conversion is required.\\
	The computation of the gray intensity from RGB levels is:
	$$
	I = \frac{6968\times (\text{float})R + 23434\times (\text{float})G + 2366\times (\text{float})B}{32768} \\
	$$
  To perform this operation, a cast is made to float on the values ​​$R$, $G$ and $B$ of the original image. This can be slow, 
  but ensures the correct image conversion. \\
  These coefficients also ensure there is no saturation in the calculation, since $\frac{6968\times 255 + 23434\times 255 + 2366\times 255}{32768} = \frac{8355840}{32768} = 255$.
\small
\begin{lstlisting}
		double *im = malloc(w*h*sizeof(double));
		if (im == NULL){
			fprintf(stderr, "Out of memory...\n");
			exit(EXIT_FAILURE);
		}
		int z;
		int zmax = w*h;	
		if (pixeldim==3){
			for(z=0;z<zmax;z++){
				im[z] =  (double)(6968*im_orig[3*z] + 23434*im_orig[3*z + 1] 
													+ 2366*im_orig[3*z + 2])/32768;
			}
			fprintf(stderr, "images converted to grayscale\n");
		} else {
			for(z=0;z<zmax;z++){
				im[z] = (double)im_orig[z];
			}
			fprintf(stderr, "images are already in grayscale\n");
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Generate Gaussian kernel using the \texttt{gaussian\_kernel} function in \texttt{gaussian\_kernel.c}: \\
\small
\begin{lstlisting}
		double *kernel = gaussian_kernel(n,sigma);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Smooth input image with the Gaussian kernel previously generated, using the \texttt{conv2d} function in \texttt{2dconvolution.c}: \\
\small
\begin{lstlisting}
		double *im_smoothed = conv2d(im, w, h, kernel, n, padding_method);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Computation of the Laplacian of the smoothed image: \\
	A $3\times 3$ approximation of the laplacian operator is used: \\
	$$
	\begin{bmatrix}
		1 &  1 & 1 \\
		1 & -8 & 1 \\
		1 &  1 & 1 
	\end{bmatrix}
	$$
	Using the \texttt{conv2d} function, the \texttt{laplacian} image is obtained. \\
\small
\begin{lstlisting}
		double operator[9] = {1, 1, 1, 1, -8, 1, 1, 1, 1};
		double *laplacian = conv2d(im_smoothed, w+n-1, h+n-1, operator, 3, padding_method);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Now the maximum absolute value of the \texttt{laplacian} image is calculated. This value is required
	for thresholding in zero-crossing calculation. \\
\small
\begin{lstlisting}
		double max_l = 0;
		int p;
		int pmax = (w+n+1)*(h+n+1);
		for (p=0;p<pmax;p++){
			if (abs(laplacian[p])>max_l){
				max_l = abs(laplacian[p]);
			}
		}
\end{lstlisting}
\vspace{1ex}
\normalsize
	Zero-crossing: \\
	The image is explored, looking in every pixel a change of sign between neighboring opposite pixels.
 	In every pixel $p$ the funcion \texttt{get\_neighborhood} (in \texttt{2dconvolution.c}) is used to get the
	9 pixels of its neighborhood:
	$$
	\begin{bmatrix}
		p_{up,left}		& p_{up,middle}		& p_{up,right}		\\
		p_{middle,left}	& p					& p_{middle,right}	\\
		p_{down,left}	& p_{down,middle}	& p_{down,right}	
	\end{bmatrix}
	$$
	Then the pixel $p$ is marked as edge pixel if it occurs that:
	\begin{itemize}
	\item $sign(p_{up,left}) \neq sign(p_{down,right})$, or 
	\item $sign(p_{up,middle}) \neq sign(p_{down,middle})$, or 
	\item $sign(p_{up,right}) \neq sign(p_{down,left})$, or
	\item $sign(p_{middle,left}) \neq sign(p_{middle,right})$. \\
	\end{itemize}
\small
\begin{lstlisting}
		float *zero_cross = calloc(w*h,sizeof(float));		
						// this image will only content values 0 and 255
						// but float type is used for saving using iio.
		if (zero_cross == NULL){
			fprintf(stderr, "Out of memory...\n");
			exit(EXIT_FAILURE);
		}
		int ind_en_lapl, fila, col;
		int *offsets = get_neighbors_offset(w+n+1, 3);
		pmax = w*h;
		int dif_fila_col = (n+1)/2;
		for (p=0;p<pmax;p++){
			fila = ((int)(p/w));
			col = p-(w*fila) + dif_fila_col;
			fila += dif_fila_col;
			ind_en_lapl = col + (w+n+1)*fila;
			double *n3 = get_neighborhood(laplacian, ind_en_lapl, 3, offsets);
			if ((n3[3]*n3[5]<0)&&(abs(n3[3]-n3[5])>(tzc*max_l))) {
				// horizontal sign change
				zero_cross[p] = 255;
			} else if ((n3[1]*n3[7]<0)&&(abs(n3[1]-n3[7])>(tzc*max_l))) {
					// vertical sign change
					zero_cross[p] = 255;
				} else if ((n3[2]*n3[6]<0)&&(abs(n3[2]-n3[6])>(tzc*max_l))) {
						// +45deg sign change
						zero_cross[p] = 255;
					} else if ((n3[0]*n3[8]<0)&&(abs(n3[0]-n3[8])>(tzc*max_l))) {
							// -45deg sign change
							zero_cross[p] = 255;
						}
			free_neighborhood(n3);
		}
		free_neighbors_offsets(offsets);
\end{lstlisting}
\vspace{1ex}
\normalsize
	Save output image (using \textit{iio}): \\
\small
\begin{lstlisting}
		iio_save_image_float_vec(argv[6], zero_cross, w, h, 1);
\end{lstlisting}
\vspace{1ex}
\normalsize
	\textit{Note: the main function in \texttt{test\_mh\_log.c} is essentially the same. The only difference is that a LoG kernel is generated (instead a Gaussian kernel) using the
	\texttt{LoG\_kernel} function in \texttt{gaussian\_kernel.c}. Therefore there is no need to use the laplacian operator, so only one convolution is made.} \\

### 2012-07-23 compile_article.sh
new file:   compile_article.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+
+EXPECTED_ARGS=1
+
+if [ $# -ne $EXPECTED_ARGS ]
+then
+  echo "Usage: `basename $0` {arg}"
+  exit 1
+fi
+
+version=$1
+
+filename_tex="edge-detection-review-v$version.tex"
+filename_aux="edge-detection-review-v$version.aux"
+
+pdflatex $filename_tex
+bibtex $filename_aux
+pdflatex $filename_tex
+pdflatex $filename_tex
+
+
+rm *.aux
+rm *.bbl
+rm *.blg
+rm *.log
+rm *.out

### 2014-05-07 review/review_answer_2.tex
added review, figure sources and jfc.sty package

\documentclass[a4paper,10pt]{report}

\usepackage[english]{babel}
%\usepackage[none]{hyphenat}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[pdftex]{hyperref}
%\usepackage[tight,scriptsize]{subfigure}
%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}
\usepackage{jfc}
\hypersetup{hypertexnames, plainpages=false,colorlinks=true}
\usepackage{amsmath}

% Title Page
\title{}
\author{}

\usepackage{xr}
\externaldocument{../edge-detection-review-2nd-revision-v1}
\reversemarginpar

\newcounter{myquestion}
\newcommand{\ml}{\mathcal{L}}
\newcommand{\ms}{\mathcal{S}}

\begin{document}
\maketitle

%\begin{abstract}
%\end{abstract}

\chapter{Introduction}
In this communication we answer to the second review of the article "A Contrario Selection of Optimal Partitions for Image Segmentation", submitted to the SIAM Journal on Imaging Sciences.
In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided. In the following we address the issues raised by the reviewers. 
For each of them, an answer is provided and a cross-reference with a red tag Cn (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided. In addition, changes made to the paper are highlighted in red (when significative) in the new version of the paper. 
These red marks will be removed for the final version.

The remarks from the reviewers are marked with the tag \textcolor{blue}{QUESTION} and our answers with the tag \ans.

%\section{Brief summary of changes}
\section{General changes}
\begin{itemize}
 \item Changed the misstyped acronym IOS for OIS in some tables.
\item We added an appendix where we analize the complexity of the algorithm.
\item Removed several typos and orthographic errors.
\item Added a missing hypothesis for Theorem \ref{theo:meaningful_partitions:optimality}.
\end{itemize}

\chapter{Answers to reviewer \#1}

\section{Summary}

\que This paper reviews classic edge detectors based on the rst and second derivatives of the image. It focuses
on three first-derivative based algorithms (namely Sobel, Prewitt and Roberts) and two second-derivative
based algorithms (Marr-Hildreth and Haralick). The paper itself explains all the implementation details.
The code provides the corresponding implementation.\\

\section{Paper review}

\que In general the paper provides all the information necessary. Some parts need nevertheless rewriting either because of minor mistakes (that make some sentences really hard to read) or because of mathematical inaccuracies.\\

\ans We thank the reviewer for the comments and for contributing to improve the quality of the paper. We did a major proof-reading of the paper and we did our best effort to improve the writing\\

\que You mention a video but it was not included in the supplementary les (35-757-1-SP.zip and 35-758-1-SP.zip) so that I could not see it.\\

\ans XXX

\que p2: There is no need to define x = ih and y = ih, you should take x = i; y = j, and then add a small h, which will be equal to 1 since you are on a grid. In theory the approximation is valid for h small, thus it is really questionable to define x = ih? \\

\ans This is a confusion, we should present the taylor approximation of the derivative using h, and then define the grid sizes as $\Delta x$ and $\Delta y$, and thus define $x=i\Delta x$. This is done in Correcction \ref{correction_a:first_derivative:discretization}. Next paragraph was also rewritten, see \refc{correction_a:first_derivative:diagonal edges}\\

\que p4: The computational diference between Prewitt and Sobel masks is not important" -> You need to justify this sentence. Show and refer to an experiment of some kind. Otherwise this is an unfounded claim. \\

\ans We think we did not explain ourselves correctly. We meant that the computational cost of both filters is the same, thus we are free to pick the more accurate one. To claim that is better to filter with a gaussian filter than a moving average seems to be a reasonable claim.\\

\section{minor remarks}

\que In general I would recommend proofreading the whole paper, there are a lot of grammar
mistakes that make the paper hard to follow, I noted some of them below but this list is not exhaustive at all.\\

\ans We thank the reviewer for the comment, most of the paper has been revised thoroughly. Specially the points remarked in this section.\\

\que Algorithms based on first derivative \\
\ans The title of the section and the first paragraph of it was rewritten to improve grammar. See Correcction \ref{correction_a:first_derivative:first_deriv_rephrase}\\

\que Taylor expansion of fi	rst order with a small h

\que p3 diferences

\ans Done in Correction \ref{correction_b::remember_eqs}. Thanks for the suggestion.\\

\ans The abstract data structure used is a priority queue implemented with a binary heap, where the cost of modification of an items is $\Theta(\log n)$.
This is now made more explicit in the corresponding section. See correction \ref{correction_b:greedy:complexity}.\\

\que Section V.D, on the complexity of the MP algorithm. The one-pass 
from the root to the leaves has a linear complexity w.r.t. the number 
of nodes. But, for each node, the possible combination of i + j = k 
need to be analyzed. Consequently, this should impact also the 
complexity. Have I missed something? \\

\ans This is correct, the analysis of the complexity of the combinations of i,j such that i+j=k was missing and we discovered that it was more complex than we though. Please see the newly added Appendix \ref{sec:appendix:complexity} where a detailed analysis is included (Correction C\ref{correction_b:appendix:complexity}). We also added some conclusions regarding that issue \ref{correction_b:conclusions:complexity}.\\

\que In Section VI, I suppose that the results for other benchmark 
approaches are obtained from the paper of arbelaaez.2011.pami. But it 
seems that in the paper of Arbelaez, the method of "Multiscale 
algorithm for image segmentation" by Koepfler et al. (denoted as MS in 
the current paper) is not included. What's the source of this result? \\

\ans No they are not from arbelaaez.2011.pami. In fact, those results are taken from Arbelaez's papers published in 2006 and 2009.
In particular, the results of Koepfler and also from Watersheds are presented in the 2006 paper (see \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}).
They are not shown in a table, but in a figure. The references were already there, in the header of Table X, but we also repeated them in the caption, for the sake of clarity.
In addition, we added the numbers of the tables and figures from where we extracted the info.\\

\que In Table XIII, what's the method NS? i guess it is the Mean Shift 
method. Because the result of NS is exactly the same as the one given 
by Mean Shift presented in the aforementioned Arbelaez's paper. And in 
Table XIV, the result given by the method of MS is exactly the same as 
the one of Mean Shift in Arbelaez's paper. So i am wondering if here 
MS actually stands for Mean Shift rather than the method of Koepfler 
as stated in the current paper. Besides, the same result between the 
method of MS and CM (stands for Mean Shift in this paper) in Table X 
confirms that suspicion. On the other hand, I would guess that MS 
stands for MultiScale (Mumford-Shah) this is indeed Koepfler's 
paper. Please clarify. \\

\ans In the first place we must apologize, we mixed up the notations, mainly due to the long period between both versions. NS and MS in this Table and in the following, meant Mean Shift as the reviewer correctly points out.
In every place that MS appeared we actually meant Mean Shift, the only exception was Table X. We corrected everything, so now MS really means Mumford-Shah (Koepfler) and CM means Mean Shift.
However we must remark that MS and CM actually give the same result (0.63) just by coincidence, that is not a mistake. We re-checked that with the references.\\

\que Numbering of the Tables XI, XII, XIII, XIV. I believe it is better 
to change the position of table XII and XIV to clearly classify the 
boundary benchmark and region benchmark. Or change the position of 
table XIII and XIV to obtain the same order (boundary then region) as 
it is the case for table XI and XII. \\

\que Done. Thanks for the suggestion.\\

\que Fig. 11 and 12. It is better to inverse the saliency maps (ie, the 
darker a contour is, the more important it is). On a screen, this is 
fine, but on a white sheet of paper (or on a white pdf), the inversion 
provides a better visualisation.

\ans Done. Thanks for the suggestion.\\

\que Page 22, first column, the authors suggest using the gPb contour 
detector as the initial segmentation. This would seem a good 
idea. However, I can share some experience here, and I have to say 
that hierarchies built from gPb by Arbelaez contain few levels. It is 
thus very difficult to "enhance" them by some manipulation, as 
proposed in this paper and in other papers. In a way, the gPb 
detector, being a learned gradient, may be too good for such a line of 
work. Besides, it is slow. \\

\ans We thank the reviewer for this suggestion. For this paper we prefer to maintain the proposed approach but we will take note for our future work.\\

\section{Typos:}

\que Page 2, last line of second column: "of this ideas" -> of these ideas\\

\ans Corrected. Thank you.\\

\que Page 3, second line of first column: "contribution of this work" -> contributions of this work \\

\ans Corrected. Thank you.\\

\que Page 10, just before theorem 1: "in the following conjecture". I 
guess it is not anymore a conjecture, unless your proof does not 
stand. \\

\ans Changed conjecture for theorem. Thanks.\\

\section{ In the bibliography }
\que Ref 3 can be replaced by the more complete \verb+\cite{5427030}+
\begin{verbatim}
@article {5427030, 
title = {Region Merging Techniques Using Information Theory Statistical Measures}, 
journal = {Image Processing, IEEE Transactions on}, 
volume = {19}, 
number = {6}, 
year = {2010}, 
month = {june}, 
pages = {1567 -1586}, 
keywords = {Algorithms, Automated, Computer Simulation, Computer-Assisted, Data Interpretation, error consistency indicators, image color homogeneity, image colour analysis, Image Enhancement, Image Interpretation, image region analysis, image segmentation, image texture, image texture homogeneity, Information Theory, information theory statistical measures, Models, nonparametric region models, object-oriented segmentation, partition significance index, Pattern Recognition, Reproducibility of Results, scale-based merging order, Sensitivity and Specificity, Statistical, statistical analysis, Subtraction Technique, texture image segmentation, unsupervised region merging techniques}, 
issn = {1057-7149}, 
doi = {10.1109/TIP.2010.2043008}, 
author = {Calderero, F. and Marques, F.} 
\end{verbatim}

\ans Done. Thanks for the updated information.\\

\que ref. 6 and 38 can efficiently be merged into a single reference, without loosing precision. \\

\ans Thank you, we didn't notice they were the same reference. Corrected.\\

\chapter{Answers to reviewer \#2}

\que The authors made a great work at improving the paper and addressing the issues raised during the review. \\
\ans Thanks very much, such an improvement wouldn't be possible without the thorough work of the referees.\\

\que However, I would still suggest some clarification in section V.E, regarding the boundary term discussion in the multi-partition case. In the paper, it seems that the main reason for not including it is to avoid the introduction of a new weighting parameter (which by the way was not needed in the greedy case?). But in the answer to question 31, it seems that it is mostly a computational issue related to the scale-climbing algorithm. \\

\ans We added a comment regarding this on Correction \ref{correction_b:boundary_term:descreasing}.\\

\que Anyway, I don't think this justifies to delay the publication of the paper with an additional review, so I already recommend the paper for publication.\\

\ans Thank you very much.\\

%

% \begin{table}[ht]
% \centering
% \begin{tabular}{|c|c|}
%  \hline
%  	order (k) & LNFA (best $\mathcal{P}$)\\
%  \hline
%  	1 &    \\
%  	. &     \\
%  	. &   \\
%  	. &   \\
%  	$n_1$ &   \\
%  \hline
%  \end{tabular}
% \caption{Best partitions selected by the multipartition (MP) algorithm for the \emph{Church} image.}\label{review:table:implementation:complexity}
% %\vskip \espaciotabla
% \end{table}

\bibliographystyle{unsrt}
\bibliography{../../tools/journal_list_short,../../tools/references,../../tools/hierarchical_segmentation}
\end{document}

### 2014-05-07 edge-detection-review-2nd-revision-v1.tex
added review, figure sources and jfc.sty package

--- 2012-07-23 edge-detection-review-v10.tex
@@
 %\numberwithin{figure}{section}
+\usepackage{marginnote}
+\newcounter{mynote}% a new counter for use in margin notes
+\usepackage{jfc}
 \begin{document}
-\bibliographystyle{ieeetr}
+
 \lstset{
@@
 
-\section{Algorithms based on first derivative}
+\section{Algorithms based on the first derivative}
 \label{sec:first}
 
-Algorithms based on first derivative studied in this paper have a common scheme, with 
-the only difference in the type of filtering. Figure \ref{fig:blockdiagram1} presents
-a block diagram of these algorithms. \\
+All the algorithms based on first derivatives that are studied in this paper share a common structure, the only difference being the type of filtering used to compute those derivatives.
+Figure \ref{fig:blockdiagram1} presents a block diagram of those algorithms.
+\myn{\label{correction_a:first_derivative:first_deriv_rephrase}\\
 
@@
 
-The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with
-a small $h$. Thus $f'(x)$ is computed as
+The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with a small $h$. Thus $f'(x)$ is computed as
 \begin{equation}
@@
 \end{equation}
-The image is composed by discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
-first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
+\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid.
+If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$.
+When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
+Thus, in the discrete space the first derivative of the image is computed as:
 \begin{equation}
@@
 
-Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
-by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. But when finding 
-diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
-attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
-diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
+\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}.
+However, in order to be able to compute diagonal edges, 2-D masks are needed.
+\\
+The \textit{Roberts operators} are one of the earliest 
+attempts to use 2-D masks for this purpose.
+These operators are based on computing the diagonal \tcom{differences} \myn implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
+By convention
 in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
@@
 The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
-by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
-difference between the third and first rows approximates the derivative in the x-direction, and 
-the difference between the third and first columns approximates the derivative in the y-direction.
-These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
-These masks are called \textit{Prewitt operators}. \\
+by taking the difference between the third and first rows (or columns) of the $3\times3$ region.
+The difference between the third and first rows approximates the derivative in the x-direction, and the difference between the third and first columns approximates the derivative in the y-direction.
+These approximations can be implemented by filtering the image with the two masks shown in Figure \ref{fig:prewitt}. Those are called \textit{Prewitt operators}. \\
 
@@
 	\includegraphics[width=0.25\textwidth]{prewitt_n.pdf}
-	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
+	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$. }
 	\label{fig:prewitt}
@@
 
-These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
-smoothing properties in the direction opposite to the gradient.
+\myn{\label{correction_a:first_derivative:sobel} The left mask can be constructed (beyond normalization) as the convolution of a vertical derivative filter
+$[-1\ 0\ 1]$ with an horizontally moving average filter $[1\ 1\ 1]$.
+The same reasoning could be applied to the right mask.
+Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter.
+Thus, they show a smoothing behaviour in the direction orthogonal to the gradient.
 
@@
 
-A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
+\textit{Sobel operators} are a slight variation of the Prewitt operators which use more weight on the central coefficients of the 
 difference. It can be shown that using a $2$ in the center location provides better image smoothing. This
-variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
-\textit{Sobel operators}. \\
+variation is implemented using masks in Figure \ref{fig:sobel}.
 
@@
 %-------------------------------------------------------------------------------
-
+\bibliographystyle{siam}
 \bibliography{bibliography}

### 2014-05-07 kernels.m

@@ -1,0 +1,79 @@
+% GAUSSIAN & LoG KERNELS
+clear all;
+close all;
+
+n1 = 25;
+n2 = 31;
+sigma = 4;
+
+x1 = zeros(n1,n1); y1 = zeros(n1,n1);
+x2 = zeros(n2,n2); y2 = zeros(n2,n2);
+
+for i = 1:n1
+    x1(:,i) = i-floor(n1/2)-1;
+    y1(i,:) = i-floor(n1/2)-1;
+end
+for i = 1:n2
+    x2(:,i) = i-floor(n2/2)-1;
+    y2(i,:) = i-floor(n2/2)-1;
+end
+
+G = zeros(n1,n1);
+LoG = zeros(n2,n2);
+
+for i = 1:n1
+    for j = 1:n1
+        G(i,j) = exp(-(x1(i,j)^2 + y1(i,j)^2)/(2*sigma^2));
+    end
+end
+for i = 1:n2
+    for j = 1:n2
+        LoG(i,j) = -((x2(i,j)^2 + y2(i,j)^2 - 2*sigma^2)/(sigma^4))*exp(-(x2(i,j)^2 + y2(i,j)^2)/(2*sigma^2));
+    end
+end
+
+
+%xmin = -floor(n/2)-3;
+%xmax = floor(n/2)+3;
+
+a = figure('Name','Gaussian Kernel','Color','white');
+h1 = subplot(2,3,[1 2 4 5]); surf(x1,y1,G); grid off; axis tight; axis off;
+    hold on; plot3(zeros(n1,1),y1(:,1),G(ceil(n1/2),:),'-k','LineWidth',3); 
+    text(0,0,1.1,'Gaussian kernel, 3D view','HorizontalAlignment','center'); hold off;
+    p = get(h1, 'pos'); 
+    p(1) = p(1) - 0.1; p(2) = p(2) + 0.05;
+    p(3) = p(3) + 0.1; p(4) = p(4) + 0.1;
+    set(h1, 'pos', p);
+h2 = subplot(2,3,3); pcolor(x1,y1,G); axis square; axis off;
+    p = get(h2, 'pos'); 
+    p(1) = p(1) - 0.05; p(2) = p(2) - 0.05;
+    p(3) = p(3) + 0.1; p(4) = p(4) + 0.1;
+    set(h2, 'pos', p);
[…]

### 2014-05-07 kernels1d.m

@@ -1,0 +1,54 @@
+% GAUSSIAN & LoG KERNELS - 1D comparsion
+clear all;
+close all;
+
+sigma = 4;
+
+x = -1*sigma:0.01:5*sigma;
+G = zeros(size(x));
+LoG = zeros(size(x));
+for i = 1:length(x)
+    G(i) = exp(-(x(i)^2)/(2*sigma^2));
+    LoG(i) = -((x(i)^2 - 2*sigma^2)/(sigma^4))*exp(-(x(i)^2)/(2*sigma^2));
+end
+
+G = G/max(G);
+LoG = LoG/max(LoG);
+
+factor = 0.01;
+% ind_G e ind_LoG por amplitud:
+% ind_G = find(abs(G)>factor*max(G),1,'last');
+% ind_LoG = find(abs(LoG)>factor*max(LoG),1,'last');
+
+area_G = cumtrapz(x,abs(G)); area_G = area_G/max(area_G);
+area_LoG = cumtrapz(x,abs(LoG)); area_LoG = area_LoG/max(area_LoG);
+% ind_G e ind_LoG por area:
+ind_G = find(area_G>=(1-factor),1,'first');
+ind_LoG = find(area_LoG>=(1-factor),1,'first');
+
+h = figure('Name','Kernels comparsion','Color','white','Position',[100 100 800 500]);
+plot(x,G,'-r','Linewidth',2); hold on; 
+plot(x,LoG,'-b','Linewidth',2); axis tight;
+title('Comparison of Gaussian and LoG functions');
+plot(x(ind_G),G(ind_G),'*r','Linewidth',5);
+plot(x(ind_LoG),LoG(ind_LoG),'*b','Linewidth',5);
+% Líneas:
+line([0 0],[min(LoG) max(LoG)],'Color','k');
+line([-1*sigma 5*sigma],[0 0],'Color','k');
+line([x(ind_G) x(ind_G)],[min(LoG) max(LoG)],'Color','r','LineStyle','--');
+line([x(ind_LoG) x(ind_LoG)],[min(LoG) max(LoG)],'Color','b','LineStyle','--');
+h2 = get(h,'Children');
+set(h2,'XTick',[]); set(h2,'YTick',[]);
+line([-1*sigma 5*sigma],[factor*max(LoG) factor*max(LoG)],'Color','k','LineStyle',':');
[…]

### 2014-05-09 review/review_answer_2.tex
Added text comment to comply with review #2

@@
 \begin{itemize}
- \item Changed the misstyped acronym IOS for OIS in some tables.
-\item We added an appendix where we analize the complexity of the algorithm.
+ \item Revised general writing of the article.
+\item We added appendices to include secondary information that distracts from the main purpose of the article.
+\item In that sense we moved the numeric values of the $\mathbf{mask_i}$ matrices to the first appendix
+\item We also moved the Implementation section, which documents the code, to the second appendix.
 \item Removed several typos and orthographic errors.
-\item Added a missing hypothesis for Theorem \ref{theo:meaningful_partitions:optimality}.
 \end{itemize}
@@
 
-\que This paper reviews classic edge detectors based on the rst and second derivatives of the image. It focuses
+\que This paper reviews classic edge detectors based on the first and second derivatives of the image. It focuses
 on three first-derivative based algorithms (namely Sobel, Prewitt and Roberts) and two second-derivative
@@
 
-\ans We thank the reviewer for the comments and for contributing to improve the quality of the paper. We did a major proof-reading of the paper and we did our best effort to improve the writing\\
+\ans We thank the reviewer for the comments and for contributing to improve the quality of the paper. We did a major proof-reading of the paper and we did our best effort to improve the writing.\\
 
@@
 We meant that the computational cost of both filters is the same, thus we are free to pick the more accurate one.
-To claim that is better to filter with a gaussian filter than a moving average seems to be a reasonable claim.\\
+To claim that is better to filter with a gaussian filter than a moving average is a well established practice in the signal processing community.\\
+
+\que Algorithm 2: lines 2 and 3 need to be moved inside the else statement (between lines 7 and 8) otherwise you are computing something that is not always used.\\
+
+\ans Fixed  in \refc{correction_a:second_derivative:marr_algo}.\\
+
+\que p10: the $b_i$ you are defining are referred as $mask_i$ in algorithm 3. It is actually better to name them
+$mask_i$ to avoid confusion with $b_{i,j}$ . It is also inaccurate to state where $bi =...$ since you have not
+used the $b_i$ yet but the $b_{i,j}$ . It makes the whole thing very hard to read.\\
+
+\ans While we agree on making the notation coherent, we prefer to keep the name $b_i$.
+Naming a matrix $B$, its rows as $\mathbf{b_i}$ and its elements $b_{i,j}$ it is a quite common (and unambiguous) way to use the notation.
+However we also added $\mathbf{mask_i}$ and defined it as being $\mathbf{b_i}$ reordered.
+Thus, we kept algorithm 3 using $mask_i$ (see \refc{correction_a:second_derivative:haralick_algo}), we defined $\mathbf{mask_i}$ and we used $\mathbf{b_i}$ before saying where.
+Finally, we stated explicitely that $\mathbf{b_i}$ is a vector containing the $i_{th} $ row of matrix $B$ (See \refc{correction_a:second_derivative:haralick_bi}).
+
+\que Can you add an experiment showing the influence of the padding method on the results?\\
+
+\que Use a "Name +year" bibliography as asked in the IPOL writing guidelines.\\
+4 points to ipol, In addition to that it is not referred to in the text.\\
+
+\ans This was corrected in two ways. As new IPOL guidelines suggest, the SIAM bibliography style was applied. Reference [4] now points to the correct article.\\
 
@@
 \que Algorithms based onfirst derivative \\
-\ans The title of the section and the first paragraph of it was rewritten to improve grammar. See Correcction \ref{correction_a:first_derivative:first_deriv_rephrase}\\
+\ans The title of the section and the first paragraph of it was rewritten to improve grammar. See  \refc{correction_a:first_derivative:first_deriv_rephrase}\\
 
-\que Taylor expansion of fi	rst order with a small h
+\que Taylor expansion of first order with a small h\\
 
-\que p3 diferences
+\ans Rephrased in \refc{correction_a:first_derivative:taylor}\\
 
-\ans Done in Correction \ref{correction_b::remember_eqs}. Thanks for the suggestion.\\
+\que p3 diferences\\
 
-\ans The abstract data structure used is a priority queue implemented with a binary heap, where the cost of modification of an items is $\Theta(\log n)$.
-This is now made more explicit in the corresponding section. See correction \ref{correction_b:greedy:complexity}.\\
+\ans Fixed in \refc{correction_a:typo}. \\
 
-\que Section V.D, on the complexity of the MP algorithm. The one-pass 
-from the root to the leaves has a linear complexity w.r.t. the number 
-of nodes. But, for each node, the possible combination of i + j = k 
-need to be analyzed. Consequently, this should impact also the 
-complexity. Have I missed something? \\
+\que "normalization" do you mean "up to the normalization"\\
 
-\ans This is correct, the analysis of the complexity of the combinations of i,j such that i+j=k was missing and we discovered that it was more complex than we though.
-Please see the newly added Appendix \ref{sec:appendix:complexity} where a detailed analysis is included (Correction C\ref{correction_b:appendix:complexity}).
-We also added some conclusions regarding that issue \ref{correction_b:conclusions:complexity}.\\
+\ans Corrected in \ref{correction_a:first_derivative:sobel} and \ref{correction_a:first_derivative:sobel}.\\
 
-\que In Section VI, I suppose that the results for other benchmark 
-approaches are obtained from the paper of arbelaaez.2011.pami. But it 
-seems that in the paper of Arbelaez, the method of "Multiscale 
-algorithm for image segmentation" by Koepfler et al. (denoted as MS in 
-the current paper) is not included. What's the source of this result? \\
+\que "thus It"\\
 
-\ans No they are not from arbelaaez.2011.pami. In fact, those results are taken from Arbelaez's papers published in 2006 and 2009.
-In particular, the results of Koepfler and also from Watersheds are presented in the 2006 paper (see \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}).
-They are not shown in a table, but in a figure. The references were already there, in the header of Table X, but we also repeated them in the caption, for the sake of clarity.
-In addition, we added the numbers of the tables and figures from where we extracted the info.\\
+\ans Fixed\\
 
-\que In Table XIII, what's the method NS? i guess it is the Mean Shift 
-method. Because the result of NS is exactly the same as the one given 
-by Mean Shift presented in the aforementioned Arbelaez's paper. And in 
-Table XIV, the result given by the method of MS is exactly the same as 
-the one of Mean Shift in Arbelaez's paper. So i am wondering if here 
-MS actually stands for Mean Shift rather than the method of Koepfler 
-as stated in the current paper. Besides, the same result between the 
-method of MS and CM (stands for Mean Shift in this paper) in Table X 
-confirms that suspicion. On the other hand, I would guess that MS 
-stands for MultiScale (Mumford-Shah) this is indeed Koepfler's 
-paper. Please clarify. \\
+\que "As mentioned before this leads to ab overdetermied"
 
-\ans In the first place we must apologize, we mixed up the notations, mainly due to the long period between both versions.
-NS and MS in this Table and in the following, meant Mean Shift as the reviewer correctly points out.
-In every place that MS appeared we actually meant Mean Shift, the only exception was Table X. We corrected everything, so now MS really means Mumford-Shah (Koepfler) and CM means Mean Shift.
-However we must remark that MS and CM actually give the same result (0.63) just by coincidence, that is not a mistake. We re-checked that with the references.\\
+\ans Fixed\\
 
-\que Numbering of the Tables XI, XII, XIII, XIV. I believe it is better 
-to change the position of table XII and XIV to clearly classify the 
-boundary benchmark and region benchmark. Or change the position of 
-table XIII and XIV to obtain the same order (boundary then region) as 
-it is the case for table XI and XII. \\
+\que "Gaussian kernel is generated by sampling the"\\
+\que "unlike the case of Gaussian kernel"\\
+\que "are not independent of each one"\\
+\que "implies complete the borders with many zero valued pixels as needed"\\
+\que "It can be seen in gure 16 an area of interest of the image."\\
+\que "appearing unwanted isolated edges"\\
+\que "because they only consists"\\
+\que "more fair"\\
 
-\que Done. Thanks for the suggestion.\\
+\que "runs significative faster"\\
 
-\que Fig. 11 and 12. It is better to inverse the saliency maps (ie, the 
-darker a contour is, the more important it is). On a screen, this is 
-fine, but on a white sheet of paper (or on a white pdf), the inversion 
-provides a better visualisation.
+\ans Fixed\\
 
-\ans Done. Thanks for the suggestion.\\
-
-\que Page 22, first column, the authors suggest using the gPb contour 
-detector as the initial segmentation. This would seem a good 
-idea. However, I can share some experience here, and I have to say 
-that hierarchies built from gPb by Arbelaez contain few levels. It is 
-thus very difficult to "enhance" them by some manipulation, as 
-proposed in this paper and in other papers. In a way, the gPb 
-detector, being a learned gradient, may be too good for such a line of 
-work. Besides, it is slow. \\
-
-\ans We thank the reviewer for this suggestion. For this paper we prefer to maintain the proposed approach but we will take note for our future work.\\
-
-\section{Typos:}
-
-\que Page 2, last line of second column: "of this ideas" -> of these ideas\\
-
-\ans Corrected. Thank you.\\
-
-\que Page 3, second line of first column: "contribution of this work" -> contributions of this work \\
-
-\ans Corrected. Thank you.\\
-
-\que Page 10, just before theorem 1: "in the following conjecture". I 
-guess it is not anymore a conjecture, unless your proof does not 
-stand. \\
-
-\ans Changed conjecture for theorem. Thanks.\\
-
-\section{ In the bibliography }
-\que Ref 3 can be replaced by the more complete \verb+\cite{5427030}+
-\begin{verbatim}
-@article {5427030, 
-title = {Region Merging Techniques Using Information Theory Statistical Measures}, 
-journal = {Image Processing, IEEE Transactions on}, 
-volume = {19}, 
-number = {6}, 
-year = {2010}, 
-month = {june}, 
-pages = {1567 -1586}, 
-keywords = {Algorithms, Automated, Computer Simulation, Computer-Assisted, Data Interpretation, error consistency indicators, image color homogeneity, image colour analysis, Image Enhancement, Image Interpretation, image region analysis, image segmentation, image texture, image texture homogeneity, Information Theory, information theory statistical measures, Models, nonparametric region models, object-oriented segmentation, partition significance index, Pattern Recognition, Reproducibility of Results, scale-based merging order, Sensitivity and Specificity, Statistical, statistical analysis, Subtraction Technique, texture image segmentation, unsupervised region merging techniques},
-issn = {1057-7149}, 
-doi = {10.1109/TIP.2010.2043008}, 
-author = {Calderero, F. and Marques, F.} 
-\end{verbatim}
-
-\ans Done. Thanks for the updated information.\\
-
-\que ref. 6 and 38 can efficiently be merged into a single reference, without loosing precision. \\
-
-\ans Thank you, we didn't notice they were the same reference. Corrected.\\
+\que "polinomial"\\
+\ans Fixed\\
 
@@
 
-\que The authors made a great work at improving the paper and addressing the issues raised during the review. \\
-\ans Thanks very much, such an improvement wouldn't be possible without the thorough work of the referees.\\
+\section{General comments}
+Overall I found the article clear (only the Haralick algorithm explanations are
+confusing). The code is very clear, well documented (although not following the Doxygen conventions) and matches almost perfectly
+the description in the pdf. Finally the submitted demo works as expected with no
+more parameters than the ones presented in the paper. All this makes the
+submission a very good one for Ipol.
 
-\que However, I would still suggest some clarification in section V.E, regarding the boundary term discussion in the multi-partition case.
-In the paper, it seems that the main reason for not including it is to avoid the introduction of a new weighting parameter (which by the way was not needed in the greedy case?).
-But in the answer to question 31, it seems that it is mostly a computational issue related to the scale-climbing algorithm.
-\\
+However, I think that the wording is largely below standard. I have reported
+some of the typos, but beyond that, several sentences are wrongly formulated and
+even entire paragraphs should be restructured. I have proposed some improvements, yet
+I think that the authors should take upon themselves to globally improve the
+quality of the paper in this regard.
 
-\ans We added a comment regarding this on Correction \ref{correction_b:boundary_term:descreasing}.\\
+\section{Comments on the PDF}
+\subsection{General description of the article}
+This article describes classical algorithms for edge detection. First the
+category of first order derivative is described. The only degree of freedom
+considered in this article for this class of algorithms lies in the operator
+used to approximate the derivative. The article includes some details on the
+Robert operators, the Prewitt operators and the Sobel operators. Otherwise the
+algorithms are kept strictly identical and are composed of:
 
-\que Anyway, I don't think this justifies to delay the publication of the paper with an additional review, so I already recommend the paper for publication.\\
+- a filtering to compute the image gradient
+- a thresholding on the magnitude of the gradient to select edges
 
-\ans Thank you very much.\\
+Each step is described within the text with sufficient details and a consistent
+pseudo-code for the entire algorithms is provided at the end of the
+corresponding section.
 
-%
+In a subsequent section, two algorithms based on second order derivatives are
+described: the Marr-Hildreth algorithm as well as the alternative proposed by
+Haralick. Both algorithms are "conceptually" composed of two steps:
 
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{|c|c|}
-%  \hline
-%  	order (k) & LNFA (best $\mathcal{P}$)\\
-%  \hline
-%  	1 &    \\
-%  	. &     \\
-%  	. &   \\
-%  	. &   \\
-%  	$n_1$ &   \\
-%  \hline
-%  \end{tabular}
-% \caption{Best partitions selected by the multipartition (MP) algorithm for the \emph{Church} image.}\label{review:table:implementation:complexity}
-% %\vskip \espaciotabla
-% \end{table}
+- compute the second order derivatives
+- find the zero crossing locations and set them as edges
+
+Once again the two algorithms differs mainly in the way of computing
+derivatives. The former rely on the filtering with laplacian of gaussian or on
+the filtering with a gaussian followed by a laplacian operator. The latter
+relies on a local bicubic approximation followed by a second derivative
+analytical computation. Sufficient details are again provided along with
+algorithmic description. However some of the details could be made clearer (see
+later comments).
+
+The fourth section provides technical details on some filtering aspects and the
+last section is dedicated to the experimental discussion.
+
+\subsection{Detailed comments, typos and typesetting}
+\subsubsection{Demo}
+The url \url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/} provided in the preprint
+contains a broken demo. This is not very important though since the demo code
+provided in the submission works correctly and it should eventually replace the
+one specified in the pdf.
+
+\subsubsection{ Section 2.3: The Sobel operators}
+No upper case on the "i" of "thus It".
+
+\subsubsection{Section 2.4: Computation of the edges}
+The threshold variable "th" should be in italic.
+Missing a "." at the end of the section.
+
+\subsubsection{Algorithm 2:}
+\begin{verbatim}
+tzc" --> "th_{ZC}"
+\end{verbatim}
+
+\subsubsection{ Section 3.2.1: Bi-cubic polynomial fitting}
+
+\que I would put more emphasis on the fact that the approximation is expressed in a
+local frame, so that the pixel coordinates $x_1,y_1\dots,x_{25},y_{25}$ are
+always the same. This is implicitly hinted by the following sentence:
+"e.g. in a neighborhood of size 5 × 5, x and y take values between −2 and 2".
+But I would insist on this fact again at the end, otherwise it is not clear why
+the matrix $B$ can be precomputed once for all.\\
+
+\ans This was taken into account and the entire section was rewritten.
+Now it is said at the begginning and at the end of the section, see Corrections \ref{second_derivative:haralick:statement} and \ref{second_derivative:haralick:mask_computation}.\\
+
+Also, the explanations are badly organized:
+\que  The least-square problem is not "approximately" solved but "EXACTLY" solved with the pseudo-inverse.\\
+
+\ans Agreed. Corrected now.\\
+
+\que $- (A^TA)^{−1}A^T f=k \Rightarrow k=Bf$ should be rephrased as
+$k=Bf$ with $B=(A^TA)^{−1}A^T$\\
+
+\ans We didn't follow exactly the suggestion, but changed the text in the same spirit. See \refc{second_derivative:haralick:normal}.
+
+\que Equation 3.4 and the one just before it should be removed since they are
+simple reformulation of the matrix equation above. At best you can keep the
+notations for the entries of B that is $b_{i,j}$ for $i\in {1,\ldots,10}$ and
+$j \in {1,ldots,25}$.\\
+
+\ans While we agree that there is some superfluous and simple equations, we found that those tricks to speed up are quite hard to follow at first.
+Thus, this level of detail could help the reader to clearly understand whats going on.\\
+
+\que the introduction of $\bf{b_i}$ comes too soon, they corresponds to the masks of table 3.1\\
+
+\ans We rearranged the explanations in this section.
+In particular now $\bf{b_i}$ is well defined, and we introduced a new variable $\mathbf{mask_i}$ which corresponds to the matrix for of $\bf{b_i}$ and postponed the introduction of these masks to a point closer to table 3.1.
+However that table was somehow superfluous and thus we moved it to an appendix, to avoid adding more noise to the text.\\
+
+\que  I would rephrase all the explanations that follows the definition of $\bf{b_i}$
+focusing on the purpose of speed up; I would first explain that for each
+$i\in {1,\ldots,10}$ one can store all the $k_i$'s in an image, and that this
+image can be obtained by a convolution with a mask of the form $\bf{bi}=...$
+And then I would add that each mask $\bf{b_i}$ can be precomputed once for all and are provided in table 3.1.\\
+
+\ans We followed this suggestion, which in our opinion is a great improvement for the presentation. See \refc{second_derivative:haralick:mask_computation} Thanks to the reviewer for this one.
+
+\que Can you explain how you have obtained the entries of the mask? If you used a program for that, I think you should provide it.\\
+
+\ans XXX\\
+
+\subsubsection{Section 3.2.2: Analytical calculation of the second derivative}
+
+\que This description is unclear to me. Does the described procedures allows to
+compute the laplacian operator (under the bicubic approximation)? If yes,
+explain why this is the case. By reading the original paper by Haralick, I
+could determine that the answer is negative, and that the goal is to compute a second derivative in the direction of the gradient.
+The only sentence referring to this purpose is at the very end of the section:
+"Given the direction defined by θ, with ρ > 0, an edge will always be an
+ascending step. This indicates that the first derivative in this direction has
+a maximum and that the second derivative has a zero crossing with negative
+slope. That is, the third derivative less than zero."
+If so I would recommend to put more emphasis on this goal and the justification that I quoted from your text. And you should definitely place the
+corresponding sentences at the very beginning (before giving the mathematical details).\\
+
+\ans Changed as suggested. See \refc{second_derivative:haralick:analitical:statement}.\\
+
+\que Using the expression "polar form" is completely misleading since theta is
+fixed. Haralick simply considers the second derivative in the direction of the
+gradient and there comes the restriction $f_\theta(\rho)$ of the polynomial
+approximant to the line of angle theta.\\
+
+\ans Changed, see \refc{second_derivative:haralick:analitical:derivative}.
+We now explicitely defined the line of angle theta, and described the points lying on that line in terms of $\rho$ and $\theta$\\
+
+\que Equation 3.9 is not mathematically correct. You should add before it, $\exists
+\rho\in[0,\rho_0]$ s.t. Equation 9.\\
+
+\que I would integrate the footnote 7 within the text, just after: "This parameter
+must be greater than zero and less
+than 1." since it provides the rationale behind these constraints.\\
+
+\ans Done, see \refc{second_derivative:haralick:analitical:parameter_value}.\\
+
+You should explain what you mean by the optimal value of $\rho_0$. If you are
+simply referring to Haralick's claims, you should state it explicitly.
+
+Also, are you sure you implement the exact conditions specified in the
+original article by Haralick. The ones I could find are as follows:
+"If for some $\rho, |\rho|\leq\rho_0, [...], f"_\alpha(\rho)<0$ [eq.1],
+$f''_alpha(\rho)=0 [eq.2] and f'(\rho)\neq 0$ [eq.3] we have discovered [...]"
+I guess that [eq.1] was meant to involve a third derivative instead of a second
+derivative. Then it is exacly your condition $c_3<0$. I don't find anything
+close to [eq.3] in your description. Besides, although (as you justified it)
+[eq.2] implies your condition $|\frac{C_2}{3C_3}|\leq\rho_0$ they are not
+equivalent.
+I would say that you should add that $C_2$ must be positive as well.
+If you want to stick with your conditions only, you should at least add a discussion pointing on the fact that they differ from the one formulated by Haralick, and
+describe the differences. I actually tried changing your code to take the
+sign of $C_2$ into account and to introduce [eq.3]: it does not seem to change
+the output. Yet I would recommend that you investigate why and provide that
+information.
+
+\subsubsection{ Section 4: Common Mathematical Operations}
+"All algorithms implemented" --> "All implemented algorithms"
+
+\subsubsection{ Figure 11:}
+"Is clearly required" --> "It ..."
+
+The text in the figure is practically unreadable.
+
+\subsubsection{ Section 5: Results}
+The experimental discussion is quite succinct, I would expect Ipol articles to
+be more in-depth. For instance, it seems important that the impact of the
+algorithms parameters be studied more deeply with dedicated examples and that
+you give some hindsight on the way you choose their values (why did you pick
+$th_zc=0.1$ for one experiment and then 0.13 in the other one). I would
+also expect the claims of the original
+articles to be challenged, confirmed or denied. As an example, Haralick claimed
+that his algorithm performed way better than the one by Marr and Hildreth.
+
+\que You use too many short paragraphs!! This is true in general for your text, but
+it is especially pronounced here. I would recommend one paragraph for the synthetic example and one or two for the windmill example.\\
+
+\ans XXX\\
+
+\que Also, I don't see the point of duplicating the execution time data, once in the figures and another time in a dedicated table.\\
+
+\ans Since we don't have any space restrictions, summarizing all execution times in a table, will help the reader compare them.
+
+I am not entirely convinced by your explanation of the thicker edges in
+Haralick case. There are two reasons form my doubts:
+- My first point is that even with a smooth model, the fact that the detection
+is based on zero-crossing of second derivative should make the detection well localized. Indeed when dealing with a smooth edge, on the contrary to the
+first derivative methods, zero-crossing is supposed to fire only at the
+location of maximum gradient.
+- Then, taking the opposite view to your argument, imagine that you would fit a more regular polynomial, let's say of degree 4. Then instead of getting a
+smoother approximant, I imagine that you would witness some kind of Runge's
+phenomenon which would translate in ringing patterns that would make the edge detection less predictable.
+It may be that the reason of the thicker edges is precisely the ringing that would create many local edges.
+But I must say that
+this is only a lead and I am not sure of this explanation either. You could
+challenge this argument on the synthetic example by plotting the profiles of
+the function $f_theta(rho)$ at few edge locations.
+
+\subsubsection{ Section 6: Conclusion}
+Your conclusion is quite poor. It simply recaps what was just stated in the
+experimental section. You could try to improve it by placing the
+algorithms under study in a broader context. The only part of your conclusion
+that follows that path is the reference to Canny's approach. Much more
+works have been made on edge detection since then, for instance approaches
+based on machine learning that have several advantages to be discussed. This one option among others.
+
+Again, too many short paragraphs.
+
+\que "runs significative faster that" --> "runs significaNTLY faster thaN"\\
+
+\ans corrected\\
+
+\section{ Comments on the Code and demo}
+
+\subsubsection{General description of the code}
+The reviewed code consists of two files: edges.c contains the main function,
+while the actual algorithms are included in \verb+classic_edge_detectors.c+; This
+latter contains several routines:
+- 6 routines are dedicated to each algorithms or their different flavors.
+Accordingly one can find the following routines:
+\begin{verbatim}
++ edges_roberts
++ edges_prewitt
++ edges_sobel
++ edges_mh for the Marr-Hildreth algorithms with gaussian and laplacian
++ edges_mh_log for the same algorithm with the laplacian of gaussian
++ edges_haralick
+- another routine called conv2d implements image filtering with a squared kernel
+- eventually two other are responsible for generating a gaussian kernel or laplacian of
+gaussian
+\end{verbatim}
+
+The implementations of all the detectors match faithfully their description in
+the pdf. This is obviously a very good point. I checked as carefully as I could
+all the kernel entries. This was quite some fun concerning Haralick's mask!
+Although in this case I only checked that the entries were identical in the pdf
+and the code (I did not check the mathematical validity of those entries).
+
+Consistently to my expectation, the three routines corresponding to first order
+derivative (robert, prewitt and sobel) are strictly identical (except for
+names of variables and the kernel entries). One could argue that this property,
+that is quite clear in the paper, should have been reflected in the code. Yet I
+would understand that the authors prefer to keep the three routines independent
+from one another.
+
+Similarly, the two routines implementing the two flavors of the Marr-Hildreth
+detector are entirely identical a part from the filtering phase. Again both
+implementation could be merged into a single routine.
+
+\subsection{ Algorithm's parameters}
+From the article I could retrieve the following list of parameters:
+\begin{verbatim}
+- First order methods : th
+- Marr-Hildreth: sigma, n, $th_{ZC}$
+- Haralick: $ro_0$
+- All algorithms: padding methods
+\end{verbatim}
+They corresponds exactly to the ones present in the code and (except for the
+padding method) also those exposed in the demo. This is quite satisfying!
+
+\subsection{ Minor comments}
+\subsubsection{ In gaussian\_kernel}
+The normalization by the sum is different from the one evoked in Section
+4.1.1. You should either change the implementation or comment on this in the article. Note
+also that the comment in the code correspond to the formula in the pdf instead of the implemented one.
+
+\subsubsection{ In LoG\_kernel }
+The same problem of consistency for the normalization occurs here. It is in
+fact worse than in the gaussian case, since I suppose that the LOG kernel does not sum to one.
+Besides dividing by $sigma^4$ is pointless if you normalize by the sum!
+Of course, kernel normalization has no impact on the edge detection since the thresholds are expressed in a relative way.
+
+\subsubsection{ In the README:}
+\que you announce that:
+"More than one algorithm can be applied to the same image with the same
+command. For example:
+edges -r 0.1 -l 3 29 0.13 molino.png
+will apply Roberts' algorithm with threshold parameter 0.1, and the
+Marr-Hildreth (LoG) with parameters sigma=3.00, N=29, TZC=0.13;"
+You should also warn that each algorithm can be launched only once. For instance
+the following call will fail:
+edges -r 0.1 -r 0.2 -l 3 29 0.13 molino.png
+
+\section{ Conclusion}
+Overall I think the submission is a good candidate for IPOL but it cannot be
+accepted in the current state. My main complain relates to the poor quality of
+the PDF:
+\begin{verbatim}
+- many explanations are unstructured
+- the writing is often bad (I am under the impression that the pdf was submitted without
+proofreading)
+- Except for the algorithmic description that is sufficiently detailed, many
+aspects are lacking: specifically in the experimental part and the conclusion.
+\end{verbatim}
 

### 2014-05-12 edge-detection-review-v11.tex
added v11 and v11 and merged with changes made starting with v10.

--- 2012-07-23 edge-detection-review-v10.tex
@@
 \ipolSetTitle{A Review of Classic Edge Detectors}
-\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
+\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$, Rafael Grompone von Gioi$^3$}
 \ipolSetAffiliations{%
 $^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
-$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}
+$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})\\
+$^3$ CMLA, ENS Cachan, France (\texttt{grompone@cmla.ens-cachan.fr})}
 
@@
 \usepackage[ruled,linesnumbered]{algorithm2e}
-\usepackage{sidecap,array}
+\usepackage{sidecap}
 \usepackage[hang,center]{subfigure}
@@
 \usepackage{color}
+\usepackage{calc}
+\usepackage{array}
 %\usepackage{algorithm}
@@
 }
+% Square cells table
+\newlength\celldim \newlength\fontheight \newlength\extraheight
+\newcounter{sqcolumns}
+\newcolumntype{S}{ @{}
+  >{\centering \rule[-0.5\extraheight]{0pt}{\fontheight + \extraheight}}
+  p{\celldim} @{} }
+\newcolumntype{Z}{ @{} >{\centering} p{\celldim} @{} }
+\newenvironment{squarecells}[1]
+  {\setlength\celldim{2em}%
+   \settoheight\fontheight{A}%
+   \setlength\extraheight{\celldim - \fontheight}%
+   \setcounter{sqcolumns}{#1 - 1}%
+   \begin{tabular}{|S|*{\value{sqcolumns}}{Z|}}\hline}
+% squarecells tabular goes here
+  {\end{tabular}}
+\newcommand\nline{\tabularnewline\hline}
 
@@
 The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
-such as Sobel, Prewitt and Roberts. In the second reviewed family it is used second derivative, for example in algorithms 
+such as Sobel, Prewitt and Roberts. In the second reviewed family, second derivatives are used, for example in algorithms 
 like Marr-Hildreth and Haralick. \\
@@
 
+% No mas link, mirar como se pone en las publicaciones nuevas. (ir a buscar el codigo a la pagina de la publicacion)
+% lo mismo con la demo... sacar GIT
+
+%% VER SI HAY UN TEXTO ESTANDAR PARA ESTO
 \begin{ipolCode}
-For all the algorithms reviewed, an open source C implementation is provided and can be downloaded from 
-\href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}. You can also follow 
-this code project in this \href{git://github.com/haldos/edges.git}{GIT repository}. An 
+For all the algorithms reviewed, an open source C implementation is provided that can be downloaded from 
+the IPOL publication of this article. An 
 \href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration} is also available, where 
@@
 
-%-------------------------------------------------------------------------------
+% You can also follow this code project in this \href{git://github.com/haldos/edges.git}{GIT repository} 
+% \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}
+
+%%-------------------------------------------------------------------------------
 
@@
 
-Edge detection is one of the first approaches published in the literature for 
-segmenting images. The basic idea is to detect abrupt changes in intensity. 
+% cambiar lo de segmenting images. Decir que edge detection es una de las operaciones
+% basicas y mas antiguas en el procesamiento de imagenes.
+% Edge detection is one of the first approaches published in the literature for 
+% segmenting images. 
+The basic idea of edge detection algorithms is to detect abrupt changes in image intensity. 
 Detecting those changes in intensity for the purpose of finding edges in images 
-can be accomplished using first or second order derivatives. \\
+can be accomplished using first or second order derivatives.
 
@@
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} the 
-behavior of such algorithms is described, as an introduction to a more sophisticated 
-analysis of the edge detection process.\\
-
+first derivative of the image \cite{Gonzalez2007Digital}. Next section describes such algorithms and serves as an introduction to a more sophisticated analysis of the edge detection process.
+
+% en este parrafo, puede ser cambiar trough por valley
 In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
@@
 Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
-which is also reviewed in section \ref{sec:second}. This algorithm has the particularity of 
-proposing a model to approximate locally the image around a point. Then, using this model, 
+which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
+proposing a model to locally approximate the image around a point; then, using this model, 
 an approximation to the second derivative of the image can be calculated analytically and 
-finding edges is achieved by imposing a condition over the model parameters. \\
+finding edges is achieved by imposing a condition over the model parameters.
 
@@
 implements the described algorithms. In section \ref{sec:appendix1} some common mathematical developments are presented. 
-A commented explanation of the source code can be found in section \ref{sec:appendix2}. This work also provides tools for code documentation 
-(see section \ref{sec:appendix2}). \\
-
-Results of the implemented algorithms are presented in section \ref{sec:results}, along with examples 
-to compare their performance. Conclusions are detailed in section \ref{sec:conclusions}.
+%A commented explanation of the source code can be found in section \ref{sec:appendix2}. This work also provides tools for code documentation 
+%(see section \ref{sec:appendix2}). 
+Results of the implemented algorithms are presented in Section \ref{sec:results}, along with examples 
+to compare their performance. Conclusions are detailed in Section \ref{sec:conclusions}.
 
@@
 Algorithms based on first derivative studied in this paper have a common scheme, with 
-the only difference in the type of filtering. Figure \ref{fig:blockdiagram1} presents
+the only difference in the type of filtering. Figure \ref{fig:blockdiagram1} shows
 a block diagram of these algorithms. \\
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{blockdiagram1.pdf}
+	\includegraphics[width=\textwidth]{blockdiagram1.pdf}
 	\caption{Block diagram of first derivative edge detection algorithms.}
@@
 				\begin{bmatrix} 
-					g_x \\ g_y
+					f_x \\ f_y
 				\end{bmatrix}
@@
 	\begin{cases}
-		M(x,y) = \sqrt{g_x^2 + g_y^2} \\
-		\alpha(x,y) = \tan^{-1} \begin{bmatrix} g_x \\ g_y \end{bmatrix} .\\
+		M(x,y) = \sqrt{f_x^2 + f_y^2} \\
+		\alpha(x,y) = \tan^{-1} \left( \dfrac{f_y}{f_x} \right) .
 	\end{cases}
@@
 The direction of an edge at an arbitrary location $(x,y)$ of the image is 
-orthogonal to the direction $\alpha(x,y)$ of the gradient vector.\\
-
-To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
-at every pixel in the image is required. When dealing with digital images, numerical approximations 
+orthogonal to the direction $\alpha(x,y)$ of the gradient vector.
+
+To obtain the gradient, the partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
+need to be computed at every pixel in the image. When dealing with digital images, numerical approximations 
 of the partial derivatives are required, calculated in a neighborhood of each point. In the following,
@@
 \begin{equation}
-	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
-\end{equation}
-The image is composed by discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
-first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
+	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.
+\end{equation}
+% explicar que esta pasando con el h
+% en las operaciones de abajo, dejar explicito que se considera h=1
+The image is composed by discrete points of coordinates $(i,j)$, so taking $x=ih$ and $y=jh$, the 
+first derivative of the image is approximated, based on the intensity values ​​at points of the image, as\footnote{This approximation considerates $h=1$, so the partial derivatives can be computed as intensity difference between neighbor pixels.}:
 \begin{equation}
 \label{eq:roberts1}
-	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
+	f_x = \frac{\partial f(x,y)}{\partial x} \cong f(i+1,j) - f(i,j)
 \end{equation}
@@
 \label{eq:roberts2}
-	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
+	f_y = \frac{\partial f(x,y)}{\partial y} \cong f(i,j+1) - f(i,j).
 \end{equation}
@@
 Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
-by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. But when finding 
-diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
+by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. %But when finding 
+%diagonal edges is desired, 2-D masks are needed. 
+The \textit{Roberts operators} are one of the earliest 
 attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
@@
 in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
-usual to normalize the masks, in order to have gain equal to $1$. \\
-
-\begin{SCfigure}[3][!h]
-	\centering
-	\includegraphics[width=0.13\textwidth]{1dmasks_n.pdf}
+usual to scale the mask values, in order to have unit $L_2$ norm.
+
+% la primer figura es una forma simple de calcular las derivadas, pero roberst, en [referenfcia]
+% propone calcularlos como en la segunda figura. Nota: cambiar signos.
+
+%% VER NORMALIZACION
+
+\begin{SCfigure}[2][!h]
+	\centering
+	\begin{squarecells}{1}
+		-1   \nline
+		1  \nline
+	\end{squarecells}
+	\quad
+	\begin{squarecells}{2}
+		-1 & 1   \nline
+	\end{squarecells}
 	\caption{One-dimensional normalized masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
@@
 
-\begin{SCfigure}[3][!h]
-	\centering
-	\includegraphics[width=0.15\textwidth]{roberts_n.pdf}
+\begin{SCfigure}[1][!h]
+	\centering
+	\begin{squarecells}{2}
+		-1 &  0  \nline
+		0 & 1  \nline
+	\end{squarecells}
+	\quad
+	\begin{squarecells}{2}
+		0  & -1  \nline
+		1 & 0  \nline
+	\end{squarecells}
 	\caption{Normalized \textit{Roberts cross-gradient} 2-D masks.}
@@
 
+Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts in his article proposes
+to compute them as shown in Figure \ref{fig:roberts}.
+In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, what produces the same result.
+
 %-------------------------------------------------------------------------------
@@
 \subsection{The \textit{Prewitt} operators}
+
+% agregar referencia
 
@@
 of them being the 3x3. These masks provide more information to find the direction of the edges, because they take 
-into account information on opposite sides of the central point. \\
+into account information on opposite sides of the central point.
 
@@
 These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
-These masks are called \textit{Prewitt operators}. \\
-
-\begin{SCfigure}[2][!h]
-	\centering
-	\includegraphics[width=0.25\textwidth]{prewitt_n.pdf}
+These masks are called \textit{Prewitt operators}.
+
+\begin{SCfigure}[1][!h]
+	\centering
+	\begin{squarecells}{3}
+		$-\frac{1}{6}$ 	& $-\frac{1}{6}$ 	& $-\frac{1}{6}$	\nline
+		0 			& 0			& 0			\nline
+		$\frac{1}{6}$ 	& $\frac{1}{6}$ 	& $\frac{1}{6}$	\nline
+	\end{squarecells}
+	\quad
+	\begin{squarecells}{3}
+		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
+		$-\frac{1}{6}$	& 0	& $\frac{1}{6}$	\nline
+		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
+	\end{squarecells}
 	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
@@
 These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
-smoothing properties in the direction opposite to the gradient.
+$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]^T$, and vice versa. Hence, these operators have 
+smoothing properties in the direction orthogonal to the gradient.
 
@@
 A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
-difference. It can be shown that using a $2$ in the center location provides better image smoothing. This
+difference. It can be shown that using double weight in the center location provides better image smoothing. This
 variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
-\textit{Sobel operators}. \\
-
-\begin{SCfigure}[2][!h]
-	\centering
-	\includegraphics[width=0.25\textwidth]{sobel_n.pdf}
+\textit{Sobel operators}.
+
+Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
+$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]^T$ (closer to a Gaussian response than Prewitt), and vice versa. 
+Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph.
+
+\begin{SCfigure}[1][!h]
+	\centering
+	\begin{squarecells}{3}
+		$-\frac{1}{8}$ 	& $-\frac{1}{4}$ 	& $-\frac{1}{8}$	\nline
+		0 			& 0			& 0			\nline
+		$\frac{1}{8}$ 	& $\frac{1}{4}$ 	& $\frac{1}{8}$	\nline
+	\end{squarecells}
+	\quad
+	\begin{squarecells}{3}approximate
+		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
+		$-\frac{1}{4}$	& 0	& $\frac{1}{4}$	\nline
+		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
+	\end{squarecells}
 	\caption{Normalized \textit{Sobel} 2-D masks of size $3\times3$.}
@@
 
-Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
-Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\
-
 The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
@@
 As mentioned before, edges can be seen as abrupt changes in intensity, i.e.\ higher values of gradient.
-An example of this behavior is shown in Figures \ref{fig:profiles3d-b} and \ref{fig:profiles3d-c}. \\
+An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using the Sobel operators.
+
+Then, once the respective operators are applied, an approximation of the gradient (stored in two matrices) 
+is obtained, containing the approximation of the partial derivatives $f_x$ and $f_y$. Gradient magnitude 
+image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}).
+
+Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
+\begin{equation*}
+	\mbox{edges}[i,j] = \begin{cases} 0,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
+									1,& \mbox{if}\ M[i,j]<\mbox{th}
+					  \end{cases}
+\end{equation*}approximate
+where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}).
+
+Note that different thresholds lead to different results.
+More and thicker edges for small thresholds values, and the opposite for large thresholds values.
+This threshold, and the others mentioned below, are defined as a percentage of the maximum value
+of the gradient image\footnote{This means no loss of generality in applying the threshold, and makes the threshold adapt to the behavior of different images.}.
+The advantage of this is to adapt the threshold to the dynamic range of the image.
+Therefore, the parameter $threshold$ in the algorithms takes values ​​between $0$ and $1$
+
+\clearpage
 
@@
 
-	\subfigure[3D image representation.]{\label{fig:profiles3d-b}\includegraphics[width=0.48\textwidth]{profile3d_image.pdf}}
-	\subfigure[3D gradient representation.]{\label{fig:profiles3d-c}\includegraphics[width=0.48\textwidth]{profile3d_gradient.pdf}}
-	\caption{Example figure, 3D image and gradient representations.}
+	\subfigure[Horizontal derivative $f_x$.]{\label{fig:profiles3d-b}\includegraphics[width=0.3\textwidth]{gradient_image_x.png}}
+	\quad
+	\subfigure[Vertical derivative $f_y$.]{\label{fig:profiles3d-c}\includegraphics[width=0.3\textwidth]{gradient_image_y.png}}
+	\quad
+	\subfigure[Gradient module image $M=\sqrt{f_x^2+f_y^2}$.]{\label{fig:profiles3d-d}\includegraphics[width=0.3\textwidth]{gradient_image.png}}
+	\caption{Example computation of the gradient. Original image and gradient image computed using Sobel operators (Operators in Figure \ref{fig:sobel}).}
 	\label{fig:profiles3d}
@@
 
-\begin{figure}[b!]
-	\centering
-	\subfigure[$th = 0.1*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_1.png}}
-	\quad
-	\subfigure[$th = 0.2*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_2.png}}
-	\quad
-	\subfigure[$th = 0.4*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
+\begin{figure}[t!]
+	\centering
+	\subfigure[$th = 0.1*\max(M)$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_1.png}}
+	\quad
+	\subfigure[$th = 0.2*\max(M)$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_2.png}}
+	\quad
+	\subfigure[$th = 0.4*\max(M)$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
 	\caption{Thresholded images. Different values of the threshold applied to the magnitude of the gradient.}
@@
 
-Then, once the respective operators are applied, an approximation of the gradient (stored in two images) 
-is obtained, containing the approximation of the partial derivatives $g_x$ and $g_y$. Gradient magnitude 
-image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}). \\
-
-Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
-\begin{equation*}
-	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{th} \\
-									1,& \mbox{if}\ M[i]<\mbox{th}
-					  \end{cases}
-\end{equation*}
-where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
-for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
-of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
-Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
-
 %-------------------------------------------------------------------------------
@@
 The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
-\clearpage
 
@@
 \REQUIRE input image, threshold $th$.
-\STATE $im_{ORIG} \leftarrow$ input image
-\IF{$im_{ORIG}$ is grayscale}
-	\STATE $im \leftarrow im_{ORIG}$
-\ELSE 
-	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
-\ENDIF
+\STATE $im \leftarrow$ input image
 \STATE Define $operator_x$ and $operator_y$ \COMMENT{Roberts, Prewitt or Sobel.}
@@
 image with different masks, without taking into account the characteristics of the edges or 
-noise in the image. \\
+noise in the image.
 
@@
 of the image. This can be done in several ways. Two different ways of doing 
-this are implemented implemented in this work (see block diagram in Figure 
-\ref{fig:blockdiagram2}); convolving the image with a Gaussian kernel and then 
+this are implemented in this work (see block diagram in Figure 
+\ref{fig:blockdiagram2}): convolving the image with a Gaussian kernel and then 
 approximating the second derivative (Laplacian) with a 3x3 kernel, or 
@@
 Gaussian function. There are more ways to do so, for example, using 
-recursive Gaussian filters \cite{Deriche1993Recursively}. \\
-\clearpage
-
-\begin{figure}[!t]
-	\centering
-	\includegraphics[width=0.9\textwidth]{blockdiagram2.pdf}
+recursive Gaussian filters \cite{Deriche1993Recursively}.
+
+\begin{figure}[!b]
+	\centering
+	\includegraphics[width=0.92\textwidth]{blockdiagram2.pdf}
 	\caption{Block diagram of Marr-Hildreth algorithm.}
@@
 
-The algorithm is divided in three steps, each one described later:
+The algorithm is divided in two steps, each one described later:
 \begin{enumerate}
-	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
 	\item Convolution of the image with:
 	\begin{itemize}
-		\item a Laplacian of Gaussian (LoG) kernel. (or)
+		\item a Laplacian of Gaussian (LoG) kernel, (or)
 		\item a Gaussian kernel and then a Laplacian operator.
@@
 
-Some auxiliary functions were implemented for computing the operations
-needed in the algorithm, such as Gaussian kernel and Laplacian of a Gaussian 
+Some auxiliary functions are needed such as Gaussian kernel and Laplacian of a Gaussian 
 kernel generation, and 2-D convolution of an image with a given kernel, 
-with different boundary conditions (these operations will be discussed 
-in detail in section \ref{sec:appendix1}).
+with different boundary conditions. These operations will be discussed 
+in detail in section \ref{sec:appendix1}.
 
@@
 \end{equation}
-indicating that image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result.\\
-
-Then, generation of both Gaussian and LoG kernels is required (see section \ref{sec:appendix1}).\\
+indicating that the image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result\footnote{The difference between these approaches lies in the compromise between the accuracy in the calculation and the computational cost (see Section \ref{kernelcomparison}).} .
+
+% agregar comentario de que la diferencia entre ellos radica en un compromiso entre precision en el calculo
+% y costo computacional. Y que se va a explicar mas adelante.
 
@@
 \begin{enumerate}
-	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling equation \ref{eq:gaussian_function}.
-	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask:
-			$\begin{bmatrix}
+	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling the Gaussian kernel (see equation \ref{eq:gaussian_function}).
+% reescribir (sampleo de un kernel gaussiano)
+	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}:
+	\begin{equation*}
+		\begin{bmatrix}
 			1 &  1 & 1 \\
@@
 			1 &  1 & 1 \\
-			\end{bmatrix}$\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
+		\end{bmatrix}
+	\end{equation*}
 	\item Find the zero crossings of the image from step 2.
@@
 
+% Sacar la footnote y explicar de otra manera las dos formas de Marr-Hildreth
+
 %-------------------------------------------------------------------------------
@@
 \subsubsection{Zero crossing}
+
+% cambiar la primer frase de este parrafo. Explicar de otra forma lo de buscar que haya al menos un cambio de signo en los pixeles vecinos opuestos.
 
@@
 a threshold is used, so that not only the signs of the opposite pixels must differ, also their 
-difference in absolute value must be greater than a certain threshold. \\
+difference in absolute value must be greater than a certain threshold.
 
@@
 \REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
-\STATE $im_{ORIG} \leftarrow$ input image
-\IF{$im_{ORIG}$ is grayscale}
-	\STATE $im \leftarrow im_{ORIG}$
-\ELSE 
-	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
-\ENDIF
+\STATE $im \leftarrow$ input image
 \STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}
@@
 \IF{Gaussian kernel}
-	\STATE Define Laplacian approx. operator $laplacian$
+	\STATE Define Laplacian operator $laplacian$
 	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
@@
 In this section the original work of Haralick \cite{bb20239} on edge detection is presented in detail.
-The main idea behind this algorithm is identical to that of the previous method; find zeros in 
-the second derivative of the image. In this method, the input image is smoothly approximated through local bi-cubic
+The main idea behind this algorithm is identical to that of the previous method: find zeros in 
+the second derivative of the image. In this method, however, the input image is smoothly approximated through local bi-cubic
 polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
@@
 
-Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}. \\
+Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}.
 
@@
 \end{equation}
+where $(x,y)$ are the offsets of each neighborhood point relative to the center point (e.g.\ in a neighborhood of size $5\times5$, $x$ and $y$ take values ​​between $-2$ and $2$).
 
@@
 neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
-points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
-In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
-computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
+points leads to an overdetermined system, which can be solved using least squares or any other data fitting technique.
+In Haralick's work, this approximation is computed using least squares.
+
+%, and to speed up the calculations the solution is 
+%computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
 
@@
 to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
-By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
+By substituting the 25 data points into the polynomial equation (\ref{eq:bicubic}), the following system of 
 equations is obtained,
-
 \begin{equation*}
@@
 
-Then, the least squares problem can be solved by using the normal equations,
-
+Then, the least squares problem can be approximately solved by computing the pseudo inverse of $\mathbf{A}$,
 \begin{equation*}
@@
 \end{equation*}
-
 $\mathbf{B}$ is a $10\times25$ matrix, 
-
 \begin{equation*}
 	\begin{bmatrix}
-		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
-		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
+		b_{1,1}		& b_{1,2}	& \cdots	& b_{1,25}	\\
+		b_{2,1}		& b_{2,2}	& \cdots	& b_{2,25}	\\
 		\vdots		& \vdots	& \ddots	& \vdots	\\
-		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
+		b_{10,1}	& b_{10,2}	& \cdots	& b_{10,25}
 	\end{bmatrix}
@@
 \end{equation*}
-
 For each coefficient ($i$ from $1$ to $10$), 
-
 \begin{equation}
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
-\end{equation}
-
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \cdots + b_{i,25}f_{25}% \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
+\end{equation}
 where
-
 \begin{equation}
 	\label{eq:b_i}
-	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
-									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
+	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \cdots	& b_{i,5}	\\
+									b_{i,6}		& b_{i,7}	& \cdots	& b_{i,10}	\\
 									\vdots		& \vdots	& \ddots	& \vdots	\\
-									b_{i,21}	& b_{i,22}	& \hdots	& b_{i,25}	\\
+									b_{i,21}	& b_{i,22}	& \cdots	& b_{i,25}	\\
 					\end{bmatrix}.\\
 \end{equation}
-
-\vspace{1ex}
+%% VER UNA FORMA DE ESCRIBIR TAMBIEN LOS f1...f25 PARA MOSTRAR POR QUE LOS bi,1 ...
+bi,25 ESTAN DISPUESTOS COMO UNA MATRIZ.
+ES PORQUE LOS f SON UN ENTORNO 5X5 DEL PIXEL, Y SE MULTIPLIAN PUNTO A PUNTO POR LOS b.
+
 Using equation \ref{eq:coefficients} and the masks given in Table \ref{table:b_i}, it is possible to compute 
 the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
-are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.\\
+are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.
+
+In order to speed up the calculations, the solution is computed by convolving the image with some precomputed 
+masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$.
 
@@
 expression in equation \ref{eq:bicubic}. If just the first order terms of this 
-polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as
-
+polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approximated as
 \begin{align}
+	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
+	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}.  
 \label{eq:sincos}
-	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
-	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
 \end{align}
-
 Now substituting the variables $x$ and $y$ in polar form as
-
 \begin{equation*}
-	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
+	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} 
 \end{equation*}
-
 in the bi-cubic polynomial, we obtain
-
 \begin{equation}
@@
 \end{equation}
-
 where
-
 \begin{align}
@@
 	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
-	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). \nonumber \\
+	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). 
 \end{align}
-
-The derivatives are obtained as follows.
-
+The derivatives are obtained as follows:
 \begin{align}
@@
 	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
-	f'''_{\theta}(\rho) = 6C_3 .\nonumber \\
+	f'''_{\theta}(\rho) = 6C_3 .\nonumber 
 \end{align}
-
 Then, the condition that the second derivative is equal to zero becomes
-
 \begin{equation}
@@
 \end{equation}
-
 where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
+should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial to become zero in a neighborhood 
 of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
-value is between $0.4$ and $0.6$.\\
+value is between $0.4$ and $0.6$.
 
 The other condition required is that the third derivative to be negative, i.e.\
-
 \begin{equation}
@@
 \end{equation}
-
-Given the direction defined by $\theta$, with $\rho>0$, one edge will always be an ascending step. This 
-indicates that first derivative in this direction has a maximum and that second derivative has a zero 
-crossing with negative slope. That is, third derivative less than zero.
+Given the direction defined by $\theta$, with $\rho>0$, an edge will always be an ascending step. This 
+indicates that the first derivative in this direction has a maximum and that the second derivative has a zero 
+crossing with negative slope. That is, the third derivative less than zero.
 
@@
 \begin{enumerate}
-	\item Find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
+	\item For each pixel in the image, find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
 	\item Compute $\sin(\theta)$ and $\cos(\theta)$ (equations \ref{eq:sincos}).
@@
 \REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
-\STATE $im_{ORIG} \leftarrow$ input image
-\IF{$im_{ORIG}$ is grayscale}
-	\STATE $im \leftarrow im_{ORIG}$
-\ELSE 
-	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
-\ENDIF
-\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
+\STATE $im \leftarrow$ input image
+\STATE Define the ten $5x5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
 \FORALL{pixel $i$ in image $im$}
@@
 	\FOR{$j=1$ \TO $10$}
-		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
+		\STATE $k_j \leftarrow$ $\sum_{n,m}neighbors[n,m]mask_j[n,m]$ % convolution($neighbors$,$mask[j]$)
 	\ENDFOR
@@
 
-\subsection{Grayscale conversion}
-\label{sec:grayscale}
-
-The first step in every algorithm presented above is to convert color images to gray intensity images. 
-For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
-\begin{equation}
-    Y = (6968 R + 23434 G + 2366 B) / 32768
-\end{equation}
-where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float} to avoid truncated results.
+%\subsection{Grayscale conversion}
+%\label{sec:grayscale}
+
+%The first step in every algorithm presented above is to convert color images to gray intensity images. 
+%For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
+%\begin{equation}
+%    Y = (6968 R + 23434 G + 2366 B) / 32768
+%\end{equation}
+%where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float} to avoid truncated results.
 See code in section \ref{app:marr-hildreth}.}.
@@
 
+% Explicar que voy a crear un patch centrado en [x,y]=[0,0], y un poco mas de la generacion del patch.
+% Explicar que se samplea en el centro de los pixeles (hay varias formas de generar el patch, por ejemplo,
+% que cada valor sea la media de la guncion en el pixel, etc).
+
 %-------------------------------------------------------------------------------
@@
 
-Gaussian kernel is generated by sampling the 2-D Gaussian function
+Gaussian kernel is generated by sampling the 2-D Gaussian function (centered at $(0,0)$)
 %\footnote{Note that for simplicity the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
@@
 \end{equation}
-where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\
+where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).
 
@@
 To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$.
-Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\
+Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.
 
@@
 one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
-have to be performed, one with the kernel and another with the Laplacian operator).\\
+have to be performed, one with the kernel and another with the Laplacian operator), but the kernel 
+support must be greater in order to obtain the same precision..
+
+% Una linea explicando que para la misma presicion se requiere un soporte mas grande (y linkar a la seccion siguiente).
 
@@
 \begin{equation} 
-	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
-\end{equation}
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
-\end{equation}
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
-\end{equation}
-
-where the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted. Therefore we obtain
-
+	\frac{\partial}{\partial x}G(x,y)=-\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
+\end{equation}
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
+\end{equation}
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
+\end{equation}
+Therefore we obtain
 \begin{equation}
@@
 \subsubsection{Gaussian and LoG functions comparison}
+\label{kernelcomparison}
 
@@
 example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
-first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
+first odd integer greater than $6\sigma$, which is $25$, and for the case of LoG function, would 
 be $29$.
+% Especificar cual es el criterio para el tamaño del LoG.
+% mirar en el software si no es mejor quitar la dependencia con n, y que se decida internamente el tamaño
+% del kernel en funcion de sigma.
 
@@
 around the edges of the image, and to get a valid output the same size as input image. In this paper, 
-two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
+two methods were implemented: zero-padding and boundary reflection. Zero-padding implies complete 
 the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
 kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
-relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
-or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.
-
-%-------------------------------------------------------------------------------
+relative to the edge of the image.
+Direct convolution is not the only way of filtering an image with a kernel.
+There are other methods such as FFT convolution or recursive filtering\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
+
+%-------------------------------------------------------------------------------
+
+\begin{comment}
 
@@
 
-To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
+To the best of our knowledge, the only freely available codes implementing these algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
 There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
@@
 
-This code documentation was generated using a perl script to extract some parts of the 
-code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
-
-Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
+%This code documentation was generated using a perl script to extract some parts of the 
+%code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
+
+%Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
 
@@
 
+\end{comment}
+
 %-------------------------------------------------------------------------------
@@
 
-The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used. \\
+The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used.
 
@@
 Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
-Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
-\\
+The execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
 
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
+	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (127$\times$127 pixels, 1 channel).}
 	\label{exectime1}
@@
 Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.
-\\
-
-Now the performance of each algorithm is analyzed, using a more realistic test image shown in Figure \ref{fig:original2}. \\
+
+Now the performance of each algorithm is analyzed, using a natural image shown in Figure \ref{fig:original2}.
 
@@
 	\centering
-	\subfigure[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
-	\quad
-	\subfigure[Image converted to greyscale.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
+	\includegraphics[width=0.48\textwidth]{results/molino_bw.png}
 	\caption{Test image 2: windmill (1000$\times$563 pixels).}
@@
 Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm.
-Table \ref{exectime2} shows the execution time for each one of the algorithms.
-\\
+Table \ref{exectime2} shows the execution time for each of the algorithms.
 
@@
 However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.
-\\
-
-Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes. \\
-
-\begin{figure}[h!]
+
+Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes.
+
+\begin{figure}[t!]
 	\centering
@@
 
-\clearpage
-
 %-------------------------------------------------------------------------------
@@
 More examples can be found in the online \href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{demo}.
-\\
 
@@
 	\centering
-	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
-	\quad
-	\subfigure[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}
+	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}
 
@@
 \end{figure}
-
-\clearpage
 
@@
 Canny edge detector \cite{Canny1986}).
-\\
 	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
@@
 This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}).
-\\
-	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order).
+	\item As mentioned earlier, Haralick's algorithm is the first to assume higher degree of regularity in the neighborhood of a pixel (third order).
 This causes thicker edges, as shown in both examples.
@@
 These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods.
-\\
 \end{itemize}
@@
 	\centering
-	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
-	\quad
-	\subfigure[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}
+	\subfigure[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}
 
@@
 
-This is an example of applying the edge detection algorithms described here, frame by frame, to a video: 
-\begin{itemize}
-	\centering
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (43 Mb).
-\end{itemize}
+This is an example of applying the edge detection algorithms, frame by frame, to a video: 
+%\begin{itemize}
+%	\centering
+%	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (43 Mb).
+%\end{itemize}
 \begin{multicols}{2}
 \begin{itemize}
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (43 Mb).
 	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts-wide_0.1.mov}{roberts version} (7.9 Mb).
@@
 \clearpage
-
 %-------------------------------------------------------------------------------
@@
 The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
-\\
 
@@
 Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make more fair the comparison with the second derivative methods.
-\\
 
@@
 In real images, they have better behavior than first derivative algorithms.
-\\
 
@@
 It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
-\\
 
@@
 It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function.
-This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
-\\
+This supposes a certain regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
 
@@
 Choosing one or the other may depend on requirements of the application.
-Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.\\
+Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.
 
@@
 available in C language, along with an online demo where users can run the algorithms 
-with arbitrary images uploaded by themselves. Also some code documentation tools are supplied, 
-which were used to generate the detailed code description presented in section \ref{sec:appendix2}. We think this tool could be useful for future IPOL articles.
+with arbitrary images uploaded by themselves.
 

### 2014-05-26 review/review_answer_2.tex
added v11 and v11 and merged with changes made starting with v10.
Completed second round of review

@@
 \usepackage{xr}
-\externaldocument{../edge-detection-review-2nd-revision-v1}
+\externaldocument{../edge-detection-review-2nd-revision-v2}
 \reversemarginpar
@@
 \chapter{Introduction}
-In this communication we answer to the second review of the article "A Contrario Selection of Optimal Partitions for Image Segmentation", submitted to the SIAM Journal on Imaging Sciences.
-In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided. In the following we address the issues raised by the reviewers. 
-For each of them, an answer is provided and a cross-reference with a red tag Cn (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided.
+In this communication we answer to the second rounds of review of the article "A Contrario Selection of Optimal Partitions for Image Segmentation", submitted to the SIAM Journal on Imaging Sciences.
+In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided.
+In the following we address the issues raised by the reviewers.
+For each issue, an answer is provided and a cross-reference with a red tag $C_n$ (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided.
 In addition, changes made to the paper are highlighted in red (when significative) in the new version of the paper.
@@
 \item In that sense we moved the numeric values of the $\mathbf{mask_i}$ matrices to the first appendix
-\item We also moved the Implementation section, which documents the code, to the second appendix.
 \item Removed several typos and orthographic errors.
+\item improved code readme and demo.
 \end{itemize}
@@
 
-\ans XXX
+\ans we removed the videos from the submission, they are nice to see, but not so relevant.\\
 
@@
 \que "As mentioned before this leads to ab overdetermied"
-
 \ans Fixed\\
@@
 \que "Gaussian kernel is generated by sampling the"\\
+\ans YY\\
 \que "unlike the case of Gaussian kernel"\\
+\ans Suggested a partial correction\\
+
 \que "are not independent of each one"\\
+\ans  Rephrased\\
 \que "implies complete the borders with many zero valued pixels as needed"\\
-\que "It can be seen in gure 16 an area of interest of the image."\\
+\ans Rephrased\\
+\que "It can be seen in Figure 16 an area of interest of the image."\\
+\ans Fixed\\
 \que "appearing unwanted isolated edges"\\
+\ans Fixd\\
 \que "because they only consists"\\
+\ans Fixed\\
 \que "more fair"\\
+\ans Fixed\\
 
 \que "runs significative faster"\\
-
 \ans Fixed\\
@@
 \subsubsection{ Section 2.3: The Sobel operators}
-No upper case on the "i" of "thus It".
+\que No upper case on the "i" of "thus It".
+
+\ans Corrected.\\
 
 \subsubsection{Section 2.4: Computation of the edges}
-The threshold variable "th" should be in italic.
-Missing a "." at the end of the section.
+\que The threshold variable "th" should be in italic.
+Missing a "." at the end of the section.\\
+
+\ans Corrected.\\
 
 \subsubsection{Algorithm 2:}
+\que
 \begin{verbatim}
@@
 \end{verbatim}
+
+\ans Changed as suggested.\\
 
@@
 
-\ans We followed this suggestion, which in our opinion is a great improvement for the presentation. See \refc{second_derivative:haralick:mask_computation} Thanks to the reviewer for this one.
+\ans We followed this suggestion, which in our opinion is a great improvement for the presentation. See \refc{second_derivative:haralick:mask_computation} Thanks to the reviewer for this one. \\
 
 \que Can you explain how you have obtained the entries of the mask? If you used a program for that, I think you should provide it.\\
-
-\ans XXX\\
 
@@
 
+\ans We rewrote the whole section, taking this comment into account.\\
+
 \que I would integrate the footnote 7 within the text, just after: "This parameter
@@
 
-You should explain what you mean by the optimal value of $\rho_0$. If you are
-simply referring to Haralick's claims, you should state it explicitly.
+\que You should explain what you mean by the optimal value of $\rho_0$. If you are
+simply referring to Haralick's claims, you should state it explicitly.\\
 
-Also, are you sure you implement the exact conditions specified in the
+\ans We actually used those values inspired on other works, and tested them experimentally.
+To do this properly we should benchmark this parameter against a proper database, and see what's best.
+From our point of view, that kind of work is outside the scope of this paper.\\
+
+\que Also, are you sure you implement the exact conditions specified in the
 original article by Haralick. The ones I could find are as follows:
@@
 the output. Yet I would recommend that you investigate why and provide that
-information.
+information.\\
+
+\ans Yes, we verified again the conditions and we are implementing exactly Haralick's conditions.
+Regarding [eq.
+1], indeed there is a typo in Haralick's paper.
+We explained that in \refc{second_derivative:haralick:analitical:condition} in a footnote.
+Regarding [eq.
+2], I think the explanation was confusing and we also made a mistake assuming $\rho>0$.
+if you lift that assumption and taking into account that the exact solution of $f''=0$ is $\rho^*=-\frac{C_2}{3C_3}$, our condition is exactly what it reads in Haralick's paper $\rho<\rho_0$.
+I think with the new presentation this should be clearer.
+Starting with \refc{second_derivative:haralick:analitical:conditions}, we rewrote the whole section.\\
 
 \subsubsection{ Section 4: Common Mathematical Operations}
-"All algorithms implemented" --> "All implemented algorithms"
+\que "All algorithms implemented" --> "All implemented algorithms"''\\
+\ans Corrected.\\
 
 \subsubsection{ Figure 11:}
-"Is clearly required" --> "It ..."
+\que "Is clearly required" --> "It ..."\\
 
-The text in the figure is practically unreadable.
+\ans Rephrased.\\
+
+\que The text in the figure is practically unreadable.\\
+\ans Enlarged figure.\\
 
@@
 
-\ans XXX\\
+\ans We corrected this, thanks.\\
 
@@
 
-\ans Since we don't have any space restrictions, summarizing all execution times in a table, will help the reader compare them.
+\ans Since we don't have any space restrictions, summarizing all execution times in a table, will help the reader compare them directly. \\
 
@@
 \subsubsection{ Section 6: Conclusion}
-Your conclusion is quite poor. It simply recaps what was just stated in the
+\que Your conclusion is quite poor. It simply recaps what was just stated in the
 experimental section. You could try to improve it by placing the
@@
 works have been made on edge detection since then, for instance approaches
-based on machine learning that have several advantages to be discussed. This one option among others.
+based on machine learning that have several advantages to be discussed. This one option among others.\\
 
-Again, too many short paragraphs.
+\ans We disagree on this.
+The conclusions depend on the focus you want to give to the article.
+This article is about reproducing basic edge detectors, and briefly compare them.
+It is not a survey not a benchmark of every possible edge detector.
+As contribution for IPOL, we think we might add some basic blocks like this one, and once there is a considerable base of papers devoted to edge detection, we might add one publication devoted to put them into a broader context.
+We'd rather keep this publication narrow focused on the subject.\\
+
+\que Again, too many short paragraphs.\\
+
+\ans Tried to improve that along the paper.\\
 
@@
 
-Consistently to my expectation, the three routines corresponding to first order
+\que Consistently to my expectation, the three routines corresponding to first order
 derivative (robert, prewitt and sobel) are strictly identical (except for
@@
 would understand that the authors prefer to keep the three routines independent
-from one another.
+from one another.\\
 
@@
 detector are entirely identical a part from the filtering phase. Again both
-implementation could be merged into a single routine.
+implementation could be merged into a single routine.\\
+
+\ans Indeed as they are different algorithms we rather keep separate implementations, although we agree that it might not be the best practice from the software point of view.\\
 
 \subsection{ Algorithm's parameters}
-From the article I could retrieve the following list of parameters:
+\que From the article I could retrieve the following list of parameters:
 \begin{verbatim}
@@
 
+\ans Thanks for the comment. We added the padding method as an input parameter of the algorithm in the code (not in the demo).\\
+
 \subsection{ Minor comments}
 \subsubsection{ In gaussian\_kernel}
-The normalization by the sum is different from the one evoked in Section
+\que The normalization by the sum is different from the one evoked in Section
 4.1.1. You should either change the implementation or comment on this in the article. Note
-also that the comment in the code correspond to the formula in the pdf instead of the implemented one.
+also that the comment in the code correspond to the formula in the pdf instead of the implemented one.\\
+
+\ans Corrected, now  the code uses the PDF formula\\
 
@@
 Of course, kernel normalization has no impact on the edge detection since the thresholds are expressed in a relative way.
+
+\ans Corrected, now  the code uses the PDF formula\\
 
@@
 the following call will fail:
-edges -r 0.1 -r 0.2 -l 3 29 0.13 molino.png
+edges -r 0.1 -r 0.2 -l 3 29 0.13 molino.png\\
+
+\ans Explicitely said in documentation that you can't execute two algorithms at the same time,\\
 

### 2014-05-26 edge-detection-review-2nd-revision-v2.tex
Completed second round of review

--- 2014-05-12 edge-detection-review-v11.tex
@@
 \ipolSetTitle{A Review of Classic Edge Detectors}
-\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$, Rafael Grompone von Gioi$^3$}
+\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
 \ipolSetAffiliations{%
 $^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
-$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})\\
-$^3$ CMLA, ENS Cachan, France (\texttt{grompone@cmla.ens-cachan.fr})}
+$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}
 
@@
 In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
-behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
+of edge detection algorithms is to find where abrupt changes in the intensity of an image have occurred. 
 The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
@@
 
-% No mas link, mirar como se pone en las publicaciones nuevas. (ir a buscar el codigo a la pagina de la publicacion)
-% lo mismo con la demo... sacar GIT
-
-%% VER SI HAY UN TEXTO ESTANDAR PARA ESTO
 \begin{ipolCode}
@@
 the IPOL publication of this article. An 
-\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration} is also available, where 
+\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration}\footnote{\url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}} is also available, where 
 you can test and reproduce our results.
@@
 
-% cambiar lo de segmenting images. Decir que edge detection es una de las operaciones
-% basicas y mas antiguas en el procesamiento de imagenes.
-% Edge detection is one of the first approaches published in the literature for 
-% segmenting images. 
-The basic idea of edge detection algorithms is to detect abrupt changes in image intensity. 
-Detecting those changes in intensity for the purpose of finding edges in images 
-can be accomplished using first or second order derivatives.
+Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms.  for 
+. The basic idea of edge detection is to detect abrupt changes in image intensity. 
+Detecting those changes can be accomplished using first or second order derivatives. \\
 
@@
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image \cite{Gonzalez2007Digital}. Next section describes such algorithms and serves as an introduction to a more sophisticated analysis of the edge detection process.
+first derivative of the image \cite{Gonzalez2007Digital}. Section ~\ref{sec:first} describes such algorithms 
+and serves as an introduction to a more sophisticated 
+analysis of the edge detection process.
 
@@
 operators. They also argued that a sudden intensity change will be seen 
-as a peak (or trough) in the first derivative or, equivalently, as a zero 
+as a peak (or valley) in the first derivative or, equivalently, as a zero 
 crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
@@
 
-\section{Algorithms based on first derivative}
+\section{Algorithms based on the first derivative}
 \label{sec:first}
 
-Algorithms based on first derivative studied in this paper have a common scheme, with 
-the only difference in the type of filtering. Figure \ref{fig:blockdiagram1} shows
-a block diagram of these algorithms. \\
+Algorithms based on first derivatives studied in this paper share a common structure, with the only difference  in the type of filtering used to compute those derivatives.
+Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure.
+\myn{\label{correction_a:first_derivative:first_deriv_rephrase}\\
 
@@
 
-The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with
-a small $h$. Thus $f'(x)$ is computed as
-\begin{equation}
-	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.
-\end{equation}
-% explicar que esta pasando con el h
-% en las operaciones de abajo, dejar explicito que se considera h=1
-The image is composed by discrete points of coordinates $(i,j)$, so taking $x=ih$ and $y=jh$, the 
-first derivative of the image is approximated, based on the intensity values ​​at points of the image, as\footnote{This approximation considerates $h=1$, so the partial derivatives can be computed as intensity difference between neighbor pixels.}:
+\myn{\label{correction_a:first_derivative:taylor}}The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$.
+Thus $f'(x)$ is computed as
+\begin{equation}
+	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
+\end{equation}
+\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid.
+If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$.
+When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
+Thus, in the discrete space the first derivative of the image is computed as:
 \begin{equation}
@@
 
-Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
-by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. %But when finding 
-%diagonal edges is desired, 2-D masks are needed. 
+\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}.
+However, in order to be able to compute diagonal edges, 2-D masks are needed.
+\\
 The \textit{Roberts operators} are one of the earliest 
-attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
-diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
+attempts to use 2-D masks for this purpose.
+These operators are based on computing the diagonal \tcom{differences} \myn{\label{correction_a:typo}} implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
+By convention
 in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
-usual to scale the mask values, in order to have unit $L_2$ norm.
+usual to scale the mask values, in order to have gain equal to $1$. \\
 
@@
 	\end{squarecells}
-	\caption{One-dimensional normalized masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
+	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
 	\label{fig:1dmasks}
@@
 	\end{squarecells}
-	\caption{Normalized \textit{Roberts cross-gradient} 2-D masks.}
+	\caption{\textit{Roberts cross-gradient} 2-D masks.}
 	\label{fig:roberts}
@@
 
-Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts in his article proposes
+Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{roberts1963machine} proposes
 to compute them as shown in Figure \ref{fig:roberts}.
@@
 These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
-These masks are called \textit{Prewitt operators}.
+These masks are called \textit{Prewitt operators} \cite{prewitt1970object}.
 
@@
 	\begin{squarecells}{3}
-		$-\frac{1}{6}$ 	& $-\frac{1}{6}$ 	& $-\frac{1}{6}$	\nline
+		$-\frac{1}{3}$ 	& $-\frac{1}{3}$ 	& $-\frac{1}{3}$	\nline
 		0 			& 0			& 0			\nline
-		$\frac{1}{6}$ 	& $\frac{1}{6}$ 	& $\frac{1}{6}$	\nline
+		$\frac{1}{3}$ 	& $\frac{1}{3}$ 	& $\frac{1}{3}$	\nline
 	\end{squarecells}
@@
 	\begin{squarecells}{3}
-		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
-		$-\frac{1}{6}$	& 0	& $\frac{1}{6}$	\nline
-		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
+		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
+		$-\frac{1}{3}$	& 0	& $\frac{1}{3}$	\nline
+		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
 	\end{squarecells}
@@
 
-These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]^T$, and vice versa. Hence, these operators have 
-smoothing properties in the direction orthogonal to the gradient.
+\myn{\label{correction_a:first_derivative:sobel}} The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
+$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]^T$.
+The same reasoning could be applied to the right mask.
+Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter.
+Thus, they show smoothing properties in the direction orthogonal to the gradient.
 
@@
 
-A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
-difference. It can be shown that using double weight in the center location provides better image smoothing. This
-variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
+A slight variation of the Prewitt operators which use more weight on the central coefficients of the 
+difference. It can be shown that using more weight in the center location provides better image smoothing. This
+variation is implemented using masks shown in Figure \ref{fig:sobel}. These operators are called 
 \textit{Sobel operators}.
 
-Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]^T$ (closer to a Gaussian response than Prewitt), and vice versa. 
-Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph.
-
 \begin{SCfigure}[1][!h]
@@
 	\begin{squarecells}{3}
-		$-\frac{1}{8}$ 	& $-\frac{1}{4}$ 	& $-\frac{1}{8}$	\nline
+		$-\frac{1}{4}$ 	& $-\frac{1}{2}$ 	& $-\frac{1}{4}$	\nline
 		0 			& 0			& 0			\nline
-		$\frac{1}{8}$ 	& $\frac{1}{4}$ 	& $\frac{1}{8}$	\nline
+		$\frac{1}{4}$ 	& $\frac{1}{2}$ 	& $\frac{1}{4}$	\nline
 	\end{squarecells}
 	\quad
-	\begin{squarecells}{3}approximate
-		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
+	\begin{squarecells}{3}
 		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
-		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
+		$-\frac{1}{2}$	& 0	& $\frac{1}{2}$	\nline
+		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
 	\end{squarecells}
-	\caption{Normalized \textit{Sobel} 2-D masks of size $3\times3$.}
+	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
 	\label{fig:sobel}
@@
 
-The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
-use Sobel operators, because they are better localized, and their filter are also less aliased, 
-because of the weighted shape of $[1\ 2\ 1]$.
+Sobel masks can be seen (up to a  scale factor) as the convolution of a horizontal derivation mask 
+$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
+Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\
+
+The computational cost of applying Prewitt and Sobel masks is exactly the same. Thus, it is preferable to 
+use Sobel operators, because edges are better localized, and their filter are also less aliased, because of the weighted shape of $[1\ 2\ 1]$.
 
@@
 As mentioned before, edges can be seen as abrupt changes in intensity, i.e.\ higher values of gradient.
-An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using the Sobel operators.
+An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using Sobel operators.
 
@@
 \begin{equation*}
-	\mbox{edges}[i,j] = \begin{cases} 0,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
-									1,& \mbox{if}\ M[i,j]<\mbox{th}
+	\mbox{edges}[i,j] = \begin{cases} 1,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
+									0,& \mbox{if}\ M[i,j]<\mbox{th}
 					  \end{cases}
-\end{equation*}approximate
+\end{equation*}
 where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-
-Note that different thresholds lead to different results.
-More and thicker edges for small thresholds values, and the opposite for large thresholds values.
-This threshold, and the others mentioned below, are defined as a percentage of the maximum value
-of the gradient image\footnote{This means no loss of generality in applying the threshold, and makes the threshold adapt to the behavior of different images.}.
-The advantage of this is to adapt the threshold to the dynamic range of the image.
-Therefore, the parameter $threshold$ in the algorithms takes values ​​between $0$ and $1$
-
-\clearpage
+It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
+for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
+of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
+Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
 
@@
 The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
+
+\clearpage
 
@@
 
-The two algorithms presented in this section (Marr-Hildreth \cite{AIM-518} [1980] and Haralick \cite{bb20239} [1987]) 
+The two algorithms presented in this section (Marr-Hildreth \cite{AIM-518} and Haralick \cite{bb20239}) 
 are based on the second derivative of the image, and both take steps to reduce noise before 
@@
 \begin{enumerate}
+	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
 	\item Convolution of the image with:
@@
 
-% agregar comentario de que la diferencia entre ellos radica en un compromiso entre precision en el calculo
-% y costo computacional. Y que se va a explicar mas adelante.
+Then, generation of both Gaussian and LoG kernels is required (see section \ref{sec:appendix1}).\\
 
@@
 \begin{algorithmic}[1]
-\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
+\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
 \STATE $im \leftarrow$ input image
-\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}
-\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
 \IF{Gaussian kernel}
@@
 \ELSE
+	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}\mynf{\label{correction_a:second_derivative:marr_algo}}.
+	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
 	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
@@
 \end{equation}
-where $(x,y)$ are the offsets of each neighborhood point relative to the center point (e.g.\ in a neighborhood of size $5\times5$, $x$ and $y$ take values ​​between $-2$ and $2$).
-
-To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
-neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
-points leads to an overdetermined system, which can be solved using least squares or any other data fitting technique.
-In Haralick's work, this approximation is computed using least squares.
-
-%, and to speed up the calculations the solution is 
-%computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
+To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted.
+As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$.
+Having $10$ coefficients and $25$ data
+points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\
+
+\myn{\label{second_derivative:haralick:statement}} In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel.
+In this way, the center of the neighboorhood is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g.
+in a neighborhood of size 5x5, x and y take values between -2 and +2).\\
+Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficientes are computed in the same way for every pixel in the image.
+Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which is a great speed-up of the algorithm.
+In the following we will state the approximation problem and show how to derive those masks from its solution.\\
 
 Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
-to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
-By substituting the 25 data points into the polynomial equation (\ref{eq:bicubic}), the following system of 
+to find 10 coefficients. As mentioned before this leads to an \tcom{overdetermined} system which can be solved by least squares. 
+By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
 equations is obtained,
+
 \begin{equation*}
@@
 
-Then, the least squares problem can be approximately solved by computing the pseudo inverse of $\mathbf{A}$,
+Then, the least squares problem can be solved by using the normal equations,
+
 \begin{equation*}
-	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
+	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} 
 \end{equation*}
-$\mathbf{B}$ is a $10\times25$ matrix, 
+
+\myn{\label{second_derivative:haralick:normal}} Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
+
+$$ \mathbf{k} = \mathbf{B}\mathbf{f} $$
+
 \begin{equation*}
 	\begin{bmatrix}
-		b_{1,1}		& b_{1,2}	& \cdots	& b_{1,25}	\\
-		b_{2,1}		& b_{2,2}	& \cdots	& b_{2,25}	\\
+		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
+		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
 		\vdots		& \vdots	& \ddots	& \vdots	\\
-		b_{10,1}	& b_{10,2}	& \cdots	& b_{10,25}
+		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
 	\end{bmatrix}
@@
 \end{equation*}
+
 For each coefficient ($i$ from $1$ to $10$),
+
 \begin{equation}
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \cdots + b_{i,25}f_{25}% \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
-\end{equation}
-where
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f} . \mathbf{b_i},\\
+\end{equation}
+
+where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.\myn{\label{correction_a:second_derivative:haralick_bi}}
+
+\begin{equation}
+	\label{eq:b_i_v}
+	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,25}
+					\end{bmatrix}.\\
+\end{equation}
+
+Let us rearrange $b_i$ as a 5x5 matrix and denote it by $\mathbf{mask_i}$ as follows
 \begin{equation}
 	\label{eq:b_i}
-	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \cdots	& b_{i,5}	\\
-									b_{i,6}		& b_{i,7}	& \cdots	& b_{i,10}	\\
+	\mathbf{mask_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
+									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
 									\vdots		& \vdots	& \ddots	& \vdots	\\
-									b_{i,21}	& b_{i,22}	& \cdots	& b_{i,25}	\\
+									b_{i,21}	 & b_{i,22}	& \hdots	& b_{i,25}	\\
 					\end{bmatrix}.\\
 \end{equation}
-%% VER UNA FORMA DE ESCRIBIR TAMBIEN LOS f1...f25 PARA MOSTRAR POR QUE LOS bi,1 ...
-bi,25 ESTAN DISPUESTOS COMO UNA MATRIZ.
-ES PORQUE LOS f SON UN ENTORNO 5X5 DEL PIXEL, Y SE MULTIPLIAN PUNTO A PUNTO POR LOS b.
-
-Using equation \ref{eq:coefficients} and the masks given in Table \ref{table:b_i}, it is possible to compute 
-the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
-are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.
-
-In order to speed up the calculations, the solution is computed by convolving the image with some precomputed 
-masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$.
-
-\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
-\begin{table}
-\centering
-\subfigure[$\mathbf{b_1}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-425 & 275 & 225 & 275 & 425 \\
-\hline
-275 & 125 &  75 & 125 & 275 \\ 
-\hline
-225 &  75 &  25 &  75 & 225 \\
-\hline
-275 & 125 &  75 & 125 & 275 \\
-\hline
-425 & 275 & 225 & 275 & 425 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_2}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--2260 & -620 & 0 & 620 & 2260 \\ 
-\hline
--1660 & -320 & 0 & 320 & 1660 \\ 
-\hline
--1460 & -220 & 0 & 220 & 1460 \\
-\hline
--1660 & -320 & 0 & 320 & 1660 \\ 
-\hline
--2260 & -620 & 0 & 620 & 2260 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_3}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-2260  &  1660 &  1460 &  1660 &  2260 \\ 
-\hline
-620   &   320 &   220 &   320 &   620 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline
--620  &  -320 &  -220 &  -320 &  -620 \\ 
-\hline
--2260 & -1660 & -1460 & -1660 & -2260 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_4}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-1130 & 620 & 450 & 620 & 1130 \\ 
-\hline
-830  & 320 & 150 & 320 &  830 \\ 
-\hline
-730  & 220 &  50 & 220 &  730 \\
-\hline
-830  & 320 & 150 & 320 &  830 \\ 
-\hline
-1130 & 620 & 450 & 620 & 1130 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_5}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--400 & -200 & 0 &  200 &  400 \\ 
-\hline
--200 & -100 & 0 &  100 &  200 \\ 
-\hline
-0    &    0 & 0 &    0 &    0 \\
-\hline
-200  &  100 & 0 & -100 & -200 \\ 
-\hline
-400  &  200 & 0 & -200 & -400 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_6}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-1130 & 830 & 730 & 830 & 1130 \\ 
-\hline
-620  & 320 & 220 & 320 &  620 \\ 
-\hline
-450  & 150 &  50 & 150 &  450 \\
-\hline
-620  & 320 & 220 & 320 &  620 \\ 
-\hline
-1130 & 830 & 730 & 830 & 1130 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_7}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--8260 & -2180 & 0 & 2180 & 8260 \\ 
-\hline
--6220 & -1160 & 0 & 1160 & 6220 \\ 
-\hline
--5540 &  -820 & 0 &  820 & 5540 \\
-\hline
--6220 & -1160 & 0 & 1160 & 6220 \\ 
-\hline
--8260 & -2180 & 0 & 2180 & 8260 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_8}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-5640  &  3600 &  2920 &  3600 &  5640 \\ 
-\hline
-1800  &   780 &   440 &   780 &  1800 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline
--1800 &  -780 &  -440 &  -780 & -1800 \\ 
-\hline
--5640 & -3600 & -2920 & -3600 & -5640 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_9}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--5640 & -1800 & 0 & 1800 & 5640 \\ 
-\hline
--3600 &  -780 & 0 &  780 & 3600 \\ 
-\hline
--2920 &  -440 & 0 &  440 & 2920 \\
-\hline
--3600 &  -780 & 0 &  780 & 3600 \\ 
-\hline
--5640 & -1800 & 0 & 1800 & 5640 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_{10}}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-8260  &  6220 &  5540 &  6220 &  8260 \\ 
-\hline
-2180  &  1160 &   820 &  1160 &  2180 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline	
--2180 & -1160 &  -820 & -1160 & -2180 \\ 
-\hline
--8260 & -6220 & -5540 & -6220 & -8260 \\
-\hline
-\end{tabular}} \\
-\caption{Masks to compute the coefficients of the bicubic fit.}
-\label{table:b_i}
-\end{table}
+
+\vspace{1ex}
+\myn{\label{second_derivative:haralick:mask_computation}} We will compute one value of $k_i$ for each pixel on the image.
+Thus we can store these coefficientes in a matrix $\bf{K_i}$ of the same size as the original image.
+Then, using the mask in shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$.
+\\
+\begin{equation}
+	\label{eq:convolution_b}
+	\mathbf{K_i} = f \star \mathbf{mask_i} \: \mathtt{for} \: i \in \{1 \ldots 10\}\\
+\end{equation}
+
+As we expressed the approximation in a local coordinate system, the coefficients $b_{i,j}$ will be the same for every pixel, thus, this convolution is performed against a fixed and precomputed mask $\mathbf{mask_i}$.
+The numeric values of these 5x5 masks used in our experiments are shown in Table \ref{table:b_i}.
 
@@
 
-The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
-expression in equation \ref{eq:bicubic}. If just the first order terms of this 
-polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approximated as
+\myn{\label{second_derivative:haralick:analitical:condition}} The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
+Thus, and pixel will lie on an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
+The statement of the idea is presented at the beggining of Section III, in the first column of page 42.
+The three conditions are presented at the end of the same section, in the second column.
+There it reads $f''_{\alpha}(\rho)<0$, $f''_{\alpha}(\rho)=0$ and $f''_{\alpha}(\rho)<0$.
+Clearly, the first and second conditions can't be true at the same time, thus the first one should be $f'''_{\alpha}(\rho)<0$.}:
+
+\begin{enumerate}
+\item the first derivative is non zero.
+\item the second derivative is zero.
+\item the third derivative is negative.
+\end{enumerate}
+
+\myn{\label{second_derivative:haralick:analitical:statement}} After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea is to compute the derivatives of the image in the direction of the gradient using that approximation.\\
+
+Let $\theta$ be the gradient angle, defined with respect clockwise from the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}.
+We \tcom{approximate} it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:
+
 \begin{align}
+\label{eq:sincos}
 	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
-	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}.  
-\label{eq:sincos}
+	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \\
 \end{align}
-Now substituting the variables $x$ and $y$ in polar form as
+
+\myn{\label{second_derivative:haralick:analitical:derivative}} Let us consider the line $r$ passing by $(x,y)$ with angle $\theta$.
+The coordinates of each point $(u,v)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(x,y)$ as follows
+
 \begin{equation*}
-	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} 
+	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
 \end{equation*}
+
 in the bi-cubic polynomial, we obtain
+
 \begin{equation}
@@
 \end{equation}
+
 where
+
 \begin{align}
@@
 	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
-	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). 
+	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). \nonumber \\
 \end{align}
-The derivatives are obtained as follows:
+
+The derivatives are obtained as follows.
+
 \begin{align}
@@
 	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
-	f'''_{\theta}(\rho) = 6C_3 .\nonumber 
+	f'''_{\theta}(\rho) = 6C_3 .\nonumber \\
 \end{align}
-Then, the condition that the second derivative is equal to zero becomes
-\begin{equation}
-	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
-\end{equation}
-where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial to become zero in a neighborhood 
-of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
-value is between $0.4$ and $0.6$.
-
-The other condition required is that the third derivative to be negative, i.e.\
+
+\myn{}The simplest way to impose these conditions is starting with the third one, i.e. that the third derivative has to be negative. Thus,\
+
 \begin{equation}
@@
 \end{equation}
-Given the direction defined by $\theta$, with $\rho>0$, an edge will always be an ascending step. This 
-indicates that the first derivative in this direction has a maximum and that the second derivative has a zero 
-crossing with negative slope. That is, the third derivative less than zero.
+
+By construction $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
+Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing on $\rho=0$.
+$f'_{\theta}(0) \geq 0$.
+This implies that $C_1 \geq 0$.
+
+The second condition that the second derivative is equal to zero becomes
+
+\begin{equation}\label{eq:haralick:cond_b_orig}
+	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0
+\end{equation}
+
+Ideally, if the edge point falls exactly on a grid location, the polynomial function 
+should be zero at $\rho=0$.
+But due to the finite nature of the image grid, the edge could be located at non integer locations.
+Thus, the condition is relaxed asking that the polynomial becomes zero in a neighborhood of radius $\rho_0>0$ centered at $\rho=0$.
+With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes
+
+\begin{equation}
+\exists\rho \in [0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
+\end{equation}
+
+Thus, the solution is $\rhoˆ*=-\frac{C_2}{3C_3} $ and imposing $|\rhoˆ*|<\rho_0$ we obtain the second condition
+
+\begin{equation}
+\left| \frac{C_2}{3C_3} \right| < \rho_0,
+\end{equation}
+
+Finally,  the first condition is that the first derivative is non zero,
+
+\begin{equation}\label{eq:cond:first}
+f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \neq 0 \\
+\end{equation}
+
+If we evaluate the solution $\rhoˆ*$ on it, we obtain
+
+\begin{equation}\label{eq:cond:first}
+f'_{\theta}(\rho) = C_1 - \frac{C_2^2}{C_3}  \\
+\end{equation}
+
+as $C_1>=0$ and $C_3<0$, $f'_{\theta}(\rho)$ will always be positive, thus the first condition always hold by construction and does not need to be verified.\\
+
+In Haralick's work $\rho_0$ is a fixed threshold which acts as an argument to the algorithm.
+\myn{\label{second_derivative:haralick:analitical:parameter_value}}  This parameter must be greater than zero and and less than the size of the grid (i.e.
+$\rho_0<1$).
+In this work we obtained the best results with values between $0.4$ and $0.6$.\\
 
@@
 \begin{algorithmic}[1]
-\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
+\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.BibTex | RIS | RefWorks
 \STATE $im \leftarrow$ input image
@@
 	\FOR{$j=1$ \TO $10$}
-		\STATE $k_j \leftarrow$ $\sum_{n,m}neighbors[n,m]mask_j[n,m]$ % convolution($neighbors$,$mask[j]$)
+		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
 	\ENDFOR
@@
 
-All algorithms implemented use some common mathematical operations that are independent of the algorithms themselves. 
+\myn{\label{}} All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
 These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
@@
 
-Gaussian kernel is generated by sampling the 2-D Gaussian function (centered at $(0,0)$)
+The Gaussian convolution mask is generated by sampling the 2-D Gaussian function (centered at $(0,0)$)
 %\footnote{Note that for simplicity the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
@@
 	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
-	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
+	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of $n$ is 
 large enough to have a good approximation of the Gaussian function in the kernel.}
@@
 can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
-one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
+one convolution with the input image (unlike the Gaussian kernel case, in which two convolutions 
 have to be performed, one with the kernel and another with the Laplacian operator), but the kernel 
-support must be greater in order to obtain the same precision..
+support must be greater in order to obtain the same precision.
 
@@
 \begin{equation} 
+	\frac{\partial}{\partial y}G(x,y)=-\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
+\end{equation}
+\begin{equation} 
 	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
@@
 \begin{equation} 
-	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
+	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
 \end{equation}
@@
 
-\begin{SCfigure}[][!t]
+\begin{SCfigure}[][!b]
 	\centering
 	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
-	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
+	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of $n$ is sufficient 
 to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
@@
 As mentioned before, the size $n$ of the kernel and the standard deviation $\sigma$ of the exponential function 
-are not independent of each one. This is because the function LoG has wider support than the
-Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to 
-generate a correctly sampled LoG kernel than in the case of the Gaussian kernel, with the same $\sigma$.\\
-
-Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
-required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
+are not independent. For the same $\sigma$ value , the function LoG has wider support than the
+Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to correcly sample the LoG kernel.\\
+
+Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. It is clear that we must use a larger kernel size in the LoG function case (approximately 18\% larger). For 
 example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{kernels.pdf}
+	\includegraphics[width=0.8\textwidth]{kernels.pdf}
 	\caption{Comparison of the Gaussian and LoG functions.}
@@
 
-When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
-around the edges of the image, and to get a valid output the same size as input image. In this paper, 
-two methods were implemented: zero-padding and boundary reflection. Zero-padding implies complete 
-the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
-kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
-relative to the edge of the image.
+When making a convolution, it is necessary to define the boundary conditions used to compute it near the edges of the image, and to get a valid output the same size as input image. In this paper, 
+two methods were implemented: zero-padding and boundary reflection.
+Zero-padding completes the borders of the image with valued pixels.
+Reflection completes those pixels with the corresponding symmetric pixel value  relative to the edge of the image.
 Direct convolution is not the only way of filtering an image with a kernel.
-There are other methods such as FFT convolution or recursive filtering\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
-
-%-------------------------------------------------------------------------------
-
-\begin{comment}
-
-\section{Implementation}
-\label{sec:appendix2}
-
-In this section a detailed explanation of the source code of the different implemented algorithms
-is presented. \\
-
-To the best of our knowledge, the only freely available codes implementing these algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
-There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
-All these links were last checked in July 2012.
-\\
-
-%This code documentation was generated using a perl script to extract some parts of the 
-%code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
-
-%Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
-
-Required external libraries: 
-\begin{itemize}
-	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
-	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
-\end{itemize}
-
-\subsection{Roberts, Prewitt and Sobel}
-\input{test_fded.tex}
-
-\subsection{Marr-Hildreth}
-\label{app:marr-hildreth}
-\input{test_mh.tex}
-
-\subsection{Haralick}
-\input{test_haralick.tex}
-
-\subsection{Gaussian kernel generation}
-\input{gaussian_kernel.tex}
-
-\subsection{2D convolution}
-\input{2dconvolution.tex}
-
-\end{comment}
+There are other methods such as FFT convolution or recursive filtering.%\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
+
+%-------------------------------------------------------------------------------
 
@@
 
-The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used.
+The results of each algorithm are first shown for a simple image composed of a white square on a black background, see Figure \ref{fig:original1}.
 
@@
 
-Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
-The execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
+Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm (including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.
+Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}).
 
@@
 	\centering
-	\subfigure[Roberts. $th=0.1$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
-	\quad
-	\subfigure[Prewitt. $th=0.1$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
-	\quad
-	\subfigure[Sobel. $th=0.1$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}
-
-	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
-	\quad
-	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
-	\quad
-	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
+	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
+	\quad
+	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
+	\quad
+	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}
+
+	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$. $\text{Exec-time}:0.03s$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
+	\quad
+	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$. $\text{Exec-time}:0.05s$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
+	\quad
+	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.04s$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
 	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.
@@
 
-\begin{table}[h!]
-	\begin{center}
-	\begin{tabular}{| l | r |}
-		\hline \rule{0pt}{3ex}
-		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
-		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
-		Haralick									& $0.040 \ s$										\\
-		\hline
-	\end{tabular}
-	\end{center}
-	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (127$\times$127 pixels, 1 channel).}
-	\label{exectime1}
-\end{table}
-\vspace{0.5cm}
-
-It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster.
+In this simple case, the first derivative edge detection algorithms run better and faster.
 All algorithms show consistent results.
 Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.
-
-Now the performance of each algorithm is analyzed, using a natural image shown in Figure \ref{fig:original2}.
 
@@
 
-First the results of the first derivative edge detection algorithms is shown Figures \ref{fig:result2-a} to \ref{fig:result2-c}.
+Now the performance of each algorithm is analyzed using a natural image shown in Figure \ref{fig:original2}. 
+First the results of the first derivative edge detection algorithms are shown Figures \ref{fig:result2-a} to \ref{fig:result2-c}.
 Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm.
@@
 	\centering
-	\subfigure[Roberts. $th=0.1$.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
-	\quad
-	\subfigure[Prewitt. $th=0.1$.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}
-
-	\subfigure[Sobel. $th=0.1$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
-	\quad
-	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
-
-	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
-	\quad
-	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
+	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
+	\quad
+	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}
+
+	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
+	\quad
+	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$. $\text{Exec-time}:1.05s$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
+
+	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$. $\text{Exec-time}:1.44s$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
+	\quad
+	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.93s$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
 	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using Figure \ref{fig:original2} as input image.}
@@
 In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative.
-It can be seen in Figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail.
+For each image in Figure \ref{fig:result3}, we show an area of ​​interest, enlarging it to have a better view of the details.
 Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded).
@@
 
+\clearpage
+
 %-------------------------------------------------------------------------------
@@
 
+\clearpage
+
 Some observations: 
 \begin{itemize}
-	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity of the edges.
-They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g.
+	\item The first derivative algorithms, although they work properly and run fast, have some problems like discontinuity of the edges.
+They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to avoid and improve this, e.g.
 Canny edge detector \cite{Canny1986}).
+\\
 	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
@@
 This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}).
-	\item As mentioned earlier, Haralick's algorithm is the first to assume higher degree of regularity in the neighborhood of a pixel (third order).
+\\
+	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order).
 This causes thicker edges, as shown in both examples.
@@
 These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods.
+\\
 \end{itemize}
@@
 
+\begin{comment}
 \subsection{Video}
+
+As supplementary material, there are videos testing the frame by frame application of these algorithms.
 
@@
 \end{multicols}
+\end{comment}
+%-------------------------------------------------------------------------------
 
 \clearpage
+
 %-------------------------------------------------------------------------------
@@
 The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation.
-Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
+They also run extremely fast, because their main operation is a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
 The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges.
-Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make more fair the comparison with the second derivative methods.
-
-The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
+Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make the comparison fairer with the second derivative methods.
+
+The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed.
 In real images, they have better behavior than first derivative algorithms.
 
-Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one.
+Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs \tcom{significantly} faster that the LoG one.
 The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
 It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
+\\
 
@@
 
-As an addendum to this work a detailed implementation of the described algorithms is presented, 
-available in C language, along with an online demo where users can run the algorithms 
-with arbitrary images uploaded by themselves.
+%As an addendum to this work a detailed implementation of the described algorithms is presented, 
+%available in C language, along with an online demo where users can run the algorithms 
+%with arbitrary images uploaded by themselves.
 
 \clearpage
-
-%-------------------------------------------------------------------------------
-
+\appendix
+\section{Haralick convolution masks}
+\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
+\begin{table}[h]
+\small
+\centering
+\subfigure[$\mathbf{mask_1}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+425 & 275 & 225 & 275 & 425 \\
+\hline
+275 & 125 &  75 & 125 & 275 \\ 
+\hline
+225 &  75 &  25 &  75 & 225 \\
+\hline
+275 & 125 &  75 & 125 & 275 \\
+\hline
+425 & 275 & 225 & 275 & 425 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_2}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-2260 & -620 & 0 & 620 & 2260 \\ 
+\hline
+-1660 & -320 & 0 & 320 & 1660 \\ 
+\hline
+-1460 & -220 & 0 & 220 & 1460 \\
+\hline
+-1660 & -320 & 0 & 320 & 1660 \\ 
+\hline
+-2260 & -620 & 0 & 620 & 2260 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_3}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+2260  &  1660 &  1460 &  1660 &  2260 \\ 
+\hline
+620   &   320 &   220 &   320 &   620 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-620  &  -320 &  -220 &  -320 &  -620 \\ 
+\hline
+-2260 & -1660 & -1460 & -1660 & -2260 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_4}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+1130 & 620 & 450 & 620 & 1130 \\ 
+\hline
+830  & 320 & 150 & 320 &  830 \\ 
+\hline
+730  & 220 &  50 & 220 &  730 \\
+\hline
+830  & 320 & 150 & 320 &  830 \\ 
+\hline
+1130 & 620 & 450 & 620 & 1130 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_5}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-400 & -200 & 0 &  200 &  400 \\ 
+\hline
+-200 & -100 & 0 &  100 &  200 \\ 
+\hline
+0    &    0 & 0 &    0 &    0 \\
+\hline
+200  &  100 & 0 & -100 & -200 \\ 
+\hline
+400  &  200 & 0 & -200 & -400 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_6}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+1130 & 830 & 730 & 830 & 1130 \\ 
+\hline
+620  & 320 & 220 & 320 &  620 \\ 
+\hline
+450  & 150 &  50 & 150 &  450 \\
+\hline
+620  & 320 & 220 & 320 &  620 \\ 
+\hline
+1130 & 830 & 730 & 830 & 1130 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_7}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-8260 & -2180 & 0 & 2180 & 8260 \\ 
+\hline
+-6220 & -1160 & 0 & 1160 & 6220 \\ 
+\hline
+-5540 &  -820 & 0 &  820 & 5540 \\
+\hline
+-6220 & -1160 & 0 & 1160 & 6220 \\ 
+\hline
+-8260 & -2180 & 0 & 2180 & 8260 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_8}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+5640  &  3600 &  2920 &  3600 &  5640 \\ 
+\hline
+1800  &   780 &   440 &   780 &  1800 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-1800 &  -780 &  -440 &  -780 & -1800 \\ 
+\hline
+-5640 & -3600 & -2920 & -3600 & -5640 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_9}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-5640 & -1800 & 0 & 1800 & 5640 \\ 
+\hline
+-3600 &  -780 & 0 &  780 & 3600 \\ 
+\hline
+-2920 &  -440 & 0 &  440 & 2920 \\
+\hline
+-3600 &  -780 & 0 &  780 & 3600 \\ 
+\hline
+-5640 & -1800 & 0 & 1800 & 5640 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_{10}}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+8260  &  6220 &  5540 &  6220 &  8260 \\ 
+\hline
+2180  &  1160 &   820 &  1160 &  2180 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-2180 & -1160 &  -820 & -1160 & -2180 \\ 
+\hline
+-8260 & -6220 & -5540 & -6220 & -8260 \\
+\hline
+\end{tabular}} \\
+\caption{5x5 masks used to compute the coefficients of the bicubic fit.}
+\label{table:b_i}
+\end{table}
+
+%-------------------------------------------------------------------------------
+
+\clearpage
+\section{Implementation}
+\label{sec:appendix2}
+
+In this section a detailed explanation of the source code of the different implemented algorithms
+is presented. \\
+
+To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
+There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
+All these links were last checked in July 2012.
+\\
+
+This code documentation was generated using a perl script to extract some parts of the 
+code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
+
+Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
+
+Required external libraries: 
+\begin{itemize}
+	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
+	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
+\end{itemize}
+
+\subsection{Roberts, Prewitt and Sobel}
+\input{test_fded.tex}
+
+\subsection{Marr-Hildreth}
+\label{app:marr-hildreth}
+\input{test_mh.tex}
+
+\subsection{Haralick}
+\input{test_haralick.tex}
+
+\subsection{Gaussian kernel generation}
+\input{gaussian_kernel.tex}
+
+\subsection{2D convolution}
+\input{2dconvolution.tex}
+
+%-------------------------------------------------------------------------------
+\clearpage
+\bibliographystyle{siam}
 \bibliography{bibliography}

### 2014-05-26 edge-detection-review-2nd-revision-v2.tex
changed url of demo

@@
 % by Haldo Spontón & Juan Cardelino
-% IPOL 2012
+% IPOL 2014
 %-------------------------------------------------------------------------------
@@
 \documentclass{ipol}
-\ipolSetYear{2012}
-\ipolSetMonth{06}
-\ipolSetDay{06}
+\ipolSetYear{2014}
+\ipolSetMonth{05}
+\ipolSetDay{26}
 %\ipolSetID{arced}
@@
 the IPOL publication of this article. An 
-\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration}\footnote{\url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}} is also available, where 
+\href{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}{online demonstration}\footnote{\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}} is also available, where 
 you can test and reproduce our results.

### 2014-05-29 review/review_answer_2.tex
added new idol class and corrected minor issues

@@
 \item improved code readme and demo.
+\item  Changed the location of the demo, it is now located at: \url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}
 \end{itemize}
@@
 
-\que You mention a video but it was not included in the supplementary les (35-757-1-SP.zip and 35-758-1-SP.zip) so that I could not see it.\\
+\que You mention a video but it was not included in the supplementary materials (35-757-1-SP.zip and 35-758-1-SP.zip) so that I could not see it.\\
 
@@
 \que "Gaussian kernel is generated by sampling the"\\
-\ans YY\\
+\ans Corrected\\
 \que "unlike the case of Gaussian kernel"\\
@@
 \subsubsection{Demo}
-The url \url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/} provided in the preprint
+\que The url \url{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/} provided in the preprint
 contains a broken demo. This is not very important though since the demo code
 provided in the submission works correctly and it should eventually replace the
-one specified in the pdf.
+one specified in the pdf.\\
+
+\ans The new demo is located  here \url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}\\
 

### 2014-11-03 review/second_round/review_answer_3.tex
First round of corrections for 3rd review. Changed almost every text change required. code still needs to be fixed

\documentclass[a4paper,10pt]{report}

\usepackage[english]{babel}
%\usepackage[none]{hyphenat}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[pdftex]{hyperref}
%\usepackage[tight,scriptsize]{subfigure}
%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}
\usepackage{jfc}
\hypersetup{hypertexnames, plainpages=false,colorlinks=true}
\usepackage{amsmath}

% Title Page
\title{}
\author{}

\usepackage{xr}
\externaldocument{../../edge-detection-review-3rd-revision-v1}
\reversemarginpar

\newcounter{myquestion}
\newcommand{\ml}{\mathcal{L}}
\newcommand{\ms}{\mathcal{S}}

\begin{document}
\maketitle

%\begin{abstract}
%\end{abstract}

\chapter{Introduction}
In this communication we answer to the third round of review of the article "A review of Classic edge detectors", submitted to the Image Processing Online Journal.
In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided. In the following we address the issues raised by the reviewers. For each issue, an answer is provided and a cross-reference with a red tag $C_n$ (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided. In addition, changes made to the paper are highlighted in red (when significative) in the new version of the paper. 
These red marks will be removed for the final version.

The remarks from the reviewers are marked with the tag \textcolor{blue}{QUESTION} and our answers with the tag \ans.

%\section{Brief summary of changes}
\section{General changes}
\begin{itemize}
 \item Improved instructions in makefile to compile in OSX (libpng headers not found)
 \item Quality of all synthetic images (kernels) has been improved

\end{itemize}

\chapter{Answers to reviewer \#1}

\section{Summary}

\que This paper reviews classic edge detectors based on the first and second derivatives of the image. It focuses
on three first-derivative based algorithms (namely Sobel, Prewitt and Roberts) and two second-derivative
based algorithms (Marr-Hildreth and Haralick). The paper itself explains all the implementation details.
The code provides the corresponding implementation.\\

\section{Paper review}

\que In comparison with the original submission the paper has improved a lot: many mistakes have been corrected
and ambiguities clarified. I have no remark on the scientific content anymore. YET the paper has still a
presentation issue. It was still not rigorously proofread, there are a lot of mistakes still. The presentation of the figures also needs to be improved.\\

\ans Thank you very much, both reviewers contributed a lot on improving the paper.

\subsection{Minor remarks}
 Again I would recommend proofreading the whole paper, there are still a lot of spelling
mistakes. I stopped writing them all down, but this list is not at all exhaustive, do not restrict to only
correcting these mistakes.

\que C3: continuos -> continuous\\
\ans fixed\\
\que C3: where a pixel is evenly sample on a grid \\
\ans fixed\\
\que p5: take value -> takes value \\
\ans fixed.\\
\que  p10: polinomial \\
\ans fixed in several places\\
\que p10 coefficientes \\
\ans fixed in several places\\
\que p12 Thus, and pixel\\
\ans fixed.\\
\que Algorithm 3: BibTex RIS RefWorks ???\\
\ans fixed.\\
\que Algorithm 3, line 2: 5x5 please write 5 x 5
\ans fixed\\
\que Fig. 9,10,11: the positioning of the caption is really strange, besides the resolution of these images is not sufficient for a display at this size, the text in the images appears blurry. must be greater"\\
\ans Changed captions to standard places. Changed to vectorial formats, now size is correct.\\
\que instead a Gaussian kernel\\
\ans fixed\\
\que auxiliar\\
\ans fixed\\
\section{Code review}

\que Compared to the first submission, I was unable to compile the new code (ubuntu 14.04, gcc 4.8.2)
\begin{verbatim}
$ make
c99 -o edges edges . c io png . c c l a s s i c e d g e d e t e c t o r s . c -lpng -lm
c l a s s i c e d g e d e t e c t o r s . c : In func t i on g a u s s i a n k e r n e l :
c l a s s i c e d g e d e t e c t o r s . c : 7 0 : 5 0 : e r r o r : M P I unde c lar ed ( f i r s t use in t h i s func t i on )
k e rne l [ i ] = k e rne l [ i ] / ( 2 . 0 * sigma * sigma *M_PI ) ;
^
c l a s s i c e d g e d e t e c t o r s . c : 7 0 : 5 0 : note : each unde c lar ed i d e n t i f i e r i s r epo r t ed only once
f o r each func t i on i t appear s in
make : [ edges ] Er ror 1
\end{verbatim}

\ans This is due a stricter standard compliance in recent versions of gcc, they removed \verb+M_PI+ as standard constant. It passed a long time since first submission and thus gcc versions changed. Thus we opted for defining it ourselves. Now compiles on that ubuntu version.\\

\que This last comment was not answered and not even mentioned in your response letter:
l62-90, you can precompute $\sigma^2$ and $\sigma^2$ outside of the scopes to avoid unnecessary and inelegant code
repetition.
\ans  XXX\\

\subsection{ Demo review}
\que I was able to play with the demo and it works fine. I would only ask for a link download all results on the
result page in order to download an archive of all images.

\ans While I agree with the usefulness of this feature for a single reviewer, I doubt about it when demo goes in production. In that case (see LSD demo) the size of the archive could be considerably big. In addition, I guess this is the kind of feature that should be enabled centrally in the demo system and not in every single demo. Finally, as the demo can change after the review, I suggest to keep it like this and go on discussing the feature, eventually we could add it later.\\

\subsection{ Conclusion}
This paper is almost ready but the presentation and the compilation problem still make a revision necessary.

\chapter{Answers to reviewer \#2}

Dear authors,

\que I took notice of your corrections. I must say that I have mixed feelings
about them. I will simply state my major sources of concerns here. Please
have a look at my annotated pdf for more details.\\
\ans We numbered each of your concerns in the PDF with the tag PDFn, where n is an index and will address them in order in the PDF Comments section.\\

\que On the one hand you did improve the only troublesome theoretical part, that
was  the Haralick algorithm description. I believe you did a fine job in
this task, although there are still some minor issues (see annotations).\\

\ans Thanks, we further improved the presentation based on your annotations and a more in depth reading of those sections.\\

\que On the other hand, you did not take any account of the issues I raised
concerning the experimental part. I therefore recap the comments I made the
previous time here:

\begin{itemize}

\item lack of in-depth study of the impact of parameters

\item lack of validation or challenging of the claims of original authors

\item  poor explanation of the thickening observed in the Haralick case.
\end{itemize}

\que
I am also concerned with a new issue of importance: It seems like with the
demo I get very different results  from the one presented in your text. The
difference are particularly obvious for the haralick filter. See my comment
in figure 15(f).

\que Furthermore, I am quite disappointed by your refusal to improve your
conclusion.\\

\section{PDF comments}

\que PDF1 results in\\
\ans done.\\

\que PDF2  since $b_i$ is a row vector and $f$ a column vector, you should rather use a matrix multiplication notation
$k_i=b_i f$
\ans Agreed. Changed in that sense.\\

\que PDF3 english: remove the first in\\
\ans done.\\

\que PDF4: Much better!\\
\ans Thanks! Your contributions helped a lot.

\que PDF5: I suggest mentioning right away that all this is done to detect ascending edges. For instance: … is to find ascending steps by looking for "negative slopes …"\\
\ans I Agree, thanks for the suggestion. I phrased it slightly different, because Haralick's work actually never mentions this explicitely, it is just deduced from the way he poses the problem.\\

\que PDF6: english\\
\ans fixed.\\

\que PDF7: notation issue: $x,y$ are different from the previous sentence. Here they are generic coordinates, while in the previous sentence they were a specific point (the one obtained for $\rho=0$. You could use $x(\rho),y(\rho)$ to remove the ambiguity.\\
\ans Actually, we explained at first that we will normalize coordinates to be offsets from the central point. Thus from that paragraph and below, we can safely assume that x,y is the point in the line and (0,0) is the center. I rewrote many parts of the section to reflect this. See corrections XXX.\\

\que PDF8:  would be nice to avoid making a typo in a footnote referring to a typo, unless you want things to get really messy! This should read $f_\alpha^{'}(\rho)\neq 0$.\\
\ans Oh well! This is a shame, I fixed it. Sorry for that!\\

\que PDF9: The conditions are not mentioned in the previous sentence, so you cannot use "these" to refer to them. You should say, impose the above-mentioned conditions\\
\ans done.\\

\que PDF10: at or "in a neighborhood of"\\
\ans it is in a neighborhood. corrected it and added some missing words there\\

\que PDF11: I think you have to remove the positivity constraint. \\
\ans I don't think so, $\rho$ is always positive because it is the modulus on polar coordinates. In addition, there is no conflict because $C3<0$, thus the equation is solvable.\\

\que PDF12: we substitute $\rho^*$ in this formula'\\
\ans Here I've chosen an intermediate writing between your proposal and the previous writing. I think it takes both your suggestion and my idea into account.\\

\que PDF13: Why did this change since the previous submission? It used to  be $\frac 1{\sqrt 2}$\\
\ans Because I think the student author, who originally wrote the paper, wrote this section based on a derived work, but not directly on Haralick's paper. Reviewing this, I found no mention on the original work that the value should be $\frac 1{\sqrt 2}$. It only mentions that it should be smaller than the grid size (quote from page 42, end of section III: "If for some $\rho$, $|\rho|<\rho_0$, where $\rho_0$ is slightly smaller that the length of the side of a pixel...") , and being strict, the minimum grid size is 1 on the horizontal and vertical dimensions, thus $\rho<1$. The only explanation I find for the confusion is that they tried to use half of the diagonal length of a pixel, but that makes no sense to me.\\

\que PDF14: Caption below the figure.\\
\ans done.\\

\que PDF15: It is\\
\ans done\\

\que PDF16: The value of $n$ needs to be large enough\\
\ans done\\

\que PDF17: below the figure as usually done.
Also redo the figure with larger text.\\
\ans done, and changed to a vectorial format to improve presentation\\

\que PDF18: Uggly, put the caption below the figure as is usaually done. Also, can you improve the resolution of the figure.\\
\ans Done.\\

\que PDF19: This section is still very poor. 
You did not take any of my comments into consideration, not even the merging of small paragraphs into larger ones.
I therefore recap the comments I made the previous time here (see the original ones for more details)
-  lack of in-depth study of the impact of parameters
- lack of validation or challenging of the claims of original authors
- poor explanation of the thickening observed in the Haralick case.

I am also concerned with a new issue of importance: It seems like with the demo I get very different results  than the one presented here. The difference are particularly obvious for the haralick filter. See my comment in figure 15(f)

\ans XXX\\

\que PDF20: This experiment cannot be faithfully reproduced in the demo because figure 12 is not in the input list

\ans XXX\\

\que PDF21: running the same experiment in the demo, I get a much noiser result. See
http://dev.ipol.im/~juanc/ipol_demo/sc_edges/tmp/7A8BE1B97C40DAA903141EA7906D527E/out_haralick.png

\ans XXX\\

\que You did not answer my point here. I think I misunderstood you in the first place, but this is because your explanation lacks details. Therefore I am still not satisfied by this explanation. 
The way you put it is misleading, because you give the impression to think that the fattening of the edge is due to smoothing. If you truly believe this, then I think that you are wrong.

Another way to interpret your explanation is that the spline interpolation if applied to sharp edges (that is when the regularity assumption is at fault) then ringing occurs which creates many duplicated edges. If you agree that this is the right explanation, and that was what you meant in the first place, please improve the formulation of your explanation. And illustrate it graphically.

Otherwise, if you think that the first explanation is the right one, then  add an experiment to verify who is right and who is wrong. The experiment could consist in  plotting  the profile of a synthetic interpolated edge. Take a perfect step as input, and see what you get. You should definitely observe ringing!

\ans XXX\\

\bibliographystyle{unsrt}
\bibliography{../../tools/journal_list_short,../../tools/references,../../tools/hierarchical_segmentation}
\end{document}

### 2014-11-03 edge-detection-review-3rd-revision-v1.tex
First round of corrections for 3rd review. Changed almost every text change required. code still needs to be fixed

--- 2014-05-26 edge-detection-review-2nd-revision-v2.tex
@@
 \documentclass{ipol}
-\ipolSetYear{2014}
-\ipolSetMonth{05}
-\ipolSetDay{26}
 %\ipolSetID{arced}
@@
 
-Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms.  for 
-. The basic idea of edge detection is to detect abrupt changes in image intensity. 
+Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms.
+The basic idea of edge detection is to detect abrupt changes in image intensity.
 Detecting those changes can be accomplished using first or second order derivatives. \\
@@
 Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure.
-\myn{\label{correction_a:first_derivative:first_deriv_rephrase}\\
+\\
 
@@
 
-\myn{\label{correction_a:first_derivative:taylor}}The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$.
-Thus $f'(x)$ is computed as
+The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
 \begin{equation}
@@
 \end{equation}
-\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid.
+Digital images are commonly seen as a discretization of a continuous image, where a pixel is evenly sampled on a grid.
 If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$.
@@
 
-\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}.
+Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}.
 However, in order to be able to compute diagonal edges, 2-D masks are needed.
@@
 The \textit{Roberts operators} are one of the earliest 
-attempts to use 2-D masks for this purpose.
-These operators are based on computing the diagonal \tcom{differences} \myn{\label{correction_a:typo}} implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
-By convention
+attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal differences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
 in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
@@
 
-\begin{SCfigure}[2][!h]
+\begin{figure}[2][!h]
 	\centering
@@
 	\label{fig:1dmasks}
-\end{SCfigure}
-
-\begin{SCfigure}[1][!h]
+\end{figure}
+
+\begin{figure}[1][!h]
 	\centering
@@
 	\label{fig:roberts}
-\end{SCfigure}
+\end{figure}
 
@@
 
-\begin{SCfigure}[1][!h]
+\begin{figure}[1][!h]
 	\centering
@@
 	\label{fig:prewitt}
-\end{SCfigure}
-
-\myn{\label{correction_a:first_derivative:sobel}} The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
+\end{figure}
+
+The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
 $[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]^T$.
@@
 	\label{fig:sobel}
-\end{SCfigure}
+\end{figure}
 
@@
 
-As mentioned before, edges can be seen as abrupt changes in intensity, i.e.\ higher values of gradient.
+As mentioned before, edges can be seen as abrupt changes in intensity, i.e. higher values of gradient.
 An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using Sobel operators.
@@
 
-Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
+Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e. for pixel a at position $i,j$ in the image compute
 \begin{equation*}
@@
 of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
-Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
+Therefore, the parameter $threshold$ in the algorithms takes values ​​between $0$ and $1$.
 
@@
 \ELSE
-	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}\mynf{\label{correction_a:second_derivative:marr_algo}}.
+	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}.
 	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
@@
 
-The surrounding neighborhood of a point $(x,y)$ in the image $f$ is approximated using the following bi-cubic polynomial
-\begin{equation}
-	\label{eq:bicubic}
-	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
-\end{equation}
+The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
+\begin{align}
+	\label{eq:bicubic:long}
+	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  +\\
+	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3 \\
+\end{align}
 To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted.
@@
 
-\myn{\label{second_derivative:haralick:statement}} In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel.
-In this way, the center of the neighboorhood is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g.
-in a neighborhood of size 5x5, x and y take values between -2 and +2).\\
-Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficientes are computed in the same way for every pixel in the image.
-Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which is a great speed-up of the algorithm.
+In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel.
+In this way, the center of the neighboorhood $(x_0,y_0)$ is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g.
+in a neighborhood of size $5\times5$, x and y take values between -2 and +2).
+Thus, the expression of the polynomial is\\
+
+\begin{equation}
+	\label{eq:bicubic}
+	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
+\end{equation}
+
+Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficients are computed in the same way for every pixel in the image.
+Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which results in a great speed-up of the algorithm.
 In the following we will state the approximation problem and show how to derive those masks from its solution.\\
 
-Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
-to find 10 coefficients. As mentioned before this leads to an \tcom{overdetermined} system which can be solved by least squares. 
-By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
+Consider 25 points in a small neighborhood the point $(0,0)$.
+This gives us an equal number of equations to find 10 coefficients.
+As mentioned before this leads to an overdetermined system which can be solved by least squares.
+By substituting the 25 data points into the polynomial equation (\ref{eq:bicubic}), the following system of 
 equations is obtained,
@@
 
-\myn{\label{second_derivative:haralick:normal}} Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
+Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
 
@@
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f} . \mathbf{b_i},\\
-\end{equation}
-
-where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.\myn{\label{correction_a:second_derivative:haralick_bi}}
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{b_i}\mathbf{f} ,\\
+\end{equation}
+
+where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.
 
@@
 
-Let us rearrange $b_i$ as a 5x5 matrix and denote it by $\mathbf{mask_i}$ as follows
+Let us rearrange $b_i$ as a $5\times5$ matrix and denote it by $\mathbf{mask_i}$ as follows
 \begin{equation}
@@
 \vspace{1ex}
-\myn{\label{second_derivative:haralick:mask_computation}} We will compute one value of $k_i$ for each pixel on the image.
-Thus we can store these coefficientes in a matrix $\bf{K_i}$ of the same size as the original image.
-Then, using the mask in shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$.
+We will compute one value of $k_i$ for each pixel on the image.
+Thus we can store these coefficients in a matrix $\bf{K_i}$ of the same size as the original image.
+Then, using the mask shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$.
 \\
@@
 
-\myn{\label{second_derivative:haralick:analitical:condition}} The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
-Thus, and pixel will lie on an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
+The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
+This means that it is actually looking for ascending step edges.
+According to Haralick's definition, a pixel will lie on such an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
 The statement of the idea is presented at the beggining of Section III, in the first column of page 42.
 The three conditions are presented at the end of the same section, in the second column.
-There it reads $f''_{\alpha}(\rho)<0$, $f''_{\alpha}(\rho)=0$ and $f''_{\alpha}(\rho)<0$.
+There it reads $f''_{\alpha}(\rho)<0$, $f''_{\alpha}(\rho)=0$ and $f'_{\alpha}(\rho)\neq 0$.
 Clearly, the first and second conditions can't be true at the same time, thus the first one should be $f'''_{\alpha}(\rho)<0$.}:
@@
 
-\myn{\label{second_derivative:haralick:analitical:statement}} After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea is to compute the derivatives of the image in the direction of the gradient using that approximation.\\
-
-Let $\theta$ be the gradient angle, defined with respect clockwise from the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}.
-We \tcom{approximate} it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:
+After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea is to compute the derivatives of the image in the direction of the gradient using that approximation.\\
+
+Let $\theta$ be the gradient angle, defined clockwise with respect to the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}.
+We approximate it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:
 
@@
 
-\myn{\label{second_derivative:haralick:analitical:derivative}} Let us consider the line $r$ passing by $(x,y)$ with angle $\theta$.
-The coordinates of each point $(u,v)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(x,y)$ as follows
-
-\begin{equation*}
-	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
-\end{equation*}
-
-in the bi-cubic polynomial, we obtain
+Let us consider the line $r$ passing by $(0,0)$ with angle $\theta$.
+The coordinates of each point $(x,y)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(0,0)$ as follows
+
+\begin{align*}
+	x = \rho\cos{\theta} \\
+	y = \rho\sin{\theta} 
+\end{align*}
+
+If we substitute the above expressions in the bi-cubic polynomia of eq., we obtain
 
@@
 
-\myn{}The simplest way to impose these conditions is starting with the third one, i.e. that the third derivative has to be negative. Thus,\
+The simplest way to impose the above-mentioned conditions is starting with the third one, i.e. that the third derivative has to be negative. Thus,
 
@@
 By construction $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
-Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing on $\rho=0$.
-$f'_{\theta}(0) \geq 0$.
-This implies that $C_1 \geq 0$.
+Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$.
+This in turn implies that $C_1 \geq 0$.
 
@@
 
-If we evaluate the solution $\rhoˆ*$ on it, we obtain
+If we evaluate $f'_{\theta}$ on the solution $\rhoˆ*$ using the above formula, we obtain
 
 \begin{equation}\label{eq:cond:first}
-f'_{\theta}(\rho) = C_1 - \frac{C_2^2}{C_3}  \\
+f'_{\theta}(\rhoˆ*) = C_1 - \frac{C_2^2}{C_3}  \\
 \end{equation}
@@
 In Haralick's work $\rho_0$ is a fixed threshold which acts as an argument to the algorithm.
-\myn{\label{second_derivative:haralick:analitical:parameter_value}}  This parameter must be greater than zero and and less than the size of the grid (i.e.
-$\rho_0<1$).
-In this work we obtained the best results with values between $0.4$ and $0.6$.\\
+This parameter must be greater than zero and and less than the size of the grid (i.e. $\rho_0<1$). In this work we obtained the best results with values between $0.4$ and $0.6$.\\
 
@@
 \begin{algorithmic}[1]
-\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.BibTex | RIS | RefWorks
+\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
 \STATE $im \leftarrow$ input image
-\STATE Define the ten $5x5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
+\STATE Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
 \FORALL{pixel $i$ in image $im$}
@@
 
-\myn{\label{}} All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
+All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
 These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
@@
 input parameters, but these are not strictly independent of each other. 
-A value of $n$ large enough is needed to ensure that no information is lost when creating the kernel $G$.
+The value of $n$ needs to be large enough to ensure that no information is lost when creating the kernel $G$.
 To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$.
@@
 
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
-	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of $n$ is 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.8\textwidth]{kernel_gaussian.pdf}
+	\caption{Gaussian kernel, $\sigma=4$, $n=25$. It is easy to see that the selected value of $n$ is 
 large enough to have a good approximation of the Gaussian function in the kernel.}
 	\label{fig:gaussian_kernel}
-\end{SCfigure}
+\end{figure}
 
@@
 
-\begin{SCfigure}[][!b]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.8\textwidth]{kernel_log.pdf}
 	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of $n$ is sufficient 
-to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
+to have a good approximation of the LoG function in the kernel, but it is greater than in the case of 
 Gaussian kernel.}
 	\label{fig:log_kernel}
-\end{SCfigure}
+\end{figure}
 
@@
 
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.8\textwidth]{kernels.pdf}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=1.0\textwidth]{kernels.pdf}
 	\caption{Comparison of the Gaussian and LoG functions.}
 	\label{fig:kernels}
-\end{SCfigure}
+\end{figure}
 
@@
 
-Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs \tcom{significantly} faster that the LoG one.
+Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significantly faster that the LoG one.
 The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
@@
 Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
-It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function.
+It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polynomial function.
 This supposes a certain regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.

### 2014-11-03 kernels.m

@@ -39,7 +39,8 @@ end
 a = figure('Name','Gaussian Kernel','Color','white');
 h1 = subplot(2,3,[1 2 4 5]); surf(x1,y1,G); grid off; axis tight; axis off;
     hold on; plot3(zeros(n1,1),y1(:,1),G(ceil(n1/2),:),'-k','LineWidth',3); 
-    text(0,0,1.1,'Gaussian kernel, 3D view','HorizontalAlignment','center'); hold off;
+    %text(0,0,1.1,'Gaussian kernel, 3D view','HorizontalAlignment','center'); 
+    hold off;
     p = get(h1, 'pos'); 
     p(1) = p(1) - 0.1; p(2) = p(2) + 0.05;
     p(3) = p(3) + 0.1; p(4) = p(4) + 0.1;
@@ -50,7 +51,7 @@ h2 = subplot(2,3,3); pcolor(x1,y1,G); axis square; axis off;
     p(3) = p(3) + 0.1; p(4) = p(4) + 0.1;
     set(h2, 'pos', p);
 subplot(2,3,6); plot(x1(1,:),G(ceil(n1/2),:),'-k','LineWidth',2); 
-    title('Gaussian kernel, 2D view','Position',[0 -0.2]);
+    %title('Gaussian kernel, 2D view','Position',[0 -0.2]);
     hold on; annotation('arrow',[0.65 0.95],[0.11 0.11],'LineWidth',1);
     annotation('arrow',[0.798 0.798],[0.1 0.5],'LineWidth',1);
     %annotation('arrow',[0.45 0.72],[0.30 0.20],'Color','red');
@@ -60,7 +61,8 @@ print(a,'-depsc','kernel_gaussian.eps');
 b = figure('Name','LoG Kernel','Color','white');
 h3 = subplot(2,3,[1 2 4 5]); surf(x2,y2,LoG); grid off; axis tight; axis off;
     hold on; plot3(zeros(n2,1),y2(:,1),LoG(ceil(n2/2),:),'-k','LineWidth',3); 
-    text(0,0,0.14,'LoG kernel, 3D view','HorizontalAlignment','center');hold off;
+    %text(0,0,0.14,'LoG kernel, 3D view','HorizontalAlignment','center');
+    hold off;
     p = get(h3, 'pos'); 
     p(1) = p(1) - 0.1; p(2) = p(2) + 0.05;
[…]

### 2014-11-03 kernels1d.m

@@ -29,10 +29,10 @@ ind_LoG = find(area_LoG>=(1-factor),1,'first');
 h = figure('Name','Kernels comparsion','Color','white','Position',[100 100 800 500]);
 plot(x,G,'-r','Linewidth',2); hold on; 
 plot(x,LoG,'-b','Linewidth',2); axis tight;
-title('Comparison of Gaussian and LoG functions');
+%title('Comparison of Gaussian and LoG functions');
 plot(x(ind_G),G(ind_G),'*r','Linewidth',5);
 plot(x(ind_LoG),LoG(ind_LoG),'*b','Linewidth',5);
-% Líneas:
+% L??neas:
 line([0 0],[min(LoG) max(LoG)],'Color','k');
 line([-1*sigma 5*sigma],[0 0],'Color','k');
 line([x(ind_G) x(ind_G)],[min(LoG) max(LoG)],'Color','r','LineStyle','--');
@@ -44,11 +44,17 @@ line([-1*sigma 5*sigma],[-factor*max(LoG) -factor*max(LoG)],'Color','k','LineSty
 line([0 x(ind_G)],[0.6*max(G) 0.6*max(G)],'Color','k','Linewidth',2);
 line([0 x(ind_LoG)],[0.3*max(G) 0.3*max(G)],'Color','k','Linewidth',2);
 % Referencias
+font_size=14;
 hleg = legend('Gaussian function','LoG function');
-set(hleg,'box','Off');
+set(hleg,'box','Off','FontSize',font_size);
 % Texto:
-text(-0.95*sigma,(factor+0.03)*max(LoG),'\pm3% of maximum absolute value','BackgroundColor','white');
-text(sigma,0.6*max(G),[sprintf('~%1.2f',x(ind_G)/sigma) '\sigma'],'FontSize',12,'BackgroundColor','white');
-text(1.2*sigma,0.3*max(G),[sprintf('~%1.2f',x(ind_LoG)/sigma) '\sigma'],'FontSize',12,'BackgroundColor','white');
+
+text(-0.95*sigma,(factor+0.03)*max(LoG),'\pm3% of maximum absolute value','FontSize',font_size,'BackgroundColor','white');
+text(sigma,0.6*max(G),[sprintf('~%1.2f',x(ind_G)/sigma) '\sigma'],'FontSize',font_size,'BackgroundColor','white');
[…]

### 2014-11-11 review/second_round/review_answer_3.tex
second round of third revision, improving references

@@
 \usepackage{xr}
-\externaldocument{../../edge-detection-review-3rd-revision-v1}
+\externaldocument{../../edge-detection-review-3rd-revision-v2}
 \reversemarginpar
@@
  \item Improved instructions in makefile to compile in OSX (libpng headers not found)
- \item Quality of all synthetic images (kernels) has been improved
+ \item Quality of all synthetic images (kernels) has been improved (See corrections \ref{math:common:gauss},\ref{math:common:log} and \ref{math:common:comparison})
 
@@
 
-\ans Thank you very much, both reviewers contributed a lot on improving the paper.
+\ans Thank you very much, both reviewers contributed a lot on improving the paper. The writing has been reviewed and the figures completely redone.
 
@@
 l62-90, you can precompute $\sigma^2$ and $\sigma^2$ outside of the scopes to avoid unnecessary and inelegant code
-repetition.
-\ans  XXX\\
+repetition.\\
+
+\ans  Although I agree with the concept, I'm not sure about the benefits of this suggestion.
+Actually they are two suggestions.
+The first one has to deal with computational cost, In my experience with optimizing code (which is quite vast), I've learnt not guess before quantifying with a profiler.
+At the current state of compilers I'm quite sure that computation is already optimized out.
+On ther other hand, with respect to the code repetition, it is repeated, but only two times, and that repetition improves readability.
+So I rather keep this in its current form.
+Anyway, I think this is a very minor issue.\\
 
@@
 
+\ans These concerns are addressed  in the answer corresponding to PDF comment PDF19\\
+
 \que
@@
 difference are particularly obvious for the haralick filter. See my comment
-in figure 15(f).
+in figure 15(f).\\
 
-\que Furthermore, I am quite disappointed by your refusal to improve your
-conclusion.\\
+\ans We had two major problems with this submission.
+First, there is a bug in the imagemagick convert command in some unixes (I don't currently remember if it is in ubuntu or in mac).
+This bug made the RGB->grayscale conversion lower the contrast a lot.
+Thus, we had made the experiements for the paper with the good versions and uploaded the bad versions with the code and the demo.
+As these algorithms are not contrast invariant, this generated worst results.
+This is fixed now.
+The second problem was a bug when the other author made a major refactor of the code.
+That bug made that the estimation of the k10 coefficient in Haralick was rubbish and thus the result of this algorithm was very bad.
+This was fixed too.\\
+
+\que Furthermore, I am quite disappointed by your refusal to improve your conclusion.\\
+
+\ans In the first review I wasn't sure about the direction you suggested to improve the conclusions, however after giving them a second thoughts and in the light of the new experiments I took notice of them and added an extensive discussion of the results and a considerable number of conclusions.
+I truly hope you are satisfied now.\\
 
@@
 \que PDF4: Much better!\\
-\ans Thanks! Your contributions helped a lot.
+\ans Thanks! Your contributions helped a lot.\\
 
@@
 Thus from that paragraph and below, we can safely assume that x,y is the point in the line and (0,0) is the center.
-I rewrote many parts of the section to reflect this.
-See corrections XXX.\\
+I rewrote many parts of the section to reflect this.\\
 
@@
 I therefore recap the comments I made the previous time here (see the original ones for more details)
--  lack of in-depth study of the impact of parameters
-- lack of validation or challenging of the claims of original authors
-- poor explanation of the thickening observed in the Haralick case.
+1) lack of in-depth study of the impact of parameters\\
+2) lack of validation or challenging of the claims of original authors\\
+3) poor explanation of the thickening observed in the Haralick case.\\
 
-I am also concerned with a new issue of importance: It seems like with the demo I get very different results  than the one presented here.
+4)I am also concerned with a new issue of importance: It seems like with the demo I get very different results  than the one presented here.
 The difference are particularly obvious for the haralick filter.
-See my comment in figure 15(f)
+See my comment in figure 15(f)\\
 
-\ans XXX\\
+\ans I separated these issued in four points in order to properly answer them.\\
+1) We added two sections with figures showing the effect of various thresholds in single images and we also added a saliency map which summarizes the results of each algorithm across all scales.
+We added examples showing the effect of smoothing for the particular case of Marr.
+See corrections \ref{results:parameter:scale} and \ref{results:parameter:smooth}\\
+2) We discussed most of the most important claims of the original authors. The only exception were the quantitative results of Haralick. See correction corrections \ref{results:discussion} \\
+3) I will answer this on question PDF22.\\
 
-\que PDF20: This experiment cannot be faithfully reproduced in the demo because figure 12 is not in the input list
+\que PDF20: This experiment cannot be faithfully reproduced in the demo because figure 12 is not in the input list\\
 
-\ans XXX\\
+\ans The square image and the rest of images used in the paper have been added to the demo.\\
 
 \que PDF21: running the same experiment in the demo, I get a much noiser result. See
-http://dev.ipol.im/~juanc/ipol_demo/sc_edges/tmp/7A8BE1B97C40DAA903141EA7906D527E/out_haralick.png
+\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/tmp/7A8BE1B97C40DAA903141EA7906D527E/out_haralick.png}\\
+\ans Thanks for pointing this out, we found a bug due a refactor of the code in previous review. It is now tested and validated.\\
 
-\ans XXX\\
-
-\que You did not answer my point here. I think I misunderstood you in the first place, but this is because your explanation lacks details. Therefore I am still not satisfied by this explanation. 
+\que PDF22: You did not answer my point here.
+I think I misunderstood you in the first place, but this is because your explanation lacks details.
+Therefore I am still not satisfied by this explanation.
 The way you put it is misleading, because you give the impression to think that the fattening of the edge is due to smoothing. If you truly believe this, then I think that you are wrong.
@@
 
-\ans XXX\\
+\ans I tried to explain the phenomenon in detail, however, until now I haven't found a detailed explanation that completely satisfies me.
+My preliminary tests does not show ringing.
+I think the degree of the polynomial is too low to show it (only 3 roots).
+The most convincing conjecture I have is due to the way the third condition is relaxed.
+The idea is to allow for solution that are not exactly at $\rho=0$, thus, we \emph{search} for the solution on an neighborhood.
+However this search is naive, because we allow as solutions all values where $|\rho|<|\rho_0|$.
+I see this problem equivalent to finding the zero of a function, if you look for all the x such that $|f(x)<\de|ta$ this will give you a \emph{fat} root.
+
+Anyway, as we have passed all reasonable thresholds for the publication of this paper, and if you don't consider this issue critical, I would propose to leave it as is or remove any explanations that could be missleading.
+Otherwise, we could ask the editor for his opinion on this subject
 

### 2014-11-11 edge-detection-review-3rd-revision-v2.tex
second round of third revision, improving references

--- 2014-11-03 edge-detection-review-3rd-revision-v1.tex
@@
 \newcounter{mynote}% a new counter for use in margin notes
+\usepackage[multidot]{grffile}
 \usepackage{jfc}
@@
 % en este parrafo, puede ser cambiar trough por valley
-In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
+In 1980 Marr and Hildreth \cite{segm:edge_region:marr:80:digital_step} argued that intensity changes are not independent 
 of image scale, so edge detection requires the use of different size 
@@
 
+\myn{\label{intro:objective}}The objective of this work is to provide reproducible implementations of the abovementioned algorithms, along with a comprehensive quantitative evaluation of the behaviour algorithms with respect to the parameters.
+Quantitative benchmarks are outside the scope of this work and will be addressed in future work.
+In addition, we will focus on low-level edge detection algorithms, which means that they will output a disconnected an unordered set of edgels (individual pixel facets where boundaries occur).
+Algorithms that perform edge integration into objects boundaries are of higher level of abstraction and not considered in this work.
+
 Along with this paper, a detailed and well commented source code is presented, which 
@@
 
-The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
+\myn{\label{first:roberts}}The Roberts edge detector was introduced in 1963 by L.G.
+Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on computing the first derivative of the image.
+The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$.
+Thus $f'(x)$ is computed as
 \begin{equation}
@@
 
-Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{roberts1963machine} proposes
+Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{im_proc:segmentation:roberts:1963:boundary_extraction} proposes
 to compute them as shown in Figure \ref{fig:roberts}.
@@
 
-\subsection{The \textit{Prewitt} operators}
+\subsection{The \textit{Prewitt} operators}\label{sec:first:prewitt}
 
@@
 
-\subsection{The \textit{Sobel} operators}
+\subsection{The \textit{Sobel} operators}\label{sec:first:sobel}
 
@@
 
-\subsection{Computation of the edges}
+\subsection{Computation of the edges}\label{sec:first:sobel:computation}
 
@@
 
-The two algorithms presented in this section (Marr-Hildreth \cite{AIM-518} and Haralick \cite{bb20239}) 
+\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:80:digital_step} and Haralick \cite{bb20239}) 
 are based on the second derivative of the image, and both take steps to reduce noise before 
@@
 
-The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
+\myn{\label{second:haralick:fitting:statement}} The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
 \begin{align}
@@
 The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
-This means that it is actually looking for ascending step edges.
+\myn{\label{second:haralick:analytical:def}} This means that it is actually looking for ascending step edges.
 According to Haralick's definition, a pixel will lie on such an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
@@
 
+\myn{\label{second:haralick:analytical:coordinates}} 
 \begin{align*}
-	x = \rho\cos{\theta} \\
-	y = \rho\sin{\theta} 
+	x = \rho\sin{\theta} \\
+	y = \rho\cos{\theta} 
 \end{align*}
@@
 
-Thus, the solution is $\rhoˆ*=-\frac{C_2}{3C_3} $ and imposing $|\rhoˆ*|<\rho_0$ we obtain the second condition
+Thus, the solution is $\rho_*=-\frac{C_2}{3C_3} $ and imposing $|\rho_*|<\rho_0$ we obtain the second condition
 
@@
 
-If we evaluate $f'_{\theta}$ on the solution $\rhoˆ*$ using the above formula, we obtain
-
-\begin{equation}\label{eq:cond:first}
-f'_{\theta}(\rhoˆ*) = C_1 - \frac{C_2^2}{C_3}  \\
-\end{equation}
-
-as $C_1>=0$ and $C_3<0$, $f'_{\theta}(\rho)$ will always be positive, thus the first condition always hold by construction and does not need to be verified.\\
-
-In Haralick's work $\rho_0$ is a fixed threshold which acts as an argument to the algorithm.
-This parameter must be greater than zero and and less than the size of the grid (i.e. $\rho_0<1$). In this work we obtained the best results with values between $0.4$ and $0.6$.\\
+\myn{\label{second:haralick:analytical:first}} If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain
+
+\begin{equation}\label{eq:cond:first:optim}
+f'_{\theta}(\rho_*) = C_1 - \frac{C_2^2}{3C_3}  \\
+\end{equation}
+
+as $C_1>=0$ and $C_3<0$, $f'_{\theta}(\rho_*)$ will always be positive, thus the first condition always hold by construction and does not need to be verified.\\
+
+In Haralick's work $\rho_0$ is a fixed threshold which acts as a parameter of the algorithm.
+Its value must be greater than zero and and less than the size of the grid (i.e. $\rho_0<1$). In this work we obtained the best results with values between $0.4$ and $0.6$.\\
 
@@
 (first odd integer greater than $6\sigma=24$).
-
+\myn{\label{math:common:gauss}} 
 \begin{figure}[ht]
@@
 Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.
-
+\myn{\label{math:common:log}} 
 \begin{figure}[ht]
@@
 % del kernel en funcion de sigma.
-
+\myn{\label{math:common:comparison}} 
 \begin{figure}[ht]
@@
 
-Now the performance of each algorithm is analyzed using a natural image shown in Figure \ref{fig:original2}. 
-First the results of the first derivative edge detection algorithms are shown Figures \ref{fig:result2-a} to \ref{fig:result2-c}.
-Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm.
-Table \ref{exectime2} shows the execution time for each of the algorithms.
-
-\begin{figure}[h!]
-	\centering
-	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
-	\quad
-	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}
-
-	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
-	\quad
-	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$. $\text{Exec-time}:1.05s$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
-
-	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$. $\text{Exec-time}:1.44s$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
-	\quad
-	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.93s$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
-	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using Figure \ref{fig:original2} as input image.}
-	\label{fig:result2}
-\end{figure}
+Now the performance of each algorithm is analyzed using a natural image (\textit{Windmill}) shown in Figure \ref{fig:original2}. 
+The results of the first derivative edge detection algorithms are shown Figures \ref{fig:result3-b} to \ref{fig:result3-d}.
+Figures \ref{fig:result3-e} and \ref{fig:result3-f} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result3-g} shows the results obtained using the Haralick algorithm.
+Table \ref{exectime2} summarizes the execution times for each of the algorithms.
 
@@
 
-In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative.
+In this example the difference in performance between the first derivative and the second derivative algorithms is more meaningful.
 For each image in Figure \ref{fig:result3}, we show an area of ​​interest, enlarging it to have a better view of the details.
@@
 
-\begin{figure}[t!]
+\begin{figure}[h!]
 	\centering
@@
 
-	\subfigure[Roberts.]{\label{fig:result3-b}\includegraphics[width=0.48\textwidth]{results/molino_roberts_zoom.png}}
-	\quad
-	\subfigure[Prewitt.]{\label{fig:result3-c}\includegraphics[width=0.48\textwidth]{results/molino_prewitt_zoom.png}}
-
-	\subfigure[Sobel.]{\label{fig:result3-d}\includegraphics[width=0.48\textwidth]{results/molino_sobel_zoom.png}}
-	\quad
-	\subfigure[Marr-Hildreth (Gaussian).]{\label{fig:result3-e}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth_zoom.png}}
-
-	\subfigure[Marr-Hildreth (LoG).]{\label{fig:result3-f}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth-log_zoom.png}}
-	\quad
-	\subfigure[Haralick.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
-	\caption{Results of the algorithms using Figure \ref{fig:original2} as input image, enlarging an interesting area.}
+	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result3-b}\includegraphics[width=0.48\textwidth]{results/molino_roberts_zoom.png}}
+	\quad
+	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result3-c}\includegraphics[width=0.48\textwidth]{results/molino_prewitt_zoom.png}}
+
+	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.55s$.]{\label{fig:result3-d}\includegraphics[width=0.48\textwidth]{results/molino_sobel_zoom.png}}
+	\quad
+	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$. $\text{Exec-time}:1.05s$.]{\label{fig:result3-e}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth_zoom.png}}
+
+	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$. $\text{Exec-time}:1.44s$.]{\label{fig:result3-f}\includegraphics[width=0.48\textwidth]{results/molino_marr-hildreth-log_zoom.png}}
+	\quad
+	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.93s$.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
+
+	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms on the \textit{Windmill} image.}
 	\label{fig:result3}
@@
 
+\newcommand{\factuno}{0.31}
 \begin{figure}[t!]
@@
 
-	\subfigure[Roberts. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
-	\quad
-	\subfigure[Prewitt. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
-	\quad
-	\subfigure[Sobel. $th=0.1$]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}
-
-	\subfigure[Marr-Hildreth (Gaussian). $\sigma=2$, $n=21$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
-	\quad
-	\subfigure[Marr-Hildreth (LoG). $\sigma=2$, $n=25$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
-	\quad
-	\subfigure[Haralick. $\rho=0.4$.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
+	\subfigure[Roberts. $th=0.1$]{\includegraphics[width=\factuno\textwidth]{examples/oranges_roberts.png}}
+	\quad
+	\subfigure[Prewitt. $th=0.1$]{\includegraphics[width=\factuno\textwidth]{examples/oranges_prewitt.png}}
+	\quad
+	\subfigure[Sobel. $th=0.1$]{\includegraphics[width=\factuno\textwidth]{examples/oranges_sobel.png}}
+
+	\subfigure[Marr-Hildreth (Gaussian). $\sigma=2$, $n=21$, $th_{ZC}=0.15$.]{\includegraphics[width=\factuno\textwidth]{results/orange_mhg_s_2_n_21_th_0.15.png}}
+	\quad
+	\subfigure[Marr-Hildreth (LoG). $\sigma=2$, $n=25$, $th_{ZC}=0.15$.]{\includegraphics[width=0.3\textwidth]{results/orange_mhl_s_2_n_25_th_0.15.png}}
+	\quad
+	\subfigure[Haralick. $\rho=0.4$.]{\includegraphics[width=\factuno\textwidth]{examples/oranges_haralick.png}}
 	\caption{Example: Oranges (536$\times$480 pixels).}
@@
 
+\subsection{Effect of the scale parameter}
+\myn{\label{results:parameter:scale}}In this section we will discuss the effect of the parameters for each of the reviewed algorithms.
+As a general rule, they all take one threshold parameter which acts as a scale parameter.
+As Figure \ref{fig:results:parameters:roberts:parameter} shows for the \emph{Roberts} algorithm, the bigger that threshold is, the coarser is the edge map obtained.
+
+In Figure \ref{fig:results:parameters:roberts:saliency} we show a summary of the threshold's behavior for this algorithm, in the whole range of possible values $[0,1]$.
+For each image, this information is summarized in a saliency map which serves to asses the quality of the hierarchical set of edges obtained.
+The saliency map is constructed by sampling the scale range (using steps of $0.05$) and computing the edge map corresponding for each scale.
+Then each edge is labeled with its scale of disappearance $\alpha^-$, which is the maximum scale at which an edge is present in the output edge map.
+This scale is drawn with a \emph{hot} colormap (black, red, yellow, white).
+Intuitively, this means that 'hotter' edges last longer across scales and thus are more meaningful.
+These results show two important properties.
+First, the obtained stack of edges are causal, i.e.
+as the scale parameter increases the edge map gets coarser.
+This can be verified by observing the values of $\alpha^-$ assigned to each contour.
+If a contour has a high $\alpha^-$ value it means that it will survive for a long time in the merging process. On the other hand, small values mean that it will be removed at early stages.
+As the figure shows, the highest values of $\alpha^-$ are assigned to the most coarser segmentation.
+And second, the stack of edges is geometrically accurate, i.e.
+the borders present in each scale are a strict subset of the previous (finer) scale.
+If they were not included they will be seen as new contours in the saliency map.
+These results empirically show that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.
+
+In Figures \ref{fig:results:parameters:first:saliency:a} and \ref{fig:results:parameters:first:saliency:b} we compare the three first derivative algorithms by means of their saliency maps.
+As the results show, they all present a similiar behavior, however the are small differences in the results of Roberts vs the other two algorithms.
+For example, edges surrounding the segments of the orange are better defined, because they have more relative weight than other spurious borders.
+This is explained by the smoothing introduced by Prewitt and Sobel masks explained in Sections \ref{sec:first:sobel} and \ref{sec:first:sobel}, which removes theses spurious edges and thus improves the meaningfulness of the stable ones.
+
+In Figures \ref{fig:results:parameters:second:saliency:a} and \ref{fig:results:parameters:second:saliency:b} we compare the second derivative algorithms by means of their saliency maps.
+In the first place, we obtain very similar hierarchies for both Marr-Hildreth implementations, which validates the affirmation that the difference is merely about computational effort if both are correctly implemented.
+However, the LoG implementation seems to give slightly better results, this is probably because it provides a more accurate discretization of the differential operators than the Gaussian approach.
+See for instance the clothing sticks in Figure \ref{fig:results:parameters:second:saliency:b}, where the boundaries in the LoG case seem better defined.
+
+\renewcommand{\anchocinco}{2.5cm}
+\newcommand{\anchocuatro}{5cm}
+
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleHHH{test}{roberts}{0002}
+\figRowMultiscaleHHH{test}{roberts}{0003}
+\figRowMultiscaleHHH{test}{roberts}{0004}
+  \caption{Results of the \textit{Roberts} algorithm our sample images.
+For each example we show the original image on the left and the results corresponding to parameter values of 0.1, 0.2 and 0.3 respectively.}
+  \label{fig:results:parameters:roberts:parameter}
+%\vskip \espaciotabla
+\end{figure*}
+
+\renewcommand{\anchocuatro}{8cm}
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleH{test}{roberts}{0002}
+\figRowMultiscaleH{test}{roberts}{0003}
+\figRowMultiscaleH{test}{roberts}{0004}
+  \caption{Results of the \textit{Roberts} algorithm on our sample images.
+For each example we show the original image on the left and a saliency map on the right.
+The value of each contour corresponds to the scale of disappearance  $\alpha^-$ of each region.
+}
+  \label{fig:results:parameters:roberts:saliency}
+%\vskip \espaciotabla
+\end{figure*}
+
+\renewcommand{\anchocuatro}{5cm}
+
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleHH{test}{0002}
+\figRowMultiscaleHH{test}{0003}
+\figRowMultiscaleHH{test}{0004}
+\figRowMultiscaleHH{test}{0005}
+  \caption{Results of the first derivative algorithms our sample images.
+For each example we show the original image on the left and the saliency map of each algorithm (Roberts, Sobel, Prewitt) respectively.
+  The value of each contour corresponds to the scale of disappearance  $\alpha^-$ of each region.}
+  \label{fig:results:parameters:first:saliency:a}
+%\vskip \espaciotabla
+\end{figure*}
+
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleHH{test}{0006}
+\figRowMultiscaleHH{test}{0007}
+\figRowMultiscaleHH{test}{0008}
+  \caption{Results of the first derivative algorithms our sample images.
+For each example we show the original image on the left and a saliency map of each algorithm (Roberts, Sobel, Prewitt) respectively.
+  The value of each contour corresponds to the scale of disappearance  $\alpha^-$ of each region.}
+  \label{fig:results:parameters:first:saliency:b}
+%\vskip \espaciotabla
+\end{figure*}
+
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleHHHH{test}{0002}
+\figRowMultiscaleHHHH{test}{0003}
+\figRowMultiscaleHHHH{test}{0004}
+\figRowMultiscaleHHHH{test}{0005}
+  \caption{Results of the second derivative algorithms our sample images.
+For each example we show the original image on the left and a saliency map of each algorithm (Marr-Hildreth-Gaussian, Marr-Hildreth-Log, Haralick) respectively.
+  The value of each contour corresponds to the scale of disappearance  $\alpha^-$ of each region.}
+  \label{fig:results:parameters:second:saliency:a}
+%\vskip \espaciotabla
+\end{figure*}
+
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleHHHH{test}{0006}
+\figRowMultiscaleHHHH{test}{0007}
+\figRowMultiscaleHHHH{test}{0008}
+  \caption{Results of the second derivative algorithms our sample images.
+For each example we show the original image on the left and the saliency map of each algorithm (Marr-Hildreth-Gaussian, Marr-Hildreth-Log, Haralick) respectively.
+  The value of each contour corresponds to the scale of disappearance  $\alpha^-$ of each region.}
+  \label{fig:results:parameters:second:saliency:b}
+%\vskip \espaciotabla
+\end{figure*}
+
+\subsection{Effect of the smoothing parameter}
+\myn{\label{results:parameter:smooth}}As we stated in previous section, all the reviewed algorithms share a scale parameter (threshold), however, the Marr-Hildreth Algorithm also has a smoothing step which adds an additional parameter $\sigma$, which controls the amount of regularization.
+In Figure \ref{fig:results:parameters:marr:sigma} we show the results using different  $\sigma$ values.
+As sigma increases more spurious edges are removed and the boundaries of the objects look smoother.
+However, if we keep increasing it we start removing meaningful edges (see the cloth band on the hat of Lena).
+
+\subsection{Discussion}
+\myn{\label{results:discussion}}In this section we will discuss the obtained results and contrast them with the claims made by the original authors.
+In the case of the first derivative algorithms, they are very simple and the claims made by the original authors are straight forward to verify.
+The Roberts algorithm was presented in his Phd Thesis (\cite{im_proc:segmentation:roberts:1963:boundary_extraction}, Chapter IV ) as a step of a more complex 3D reconstruction system.
+For this reason, only qualitative results are shown on few images.
+The effects of the parameter or the discretization is not validated.
+The only observation made in that work is the presence of spurious edges due to noise (pg.
+31), which are removed by a line fitting postprocessing.
+With respect to Prewitt and Sobel Operators,
+
+In Haralick's work, there is an interesting quantitative validation over a synthetic grid.
+On that evaluation, sensitivity to noise as well as comparative performance with Prewitt operators.
+As in this work we didn't benchmark quantitaive
+
+The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:80:digital_step}, were he made the following observations:
+\begin{itemize}
+\item "A major difficulty with natural images is that changes can and do occur over a
+wide range of scales (Marr 1976a, b).
+No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales.
+".
+
+\item "The reason is that strict bandlimiting gives rise to sidelobes in the spatial spatial filter, and the consequence of these is that, in the zero-crossing image, strong intensity changes give rise to echoes as well as to the directly corresponding zero-crossings..."
+
+\item "Physical edges will produce roughly coincident zero-crossings in channels of nearby sizes.
+The spatial coincidence assumption asserts that the converse of this is true, that is the coincidence of zero crossings is sufficient evidence for the existence of a real physical edge.
+If the zero-crossings in one channel are not consistent with those in the others, they are probably caused by different physical phenomena, so descriptions need to be formed from both sources and kept somewhat separate."
+\end{itemize}
+
+The first claim is validated in our experiments varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges.
+The same reasoning applies to the zero crossing threshold $t_{ZC}$ .
+
+The second observation is evidenced by the halos seen in the blades of the windmill, see inside left blade and left of the top blade in Figure \ref{fig:result3-e} .
+\footnote{Please look at these issues in the electronic version or the online demo, as they are not noticeable in the printed version.}
+
+The third claim is evidenced by the structure of the saliency maps of the threshold parameter (Figures \ref{fig:results:parameters:second:saliency:a} and \ref{fig:results:parameters:second:saliency:b}) and the smoothing parameter (Figure \ref{fig:results:parameters:marr:sigma:saliency}).
+In both maps we observe that meaningful edges are coincident across scales and the spurious ones are not.
+
+\renewcommand{\anchocuatro}{3.6cm}
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleSigma{test}{mh_th_0.15}{0002}
+\figRowMultiscaleSigma{test}{mh_th_0.15}{0003}
+\figRowMultiscaleSigma{test}{mh_th_0.15}{0004}
+  \caption{Effect of the $\sigma$ parameter for the \textit{Marr-Hildreth} (Gaussian) algorithm on our sample images.
+For each example we show the original image on the left and various consecutive results with different smoothing values ($\sigma=1,2,3,4$) on the right.}
+  \label{fig:results:parameters:marr:sigma}
+%\vskip \espaciotabla
+\end{figure*}
+
+\renewcommand{\anchocuatro}{8cm}
+\begin{figure*}[ht]
+  \centering
+\figRowMultiscaleH{test}{mh_th_0.15}{0002}
+\figRowMultiscaleH{test}{mh_th_0.15}{0003}
+\figRowMultiscaleH{test}{mh_th_0.15}{0004}
+  \caption{Effect of the $\sigma$ parameter for the \textit{Marr-Hildreth} (Gaussian) algorithm on our sample images.
+For each example we show the original image on the left and the generated saliency map  on the right.
+$\sigma$ is sampled regularly in the range $[0.5,5]$ using a step of $0.5$.}
+  \label{fig:results:parameters:marr:sigma:saliency}
+%\vskip \espaciotabla
+\end{figure*}
 %-------------------------------------------------------------------------------
@@
 It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polynomial function.
-This supposes a certain regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
+This supposes a certain regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist due to ringing effects, and get some thicker edges too.
 
@@
 Choosing one or the other may depend on requirements of the application.
-Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.
+Furthermore, we must remark that we focused on low level edge detectors, and thus none of the implemented methods ensure edge connectivity.
+This algorithms only serve as the input for a further edge integration step.
+There are many proposals in the literature aiming to integrate edgels into meaningful object boundaries.
+Examples of these approaches are the Canny \cite{Canny1986} edge detector or the meanignful level lines algorithm \cite{segm:cao:03:extracting_meaningful}.
+
+With respect to the parameters, these methods require the manual determination of the parameters, and it is very hard to find a single threshold value suitable for a broad category of examples.
+In this sense, \emph{A contrario} methods provide a better approach for setting parameters.
+For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.
+This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).
+
+\subsection{Future work}
+
+The goal of this work was to provide reference implementation for some of the most classical edge detectors.
+Some of them have been published for many years and are quite simple, however, it is surprisingly hard to find trustable implementations.
+We focused on low-level edge detection, which means that we do not consider post processing steps to integrate the detected edgels into meaningul lines.
+Extensive qualitative validation has been performed, showing the effects of the parameters and the scale-space properties of the algorithms in order to faithfully compare them.
+We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
+A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
+In particular, quantitative challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation.
+From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic images.
+Even at the moment of writing (2014), there are only a dozen edge detection and segmentation works presenting a serious quantitative analysis over a standard database of images.
 
@@
 \bibliographystyle{siam}
-\bibliography{bibliography}
+\bibliography{journal_list_long,bibliography,references}
 

### 2014-11-11 old_versions/edge-detection-review-2nd-revision-v1.tex
second round of third revision, improving references

--- 2014-05-07 edge-detection-review-2nd-revision-v1.tex
@@
 
-The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with a small $h$. Thus $f'(x)$ is computed as
+\myn{\label{correction_a:first_derivative:taylor}}The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$.
+Thus $f'(x)$ is computed as
 \begin{equation}
@@
 attempts to use 2-D masks for this purpose.
-These operators are based on computing the diagonal \tcom{differences} \myn implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
+These operators are based on computing the diagonal \tcom{differences} \myn{\label{correction_a:typo}} implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
 By convention
@@
 
-\myn{\label{correction_a:first_derivative:sobel} The left mask can be constructed (beyond normalization) as the convolution of a vertical derivative filter
-$[-1\ 0\ 1]$ with an horizontally moving average filter $[1\ 1\ 1]$.
+\myn{\label{correction_a:first_derivative:sobel}} The left mask can be constructed (up to the normalization factor) as the convolution of two filters: a vertical derivative 
+$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]$.
 The same reasoning could be applied to the right mask.
@@
 
-Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
+Sobel masks can be seen (up to the normalization factor) as the convolution of a horizontal derivation mask 
 $[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
@@
 
-The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
-use Sobel operators, because they are better localized, and their filter are also less aliased, 
-because of the weighted shape of $[1\ 2\ 1]$.
+The computational cost of applying Prewitt and Sobel masks is exactly the same. Thus, it is preferable to 
+use Sobel operators, because they are better localized, and their filter are also less aliased, because of the weighted shape of $[1\ 2\ 1]$.
 
@@
 
-The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.
+The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}
 
@@
 \begin{algorithmic}[1]
-\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
+\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $th_{ZC}$.
 \STATE $im_{ORIG} \leftarrow$ input image
@@
 \ENDIF
-\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}
-\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
+
 \IF{Gaussian kernel}
@@
 \end{equation}
-
-To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
-neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
-points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
-In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
-computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\
+To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted.
+As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$.
+Having $10$ coefficients and $25$ data
+points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\
+
+\myn{\label{second_derivative:haralick:statement}} In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel.
+In this way, the center of the neighboorhood is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g.
+in a neighborhood of size 5x5, x and y take values between -2 and +2).\\
+Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficientes are computed in the same way for every pixel in the image.
+Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which is a great speed-up of the algorithm.
+In the following we will state the approximation problem and show how to derive those masks from its solution.\\
 
 Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
-to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
+to find 10 coefficients. As mentioned before this leads to an \tcom{overdetermined} system which can be solved by least squares. 
 By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
@@
 \begin{equation*}
-	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
+	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} 
 \end{equation*}
 
-$\mathbf{B}$ is a $10\times25$ matrix,
+\myn{\label{second_derivative:haralick:normal}} Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
+
+$$ \mathbf{k} = \mathbf{B}\mathbf{f} $$
 
@@
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
-\end{equation}
-
-where
-
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f} . \mathbf{b_i},\\
+\end{equation}
+
+where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.\myn{\label{correction_a:second_derivative:haralick_bi}}
+
+\begin{equation}
+	\label{eq:b_i_v}
+	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,25}
+					\end{bmatrix}.\\
+\end{equation}
+
+Let us rearrange $b_i$ as a 5x5 matrix and denote it by $\mathbf{mask_i}$ as follows
 \begin{equation}
 	\label{eq:b_i}
-	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
+	\mathbf{mask_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
 									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
@@
 \vspace{1ex}
-Using equation \ref{eq:coefficients} and the masks given in Table \ref{table:b_i}, it is possible to compute 
-the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
-are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.\\
-
-\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
-\begin{table}
-\centering
-\subfigure[$\mathbf{b_1}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-425 & 275 & 225 & 275 & 425 \\
-\hline
-275 & 125 &  75 & 125 & 275 \\ 
-\hline
-225 &  75 &  25 &  75 & 225 \\
-\hline
-275 & 125 &  75 & 125 & 275 \\
-\hline
-425 & 275 & 225 & 275 & 425 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_2}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--2260 & -620 & 0 & 620 & 2260 \\ 
-\hline
--1660 & -320 & 0 & 320 & 1660 \\ 
-\hline
--1460 & -220 & 0 & 220 & 1460 \\
-\hline
--1660 & -320 & 0 & 320 & 1660 \\ 
-\hline
--2260 & -620 & 0 & 620 & 2260 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_3}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-2260  &  1660 &  1460 &  1660 &  2260 \\ 
-\hline
-620   &   320 &   220 &   320 &   620 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline
--620  &  -320 &  -220 &  -320 &  -620 \\ 
-\hline
--2260 & -1660 & -1460 & -1660 & -2260 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_4}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-1130 & 620 & 450 & 620 & 1130 \\ 
-\hline
-830  & 320 & 150 & 320 &  830 \\ 
-\hline
-730  & 220 &  50 & 220 &  730 \\
-\hline
-830  & 320 & 150 & 320 &  830 \\ 
-\hline
-1130 & 620 & 450 & 620 & 1130 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_5}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--400 & -200 & 0 &  200 &  400 \\ 
-\hline
--200 & -100 & 0 &  100 &  200 \\ 
-\hline
-0    &    0 & 0 &    0 &    0 \\
-\hline
-200  &  100 & 0 & -100 & -200 \\ 
-\hline
-400  &  200 & 0 & -200 & -400 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_6}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-1130 & 830 & 730 & 830 & 1130 \\ 
-\hline
-620  & 320 & 220 & 320 &  620 \\ 
-\hline
-450  & 150 &  50 & 150 &  450 \\
-\hline
-620  & 320 & 220 & 320 &  620 \\ 
-\hline
-1130 & 830 & 730 & 830 & 1130 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_7}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--8260 & -2180 & 0 & 2180 & 8260 \\ 
-\hline
--6220 & -1160 & 0 & 1160 & 6220 \\ 
-\hline
--5540 &  -820 & 0 &  820 & 5540 \\
-\hline
--6220 & -1160 & 0 & 1160 & 6220 \\ 
-\hline
--8260 & -2180 & 0 & 2180 & 8260 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_8}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-5640  &  3600 &  2920 &  3600 &  5640 \\ 
-\hline
-1800  &   780 &   440 &   780 &  1800 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline
--1800 &  -780 &  -440 &  -780 & -1800 \\ 
-\hline
--5640 & -3600 & -2920 & -3600 & -5640 \\
-\hline
-\end{tabular}} \\
-\subfigure[$\mathbf{b_9}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
--5640 & -1800 & 0 & 1800 & 5640 \\ 
-\hline
--3600 &  -780 & 0 &  780 & 3600 \\ 
-\hline
--2920 &  -440 & 0 &  440 & 2920 \\
-\hline
--3600 &  -780 & 0 &  780 & 3600 \\ 
-\hline
--5640 & -1800 & 0 & 1800 & 5640 \\
-\hline
-\end{tabular}}
-\qquad\qquad
-\subfigure[$\mathbf{b_{10}}$.] {
-\begin{tabular}{|*{5}{C|}}
-\hline
-8260  &  6220 &  5540 &  6220 &  8260 \\ 
-\hline
-2180  &  1160 &   820 &  1160 &  2180 \\ 
-\hline
-0     &     0 &     0 &     0 &     0 \\
-\hline
--2180 & -1160 &  -820 & -1160 & -2180 \\ 
-\hline
--8260 & -6220 & -5540 & -6220 & -8260 \\
-\hline
-\end{tabular}} \\
-\caption{Masks to compute the coefficients of the bicubic fit.}
-\label{table:b_i}
-\end{table}
+\myn{\label{second_derivative:haralick:mask_computation}} We will compute one value of $k_i$ for each pixel on the image.
+Thus we can store these coefficientes in a matrix $\bf{K_i}$ of the same size as the original image.
+Then, using the mask in shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$.
+\\
+\begin{equation}
+	\label{eq:convolution_b}
+	\mathbf{K_i} = f \star \mathbf{mask_i} \mathtt{for} i \in {1 \ldots 10}\\
+\end{equation}
+
+As we expressed the approximation in a local coordinate system, the coefficients $b_{i,j}$ will be the same for every pixel, thus, this convolution is performed against a fixed and precomputed mask $\mathbf{mask_i}$.
+The numeric values of these 5x5 masks used in our experiments are shown in Table \ref{table:b_i}.
 
@@
 
-The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
-expression in equation \ref{eq:bicubic}. If just the first order terms of this 
-polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as
+\myn{\label{second_derivative:haralick:analitical:statement}} After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea of Haralick's algorithm is to compute the second derivative of the image in the direction of the gradient using that approximation.\\
+
+Let $\theta$ be the gradient angle, defined with respect to the negative x-axis, and it is \tcom{approximated} using the first order terms of the 
+polynomial \ref{eq:bicubic} obtaining the following expression:
 
@@
 
-Now substituting the variables $x$ and $y$ in polar form as
+\myn{\label{second_derivative:haralick:analitical:derivative}} Let us consider the (half) line $r$ passing by $(x,y)$ with angle $\theta$.
+Each point $(u,v)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(x,y)$ as follows
 
@@
 
-\begin{equation}
-	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
-\end{equation}
-
-where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
-of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
-value is between $0.4$ and $0.6$.\\
+\begin{equation}\label{eq:haralick:cond_b_orig}
+	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0
+\end{equation}
+
+ Ideally, to hold this condition, the polynomial function 
+should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial becomes zero in a neighborhood 
+of $\rho=0$ ($\rho\leq\rho_0$). \\
+With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes
+
+\begin{equation}
+\left| \frac{C_2}{3C_3} \right| < \rho_0,
+\end{equation}
+
+where $\rho_0$ is a fixed threshold, which passed as an argument to the algorithm.
+\myn{\label{second_derivative:haralick:analitical:parameter_value}}  This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal  value is between $0.4$ and $0.6$.\\
 
@@
 \ENDIF
-\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
+\STATE Define the ten $5x5$ masks ($mask[1] \dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.} \mynf{\label{correction_a:second_derivative:haralick_algo}}
 \FORALL{pixel $i$ in image $im$}
@@
 %-------------------------------------------------------------------------------
-
-\section{Implementation}
-\label{sec:appendix2}
-
-In this section a detailed explanation of the source code of the different implemented algorithms
-is presented. \\
-
-To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
-There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
-All these links were last checked in July 2012.
-\\
-
-This code documentation was generated using a perl script to extract some parts of the 
-code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
-
-Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
-
-Required external libraries: 
-\begin{itemize}
-	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
-	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
-\end{itemize}
-
-\subsection{Roberts, Prewitt and Sobel}
-\input{test_fded.tex}
-
-\subsection{Marr-Hildreth}
-\label{app:marr-hildreth}
-\input{test_mh.tex}
-
-\subsection{Haralick}
-\input{test_haralick.tex}
-
-\subsection{Gaussian kernel generation}
-\input{gaussian_kernel.tex}
-
-\subsection{2D convolution}
-\input{2dconvolution.tex}
 
@@
 
-Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one.
+Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs \tcom{significantly} faster that the LoG one.
 The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
@@
 Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
-It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function.
+It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic \tcom{polynomial} function.
 This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
@@
 \clearpage
-
-%-------------------------------------------------------------------------------
+\appendix
+\section{Haralick convolution masks}
+\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
+\begin{table}[h]
+\small
+\centering
+\subfigure[$\mathbf{mask_1}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+425 & 275 & 225 & 275 & 425 \\
+\hline
+275 & 125 &  75 & 125 & 275 \\ 
+\hline
+225 &  75 &  25 &  75 & 225 \\
+\hline
+275 & 125 &  75 & 125 & 275 \\
+\hline
+425 & 275 & 225 & 275 & 425 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_2}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-2260 & -620 & 0 & 620 & 2260 \\ 
+\hline
+-1660 & -320 & 0 & 320 & 1660 \\ 
+\hline
+-1460 & -220 & 0 & 220 & 1460 \\
+\hline
+-1660 & -320 & 0 & 320 & 1660 \\ 
+\hline
+-2260 & -620 & 0 & 620 & 2260 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_3}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+2260  &  1660 &  1460 &  1660 &  2260 \\ 
+\hline
+620   &   320 &   220 &   320 &   620 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-620  &  -320 &  -220 &  -320 &  -620 \\ 
+\hline
+-2260 & -1660 & -1460 & -1660 & -2260 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_4}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+1130 & 620 & 450 & 620 & 1130 \\ 
+\hline
+830  & 320 & 150 & 320 &  830 \\ 
+\hline
+730  & 220 &  50 & 220 &  730 \\
+\hline
+830  & 320 & 150 & 320 &  830 \\ 
+\hline
+1130 & 620 & 450 & 620 & 1130 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_5}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-400 & -200 & 0 &  200 &  400 \\ 
+\hline
+-200 & -100 & 0 &  100 &  200 \\ 
+\hline
+0    &    0 & 0 &    0 &    0 \\
+\hline
+200  &  100 & 0 & -100 & -200 \\ 
+\hline
+400  &  200 & 0 & -200 & -400 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_6}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+1130 & 830 & 730 & 830 & 1130 \\ 
+\hline
+620  & 320 & 220 & 320 &  620 \\ 
+\hline
+450  & 150 &  50 & 150 &  450 \\
+\hline
+620  & 320 & 220 & 320 &  620 \\ 
+\hline
+1130 & 830 & 730 & 830 & 1130 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_7}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-8260 & -2180 & 0 & 2180 & 8260 \\ 
+\hline
+-6220 & -1160 & 0 & 1160 & 6220 \\ 
+\hline
+-5540 &  -820 & 0 &  820 & 5540 \\
+\hline
+-6220 & -1160 & 0 & 1160 & 6220 \\ 
+\hline
+-8260 & -2180 & 0 & 2180 & 8260 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_8}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+5640  &  3600 &  2920 &  3600 &  5640 \\ 
+\hline
+1800  &   780 &   440 &   780 &  1800 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-1800 &  -780 &  -440 &  -780 & -1800 \\ 
+\hline
+-5640 & -3600 & -2920 & -3600 & -5640 \\
+\hline
+\end{tabular}} \\
+\subfigure[$\mathbf{mask_9}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+-5640 & -1800 & 0 & 1800 & 5640 \\ 
+\hline
+-3600 &  -780 & 0 &  780 & 3600 \\ 
+\hline
+-2920 &  -440 & 0 &  440 & 2920 \\
+\hline
+-3600 &  -780 & 0 &  780 & 3600 \\ 
+\hline
+-5640 & -1800 & 0 & 1800 & 5640 \\
+\hline
+\end{tabular}}
+\qquad\qquad
+\subfigure[$\mathbf{mask_{10}}$.] {
+\begin{tabular}{|*{5}{C|}}
+\hline
+8260  &  6220 &  5540 &  6220 &  8260 \\ 
+\hline
+2180  &  1160 &   820 &  1160 &  2180 \\ 
+\hline
+0     &     0 &     0 &     0 &     0 \\
+\hline
+-2180 & -1160 &  -820 & -1160 & -2180 \\ 
+\hline
+-8260 & -6220 & -5540 & -6220 & -8260 \\
+\hline
+\end{tabular}} \\
+\caption{5x5 masks used to compute the coefficients of the bicubic fit.}
+\label{table:b_i}
+\end{table}
+
+%-------------------------------------------------------------------------------
+
+\clearpage
+\section{Implementation}
+\label{sec:appendix2}
+
+In this section a detailed explanation of the source code of the different implemented algorithms
+is presented. \\
+
+To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}.
+There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}.
+All these links were last checked in July 2012.
+\\
+
+This code documentation was generated using a perl script to extract some parts of the 
+code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\
+
+Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\
+
+Required external libraries: 
+\begin{itemize}
+	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
+	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
+\end{itemize}
+
+\subsection{Roberts, Prewitt and Sobel}
+\input{test_fded.tex}
+
+\subsection{Marr-Hildreth}
+\label{app:marr-hildreth}
+\input{test_mh.tex}
+
+\subsection{Haralick}
+\input{test_haralick.tex}
+
+\subsection{Gaussian kernel generation}
+\input{gaussian_kernel.tex}
+
+\subsection{2D convolution}
+\input{2dconvolution.tex}
+
+%-------------------------------------------------------------------------------
+\clearpage
 \bibliographystyle{siam}

### 2014-11-11 old_versions/edge-detection-review-3rd-revision-v1.tex
second round of third revision, improving references

--- 2014-11-11 edge-detection-review-3rd-revision-v2.tex
@@
 
-Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{im_proc:segmentation:roberts:1963:boundary_extraction} proposes
+Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{roberts1963machine} proposes
 to compute them as shown in Figure \ref{fig:roberts}.
@@
 \bibliographystyle{siam}
-\bibliography{journal_list_long,bibliography,references}
+\bibliography{bibliography}
 

### 2014-11-11 old_versions/edge-detection-review-v12.tex
second round of third revision, improving references

--- 2014-05-12 edge-detection-review-v11.tex
@@
 %-------------------------------------------------------------------------------
-% A Review of Classic Edge Detectors - v10
+% A Review of Classic Edge Detectors - v12
 % by Haldo Spontón & Juan Cardelino
@@
 The basic idea of edge detection algorithms is to detect abrupt changes in image intensity. 
-Detecting those changes in intensity for the purpose of finding edges in images 
-can be accomplished using first or second order derivatives.
+Detecting those changes can be accomplished using first or second order derivatives.
 
@@
 in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
-usual to scale the mask values, in order to have unit $L_2$ norm.
+usual to scale the mask values, in order to have unit derivative when unitary steps in intensity occurs.
 
@@
 difference. It can be shown that using double weight in the center location provides better image smoothing. This
-variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
+variation is implemented using the masks in Figure \ref{fig:sobel}. These operators are called 
 \textit{Sobel operators}.
@@
 
-The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
-use Sobel operators, because they are better localized, and their filter are also less aliased, 
+The computational difference between Prewitt and Sobel masks is not important, thus It is preferable to 
+use Sobel operators, because edges are better localized, and their filter are also less aliased, 
 because of the weighted shape of $[1\ 2\ 1]$.
@@
 \begin{equation*}
-	\mbox{edges}[i,j] = \begin{cases} 0,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
-									1,& \mbox{if}\ M[i,j]<\mbox{th}
+	\mbox{edges}[i,j] = \begin{cases} 1,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
+									0,& \mbox{if}\ M[i,j]<\mbox{th}
 					  \end{cases}
@@
 
-\clearpage
-
 \begin{figure}[t!]
@@
 \end{equation}
-where
+where\footnote{In this matrix the coefficients $b_i$ are simply spatially arranged with the same topology as the pixels they multiply.}
 \begin{equation}
@@
 where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial to become zero in a neighborhood 
-of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
+should be zero at $\rho=0$, but this condition is relaxed\footnote{This polynomial being zero at $\rho = 0$ means that the edge passes right through the exact coordinates of the pixel.
+Since the edge may be slightly displaced, other $\rho$ values are permitted as long as the edge is inside the pixel, and for this, the maximum distance from the center of the pixel may be $1/\sqrt{2}$.}, allowing the polynomial to become zero in a neighborhood
+of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\frac{1}{\sqrt{2}}$. Its optimal 
 value is between $0.4$ and $0.6$.
@@
 have to be performed, one with the kernel and another with the Laplacian operator), but the kernel 
-support must be greater in order to obtain the same precision..
+support must be greater in order to obtain the same precision.
 
@@
 Direct convolution is not the only way of filtering an image with a kernel.
-There are other methods such as FFT convolution or recursive filtering\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
+There are other methods such as FFT convolution or recursive filtering.%\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
 
@@
 
+As supplementary material, there are videos testing the frame by frame application of these algorithms.
+
+\begin{comment}
 This is an example of applying the edge detection algorithms, frame by frame, to a video: 
@@
 \end{multicols}
+\end{comment}
+%-------------------------------------------------------------------------------
 
 \clearpage
-%-------------------------------------------------------------------------------
 
@@
 
-The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
+The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed.
 In real images, they have better behavior than first derivative algorithms.
@@
 
-As an addendum to this work a detailed implementation of the described algorithms is presented, 
-available in C language, along with an online demo where users can run the algorithms 
-with arbitrary images uploaded by themselves.
+%As an addendum to this work a detailed implementation of the described algorithms is presented, 
+%available in C language, along with an online demo where users can run the algorithms 
+%with arbitrary images uploaded by themselves.
 
@@
 
+\bibliographystyle{abbvr}
 \bibliography{bibliography}

### 2014-11-11 old_versions/edge-detection-review-v2.tex
second round of third revision, improving references

--- 2012-07-23 edge-detection-review-v3.tex
@@
 \usepackage{sidecap,subfig,array}
-\usepackage[table]{xcolor}
-\usepackage{multicol}
 \newtheorem{theorem}{Theorem}
@@
 \numberwithin{table}{section}
-\numberwithin{figure}{section}
 \begin{document}
@@
 \begin{ipolAbstract}
-In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
+In this paper we study different alternatives for edge detection in digital images. The main idea 
 behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
@@
 
-Edge detection is one of the first approaches published in the literature for 
-segmenting images. The basic idea is to detect abrupt changes in intensity. 
-Detecting those changes in intensity for the purpose of finding edges in images 
-can be accomplished using first or second order derivatives.
+Edge detection is the first approach for segmenting images
+based on abrupt changes in intensity. Detecting those changes in intensity 
+for the purpose of finding edges in images can be accomplished using first 
+or second order derivatives.
 
@@
 first derivative of the image. In section ~\ref{sec:first} we study the 
-performance of such algorithms, as an introduction to a more sophisticated 
+performance of these algorithms, as an introduction to a more sophisticated 
 analysis into the edge finding process.\\
@@
 	\centering
-	\includegraphics[width=0.14\textwidth]{1dmasks.pdf}
+	\includegraphics[width=0.14\textwidth]{1dmasks.eps}
 	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
@@
 	\centering
-	\includegraphics[width=0.15\textwidth]{roberts.pdf}
+	\includegraphics[width=0.15\textwidth]{roberts.eps}
 	\caption{\textit{Roberts cross-gradient} 2-D masks.}
@@
 	\centering
-	\includegraphics[width=0.25\textwidth]{prewitt.pdf}
+	\includegraphics[width=0.25\textwidth]{prewitt.eps}
 	\caption{\textit{Prewitt} 2-D masks of size $3\times3$.}
@@
 	\centering
-	\includegraphics[width=0.25\textwidth]{sobel.pdf}
+	\includegraphics[width=0.25\textwidth]{sobel.eps}
 	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
@@
 
+%\begin{align}
+%	\label{eq:haralick_masks}
+%	\mathbf{b_1} & = 	\begin{bmatrix} 425 & 275 & 225 & 275 & 425 \\ 
+%										275 & 125 &  75 & 125 & 275 \\ 
+%										225 &  75 &  25 &  75 & 225 \\
+%										275 & 125 &  75 & 125 & 275 \\
+%										425 & 275 & 225 & 275 & 425 
+%						\end{bmatrix} &
+%	\mathbf{b_2} & =	\begin{bmatrix} -2260 & -620 & 0 & 620 & 2260 \\ 
+%										-1660 & -320 & 0 & 320 & 1660 \\ 
+%										-1460 & -220 & 0 & 220 & 1460 \\
+%										-1660 & -320 & 0 & 320 & 1660 \\ 
+%										-2260 & -620 & 0 & 620 & 2260 
+%						\end{bmatrix} \nonumber \\
+%	\mathbf{b_3} & =	\begin{bmatrix} 2260  &  1660 &  1460 &  1660 &  2260 \\ 
+%									  	620   &   320 &   220 &   320 &   620 \\ 
+%										0     &     0 &     0 &     0 &     0 \\
+%										-620  &  -320 &  -220 &  -320 &  -620 \\ 
+%										-2260 & -1660 & -1460 & -1660 & -2260
+%						\end{bmatrix} &
+%	\mathbf{b_4} & =	\begin{bmatrix} 1130 & 620 & 450 & 620 & 1130 \\ 
+%									  	830  & 320 & 150 & 320 &  830 \\ 
+%										730  & 220 &  50 & 220 &  730 \\
+%										830  & 320 & 150 & 320 &  830 \\ 
+%										1130 & 620 & 450 & 620 & 1130
+%						\end{bmatrix} \nonumber \\
+%	\mathbf{b_5} & =	\begin{bmatrix} -400 & -200 & 0 &  200 &  400 \\ 
+%									  	-200 & -100 & 0 &  100 &  200 \\ 
+%										0    &    0 & 0 &    0 &    0 \\
+%										200  &  100 & 0 & -100 & -200 \\ 
+%										400  &  200 & 0 & -200 & -400
+%						\end{bmatrix} &
+%	\mathbf{b_6} & =	\begin{bmatrix} 1130 & 830 & 730 & 830 & 1130 \\ 
+%									  	620  & 320 & 220 & 320 &  620 \\ 
+%										450  & 150 &  50 & 150 &  450 \\
+%										620  & 320 & 220 & 320 &  620 \\ 
+%										1130 & 830 & 730 & 830 & 1130
+%						\end{bmatrix} \nonumber \\
+%	\mathbf{b_7} & =	\begin{bmatrix} -8260 & -2180 & 0 & 2180 & 8260 \\ 
+%									  	-6220 & -1160 & 0 & 1160 & 6220 \\ 
+%										-5540 &  -820 & 0 &  820 & 5540 \\
+%										-6220 & -1160 & 0 & 1160 & 6220 \\ 
+%										-8260 & -2180 & 0 & 2180 & 8260
+%						\end{bmatrix} &
+%	\mathbf{b_8} & =	\begin{bmatrix} 5640  &  3600 &  2920 &  3600 &  5640 \\ 
+%									  	1800  &   780 &   440 &   780 &  1800 \\ 
+%										0     &     0 &     0 &     0 &     0 \\
+%										-1800 &  -780 &  -440 &  -780 & -1800 \\ 
+%										-5640 & -3600 & -2920 & -3600 & -5640
+%						\end{bmatrix} \nonumber \\
+%	\mathbf{b_9} & =	\begin{bmatrix} -5640 & -1800 & 0 & 1800 & 5640 \\ 
+%									  	-3600 &  -780 & 0 &  780 & 3600 \\ 
+%										-2920 &  -440 & 0 &  440 & 2920 \\
+%										-3600 &  -780 & 0 &  780 & 3600 \\ 
+%										-5640 & -1800 & 0 & 1800 & 5640
+%						\end{bmatrix} &
+%	\mathbf{b_{10}} & =	\begin{bmatrix} 8260  &  6220 &  5540 &  6220 &  8260 \\ 
+%									  	2180  &  1160 &   820 &  1160 &  2180 \\ 
+%										0     &     0 &     0 &     0 &     0 \\
+%										-2180 & -1160 &  -820 & -1160 & -2180 \\ 
+%										-8260 & -6220 & -5540 & -6220 & -8260
+%						\end{bmatrix} \nonumber \\
+%\end{align}
+
 \subsubsection{Analytical calculation of the second derivative}
@@
 
-First we see the results of each of the algorithms on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\
+Let us analyze the performance of each of the methods described above using some test images. The first image 
+is a simple example, a white square on black background. In figure \ref{fig:result1} we can see the results 
+of applying each of the algorithms to this image.
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.2\textwidth]{results/square.png}
-	\caption{Test image 1: white square on black background.}
-	\label{fig:original1}
-\end{figure}
-
-The figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
-The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm.
-Execution time\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
-\\
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Roberts.]{\label{fig:result1-a}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_roberts.png}\hspace*{0.5cm}}
-	\quad
-	\subfloat[Prewitt.]{\label{fig:result1-b}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_prewitt.png}\hspace*{0.5cm}}
-	\quad
-	\subfloat[Sobel.]{\label{fig:result1-c}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_sobel.png}\hspace*{0.5cm}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result1-d}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth.png}\hspace*{0.5cm}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result1-e}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth-log.png}\hspace*{0.5cm}}
-	\quad
-	\subfloat[Haralick.]{\label{fig:result1-f}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_haralick.png}\hspace*{0.5cm}}
-	\caption{Results of the algorithms using the figure \ref{fig:original1} as input image.}
-	\label{fig:result1}
-\end{figure}
-
-\begin{table}[t!]
-	\begin{center}
-	\begin{tabular}{| l | r |}
-		\hline \rule{0pt}{3ex}
-		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
-		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
-		Haralick									& $0.040 \ s$										\\
-		\hline
-	\end{tabular}
-	\end{center}
-	\caption{Execution time of the algorithms using the figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
-	\label{exectime1}
-\end{table}
-\vspace{0.5cm}
-
-It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster.
-Let us now analyze the performance of each of the algorithmos using a more realistic test image, shown in figure \ref{fig:original2}.
-\\
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=0.5\textwidth]{results/molino.jpg}
-	\caption{Test image 2: windmill.}
-	\label{fig:original2}
-\end{figure}
-
-First we show in figure \ref{fig:result2} the results of the first derivative edge detection algorithms.
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Roberts.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}
-
-	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
-	\caption{Results of the first derivative algorithms using the figure \ref{fig:original2} as input image.}
-	\label{fig:result2}
-\end{figure}
-
-Now we show in figures \ref{fig:result3-a} and \ref{fig:result3-b} the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and in figure \ref{fig:result3-c} the results obtained using the Haralick algorithm.
-In table \ref{exectime2} we show the execution time for each one of the algorithms.
-\\
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-a}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-b}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
-
-	\subfloat[Haralick.]{\label{fig:result3-c}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
-	\caption{Results of the Marr-Hildreth and the Haralick algorithms, using the figure \ref{fig:original2} as input image.}
-	\label{fig:result3}
-\end{figure}
-
-\begin{table}[h!]
-	\begin{center}
-	\begin{tabular}{| l | r |}
-		\hline \rule{0pt}{3ex}
-		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
-		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
-		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
-		Haralick									& $0.930 \ s$										\\
-		\hline
-	\end{tabular}
-	\end{center}
-	\caption{Execution time of the algorithms using the figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
-	\label{exectime2}
-\end{table}
-\vspace{0.5cm}
-
-It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. \\
-
-\clearpage
 %-------------------------------------------------------------------------------
@@
 
-\clearpage
 %-------------------------------------------------------------------------------
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
+	\includegraphics[width=0.5\textwidth]{kernel_gaussian.eps}
 	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
+	\includegraphics[width=0.5\textwidth]{kernel_log.eps}
 	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
@@
 
-\begin{SCfigure}[][!t]
+\begin{SCfigure}[][!b]
 	\centering
-	\includegraphics[width=0.5\textwidth]{kernels.pdf}
+	\includegraphics[width=0.5\textwidth]{kernels.eps}
 	\caption{Comparison of the Gaussian and LoG functions.}
@@
 
-\subsection{Images}
-
-Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} to \ref{fig:example3}. More examples can be found in the online demo. \\
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
-	\caption{Example: Lena.}
-	\label{fig:example1}
-\end{figure}
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
-	\caption{Example: Oranges.}
-	\label{fig:example2}
-\end{figure}
-
-\vspace{2cm}
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
-	\caption{Example: Shapes.}
-	\label{fig:example3}
-\end{figure}
-
 %-------------------------------------------------------------------------------
@@
 
-%TODO: videos de ejemplo con todos los métodos.
-
-This is an example of applying the edge detection algorithms described here, frame by frame, to a video: 
+This is an example of applying the Marr-Hildreth edge detector, frame by frame, to a video:
 \begin{itemize}
-	\centering
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (xxxMb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original}(xxxMb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth.mov}{marr-hildreth version}(xxxMb).
 \end{itemize}
-\begin{multicols}{2}
-\begin{itemize}
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts.mov}{roberts version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt.mov}{prewitt version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel.mov}{sobel version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian.mov}{marr-hildreth-gaussian version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log.mov}{marr-hildreth-log version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick.mov}{haralick version} (xxxMb).
-\end{itemize}
-\end{multicols}
 

### 2014-11-11 old_versions/edge-detection-review-v5.tex
second round of third revision, improving references

--- 2012-07-23 edge-detection-review-v4.tex
@@
 %-------------------------------------------------------------------------------
-% Edge detectors review
+% Edge detectors review v5
 % by Haldo Spontón & Juan Cardelino
@@
 \lstset{
-	backgroundcolor=\color[rgb]{0.9,0.9,0.9},
+	backgroundcolor=\color[rgb]{0.94,0.94,0.94},
     tabsize=4,
@@
          extendedchars=\true,
-         language=C++,
+         language=C,
          basicstyle=\scriptsize,
@@
 In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
-behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
+behind edge detection is to find where there are abrupt changes in the intensity of an image. 
 As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
-such as Sobel, Prewitt and Roberts \cite{Gonzalez2007Digital}. The second derivative is also used, for example in algorithms 
-like Marr-Hildreth \cite{AIM-518} and Haralick \cite{bb20239}.\\
+such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
+like Marr-Hildreth and Haralick.\\
 We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
@@
 image with a kernel (step required in some of the algorithms).\\
-
-%The Marr-Hildreth algorithm is a method of detecting edges in digital 
-%images. It's based on convolving the image with the Laplacian of the 
-%Gaussian function. Then, zero crossings are detected in the filtered 
-%result to obtain the edges points.
 \end{ipolAbstract}
@@
 downloaded from \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}.
-%\begin{itemize}
-%	\item Roberts, Prewitt and Sobel operators: \href{http://iie.fing.edu.uy/~haldos/ipol/fded.tar.gz}{here}.
-%	\item Marr-Hildreth algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/marr-hildreth.tar.gz}{here}.
-%	\item Haralick algorithm: \href{http://iie.fing.edu.uy/~haldos/ipol/haralick.tar.gz}{here}.
-%\end{itemize}
-% TODO referencia a github
+You can also follow this code project in this GIT repository: 
+\begin{center}
+\url{git://github.com/haldos/edges.git}
+\end{center}
 \end{ipolCode}
@@
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image. In section ~\ref{sec:first} we study the 
+first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} we study the 
 performance of such algorithms, as an introduction to a more sophisticated 
@@
 
-In 1980 Marr and Hildreth argued that intensity changes are not independent 
+In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
 of image scale, so the edge detection requires the use of different size 
@@
 study of this algorithm, presented in Section ~\ref{sec:second}. In this section 
-we also present the Haralick algorithm, which is also based on the second derivative.
+we also present the Haralick algorithm \cite{bb20239}, which is also based on the second derivative.
 This algorithm has the particularity of proposing a model to locally approximate 
-the image around a point. Then, with this model, second derivative can be calculated 
+the image around a point. Then, with this model, the second derivative can be calculated 
 analytically (rather approximate) and finding edges is done by imposing a 
@@
 \ref{sec:conclusions} respectively. In section ~\ref{sec:examples} we present some examples to compare the 
-performance of the algorithms studied in this work. We also present a 
-video created using the implemented algorithms, applied frame by frame.\\
-
+performance of the algorithms studied in this work. \\
+
+In conjunction with this paper we present a detailed and well commented source code, which 
+implements the described algorithms. Also in the appendices (sections \ref{sec:appendix1} 
+and \ref{sec:appendix2}) we show some common mathematical developments, and a commented 
+explanation of the source code. This work also provides tools for code documentation 
+(see section \ref{sec:appendix2}). \\
+
+\nocite{IPOL}
 %-------------------------------------------------------------------------------
@@
 
-The two algorithms that we will see in this section (Marr-Hildreth [1980] and Haralick [1987]) 
+The two algorithms that we will see in this section (Marr-Hildreth \cite{AIM-518} [1980] and Haralick \cite{bb20239} [1987]) 
 are based on the second derivative of the image, and both take steps to reduce noise before 
@@
 Gaussian function. There are more ways to do so, for example, using 
-recursive Gaussian filters.\\ % poner referencia!
+recursive Gaussian filters \cite{Deriche1993Recursively}.\\
 
@@
 equivalent to cancel the second derivative of the polynomial as a function of the parameters of it.\footnote{This implementation is slightly 
-different to the traditional implementation of the Haralick algorithm. We do not use the condition concerning the third derivative.}\\
+different to the traditional implementation of the Haralick algorithm.
+We do not use the condition concerning the third derivative (to be negative), in order to implement two edge detection methods using only the second derivative (along the Marr-Hildreth algorithm).}\\
 
@@
 
-First we see the results of each of the algorithms on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\
+First we show the results of each algorithm on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\
 
@@
 	\subfloat[Haralick.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
-	\caption{Results of the first derivative, Marr-Hildreth and Haralick algorithms using the figure \ref{fig:original2} as input image.}
+	\caption{Results of the first derivative, Marr-Hildreth and Haralick algorithms using figure \ref{fig:original2} as input image.}
 	\label{fig:result2}
@@
 	\end{center}
-	\caption{Execution time of the algorithms using the figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
+	\caption{Execution time of the algorithms using figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
 	\label{exectime2}
 \end{table}
-\vspace{0.5cm}
-
-It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones.
-We can see in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to better view of the detail.
+
+In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative.
+We can see in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail.
 Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded).
@@
 
-Another observation is that the edges detected by the Haralick algorithm appear to be thicker than those detected with the other algorithms.
-This may be due the regularity that Haralick assumes, previously mentioned.
-\\
+Another observation is that edges detected by Haralick algorithm appear to be thicker than those detected with the other algorithms. This may be due the regularity that Haralick assumes, we mentioned 
+before. \\
 
@@
 	\subfloat[Haralick.]{\label{fig:result3-g}\includegraphics[width=0.48\textwidth]{results/molino_haralick_zoom.png}}
-	\caption{Results of the algorithms using the figure \ref{fig:original2} as input image, enlarging an interesting area.}
+	\caption{Results of the algorithms using figure \ref{fig:original2} as input image, enlarging an interesting area.}
 	\label{fig:result3}
@@
 
-We have discussed and carefully implemented some of the more traditional methods of edge detection in digital images.
+We have discussed and carefully implemented some of the most traditional methods of edge detection in digital images.
 We have tested the implemented algorithms with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
@@
 
-Comparing the two versions of the Marr-Hildreth algorithm, is significantly faster the version with Gaussian kernel and numerical approximation of the Laplacian.
-The LoG kernel version, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
+Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one.
+The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
 It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
@@
 
-All algorithms have their advantages and disadvantages. Choosing one or other may depend on requirements after edge detection. \\
-
-% At the end:
-Finally, this paper is a summary of some classical edge detection algorithms.
+All algorithms have their advantages and disadvantages. Choosing one or other may depend on requirements of the application. \\
+
+To conclude, this paper is a summary of some classical edge detection algorithms.
 In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods.
 \\
+
+As an addendum to this work we present a detailed implementation of the described algorithms, 
+available in C language, along with an online demo where users can run the algorithms 
+with arbitrary images uploaded by themselves. We also supply some code documentation tools, 
+which were used to generate the detailed code description presented in section \ref{sec:appendix2} (Appendix 2). We think this tool could be useful for future IPOL articles. \\
 
@@
 	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity in the edges.
-They are also very affected by noise in images, appearing unwanted isolated edges.
+They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g.
+Canny edge detector \cite{Canny1986}).
 \\
@@
 This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel.
-This behavior can be seen in Lena's hair (figure \ref{fig:example1}), or in the inside of the oranges (example \ref{fig:example2}).
+This behavior can be seen in Lena's hair (figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}).
 \\
@@
 
-%\vspace{2cm}
-
-%\begin{figure}[t!]
-%	\centering
-%	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}
-
-%	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
-%	\quad
-%	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
-%	\quad
-%	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}
-
-%	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
-%	\quad
-%	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
-%	\quad
-%	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
-%	\caption{Example: Shapes.}
-%	\label{fig:example3}
-%\end{figure}
-
 %-------------------------------------------------------------------------------
 \subsection{Video}
-
-%TODO: videos de ejemplo con todos los métodos.
 
@@
 	\centering
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (xxxMb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (43 Mb).
 \end{itemize}
@@
 \begin{itemize}
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts.mov}{roberts version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt.mov}{prewitt version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel.mov}{sobel version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian.mov}{marr-hildreth-gaussian version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log.mov}{marr-hildreth-log version} (xxxMb).
-	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick.mov}{haralick version} (xxxMb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts-wide_0.1.mov}{roberts version} (7.9 Mb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt-wide_0.1.mov}{prewitt version} (9.1 Mb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel-wide_0.1.mov}{sobel version} (9.1 Mb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian-wide_3_19_0.04.mov}{marr-hildreth-gaussian version} (12 Mb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log-wide_3_25_0.04.mov}{marr-hildreth-log version} (14 Mb).
+	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick-wide_0.5.mov}{haralick version} (12 Mb).
 \end{itemize}
@@
 \subsection{Kernel generation}
-\label{sec:appendix2}
 
@@
 
+This code documentation was generated using a perl script to extract some parts of the 
+code and comments into latex files. This code documentation tool is available as a GIT
+project: 
+\begin{center} 
+\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
+\end{center}
+Also Doxygen code documentation is available:
+\begin{center}
+\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}
+\end{center}
+
 \subsection{Roberts, Prewitt and Sobel}

### 2014-11-11 old_versions/edge-detection-review-v6.tex
second round of third revision, improving references

--- 2014-11-11 old_versions/edge-detection-review-v5.tex
@@
 %-------------------------------------------------------------------------------
-% Edge detectors review v5
+% Edge detectors review v6
 % by Haldo Spontón & Juan Cardelino
@@
 %-------------------------------------------------------------------------------
+
 \documentclass{ipol}
 \ipolSetYear{2012}
-\ipolSetDOI{10.XXXX/ipol.2012.sc-red}
+\ipolSetDOI{10.XXXX/ipol.2012.sc-arced}
 
@@
 \usepackage{color}
+%\usepackage{algorithm}
+\usepackage{algorithmic}
 \newtheorem{theorem}{Theorem}
@@
 %-------------------------------------------------------------------------------
-\title{Review of Edge Detectors}
+
+\title{A Review of Classic Edge Detectors}
 \author{Haldo Spont\'on,\\
@@
 %-------------------------------------------------------------------------------
+
 \begin{ipolAbstract}
-In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
+In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
 behind edge detection is to find where there are abrupt changes in the intensity of an image. 
@@
 like Marr-Hildreth and Haralick.\\
-We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
-point of view (number of operations, execution time), considering different forms to convolve an 
-image with a kernel (step required in some of the algorithms).\\
+Results obtained from a qualitative point of view (perceptual) and from a quantitative 
+point of view (number of operations, execution time) are compared, considering different forms to convolve an 
+image with a kernel (step required in some of the algorithms).
 \end{ipolAbstract}
@@
 %-------------------------------------------------------------------------------
+
 \begin{ipolCode}
 The C implementation of the algorithms implemented in this work (beta version) can be
-downloaded from \href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}.
+downloaded from:
+\begin{center}
+\href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{\url{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}}.
+\end{center}
 You can also follow this code project in this GIT repository: 
@@
 %-------------------------------------------------------------------------------
+
 %\begin{ipolSupp}
@@
 %-------------------------------------------------------------------------------
+
 \section{Introduction}
@@
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} we study the 
-performance of such algorithms, as an introduction to a more sophisticated 
-analysis into the edge finding process.\\
+first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} the 
+behavior of such algorithms is described, as an introduction to a more sophisticated 
+analysis into the edge detection process.\\
 
 In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
-of image scale, so the edge detection requires the use of different size 
+of image scale, so edge detection requires the use of different size 
 operators. They also argued that a sudden intensity change will be seen 
@@
 study of this algorithm, presented in Section ~\ref{sec:second}. In this section 
-we also present the Haralick algorithm \cite{bb20239}, which is also based on the second derivative.
+the Haralick's algorithm \cite{bb20239} is also presented, which is based on second derivative too.
 This algorithm has the particularity of proposing a model to locally approximate 
-the image around a point. Then, with this model, the second derivative can be calculated 
+the image around a point. Then, with this model, second derivative can be calculated 
 analytically (rather approximate) and finding edges is done by imposing a 
@@
 
-In conjunction with this paper we present a detailed and well commented source code, which 
-implements the described algorithms. Also in the appendices (sections \ref{sec:appendix1} 
-and \ref{sec:appendix2}) we show some common mathematical developments, and a commented 
-explanation of the source code. This work also provides tools for code documentation 
-(see section \ref{sec:appendix2}). \\
+Within this paper, a detailed and well commented source code is presented, which 
+implements the described algorithms. In sections \ref{sec:appendix1} 
+and \ref{sec:appendix2} some common mathematical developments, and a commented 
+explanation of the source code are shown. This work also provides tools for code documentation 
+(see section \ref{sec:appendix2}).
 
 \nocite{IPOL}
-%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+
 \section{Algorithms based on first derivative}
@@
 intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
-as the vector:
+as the vector
 \begin{equation}
@@
 					\cfrac{\partial f}{\partial x} \\ \cfrac{\partial f}{\partial y} \\
-				\end{bmatrix}
+				\end{bmatrix}.
 \end{equation}
@@
 The magnitude ($M$) and direction ($\alpha$) of the gradient vector $\nabla f$ at location $(x,y)$
-are calculated as:
-\begin{equation}
+are calculated as
+\begin{equation}
+\label{eq:mag_alpha}
 	\begin{cases}
 		M(x,y) = \sqrt{g_x^2 + g_y^2} \\
-		\alpha(x,y) = \tan^{-1} \begin{bmatrix} g_x \\ g_y \end{bmatrix} \\
+		\alpha(x,y) = \tan^{-1} \begin{bmatrix} g_x \\ g_y \end{bmatrix} .\\
 	\end{cases}
@@
 
-We need to compute partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
-at every pixel in the image. Since we are dealing with digital images, numerical approximations 
-of the partial derivatives over a neighborhood about a point are required. In the following we will study 
-the methods of Roberts, Prewitt and Sobel, whose main difference is how they perform this calculation.\\
+The computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
+at every pixel in the image is required. Dealing with digital images, numerical approximations 
+of the partial derivatives over a neighborhood about a point are required. In the following,
+the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.
+
+%-------------------------------------------------------------------------------
 
@@
 The most usual approach of the first derivative is to take the Taylor expansion of first order with
-a small $h$. Approximating the first derivative:
-\begin{equation}
-	f'(x) \simeq \frac{f(x+h) - f(x)}{h}\\
-\end{equation}
-The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, we approximate the 
-first derivative of the image based on the intensity values ​​at points of the image as:
+a small $h$. Approximating the first derivative
+\begin{equation}
+	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
+\end{equation}
+The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
+first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
 \begin{equation}
@@
 \label{eq:roberts2}
-	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j)\\
+	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
 \end{equation}
@@
 Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
-by filtering the image $f(x,y)$ with the 1-D masks in figure \ref{fig:1dmasks}. But when we want 
-to find diagonal edges, we need 2-D masks. The \textit{Roberts operators} are one of the earliest 
+by filtering the image $f(x,y)$ with the 1-D masks in figure \ref{fig:1dmasks}. But when finding 
+diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
 attempts to use 2-D masks for this purpose. These operators are based on implementing the diagonal 
-diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}.\\
+diferences implemented by filtering an image with the masks in figure \ref{fig:roberts}. By convention 
+in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. \\
 
@@
 \end{SCfigure}
+
+%-------------------------------------------------------------------------------
 
@@
 
+%-------------------------------------------------------------------------------
+
 \subsection{The \textit{Sobel} operators}
 
-A slight variation of the Prewitt operators use more weight on the central coefficients of the 
+A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
 difference. It can be shown that using a $2$ in the center location provides image smoothing. This
@@
 The Prewitt masks are simpler to implement than the Sobel masks, but the computational difference
-between them is not an issue. It is preferable to use Sobel operators, because they have better
-noise-suppression characteristics (smoothing). Noise suppression is an important issue when dealing
-with derivatives.\\
-
-%-------------------------------------------------------------------------------
+between them is not an issue. It is preferable to use Sobel operators, because they are better localized, 
+and their filter are also less aliased, because of the Gaussian shape of $(1\ 2\ 1)$.
+
+% they have better noise-suppression characteristics (smoothing). Noise suppression is an important issue when dealing with derivatives.
+
+%-------------------------------------------------------------------------------
+
+\subsection{Edges image}
+
+Once the respective operators are applied, two images per method are obtained, containing the approximation of horizontal
+and vertical components of the first derivative, respectively. Edges image is obtained by applying a 
+threshold to the first derivative magnitude image, calculated using the equation \ref{eq:mag_alpha}. 
+A black and white image is obtained, in which edge points are indicated in white.
+
+%-------------------------------------------------------------------------------
+
+\subsection{Pseudo-code}
+
+Pseudo-code of implemented algorithms is shown in algorithm \ref{algo:fded}. \\
+
+\begin{algorithm}
+\caption{First derivative edge detection algorithms.}
+\label{algo:fded}
+\begin{algorithmic}[1]
+\REQUIRE input image, threshold $th$.
+\STATE $im_{ORIG} \leftarrow$ input image
+\IF{$im_{ORIG}$ is grayscale}
+	\STATE $im \leftarrow im_{ORIG}$
+\ELSE 
+	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
+\ENDIF
+\STATE Define $operator_x$ and $operator_y$ \COMMENT{Roberts, Prewitt or Sobel.}
+\STATE $g_x \leftarrow$ convolution($im$,$operator_x$)
+\STATE $g_y \leftarrow$ convolution($im$,$operator_y$)
+\STATE $max_M \leftarrow 0$
+\FORALL{pixel $i$ in image}
+	\STATE $M[i] \leftarrow \sqrt{g_x^2+g_y^2}$
+	\IF{$M[i]>max_M$}
+		\STATE $max_M \leftarrow M[i]$
+	\ENDIF
+\ENDFOR
+\FORALL{pixel $i$ in image}
+	\IF{$M[i] \geq th \times max_M$}
+		\STATE $im_{OUT}[i] \leftarrow 255$
+	\ELSE
+		\STATE $im_{OUT}[i] \leftarrow 0$
+	\ENDIF
+\ENDFOR
+\RETURN output image $\leftarrow im_{OUT}$
+\end{algorithmic}
+\end{algorithm}
+
+%-------------------------------------------------------------------------------
+
 \section{Algorithms based on second derivative}
@@
 
-The two algorithms that we will see in this section (Marr-Hildreth \cite{AIM-518} [1980] and Haralick \cite{bb20239} [1987]) 
+The two algorithms that are going to be shown in this section (Marr-Hildreth \cite{AIM-518} [1980] and Haralick \cite{bb20239} [1987]) 
 are based on the second derivative of the image, and both take steps to reduce noise before 
-detecting edges in the image.\\
+detecting edges in the image.
+
+%-------------------------------------------------------------------------------
 
@@
 The Marr-Hildreth algorithm is a method of detecting edges in digital 
-images. It's based on find zero crossing points of the second derivative
-of the image. This can be done in several ways. In this work we implemented 
-two different ways; convolving the image with a Gaussian kernel and then 
+images. It is based on find zero crossing points of the second derivative
+of the image. This can be done in several ways. Two different ways of doing 
+this are implemented implemented in this work; convolving the image with a Gaussian kernel and then 
 approximating the second derivative (Laplacian) with a 3x3 kernel, or 
@@
 	\end{itemize}
-	\item Search of zero crossing points in the filtered image.\\
+	\item Search of zero crossing points in the filtered image.
 \end{enumerate}
@@
 kernel generation, and 2-D convolution of an image with a given kernel, 
-with zero padding.\\
-
-%-------------------------------------------------------------------------------
+with zero padding.
+
+%-------------------------------------------------------------------------------
+
 \subsubsection{Gaussian and LoG kernels}
 
-The Marr-Hildreth algorithm consists of convolving the LoG filter with an input image, $f(x,y)$.
+The Marr-Hildreth algorithm consists of convolving the input image $f(x,y)$ with a LoG filter;
 \begin{equation}\label{eq:log}
-  g(x,y) = [\nabla^2G(x,y)]\star f(x,y)
+  g(x,y) = [\nabla^2G(x,y)]\star f(x,y), 
 \end{equation}
@@
 \end{equation}
-indicating that we can smooth the image first with a Gaussian filter and then compute the Laplacian of the result.\\
-
-Then we need to generate both Gaussian kernel and LoG kernel (see section \ref{sec:appendix2}).\\
-
-The Marr-Hildreth edge-detection algorithm may be summarized as follows:
+indicating that image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result.\\
+
+Then, generation of both Gaussian and LoG kernel is required (see section \ref{sec:appendix2}).\\
+
+Marr-Hildreth edge-detection algorithm may be summarized as follows:
 \begin{enumerate}
@@
 %-------------------------------------------------------------------------------
+
 \subsubsection{Zero crossing}
@@
 different. There are four cases to test: left/right, up/down, and the two diagonals. In this case 
-we work with a threshold, so that not only the signs of the opposite pixels must differ, also their 
+a threshold is used, so that not only the signs of the opposite pixels must differ, also their 
 difference in absolute value must be greater than a certain threshold.\\
@@
 finding the zero crossings is needed, methods for finding zero crossings with subpixel accuracy 
-should be used.\\
-
-%-------------------------------------------------------------------------------
+should be used.
+
+%-------------------------------------------------------------------------------
+
+\subsubsection{Pseudo-code}
+
+Pseudo-code of Marr-Hildreth's algorithm is shown in algorithm \ref{algo:mh}.
+
+\begin{algorithm}[t]
+\caption{Marr-Hildreth edge detection algorithm.}
+\label{algo:mh}
+\begin{algorithmic}[1]
+\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
+\STATE $im_{ORIG} \leftarrow$ input image
+\IF{$im_{ORIG}$ is grayscale}
+	\STATE $im \leftarrow im_{ORIG}$
+\ELSE 
+	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
+\ENDIF
+\STATE $kernel \leftarrow$ Generated Gaussian or LoG kernel
+\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
+\IF{Gaussian kernel}
+	\STATE Define Laplacian approx. operator $laplacian$
+	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
+\ELSE
+	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
+\ENDIF
+\STATE $max_L \leftarrow 0$
+\FORALL{pixel $i$ in image $im_{LAPL}$}
+	\IF{$im_{LAPL}[i]>max_L$}
+		\STATE $max_L \leftarrow im_{LAPL}[i]$
+	\ENDIF
+\ENDFOR
+\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
+	\IF{opposite neighbors of pixel $i$ have opposite signs in $im_{LAPL}$}
+		\STATE $im_{OUT}[i] \leftarrow 255$
+	\ELSE
+		\STATE $im_{OUT}[i] \leftarrow 0$
+	\ENDIF
+\ENDFOR
+\RETURN output image $\leftarrow im_{OUT}$
+\end{algorithmic}
+\end{algorithm}
+
+%-------------------------------------------------------------------------------
+
 \subsection{The \textit{Haralick} algorithm}
@@
 The idea behind the Haralick edge detector is identical to that of the previous method; find zeros in 
-the second derivative of the image. In this method, we smooth the input image through local bi-cubic
-polynomial fitting. Then, calculating the second derivative analytically, we can find an expression 
-equivalent to cancel the second derivative of the polynomial as a function of the parameters of it.\footnote{This implementation is slightly 
-different to the traditional implementation of the Haralick algorithm.
-We do not use the condition concerning the third derivative (to be negative), in order to implement two edge detection methods using only the second derivative (along the Marr-Hildreth algorithm).}\\
+the second derivative of the image. In this method, the input image is smoothed through local bi-cubic
+polynomial fitting.
+Then, calculating the second derivative analytically, it is possible to find an equivalent expression to cancel the second derivative of the polynomial as a function of its parameters.\footnote{This implementation is slightly
+different from the traditional implementation of the Haralick algorithm.
+We do not use the condition concerning the third derivative (to be negative), in order to implement two edge detection methods using only the second derivative (along the Marr-Hildreth algorithm).}
+
+%-------------------------------------------------------------------------------
 
@@
 
-We adjust the surrounding neighborhood of a point $(x,y)$ in the image $f$ using the following bi-cubic polynomial:
+The surrounding neighborhood of a point $(x,y)$ in the image $f$ is adjusted using the following bi-cubic polynomial
 \begin{equation}
 	\label{eq:bicubic}
-	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3\\
-\end{equation}
-
-To solve this problem, we need to take more neighbors than coefficients to be adjusted. The smallest 
+	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3.\\
+\end{equation}
+
+To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. The smallest 
 neighborhood of odd dimensions has size $5\times5$. Now it is possible to convolve the input image
@@
 
-Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. We use a least square
-formulation to obtain the coefficients (25 points give equal number of equations):
+Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. A least square
+formulation is used to obtain the coefficients (25 points give equal number of equations),
 
@@
 		\vdots \\
-		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3\\
+		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3 .\\
 	\end{array}
@@
 
-Using matrix notation:
+Using matrix notation,
 
@@
 	\end{bmatrix}
-	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k}\\
+	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k} .\\
 \end{equation*}
 
-Then:
+Then,
 
 \begin{equation*}
-	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f}\\
+	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
 \end{equation*}
 
-$\mathbf{B}$ is a $10\times25$ matrix:
+$\mathbf{B}$ is a $10\times25$ matrix,
 
@@
 		k_{10}
-	\end{bmatrix}\\
+	\end{bmatrix} .\\
 \end{equation*}
 
-For each coefficient ($i$ from $1$ to $10$):
+For each coefficient ($i$ from $1$ to $10$),
 
@@
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i}\\
-\end{equation}
-
-where:
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
+\end{equation}
+
+where
 
@@
 									b_{i,21}	& b_{i,22}	& \hdots	& b_{i,25}	\\
-					\end{bmatrix}\\
-\end{equation}
-
-Using equation \ref{eq:coefficients} and the masks given in table \ref{table:b_i}, we can compute 
+					\end{bmatrix}.\\
+\end{equation}
+
+\vspace{1ex}
+Using equation \ref{eq:coefficients} and the masks given in table \ref{table:b_i}, it is possible to compute 
 the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
@@
 \end{table}
-\clearpage
+
+%-------------------------------------------------------------------------------
 
@@
 
-We approximate the neighborhood of each point of the image using the bi-cubic polynomial 
-expression in equation \ref{eq:bicubic}. If we just take the first order terms of this 
-polynomial, the Gradient angle, defined with positive y-axis is defined as:
+The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
+expression in equation \ref{eq:bicubic}. If just the first order terms of this 
+polynomial are taken, the gradient angle, defined with positive y-axis, is defined as
 
@@
 	\sin(\theta) & = \frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
-	\cos(\theta) & = \frac{k_3}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
+	\cos(\theta) & = \frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
 \end{align}
 
-Now substituting the variables $x$ and $y$ in polar form as:
+Now substituting the variables $x$ and $y$ in polar form as
 
 \begin{equation*}
-	x = \rho\cos{\theta} \: , \: y = \rho\sin(\theta) \\
+	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin(\theta) \\
 \end{equation*}
 
-in the bi-cubic polynomial, we obtain:
-
-\begin{equation}
-	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3
-\end{equation}
-
-where:
+in the bi-cubic polynomial, we obtain
+
+\begin{equation}
+	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3 ,
+\end{equation}
+
+where
 
@@
 
-The derivatives are obtained as follows:
+The derivatives are obtained as follows.
 
@@
 	f'{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
-	f''{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
+	f''{\theta}(\rho) = 2C_2 + 6C_3\rho .\nonumber \\
 \end{align}
 
-Then, the condition of the second derivative equal to zero becomes:
-
-\begin{equation}
-	f''{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0.
-\end{equation}
+Then, the condition that the second derivative is equal to zero becomes
+
+\begin{equation}
+	f''{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0 .
+\end{equation}
+
+%-------------------------------------------------------------------------------
 
@@
 %-------------------------------------------------------------------------------
+
+\subsubsection{Pseudo-code}
+
+Pseudo-code of Haralick's algorithm is shown in algorithm \ref{algo:haralick}.
+
+\begin{algorithm}[t]
+\caption{Haralick edge detection algorithm.}
+\label{algo:haralick}
+\begin{algorithmic}[1]
+\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
+\STATE $im_{ORIG} \leftarrow$ input image
+\IF{$im_{ORIG}$ is grayscale}
+	\STATE $im \leftarrow im_{ORIG}$
+\ELSE 
+	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
+\ENDIF
+\STATE Define the ten $5x5$ masks used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See table \ref{table:b_i}.}
+\FORALL{pixel $i$ in image $im$}
+	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
+	\FOR{$j=1$ \TO $10$}
+		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
+	\ENDFOR
+	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
+	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
+	\IF{$|C_2/3C_3|\leq\rho_0$}
+		\STATE $im_{OUT}[i] \leftarrow 255$
+	\ELSE
+		\STATE $im_{OUT}[i] \leftarrow 0$
+	\ENDIF
+\ENDFOR
+\RETURN output image $\leftarrow im_{OUT}$
+\end{algorithmic}
+\end{algorithm}
+
+%-------------------------------------------------------------------------------
+
+\section{Common Operations}
+\label{sec:appendix1}
+
+All algorithms implemented use some common operations that are independent of the algorithms themselves. 
+These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
+implemented with care.
+
+%-------------------------------------------------------------------------------
+
+\subsection{Grayscale conversion}
+
+The first step in every algorithm presented above is to convert color images to gray intensity images. 
+We use the libpng coefficients to do this:
+\begin{equation}
+    Y = (6968 R + 23434 G + 2366 B) / 32768
+\end{equation}
+where $R$, $G$ and $B$ are the red, green and blue components respectively.
+
+%-------------------------------------------------------------------------------
+
+\subsection{Kernel generation}
+
+Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
+kernels are generated by sampling the corresponding analytical function, which in each case depends 
+on the parameter $\sigma$ (standard deviation of the Gaussian function). The result is an array of 
+size $n$ by $n$.
+
+%-------------------------------------------------------------------------------
+
+\subsubsection{Gaussian kernel}
+
+Gaussian kernel is generated by sampling the 2-D Gaussian function\footnote{Note that for simplicity
+the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
+\begin{equation}
+	\label{eq:gaussian_function}
+	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
+\end{equation}
+where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\
+
+The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
+input parameters, but these are not strictly independent of each other. 
+A large enough value of $n$ is needed to ensure that no information is lost when creating the kernel $G$.
+To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$.
+Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\
+
+Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
+(first odd integer greater than $6\sigma=24$).
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
+	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
+sufficient to have a good approximation of the Gaussian function in the kernel.}
+	\label{fig:gaussian_kernel}
+\end{SCfigure}
+
+%-------------------------------------------------------------------------------
+
+\subsubsection{LoG kernel}
+
+The Laplacian of Gaussian $\nabla^2G(x,y)$ can be analytically obtained first, and a kernel can be generated
+with the result function. Use this kernel for edge detection involves only one convolution with 
+the input image (unlike the case of Gaussian kernel, in which two convolutions have to be maded, one 
+with the kernel and another with the Laplacian operator).\\
+
+\begin{equation}
+	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
+\end{equation}
+
+We first obtain
+
+\begin{equation} 
+	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2}
+\end{equation}
+
+and then
+
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
+\end{equation}
+
+Similarly,
+
+\begin{equation} 
+	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
+\end{equation}
+
+Therefore we obtain
+
+\begin{equation}
+	\label{eq:log_function}
+	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
+\end{equation}\\
+
+Now the LoG kernel is generated by sampling this function. Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
+	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
+to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
+Gaussian kernel.}
+	\label{fig:log_kernel}
+\end{SCfigure}
+
+%-------------------------------------------------------------------------------
+
+\subsubsection{Gaussian and LoG functions comparison}
+
+As mentioned before, the size of the kernel $n$ and the standard deviation of the exponential function 
+$\sigma$ are not independent of each one. This is because the function LoG is "wider" that the 
+Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to 
+generate a correctly sampled LoG kernel than in the case of the Gaussian kernel, with the same $\sigma$.\\
+
+Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
+required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
+example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
+first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
+be $29$.
+
+\begin{SCfigure}[][!t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{kernels.pdf}
+	\caption{Comparison of the Gaussian and LoG functions.}
+	\label{fig:kernels}
+\end{SCfigure}
+
+%-------------------------------------------------------------------------------
+
+\subsection{Convolution}
+
+When making a convolution, it is necessary to define the padding method to calculate such convolution 
+around the edges of the image, and to get a valid output the same size as input image. In this paper, 
+two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
+the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
+kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
+relative to the edge of the image. \\
+
+%-------------------------------------------------------------------------------
+
+\section{Implementation}
+\label{sec:appendix2}
+
+This code documentation was generated using a perl script to extract some parts of the 
+code and comments into latex files. This code documentation tool is available as a GIT
+project: 
+\begin{center} 
+\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
+\end{center}
+Also Doxygen code documentation is available:
+\begin{center}
+\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}
+\end{center}
+
+\subsection{Roberts, Prewitt and Sobel}
+\input{test_fded.tex}
+
+\subsection{Marr-Hildreth}
+\label{app:marr-hildreth}
+\input{test_mh.tex}
+
+\subsection{Haralick}
+\input{test_haralick.tex}
+
+\subsection{Gaussian kernel generation}
+\input{gaussian_kernel.tex}
+
+\subsection{2D convolution}
+\input{2dconvolution.tex}
+
+%-------------------------------------------------------------------------------
+
 \section{Results}
@@
 
-First we show the results of each algorithm on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\
+First the results of each algorithm are shown, using a simple test input image, the white square on black background shown in figure \ref{fig:original1}. \\
 
@@
 
-Let us now analyze the performance of each of the algorithms using a more realistic test image, shown in figure \ref{fig:original2}. \\
+Now the performance of each algorithm is analyzed, using a more realistic test image shown in figure \ref{fig:original2}. \\
 
@@
 
-First we show in figures \ref{fig:result2-a} to \ref{fig:result2-c} the results of the first derivative edge detection algorithms.
+First the results of the first derivative edge detection algorithms is shown figures \ref{fig:result2-a} to \ref{fig:result2-c}.
 Figures \ref{fig:result2-d} and \ref{fig:result2-e} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and figure \ref{fig:result2-f} shows the results obtained using the Haralick algorithm.
-In table \ref{exectime2} we show the execution time for each one of the algorithms.
+Table \ref{exectime2} shows the execution time for each one of the algorithms.
 \\
+
 \begin{figure}[h!]
@@
 	\subfloat[Haralick.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
-	\caption{Results of the first derivative, Marr-Hildreth and Haralick algorithms using figure \ref{fig:original2} as input image.}
+	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using figure \ref{fig:original2} as input image.}
 	\label{fig:result2}
@@
 In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative.
-We can see in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail.
+It can be seen in figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail.
 Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded).
-Neither Haralick algorithm detects it.
+Neither the Haralick algorithm detects it.
 However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.
@@
 
-Another observation is that edges detected by Haralick algorithm appear to be thicker than those detected with the other algorithms. This may be due the regularity that Haralick assumes, we mentioned 
-before. \\
+Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes. \\
 
@@
 \clearpage
-%-------------------------------------------------------------------------------
-\section{Conclusions}
-\label{sec:conclusions}
-
-We have discussed and carefully implemented some of the most traditional methods of edge detection in digital images.
-We have tested the implemented algorithms with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
-\\
-
-The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation.
-Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
-The results obtained with these methods are quite good, considering its simplicity, but they have problems such as noise and discontinuity at the edges.
-\\
-
-The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
-In real images, they have better behavior towards first derivative algorithms.
-\\
-
-Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one.
-The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
-It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
-\\
-
-Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
-It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function.
-This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
-\\
-
-All algorithms have their advantages and disadvantages. Choosing one or other may depend on requirements of the application. \\
-
-To conclude, this paper is a summary of some classical edge detection algorithms.
-In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods.
-\\
-
-As an addendum to this work we present a detailed implementation of the described algorithms, 
-available in C language, along with an online demo where users can run the algorithms 
-with arbitrary images uploaded by themselves. We also supply some code documentation tools, 
-which were used to generate the detailed code description presented in section \ref{sec:appendix2} (Appendix 2). We think this tool could be useful for future IPOL articles. \\
-
-\clearpage
-%-------------------------------------------------------------------------------
-\section{Examples}
+
+%-------------------------------------------------------------------------------
+
+\subsection{Further examples}
 \label{sec:examples}
 
-\subsection{Images}
-
 Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} and \ref{fig:example2}. More examples can be found in the online \href{http://link_to_demo}{demo}. \\
-
-\begin{figure}[h!]
-	\centering
-	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
-	\quad
-	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
-	\caption{Example: Lena (512$\times$512 pixels).}
-	\label{fig:example1}
-\end{figure}
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
-	\quad
-	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}
-
-	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
-	\quad
-	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
-	\quad
-	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}
-
-	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
-	\quad
-	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
-	\quad
-	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
-	\caption{Example: Oranges (536$\times$480 pixels).}
-	\label{fig:example2}
-\end{figure}
 
@@
 
-%-------------------------------------------------------------------------------
+\begin{figure}[h!]
+	\centering
+	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}
+	\quad
+	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/lena_bw.png}}
+
+	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
+	\quad
+	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
+	\quad
+	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}
+
+	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
+	\quad
+	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
+	\quad
+	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
+	\caption{Example: Lena (512$\times$512 pixels).}
+	\label{fig:example1}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[t!]
+	\centering
+	\subfloat[Original.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}
+	\quad
+	\subfloat[Grayscale.]{\includegraphics[width=0.3\textwidth]{examples/oranges_bw.png}}
+
+	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
+	\quad
+	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
+	\quad
+	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}
+
+	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
+	\quad
+	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
+	\quad
+	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
+	\caption{Example: Oranges (536$\times$480 pixels).}
+	\label{fig:example2}
+\end{figure}
+
+%-------------------------------------------------------------------------------
+
 \subsection{Video}
@@
 %-------------------------------------------------------------------------------
-%\begin{thebibliography}{9}
-%\end{thebibliography}
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+Some of the most traditional methods of edge detection in digital images are discussed and carefully implemented in this work.
+The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
+\\
+
+The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation.
+Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
+The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity at the edges.
+\\
+
+The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
+In real images, they have better behavior towards first derivative algorithms.
+\\
+
+Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one.
+The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
+It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
+\\
+
+Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
+It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function.
+This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.
+\\
+
+All algorithms have their advantages and disadvantages. Choosing one or the other may depend on requirements of the application. \\
+
+%To conclude, this paper is a summary of some classical edge detection algorithms.
+In the preparation of this paper, no similar material (concentrated in one text) was found, so this paper appears to be useful material for whom starts studying edge detection methods.
+\\
+
+As an addendum to this work a detailed implementation of the described algorithms is presented, 
+available in C language, along with an online demo where users can run the algorithms 
+with arbitrary images uploaded by themselves. Also some code documentation tools are supplied, 
+which were used to generate the detailed code description presented in section \ref{sec:appendix2} (Appendix 2). We think this tool could be useful for future IPOL articles.
+
+\clearpage
+
+%-------------------------------------------------------------------------------
 
@@
 
-\clearpage
-%-------------------------------------------------------------------------------
-\section{Appendix 1: Common operations}
-\label{sec:appendix1}
-
-All algorithms implemented use some common operations that are independent of the algorithms themselves. 
-These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
-implemented with care.\\
-
-\subsection{Grayscale conversion}
-
-The first step in every algorithm presented above is to convert color images to gray intensity images. 
-We use the libpng coefficients to do this:
-\begin{equation}
-    Y = (6968 R + 23434 G + 2366 B) / 32768
-\end{equation}
-where $R$, $G$ and $B$ are the red, green and blue components respectively.\\
-
-%-------------------------------------------------------------------------------
-\subsection{Kernel generation}
-
-Some of the algorithms presented above require the use of a Gaussian kernel or a LoG kernel. These 
-kernels are generated by sampling the corresponding analytical function, which in each case depends 
-on the parameter $\sigma$ (standard deviation of the Gaussian function). The result is an array of 
-size $n$ by $n$.\\
-
-\subsubsection{Gaussian kernel}
-
-We generate the Gaussian kernel by sampling the 2-D Gaussian function\footnote{Note that for simplicity 
-we omitted the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$.}
-\begin{equation}
-	\label{eq:gaussian_function}
-	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
-\end{equation}
-where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\
-
-The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
-input parameters, but these are not strictly independent of each other. We need a value 
-of $n$ large enough to ensure that no information is lost when creating the kernel $G$. To ensure this, we take 
-$n$ equal to the first odd integer greater than $6\sigma$. Larger values of $n$ does not add more 
-significant samples ​​of $G$, and increases the number of operations in the convolution.\\
-
-Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
-(first odd integer greater than $6\sigma=24$).\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
-	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
-sufficient to have a good approximation of the Gaussian function in the kernel.}
-	\label{fig:gaussian_kernel}
-\end{SCfigure}
-
-\subsubsection{LoG kernel}
-
-We can obtain analytically the Laplacian of Gaussian $\nabla^2G(x,y)$ first, and generate a kernel 
-with the result function. Use this kernel for edge detection involves only one convolution with 
-the input image (unlike the case of Gaussian kernel, where we have to make two convolutions, one 
-with the kernel and another with the Laplacian operator).\\
-
-\begin{equation}
-	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
-\end{equation}
-
-We first obtain:
-
-\begin{equation} 
-	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2}
-\end{equation}
-
-And then:
-
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
-\end{equation}
-
-Similarly:
-
-\begin{equation} 
-	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2} 
-\end{equation}
-
-Therefore we obtain:
-
-\begin{equation}
-	\label{eq:log_function}
-	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}
-\end{equation}\\
-
-Now we generate the LoG kernel by sampling this function. Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the 
-values $\sigma=4$ and $n=31$.\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
-	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
-to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
-Gaussian kernel.}
-	\label{fig:log_kernel}
-\end{SCfigure}
-
-\subsubsection{Gaussian and LoG functions comparison}
-
-As we said before, the size of the kernel $n$ and the standard deviation of the exponential function 
-$\sigma$ are not independent of each one. This is because the function LoG is "wider" that the 
-Gaussian function (i.e. LoG function has a slower decay), so we need a greater value of $n$ to 
-generate a correctly sampled LoG kernel than the needed in the case of the Gaussian kernel, with the same $\sigma$.\\
-
-Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. Is clearly 
-required a larger kernel size in the case of the LoG function (approximately 18\% more). For 
-example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
-first odd integer greater than $6*\sigma$, which is $25$, and for the case of LoG function, would 
-be $29$.\\
-
-\begin{SCfigure}[][!t]
-	\centering
-	\includegraphics[width=0.5\textwidth]{kernels.pdf}
-	\caption{Comparison of the Gaussian and LoG functions.}
-	\label{fig:kernels}
-\end{SCfigure}
-
-\clearpage
-%-------------------------------------------------------------------------------
-\section{Appendix 2: Implementation}
-\label{sec:appendix2}
-
-This code documentation was generated using a perl script to extract some parts of the 
-code and comments into latex files. This code documentation tool is available as a GIT
-project: 
-\begin{center} 
-\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
-\end{center}
-Also Doxygen code documentation is available:
-\begin{center}
-\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}
-\end{center}
-
-\subsection{Roberts, Prewitt and Sobel}
-\input{test_fded.tex}
-
-\subsection{Marr-Hildreth}
-\label{app:marr-hildreth}
-\input{test_mh.tex}
-
-\subsection{Haralick}
-\input{test_haralick.tex}
-
-\subsection{Gaussian kernel generation}
-\input{gaussian_kernel.tex}
-
-\subsection{2D convolution}
-\input{2dconvolution.tex}
-
-%-------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------
+
 \end{document}

### 2014-11-11 old_versions/edge-detection-review-v7.tex
second round of third revision, improving references

--- 2012-07-23 edge-detection-review-v8.tex
@@
 %-------------------------------------------------------------------------------
-% A Review of Classic Edge Detectors - v8
+% A Review of Classic Edge Detectors - v7
 % by Haldo Spontón & Juan Cardelino
@@
 behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
-The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
+The first family of algorithms reviews the use of first derivative to find the changes of intensity, 
 such as Sobel, Prewitt and Roberts. In the second family reviewed, second derivative is used, for example in algorithms 
@@
 Results obtained from a qualitative point of view (perceptual) and from a quantitative 
-point of view (number of operations, execution time) are compared, considering different ways to convolve an 
+point of view (number of operations, execution time) are compared, considering different forms to convolve an 
 image with a kernel (step required in some of the algorithms).
@@
 
-Results of the implemented algorithms are presented in section \ref{sec:results}, along with examples 
+Results of the implemented algorithms are presented in section \ref{sec:results}, along some examples 
 to compare their performance. Conclusions are detailed in section \ref{sec:conclusions}.
@@
 
-To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
+The computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
 at every pixel in the image is required. When dealing with digital images, numerical approximations 
@@
 
-The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with
-a small $h$. Thus $f'(x)$ is computed as
+The most usual approach of the first derivative is to take the Taylor expansion of first order with
+a small $h$. Approximating the first derivative
 \begin{equation}
@@
 \end{equation}
-The image is composed by discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
+The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
 first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
@@
 diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
-in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
-usual to normalize the masks, in order to have gain equal to $1$. \\
+in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. \\
 
@@
 Masks of size $2\times2$, although conceptually simple, are not symmetrical with respect to the central points. 
-Having symmetrical edges is a desirable property and can only be achieved with oddly sized masks, the smallest 
+Having symmetrical edges is a desirable property and can only be achieved with oddly sized mask, the smallest 
 of them being the 3x3. These masks provide more information to find the direction of the edges, because they take 
@@
 These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
-smoothing properties in the direction opposite to the gradient.
+$[-1\ 0\ 1]$ with vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
+smoothing properties.
 
@@
 A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
-difference. It can be shown that using a $2$ in the center location provides better image smoothing. This
+difference. It can be shown that using a $2$ in the center location provides image smoothing. This
 variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
@@
 
-Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
-Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\
-
-The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
+Sobel masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
+$[-1\ 0\ 1]$ with vertically filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
+Hence, these operators have smoothing properties, as mentioned in the preceding paragraph. \\
+
+Computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
 use Sobel operators, because they are better localized, and their filter are also less aliased, 
-because of the weighted shape of $[1\ 2\ 1]$.
+because of the Gaussian shape of $[1\ 2\ 1]$.
 
@@
 	\subfigure[$th = 0.4*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
-	\caption{Thresholded images. Different values of the threshold applied to the magnitude of the gradient.}
+	\caption{Thresholded images.}
 	\label{fig:thresholding}
@@
 Then, once the respective operators are applied, an approximation of the gradient (stored in two images) 
-is obtained, containing the approximation of the partial derivatives $g_x$ and $g_y$. Gradient magnitude 
-image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}). \\
-
-Finally, the edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
+is obtained, containing the approximation of horizontal and vertical components of the first derivative, 
+$g_x$ and $g_y$ respectively. Gradient magnitude image $M$ is calculated using the equation \ref{eq:mag_alpha} 
+(Figure \ref{fig:profiles3d-c}). \\
+
+Finally, edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
 \begin{equation*}
@@
 It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
-for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
-of the image to be processed. The advantage of this is to adapt the threshold to the dynamic range of the image. 
-Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
+for larger thresholds.
 
@@
 
-The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
+Pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
 \clearpage
@@
 The Marr-Hildreth algorithm is a method of detecting edges in digital 
-images. It is based on finding zero crossing points of the second derivative
+images. It is based on find zero crossing points of the second derivative
 of the image. This can be done in several ways. Two different ways of doing 
@@
 computational cost. In general it is a technique that yields good results, but if more precision in 
-finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
-accuracy could be used (e.g.\ marching squares \cite{marching_cubes}).
+finding the zero crossings is needed, methods for finding zero crossings with subpixel accuracy 
+should be used (e.g.\ marching squares \cite{marching_cubes}).
 
@@
 
-The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.
+Pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.
 
@@
 
-In this section the original work of Haralick \cite{bb20239} on edge detection is presented in detail.
-The main idea behind this algorithm is identical to that of the previous method; find zeros in 
-the second derivative of the image. In this method, the input image is smoothly approximated through local bi-cubic
+This section explains in detail the original work by Haralick, particularly is applied the same 
+technique bicubic polynomial fitting (see \ref{sec:bicubic}). There are other options for image 
+interpolation \cite{getreuer}. \\
+
+The idea behind the Haralick edge detector is identical to that of the previous method; find zeros in 
+the second derivative of the image. In this method, the input image is smoothed through local bi-cubic
 polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
 an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
-its parameters.%\footnote{This implementation is slightly different from the traditional implementation 
-%of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
-%in order to implement two edge detection methods using only the second derivative (along the 
-%Marr-Hildreth algorithm).}
+its parameters.\footnote{This implementation is slightly different from the traditional implementation 
+of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
+in order to implement two edge detection methods using only the second derivative (along the 
+Marr-Hildreth algorithm).}
 
@@
 \label{sec:bicubic}
-
-Here, the interpolation method used by Haralick is presented, although there are other options to do this \cite{getreuer}. \\
 
@@
 to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
-By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
+By substituting the 25 data points into the polinomial equation (equation \ref{eq:bicubic}), the following system of 
 equations is obtained,
@@
 
-Using matrix notation, the system can be rewritten as
+Using matrix notation,
 
@@
 expression in equation \ref{eq:bicubic}. If just the first order terms of this 
-polynomial are taken, the gradient angle $\theta$, defined with negative y-axis, can be approxmated as
+polynomial are taken, the gradient angle $\theta$, defined with positive y-axis, is defined as
 
@@
 \label{eq:sincos}
-	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
-	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
+	\sin(\theta) & = \frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
+	\cos(\theta) & = \frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
 \end{align}
@@
 \begin{equation*}
-	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
+	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin(\theta) \\
 \end{equation*}
@@
 	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
-	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
-	f'''_{\theta}(\rho) = 6C_3 .\nonumber \\
+	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho .\nonumber \\
 \end{align}
@@
 \begin{equation}
-	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
-\end{equation}
-
-where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. \\
-
-The other condition required is that the third derivative to be negative, i.e.\
-
-\begin{equation}
-	f'''_{\theta}(\rho) = 6C_3 < 0 \ \ \Rightarrow \ \ C_3 < 0.
+	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0 .
 \end{equation}
@@
 	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
-	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$ and $C_3 < 0$, then that point is an edge point.
+	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$, then that point is an edge point.
 \end{enumerate}
@@
 
-The pseudo-code of Haralick's algorithm is shown in Algorithm \ref{algo:haralick}.
+Pseudo-code of Haralick's algorithm is shown in Algorithm \ref{algo:haralick}.
 
@@
 These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
-implemented with care. In this section those operations are reviewed.
+implemented with care.
 
@@
 \end{equation}
-where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float}. See code in section \ref{app:marr-hildreth}.}.
+where $R$, $G$ and $B$ are the red, green and blue components respectively.
 
@@
 input parameters, but these are not strictly independent of each other. 
-A value of $n$ big enough is needed to ensure that no information is lost when creating the kernel $G$.
+A large enough value of $n$ is needed to ensure that no information is lost when creating the kernel $G$.
 To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$.
@@
 one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
-have to be performed, one with the kernel and another with the Laplacian operator).\\
+have to be maded, one with the kernel and another with the Laplacian operator).\\
 
@@
 kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
-relative to the edge of the image. Convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
-or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.
+relative to the edge of the image. \\
 
@@
 \end{center}
-Required external libraries: 
-\begin{center}
-\texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}) \\
-\texttt{libpng} (\url{http://www.libpng.org/})
-\end{center}
+Required external libraries: \texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}).
 
@@
 
-The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used. \\
+First the results of each algorithm are shown, using a simple test input image, the white square on black background shown in Figure \ref{fig:original1}. \\
 
@@
 Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm.
+The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm.
 Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.
@@
 	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
-	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.
-The first three (\ref{fig:result1-a}, \ref{fig:result1-b} and \ref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\ref{fig:result1-d} and \ref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\ref{fig:result1-f}) corresponds to the Haralick algorithm.}
+	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.}
 	\label{fig:result1}
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including i/o) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
+	\caption{Execution time of the algorithms using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
 	\label{exectime1}
@@
 It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster.
-All algorithms show consistent results.
-Thicker edges appear in the results of Haralick, due to smooth edges model that is imposed in this algorithm.
+All algorithms show consistent results, but the Haralick algorithm shows some rare artifacts out of the edges.
+This may occur because this is the only algorithm that assumes a certain regularity (third degree) in the neighborhood of a pixel.
+Synthetic edges are far from this regularity.
+This may cause the adjustment of the bicubic polynomial function around these points to be quite different from reality.
 \\
@@
 	\quad
-	\subfigure[Image converted to greyscale.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
+	\subfigure[Grayscale image.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
 	\caption{Test image 2: windmill (1000$\times$563 pixels).}
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including i/o) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
+	\caption{Execution time of the algorithms using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
 	\label{exectime2}
@@
 Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded).
-Neither the Haralick algorithm is able to detect it.
+Neither the Haralick algorithm detects it.
 However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.
@@
 \begin{itemize}
-	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity of the edges.
+	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity in the edges.
 They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g.
@@
 
-Some of the most traditional methods of edge detection in digital images were discussed and carefully implemented in this work.
+Some of the most traditional methods of edge detection in digital images are discussed and carefully implemented in this work.
 The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods.
@@
 Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels).
-The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges.
+The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity at the edges.
 \\
@@
 The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed.
-In real images, they have better behavior than first derivative algorithms.
+In real images, they have better behavior towards first derivative algorithms.
 \\

### 2014-11-11 old_versions/edge-detection-review-v9.tex
second round of third revision, improving references

--- 2012-07-23 edge-detection-review-v10.tex
@@
 %-------------------------------------------------------------------------------
-% A Review of Classic Edge Detectors - v10
+% A Review of Classic Edge Detectors - v9
 % by Haldo Spontón & Juan Cardelino
@@
 
-Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
+Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
 \begin{equation*}
-	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{th} \\
-									1,& \mbox{if}\ M[i]<\mbox{th}
+	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{threshold} \\
+									1,& \mbox{if}\ M[i]<\mbox{threshold}.
 					  \end{cases}
 \end{equation*}
-where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
+A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
+It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
 for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
-of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
+of the image to be processed. The advantage of this is to adapt the threshold to the dynamic range of the image. 
 Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.
@@
 
-where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
-should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
-of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
-value is between $0.4$ and $0.6$.\\
+where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. This parameter indicates how far 
+from the center of the pixel can be a zero of the second derivative. It must be greater than zero and and 
+less than $\sqrt{2}$. Its optimal value is between $0.4$ and $0.6$.\\
 
@@
 	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.
-The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm.
+The first three (\ref{fig:result1-a}, \ref{fig:result1-b} and \ref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\ref{fig:result1-d} and \ref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\ref{fig:result1-f}) corresponds to the Haralick algorithm.
 Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
@@
 	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
-This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel.
+This algorithm provides a first step to regularity in the image, because of the filtering with a Gaussian kernel.
 This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}).
@@
 Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas.
-It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function.
+It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinimial function.
 This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too.

### 2014-11-11 old_versions/edge-detection-review.tex
second round of third revision, improving references

--- 2014-11-11 old_versions/edge-detection-review-v2.tex
@@
 \numberwithin{table}{section}
+\numberwithin{figure}{section}
 \begin{document}
@@
 \begin{ipolAbstract}
-In this paper we study different alternatives for edge detection in digital images. The main idea 
+In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
 behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
@@
 
-Edge detection is the first approach for segmenting images
-based on abrupt changes in intensity. Detecting those changes in intensity 
-for the purpose of finding edges in images can be accomplished using first 
-or second order derivatives.
+Edge detection is one of the first approaches published in the literature for 
+segmenting images. The basic idea is to detect abrupt changes in intensity. 
+Detecting those changes in intensity for the purpose of finding edges in images 
+can be accomplished using first or second order derivatives.
 
@@
 first derivative of the image. In section ~\ref{sec:first} we study the 
-performance of these algorithms, as an introduction to a more sophisticated 
+performance of such algorithms, as an introduction to a more sophisticated 
 analysis into the edge finding process.\\
@@
 
-Let us analyze the performance of each of the methods described above using some test images. The first image 
-is a simple example, a white square on black background. In figure \ref{fig:result1} we can see the results 
-of applying each of the algorithms to this image.
+Let us analyze the performance of each of the methods described above using a test image shown in figure \ref{fig:original}.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\textwidth]{results/molino.eps}
+	\caption{Test image: windmill.}
+	\label{fig:original}
+\end{figure}
+
+First we analyze the results of the first derivative edge detection algorithms.
+
+\begin{figure}[h!]
+	\centering
+	\subfloat[Roberts.]{\label{fig:result1-a}\includegraphics[width=0.5\textwidth]{results/molino.roberts.eps}}
+
+	\subfloat[Prewitt.]{\label{fig:result1-b}\includegraphics[width=0.5\textwidth]{results/molino.prewitt.eps}}
+
+	\subfloat[Sobel.]{\label{fig:result1-c}\includegraphics[width=0.5\textwidth]{results/molino.sobel.eps}}
+	\caption{Results of the algorithms using the figure \ref{fig:original} as input image}
+	\label{fig:result1}
+\end{figure}
+
+%\begin{figure}[b!]
+%	\centering
+%	\subfloat[Roberts.]{\label{fig:result1-a}\includegraphics[width=0.3\textwidth]{results/square/square.roberts.eps}}
+%	\quad
+%	\subfloat[Prewitt.]{\label{fig:result1-b}\includegraphics[width=0.3\textwidth]{results/square/square.prewitt.eps}}
+%	\quad
+%	\subfloat[Sobel.]{\label{fig:result1-c}\includegraphics[width=0.3\textwidth]{results/square/square.sobel.eps}}
+
+%	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result1-a}\includegraphics[width=0.3\textwidth]{results/square/square.marr-hildreth.eps}}
+%	\quad
+%	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result1-b}\includegraphics[width=0.3\textwidth]{results/square/square.marr-hildreth-log.eps}}
+%	\quad
+%	\subfloat[Haralick.]{\label{fig:result1-c}\includegraphics[width=0.3\textwidth]{results/square/square.haralick.eps}}
+%	\caption{Results of the algorithms using the figure \ref{fig:original1} as input image}
+%	\label{fig:result1}
+%\end{figure}
 

### 2014-11-18 edge-detection-review-3rd-revision-v2.tex
added 2nd version of review and missing files

@@
 % en este parrafo, puede ser cambiar trough por valley
-In 1980 Marr and Hildreth \cite{segm:edge_region:marr:80:digital_step} argued that intensity changes are not independent 
+In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
 of image scale, so edge detection requires the use of different size 
@@
 These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
-These masks are called \textit{Prewitt operators} \cite{prewitt1970object}.
+These masks are called \textit{Prewitt operators} \cite{im_proc:segmentation:prewitt:1970:object_enhancement}.
 
@@
 
-\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:80:digital_step} and Haralick \cite{bb20239}) 
+\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
 are based on the second derivative of the image, and both take steps to reduce noise before 
@@
 finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
-accuracy could be used (e.g.\ marching squares \cite{marching_cubes}).
+accuracy could be used (e.g.\ marching squares \cite{comp_graph:surf_recon:lorensen:87:marching_cubes}).
 
@@
 
-If we substitute the above expressions in the bi-cubic polynomia of eq., we obtain
+If we substitute the above expressions in the bi-cubic polynomial of Eq. \ref{eq:bicubic}, we obtain
 
@@
 \begin{equation}
-\exists\rho \in [0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
+\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
 \end{equation}
@@
 In Figure \ref{fig:results:parameters:roberts:saliency} we show a summary of the threshold's behavior for this algorithm, in the whole range of possible values $[0,1]$.
-For each image, this information is summarized in a saliency map which serves to asses the quality of the hierarchical set of edges obtained.
+For each image, this information is summarized in a saliency map which serves to assess the quality of the obtained hierarchical set of edges.
 The saliency map is constructed by sampling the scale range (using steps of $0.05$) and computing the edge map corresponding for each scale.
@@
 If they were not included they will be seen as new contours in the saliency map.
-These results empirically show that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.
+These results show empirical evidence that the obtained hierarchy holds the strong causality property, introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.
 
@@
 \renewcommand{\anchocinco}{2.5cm}
-\newcommand{\anchocuatro}{5cm}
+\newcommand{\anchocuatro}{4.5cm}
 
@@
 
-\renewcommand{\anchocuatro}{5cm}
+\renewcommand{\anchocuatro}{4.5cm}
 
@@
 31), which are removed by a line fitting postprocessing.
-With respect to Prewitt and Sobel Operators,
+With respect to \emph{Prewitt} and \emph{Sobel} operators, they are arise from an in-depth study of different ways to discretize the gradient.
+In \cite{im_proc:segmentation:prewitt:1970:object_enhancement} these discretizations are discussed and compared qualitatively.
+In this work we tested only two of the possible discretizations, and we obtained very similar results using both operators.
+On the other hand, they show better performance that the \emph{Roberts} mask;
 
 In Haralick's work, there is an interesting quantitative validation over a synthetic grid.
-On that evaluation, sensitivity to noise as well as comparative performance with Prewitt operators.
-As in this work we didn't benchmark quantitaive
-
-The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:80:digital_step}, were he made the following observations:
+On that evaluation, he shows results of noise sensitivity as well as comparative performance with Prewitt operators.
+As in this work we didn't focus on quantitative benchmarks, we didn't reproduce those results.
+
+The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:84:digital_step}, where he made the following observations:
 \begin{itemize}
@@
 
-The first claim is validated in our experiments varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges.
+The first claim is validated in our experiments by varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges.
 The same reasoning applies to the zero crossing threshold $t_{ZC}$ .
@@
 For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.
-This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).
+This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).
 
@@
 
-The goal of this work was to provide reference implementation for some of the most classical edge detectors.
+The goal of this work was to provide reference implementations for some of the most classical edge detectors.
 Some of them have been published for many years and are quite simple, however, it is surprisingly hard to find trustable implementations.
@@
 We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
-A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
+A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
 In particular, quantitative challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation.
-From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic images.
+From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic image.
 Even at the moment of writing (2014), there are only a dozen edge detection and segmentation works presenting a serious quantitative analysis over a standard database of images.

### 2014-11-18 figuras_zoom.m

@@ -1,0 +1,15 @@
+
+clear all
+close all
+home
+
+fname='results/windmill_mhg_s_3_n_29_th_0.09.png';
+h=figure;
+I=imread(fname);
+imshow(I,[]);
+magnifyOnFigure(h, 'displayLinkStyle', 'straight',...
+                    'EdgeColor', 'red',...
+                    'magnifierShape', 'rectangle',...
+                    'frozenZoomAspectratio', 'on',...
+                    'secondaryAxesFaceColor', [1 0 0],...
+                    'edgeWidth', 2);

### 2015-02-13 review/third_round/review_answer_4.tex
last review

\documentclass[a4paper,10pt]{report}

\usepackage[english]{babel}
%\usepackage[none]{hyphenat}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[pdftex]{hyperref}
%\usepackage[tight,scriptsize]{subfigure}
%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}
\usepackage{jfc}
\hypersetup{hypertexnames, plainpages=false,colorlinks=true}
\usepackage{amsmath}

% Title Page
\title{}
\author{}

\usepackage{xr}
\externaldocument{../../edge-detection-review-3rd-revision-v3}
\reversemarginpar

\newcounter{myquestion}
\newcommand{\ml}{\mathcal{L}}
\newcommand{\ms}{\mathcal{S}}

\begin{document}
\maketitle

%\begin{abstract}
%\end{abstract}

\chapter{Introduction}
In this communication we answer to the fourth round of review of the article "A review of Classic edge detectors", submitted to the Image Processing Online Journal.
In the first place, we would like to thank again the reviewers for the detailed and thoughtful review provided. In the following we address the issues raised by the reviewers. For each issue, an answer is provided and a cross-reference with a red tag $C_n$ (where n is the number of the correction) to the corresponding (corrected) part of the paper is provided. In addition, changes made to the paper are highlighted in red (when significative) in the new version of the paper. 
These red marks will be removed for the final version.

The remarks from the reviewers are marked with the tag \textcolor{blue}{QUESTION} and our answers with the tag \ans.

%\section{Brief summary of changes}
\section{General changes}
\begin{itemize}
 \item Improved instructions in makefile to compile in OSX (libpng headers not found)
 \item Quality of all synthetic images (kernels) has been improved (See corrections \ref{math:common:gauss},\ref{math:common:log} and \ref{math:common:comparison})

\end{itemize}

\chapter{Answers to reviewer \#1}

\que I think the paper scientific content is now ready.\\

\que Apart from that, I had some problems with an experiment (see below) and I
think the authors did not follow the reviewers advice to reread the paper.
There were numerous grammatical errors, as well as typos or presentation
problems, which could have been easily avoided and made this review very
painful.
I'll leave it to the editor to decide whether he prefers another revision or
move directly to editing.\\

\section{CODE}

\begin{verbatim}

The code compiles fine, and "make test" also works. But when I run it with
"oranges.png" it fails (tried several parameters and both prewitt and
roberts). The problem seems to happen when reading the image

 ./edges input/oranges.png -p 0.2
A review of classic edge detection algorithms
Haldo Sponton & Juan Cardelino, IPOL 2014

Input image: 0.2
Size: 140655700885148 x 140733592255632

Number of algorithms to run: 1
Error: xmalloc: out of memory
Running Prewitt, threshold=0.20"

On the other hand, the image size is ok when I don't try to run any
algorithm:
 ./edges input/oranges.png
A review of classic edge detection algorithms
Haldo Sponton & Juan Cardelino, IPOL 2014

Input image: input/oranges.png
Size: 536 x 480

Number of algorithms to run: 0

Can you check it out?

\end{verbatim}

\ans the problem is related to the order of the arguments. The input images must be the last parameter in every case. I noted that on the README.txt.\\

\que Concerning my requirement for the demo, I was not asking for an archive of
all demo results, which would be meaningless, but for each experiment an
archive containing the results of  the different detections instead of
having to download each detection result separately.\\

\ans I agree it would be useful, but I propose to leave it to a future version of the demo. I think this can be done in the IPOL framework without another round of review.\\

\section{Paper}

\que Algorithm 2 why not merge lines 7-8?\\

\ans done.\\

Then a list of errors (again). AGAIN I did not write all mistakes down, so I
strongly encourage you to proofread your paper.\\

\que abstract: results obtained from a qualitative point of view... -> rephrase\\
\ans rephrased.\\

\que p1: "operations operations"\\
\ans done\\

\que p1: two footnotes with the same url address.\\
\ans corrected.\\

\que p2: "a disconnected an unordered "???\\
\ans corrected.\\

\que p4:  a slight variation of [...] difference -> rephrase, no main verb in
that sentence.\\
\ans rephrased.\\

\que p4: "their filter are" -> singular or plural?\\
\ans corrected.\\

\que p4: different thresholds lead to different results. Thicker... -> results:
thicker...\\
\ans corrected.\\

\que p5: [h!] in front of each figure.\\
\ans corrected.\\

\que p8: consists on convolving\\
\ans corrected.\\

\que p8 not only... BUT also\\
\ans rephrased.\\

\que p24: "are arise"\\
\ans corrected.\\

\que p25: ".\\
\ans corrected.\\

\que p29: "results [...] algorithms our sample images." fig 20-21-22-23\\
\ans corrected.\\

\chapter{Answers to reviewer \#2}

The comments of this reviewer were addressed in the previous version.\\

\bibliographystyle{unsrt}
\bibliography{../../tools/journal_list_short,../../tools/references,../../tools/hierarchical_segmentation}
\end{document}

### 2015-02-13 edge-detection-review-3rd-revision-v3.tex
last review

--- 2014-11-18 edge-detection-review-3rd-revision-v2.tex
@@
 %-------------------------------------------------------------------------------
-% A Review of Classic Edge Detectors - v10
+% A Review of Classic Edge Detectors - v3.3
 % by Haldo Spontón & Juan Cardelino
-% IPOL 2014
+% IPOL 2015
 %-------------------------------------------------------------------------------
@@
 \usepackage{jfc}
+
+\newcommand{\demourl}{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}
+
 \begin{document}
@@
 like Marr-Hildreth and Haralick. \\
-Results obtained from a qualitative point of view (perceptual) and from a quantitative 
-point of view (number of operations, execution time) are compared, considering different ways to convolve an 
-image with a kernel (step required in some of the algorithms).
+Obtained results are analyzed from a qualitative point of view (perceptual) and from a quantitative 
+point of view (number of operations, execution time), considering different ways to convolve an image with a kernel (step required in some of the algorithms).
 \end{abstract}
@@
 \begin{ipolCode}
-For all the algorithms reviewed, an open source C implementation is provided that can be downloaded from 
+For all the reviewed algorithms, an open source C implementation is provided which can be downloaded from 
 the IPOL publication of this article. An 
-\href{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}{online demonstration}\footnote{\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}} is also available, where 
-you can test and reproduce our results.
+\href{\demourl}{online demonstration}%\footnote{\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}} 
+is also available, where you can test and reproduce our results.
 \end{ipolCode}
@@
 
-Edge detection is one of the oldest and most basic operations operations in image processing, which is often used as a basic building block for more complex algorithms.
+Edge detection is one of the oldest and most basic operations in image processing, which is often used as a basic building block for more complex algorithms.
 The basic idea of edge detection is to detect abrupt changes in image intensity.
@@
 
-In the 70's, edge detection methods were based on using small operators 
+In the 70's, edge detection methods were implemented using small operators 
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image \cite{Gonzalez2007Digital}. Section ~\ref{sec:first} describes such algorithms 
+first derivative of the image \cite{Gonzalez2007Digital}. Section \ref{sec:first} describes such algorithms 
 and serves as an introduction to a more sophisticated 
@@
 
-\myn{\label{intro:objective}}The objective of this work is to provide reproducible implementations of the abovementioned algorithms, along with a comprehensive quantitative evaluation of the behaviour algorithms with respect to the parameters.
+The objective of this work is to provide reproducible implementations of the abovementioned algorithms, along with a comprehensive quantitative evaluation of the behaviour algorithms with respect to the parameters.
 Quantitative benchmarks are outside the scope of this work and will be addressed in future work.
-In addition, we will focus on low-level edge detection algorithms, which means that they will output a disconnected an unordered set of edgels (individual pixel facets where boundaries occur).
+In addition, we will focus on low-level edge detection algorithms, which means that they will output a disconnected and unordered set of edgels (individual pixel facets where boundaries occur).
 Algorithms that perform edge integration into objects boundaries are of higher level of abstraction and not considered in this work.
@@
 
-Algorithms based on first derivatives studied in this paper share a common structure, with the only difference  in the type of filtering used to compute those derivatives.
+The algorithms based on first derivatives studied in this paper share a common structure, which only differs in the type of filtering used to compute those derivatives.
 Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure.
@@
 
-\begin{figure}[!h]
+\begin{figure}[h!]
 	\centering
@@
 
-The usual tool to find the amplitude and direction of changes in 
-intensity of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
+The usual tool to find the magnitude and direction of the
+intensity changes of an image $f$ is the gradient operator (denoted as $\nabla$), defined 
 as the vector
@@
 need to be computed at every pixel in the image. When dealing with digital images, numerical approximations 
-of the partial derivatives are required, calculated in a neighborhood of each point. In the following,
+of these derivatives are computed in a neighborhood of each point. In the following,
 the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.
@@
 \myn{\label{first:roberts}}The Roberts edge detector was introduced in 1963 by L.G.
-Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on computing the first derivative of the image.
+Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on the computation of the first derivative of the image.
 The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$.
@@
 
-\begin{figure}[2][!h]
+\begin{figure}[h!]
 	\centering
@@
 
-\begin{figure}[1][!h]
+\begin{figure}[h!]
 	\centering
@@
 to compute them as shown in Figure \ref{fig:roberts}.
-In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, what produces the same result.
+In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, which produces the same result.
 
@@
 
-The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
+The simplest digital approximation of the partial derivatives using masks of size $3\times3$ is obtained 
 by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
@@
 the difference between the third and first columns approximates the derivative in the y-direction.
-These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
+These approximations can be implemented by filtering the image with the two masks shown in Figure \ref{fig:prewitt}.
 These masks are called \textit{Prewitt operators} \cite{im_proc:segmentation:prewitt:1970:object_enhancement}.
 
-\begin{figure}[1][!h]
+\begin{figure}[h!]
 	\centering
@@
 The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
-$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]^T$.
+$[-1\ 0\ 1]^T$ and an horizontal moving average $[1\ 1\ 1]$.
 The same reasoning could be applied to the right mask.
-Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter.
+Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter, and viceversa.
 Thus, they show smoothing properties in the direction orthogonal to the gradient.
@@
 
-A slight variation of the Prewitt operators which use more weight on the central coefficients of the 
-difference. It can be shown that using more weight in the center location provides better image smoothing. This
-variation is implemented using masks shown in Figure \ref{fig:sobel}. These operators are called 
-\textit{Sobel operators}.
-
-\begin{figure}[1][!h]
+\textit{Sobel operators} are a slight variation of the Prewitt operators which use more weight on the central coefficients of the masks.
+It can be shown that using more weight in the center location provides better image smoothing.
+This variation is implemented using masks shown in Figure \ref{fig:sobel}.
+
+\begin{figure}[h!]
 	\centering
@@
 
-Sobel masks can be seen (up to a  scale factor) as the convolution of a horizontal derivation mask 
-$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
+Sobel masks can be seen (up to a  scale factor) as the convolution of a vertical derivation mask 
+$[-1\ 0\ 1]^T$ with an horizontal smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
 Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\
 
-The computational cost of applying Prewitt and Sobel masks is exactly the same. Thus, it is preferable to 
-use Sobel operators, because edges are better localized, and their filter are also less aliased, because of the weighted shape of $[1\ 2\ 1]$.
+The computational cost of applying Prewitt and Sobel masks is exactly the same.
+Thus, it is preferable to use Sobel operators, because edges are better localized, and its filters are also less aliased, because of the weighted shape of the kernel.
 
@@
 
-As mentioned before, edges can be seen as abrupt changes in intensity, i.e. higher values of gradient.
+As mentioned before, edges can be seen as abrupt changes in intensity, which correspond high values of the gradient.
 An example of this behavior is shown in Figure \ref{fig:profiles3d}, where gradient was computed using Sobel operators.
@@
 
-Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e. for pixel a at position $i,j$ in the image compute
+Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e. for a pixel at position $i,j$ in the image the algorithm computes
 \begin{equation*}
@@
 \end{equation*}
-where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
-for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
-of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
+where $\emph{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
+It is noted that different thresholds lead to different results: thicker or more edges for smaller thresholds, fewer and sparser edges for larger thresholds.
+This threshold, and the others mentioned below, are defined as a percentage of the maximum value
+of the gradient image. The advantage of this approach is to adapt the threshold to the dynamic range of the image. 
 Therefore, the parameter $threshold$ in the algorithms takes values ​​between $0$ and $1$.
@@
 
-The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
+The pseudo-code of the implemented algorithms is shown in Algorithm \ref{algo:fded}.
 
@@
 \begin{algorithmic}[1]
-\REQUIRE input image, threshold $th$.
-\STATE $im \leftarrow$ input image
+\REQUIRE $im$ input image, threshold $th$.
 \STATE Define $operator_x$ and $operator_y$ \COMMENT{Roberts, Prewitt or Sobel.}
@@
 image with different masks, without taking into account the characteristics of the edges or 
-noise in the image.
-
-\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
-are based on the second derivative of the image, and both take steps to reduce noise before 
+the noise in the image.
+
+Both algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
+are based on the second derivative of the image, and take steps to reduce noise before 
 detecting edges in the image.
@@
 images. It is based on finding zero crossing points of the second derivative
-of the image. This can be done in several ways. Two different ways of doing 
-this are implemented in this work (see block diagram in Figure 
-\ref{fig:blockdiagram2}): convolving the image with a Gaussian kernel and then 
+of the image.
+This computation can be done in several ways, but in this work we implement two of these ways  (see block diagram in Figure \ref{fig:blockdiagram2}): convolving the image with a Gaussian kernel and then
 approximating the second derivative (Laplacian) with a 3x3 kernel, or 
 convolving the image with a kernel calculated as the Laplacian of a 
-Gaussian function. There are more ways to do so, for example, using 
+Gaussian function. There are more ways to do perform the same tasks, for example, using 
 recursive Gaussian filters \cite{Deriche1993Recursively}.
@@
 
-Some auxiliary functions are needed such as Gaussian kernel and Laplacian of a Gaussian 
+IN order to implement this algorithm some auxiliary functions are needed, such as Gaussian kernel and Laplacian of a Gaussian 
 kernel generation, and 2-D convolution of an image with a given kernel, 
-with different boundary conditions. These operations will be discussed 
+using different boundary conditions. These operations will be discussed 
 in detail in section \ref{sec:appendix1}.
@@
 
-The Marr-Hildreth algorithm consists on convolving the input image $f(x,y)$ with a LoG kernel;
+The main idea of Marr-Hildreth's algorithm is to convolve the input image $f(x,y)$ with a LoG kernel;
 \begin{equation}\label{eq:log}
@@
 
-Then, generation of both Gaussian and LoG kernels is required (see section \ref{sec:appendix1}).\\
+Then, the algorithm requires the generation of both Gaussian and LoG kernels (see section \ref{sec:appendix1}).\\
 
@@
 different. There are four cases to test: left/right, up/down, and the two diagonals. In this case 
-a threshold is used, so that not only the signs of the opposite pixels must differ, also their 
+two conditions must hold: first, the signs of the opposite pixels must differ, and second, their 
 difference in absolute value must be greater than a certain threshold.
@@
 presented in the previous paragraph is attractive for its simplicity of implementation and its low 
-computational cost. In general it is a technique that yields good results, but if more precision in 
-finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
+computational cost. In general it yields good results, but for applications that require accurate edge localization, more advanced methods for finding zero crossings with subpixel 
 accuracy could be used (e.g.\ marching squares \cite{comp_graph:surf_recon:lorensen:87:marching_cubes}).
@@
 \begin{algorithmic}[1]
-\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
-\STATE $im \leftarrow$ input image
+\REQUIRE $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
 \IF{Gaussian kernel}
@@
 
-In this section the original work of Haralick \cite{bb20239} on edge detection is presented in detail.
-The main idea behind this algorithm is identical to that of the previous method: find zeros in 
+In this section we present a detailed description of Haralick's original work \cite{bb20239}.
+The main idea of this algorithm is identical to that of the previous method: find zeros in 
 the second derivative of the image. In this method, however, the input image is smoothly approximated through local bi-cubic
@@
 
-Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}.
+Here, we present the interpolation method used in the original work of Haralick, although there are other options to do this \cite{getreuer}.
 
@@
 Having $10$ coefficients and $25$ data
-points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\
+points leads to an overdetermined system, which can be solved using least squares or any other suitable fitting technique.\\
 
@@
 
-where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.
+where $\mathbf{b_i}$ is the vector containing the $i-th$ row of $B$, i.e.
 
@@
 
-Let us rearrange $b_i$ as a $5\times5$ matrix and denote it by $\mathbf{mask_i}$ as follows
+Let us rearrange $\mathbf{b_i}$ as a $5\times5$ matrix and denote it by $\mathbf{mask_i}$ as follows
 \begin{equation}
@@
 The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
-\myn{\label{second:haralick:analytical:def}} This means that it is actually looking for ascending step edges.
+This means that it is actually looking for ascending step edges.
 According to Haralick's definition, a pixel will lie on such an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
@@
 	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
-	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$ and $C_3 < 0$, then that point is an edge point.
+	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$ and $C_3 < 0$, then the candidate is an edge point.
 \end{enumerate}
@@
 \begin{algorithmic}[1]
-\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
-\STATE $im \leftarrow$ input image
+\REQUIRE $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
 \STATE Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
@@
 
-When making a convolution, it is necessary to define the boundary conditions used to compute it near the edges of the image, and to get a valid output the same size as input image. In this paper, 
+When performing a convolution, it is necessary to define the boundary conditions used to compute it near the edges of the image, and to get a valid output of the same size as input image.
+In this paper,
 two methods were implemented: zero-padding and boundary reflection.
-Zero-padding completes the borders of the image with valued pixels.
+Zero-padding completes the borders of the image with zero valued pixels.
 Reflection completes those pixels with the corresponding symmetric pixel value  relative to the edge of the image.
-Direct convolution is not the only way of filtering an image with a kernel.
-There are other methods such as FFT convolution or recursive filtering.%\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
+Direct convolution is not the only way of filtering an image with a kernel, there are other methods such as FFT convolution or recursive filtering.%\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.
 
@@
 	\includegraphics[width=0.48\textwidth]{results/molino_bw.png}
-	\caption{Test image 2: windmill (1000$\times$563 pixels).}
+	\caption{Test image 2: Windmill (1000$\times$563 pixels).}
 	\label{fig:original2}
@@
 	\end{center}
-	\caption{Execution time of the algorithms (including I/O) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
+	\caption{Execution time of the algorithms (including I/O) using the \emph{Windmill} image from Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
 	\label{exectime2}
@@
 
-Further examples obtained with the implemented algorithms are shown in Figures \ref{fig:example1} and \ref{fig:example2}.
-More examples can be found in the online \href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{demo}.
+Further examples obtained with the implemented algorithms are shown in Figures \ref{fig:example1} and \ref{fig:example2}. More examples can be found in the online \href{\demourl}{demo}.
 
@@
 	\item The first derivative algorithms, although they work properly and run fast, have some problems like discontinuity of the edges.
-They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to avoid and improve this, e.g.
+They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to improve this, e.g.
 Canny edge detector \cite{Canny1986}).
 \\
-	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail.
+	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results regarding edge details.
 This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel.
@@
 This can be verified by observing the values of $\alpha^-$ assigned to each contour.
-If a contour has a high $\alpha^-$ value it means that it will survive for a long time in the merging process. On the other hand, small values mean that it will be removed at early stages.
+If a contour has a high $\alpha^-$ value it means that it will survive for a long time in the detection process. On the other hand, small values mean that it will be removed at early stages.
 As the figure shows, the highest values of $\alpha^-$ are assigned to the most coarser segmentation.
@@
 If they were not included they will be seen as new contours in the saliency map.
-These results show empirical evidence that the obtained hierarchy holds the strong causality property, introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.
+These results show empirical evidence that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini in \cite{book:morel:95:variational_pde}.
 
@@
 As the results show, they all present a similiar behavior, however the are small differences in the results of Roberts vs the other two algorithms.
-For example, edges surrounding the segments of the orange are better defined, because they have more relative weight than other spurious borders.
+For example in the latter, edges surrounding the segments of the orange are better defined, because they have more relative weight than other spurious borders.
 This is explained by the smoothing introduced by Prewitt and Sobel masks explained in Sections \ref{sec:first:sobel} and \ref{sec:first:sobel}, which removes theses spurious edges and thus improves the meaningfulness of the stable ones.
@@
 \figRowMultiscaleHHH{test}{roberts}{0004}
-  \caption{Results of the \textit{Roberts} algorithm our sample images.
+  \caption{Results of the \textit{Roberts} algorithm on our sample images.
 For each example we show the original image on the left and the results corresponding to parameter values of 0.1, 0.2 and 0.3 respectively.}
@@
 \figRowMultiscaleHH{test}{0005}
-  \caption{Results of the first derivative algorithms our sample images.
+  \caption{Results of the first derivative algorithms on our sample images.
 For each example we show the original image on the left and the saliency map of each algorithm (Roberts, Sobel, Prewitt) respectively.
@@
 \figRowMultiscaleHH{test}{0008}
-  \caption{Results of the first derivative algorithms our sample images.
+  \caption{Results of the first derivative algorithms on our sample images.
 For each example we show the original image on the left and a saliency map of each algorithm (Roberts, Sobel, Prewitt) respectively.
@@
 \figRowMultiscaleHHHH{test}{0005}
-  \caption{Results of the second derivative algorithms our sample images.
+  \caption{Results of the second derivative algorithms on our sample images.
 For each example we show the original image on the left and a saliency map of each algorithm (Marr-Hildreth-Gaussian, Marr-Hildreth-Log, Haralick) respectively.
@@
 \figRowMultiscaleHHHH{test}{0008}
-  \caption{Results of the second derivative algorithms our sample images.
+  \caption{Results of the second derivative algorithms on our sample images.
 For each example we show the original image on the left and the saliency map of each algorithm (Marr-Hildreth-Gaussian, Marr-Hildreth-Log, Haralick) respectively.
@@
 \subsection{Discussion}
-\myn{\label{results:discussion}}In this section we will discuss the obtained results and contrast them with the claims made by the original authors.
+In this section we will discuss the obtained results and contrast them with the claims made by the original authors.
 In the case of the first derivative algorithms, they are very simple and the claims made by the original authors are straight forward to verify.
@@
 31), which are removed by a line fitting postprocessing.
-With respect to \emph{Prewitt} and \emph{Sobel} operators, they are arise from an in-depth study of different ways to discretize the gradient.
+With respect to \emph{Prewitt} and \emph{Sobel} operators, they arise from an in-depth study of different ways to discretize the gradient.
 In \cite{im_proc:segmentation:prewitt:1970:object_enhancement} these discretizations are discussed and compared qualitatively.
@@
 
-The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:84:digital_step}, where he made the following observations:
+Regarding the Marr-Hildreth algorithm, the obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:84:digital_step}, where he made the following observations:
 \begin{itemize}
@@
 No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales.
-".
+"
 
@@
 
-The second observation is evidenced by the halos seen in the blades of the windmill, see inside left blade and left of the top blade in Figure \ref{fig:result3-e} .
+The second observation is evidenced by the halos seen in the blades of the windmill, see inside left blade and left of the top blade in Figure \ref{fig:result3-e}.
 \footnote{Please look at these issues in the electronic version or the online demo, as they are not noticeable in the printed version.}
@@
 There are many proposals in the literature aiming to integrate edgels into meaningful object boundaries.
-Examples of these approaches are the Canny \cite{Canny1986} edge detector or the meanignful level lines algorithm \cite{segm:cao:03:extracting_meaningful}.
+Examples of these approaches are the Canny edge detector \cite{Canny1986} or the meanignful level lines algorithm \cite{segm:cao:03:extracting_meaningful}.
 
 With respect to the parameters, these methods require the manual determination of the parameters, and it is very hard to find a single threshold value suitable for a broad category of examples.
-In this sense, \emph{A contrario} methods provide a better approach for setting parameters.
+In this sense, \emph{A contrario} methods \cite{segm:cao:03:extracting_meaningful} provide a better approach for setting parameters.
 For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.

### 2015-02-13 edge-detection-review-3rd-revision-v3.tex
final version

@@
 
-\myn{\label{first:roberts}}The Roberts edge detector was introduced in 1963 by L.G.
+The Roberts edge detector was introduced in 1963 by L.G.
 Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on the computation of the first derivative of the image.
@@
 \begin{enumerate}
-	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
+	\item Grayscale conversion of the input image.
 	\item Convolution of the image with:
@@
 
-\myn{\label{second:haralick:fitting:statement}} The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
+The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
 \begin{align}
@@
 
-\myn{\label{second:haralick:analytical:coordinates}} 
 \begin{align*}
@@
 
-\myn{\label{second:haralick:analytical:first}} If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain
+If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain
 
@@
 (first odd integer greater than $6\sigma=24$).
-\myn{\label{math:common:gauss}} 
 \begin{figure}[ht]
@@
 Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.
-\myn{\label{math:common:log}} 
+
 \begin{figure}[ht]
@@
 % del kernel en funcion de sigma.
-\myn{\label{math:common:comparison}} 
+
 \begin{figure}[ht]
@@
 \subsection{Effect of the scale parameter}
-\myn{\label{results:parameter:scale}}In this section we will discuss the effect of the parameters for each of the reviewed algorithms.
+In this section we will discuss the effect of the parameters for each of the reviewed algorithms.
 As a general rule, they all take one threshold parameter which acts as a scale parameter.
@@
 \subsection{Effect of the smoothing parameter}
-\myn{\label{results:parameter:smooth}}As we stated in previous section, all the reviewed algorithms share a scale parameter (threshold), however, the Marr-Hildreth Algorithm also has a smoothing step which adds an additional parameter $\sigma$, which controls the amount of regularization.
+As we stated in previous section, all the reviewed algorithms share a scale parameter (threshold), however, the Marr-Hildreth Algorithm also has a smoothing step which adds an additional parameter $\sigma$, which controls the amount of regularization.
 In Figure \ref{fig:results:parameters:marr:sigma} we show the results using different  $\sigma$ values.
@@
 %-------------------------------------------------------------------------------
+\clearpage
+\bibliographystyle{siam}
+\bibliography{journal_list_long,bibliography,references}
+
+%-------------------------------------------------------------------------------
+
+\end{document}
+
+%-------------------------------------------------------------------------------
 
@@
 %-------------------------------------------------------------------------------
-\clearpage
-\bibliographystyle{siam}
-\bibliography{journal_list_long,bibliography,references}
-
-%-------------------------------------------------------------------------------
-
-\end{document}

### 2015-03-25 edge-detection-review-3rd-revision-v3.tex
corrected MH algorithm pseudocode
v4 corrections

@@
 \ipolSetTitle{A Review of Classic Edge Detectors}
-\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
+\ipolSetAuthors{Haldo Spont\'on\ipolAuthorMark{1}, Juan Cardelino\ipolAuthorMark{2}}
 \ipolSetAffiliations{%
-$^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
-$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}
+\ipolAuthorMark{1} Inst. Ing. Eléctrica, Universidad de la República, Uruguay (\texttt{haldos@fing.edu.uy})\\
+\ipolAuthorMark{2} Inst. Ing. Eléctrica, Universidad de la República, Uruguay (\texttt{juanc@fing.edu.uy})}
 
@@
 such as Sobel, Prewitt and Roberts. In the second reviewed family, second derivatives are used, for example in algorithms 
-like Marr-Hildreth and Haralick. \\
+like Marr-Hildreth and Haralick.
 Obtained results are analyzed from a qualitative point of view (perceptual) and from a quantitative 
@@
 the IPOL publication of this article. An 
-\href{\demourl}{online demonstration}%\footnote{\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}} 
-is also available, where you can test and reproduce our results.
+\href{\demourl}{online demonstration} is also available, where you can test and reproduce our results.
 \end{ipolCode}
@@
 which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
-proposing a model to locally approximate the image around a point; then, using this model, 
+proposing a model to approximate the image locally around a point; then, using this model, 
 an approximation to the second derivative of the image can be calculated analytically and 
@@
 implements the described algorithms. In section \ref{sec:appendix1} some common mathematical developments are presented. 
-%A commented explanation of the source code can be found in section \ref{sec:appendix2}. This work also provides tools for code documentation 
+%A commented explanation of the source code can be found in section \ref{sec:appendix2e}. This work also provides tools for code documentation 
 %(see section \ref{sec:appendix2}). 
@@
 \IF{Gaussian kernel}
+	\STATE $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
+	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
 	\STATE Define Laplacian operator $laplacian$
-	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
+	\STATE $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
 \ELSE
-	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}.
-	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
-	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
+\STATE $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
+\STATE $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
 \ENDIF

### 2015-03-25 edge-detection-review-3rd-revision-v3.tex
advanced on 4th round of corrections

@@
-%-------------------------------------------------------------------------------
+% %-------------------------------------------------------------------------------
 % A Review of Classic Edge Detectors - v3.3
@@
 \usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
-\usepackage[ruled,linesnumbered]{algorithm2e}
 \usepackage{sidecap}
@@
 %\usepackage{algorithm}
-\usepackage{algorithmic}
+%\usepackage[ruled,linesnumbered]{algorithm2e}
+%\usepackage{algorithmic,algorithmicx,algpseudocode}
+\usepackage{algorithm,algorithmicx,algpseudocode}
 \newtheorem{theorem}{Theorem}
@@
 
-The algorithms based on first derivatives studied in this paper share a common structure, which only differs in the type of filtering used to compute those derivatives.
+The algorithms based on first derivatives studied in this paper share a common structure, and only differ in the type of filtering used to compute those derivatives.
 Figure \ref{fig:blockdiagram1} shows a block diagram of that common structure.
@@
 However, in order to be able to compute diagonal edges, 2-D masks are needed.
-\\
-The \textit{Roberts operators} are one of the earliest 
-attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal differences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
-in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
+The \textit{Roberts operators} are one of the earliest attempts to use 2-D masks for this purpose.
+These operators are based on computing the diagonal differences implemented by filtering an image with the masks in Figure \ref{fig:roberts}.
+By convention
+in these figures, x-coordinates grow from left to right and y-coordinates grow from top to down. It is also 
 usual to scale the mask values, in order to have gain equal to $1$. \\
@@
 
-Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{im_proc:segmentation:roberts:1963:boundary_extraction} proposes
-to compute them as shown in Figure \ref{fig:roberts}.
 In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, which produces the same result.
@@
 
-The left mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
+The left hand mask can be constructed (up to a scale factor) as the convolution of two filters: a vertical derivative 
 $[-1\ 0\ 1]^T$ and an horizontal moving average $[1\ 1\ 1]$.
-The same reasoning could be applied to the right mask.
+The same reasoning could be applied to the right hand mask.
 Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter, and viceversa.
@@
 is obtained, containing the approximation of the partial derivatives $f_x$ and $f_y$. Gradient magnitude 
-image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}).
+image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-d}).
 
@@
 where $\emph{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
-It is noted that different thresholds lead to different results: thicker or more edges for smaller thresholds, fewer and sparser edges for larger thresholds.
+It is noted that different threshold values lead to different results: thicker or more edges for smaller thresholds, fewer and sparser edges for larger thresholds.
 This threshold, and the others mentioned below, are defined as a percentage of the maximum value
@@
 \begin{algorithmic}[1]
-\REQUIRE $im$ input image, threshold $th$.
-\STATE Define $operator_x$ and $operator_y$ \COMMENT{Roberts, Prewitt or Sobel.}
-\STATE $g_x \leftarrow$ convolution($im$,$operator_x$)
-\STATE $g_y \leftarrow$ convolution($im$,$operator_y$)
-\STATE $max_M \leftarrow 0$
-\FORALL{pixel $i$ in image}
-	\STATE $M[i] \leftarrow \sqrt{g_x^2+g_y^2}$ \COMMENT{Gradient magnitude.}
-	\IF{$M[i]>max_M$}
-		\STATE $max_M \leftarrow M[i]$
-	\ENDIF
-\ENDFOR
-\FORALL{pixel $i$ in image}
-	\IF{$M[i] \geq th \times max_M$}
-		\STATE $im_{OUT}[i] \leftarrow 255$
-	\ELSE
-		\STATE $im_{OUT}[i] \leftarrow 0$
-	\ENDIF
-\ENDFOR
-\RETURN output image $\leftarrow im_{OUT}$
+\Require $im$ input image, threshold $th$.
+\State Define $operator_x$ and $operator_y$ \Comment{Roberts, Prewitt or Sobel.}
+\State $g_x \leftarrow$ convolution($im$,$operator_x$)
+\State $g_y \leftarrow$ convolution($im$,$operator_y$)
+\State $max_M \leftarrow 0$
+\ForAll{pixel $i$ in image}
+	\State $M[i] \leftarrow \sqrt{g_x^2+g_y^2}$ \Comment{Gradient magnitude.}
+	\If{$M[i]>max_M$}
+		\State $max_M \leftarrow M[i]$
+	\EndIf
+\EndFor
+\ForAll{pixel $i$ in image}
+	\If{$M[i] \geq th \times max_M$}
+		\State $edges[i] \leftarrow 1$
+	\Else
+		\State $edges[i] \leftarrow 0$
+	\EndIf
+\EndFor
+\Return $edges$
 \end{algorithmic}
@@
 the noise in the image.
-
-Both algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
+The algorithms that will presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
 are based on the second derivative of the image, and take steps to reduce noise before 
@@
 
-The algorithm is divided in two steps, each one described later:
+The algorithm is divided in two steps, which will be described later:
 \begin{enumerate}
-	\item Grayscale conversion of the input image.
 	\item Convolution of the image with:
@@
 
-IN order to implement this algorithm some auxiliary functions are needed, such as Gaussian kernel and Laplacian of a Gaussian 
+In order to implement this algorithm some auxiliary functions are needed, such as Gaussian kernel and Laplacian of a Gaussian 
 kernel generation, and 2-D convolution of an image with a given kernel, 
@@
 \end{equation}
-indicating that the image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result\footnote{The difference between these approaches lies in the compromise between the accuracy in the calculation and the computational cost (see Section \ref{kernelcomparison}).} .
+indicating that the image can be smoothed with a Gaussian filter first, and then compute the Laplacian of the result\footnote{The difference between these approaches lies in the compromise between the accuracy in the calculation and the computational cost (see Section \ref{kernelcomparison}).}.
 
@@
 	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}:
-	\begin{equation*}
+	\begin{equation}
 		\begin{bmatrix}
@@
 		\end{bmatrix}
-	\end{equation*}
+		\label{eq:laplacian:mask}
+	\end{equation}
 	\item Find the zero crossings of the image from step 2.
@@
 The zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
-image (both Gaussian and LoG kernels). Each pixel $p$ has eight neighbors, named according to their position 
-as follows.
+image for both Gaussian and LoG kernels. Each pixel $p$ has eight neighbors, named according to their position 
+as follows:
 
@@
 \begin{algorithmic}[1]
-\REQUIRE $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
-\IF{Gaussian kernel}
-	\STATE $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
-	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
-	\STATE Define Laplacian operator $laplacian$
-	\STATE $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
-\ELSE
-\STATE $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
-\STATE $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
-\ENDIF
-\STATE $max_L \leftarrow 0$
-\FORALL{pixel $i$ in image $im_{LAPL}$}
-	\IF{$im_{LAPL}[i]>max_L$}
-		\STATE $max_L \leftarrow im_{LAPL}[i]$
-	\ENDIF
-\ENDFOR
-\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
-	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
-		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
-			\STATE $im_{OUT}[i] \leftarrow 255$
-		\ELSE
-			\STATE $im_{OUT}[i] \leftarrow 0$
-		\ENDIF
-	\ENDFOR
-\ENDFOR
-\RETURN output image $\leftarrow im_{OUT}$
+\Require $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
+\If{Gaussian kernel}
+	\State $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
+	\State $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
+	\State Define Laplacian operator $laplacian$ \Comment{See mask definition in Eq. \ref{eq:laplacian:mask}}
+	\State $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
+\Else
+\State $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
+\State $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
+\EndIf
+\State $max_L \leftarrow 0$
+\ForAll{pixel $i$ in image $im_{LAPL}$}
+	\If{$im_{LAPL}[i]>max_L$}
+		\State $max_L \leftarrow im_{LAPL}[i]$
+	\EndIf
+\EndFor
+\ForAll{pixel $i$ in image $im_{LAPL}$, except borders}
+	\ForAll{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
+		\If{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \bf{and} ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
+			\State $edges[i] \leftarrow 1$
+		\Else
+			\State $edges[i] \leftarrow 0$
+		\EndIf
+	\EndFor
+\EndFor
+\Return $edges$
 \end{algorithmic}
@@
 
-The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
+The surrounding neighborhood of a point $(x_0,y_0)$ in the image $I$ is approximated using the following bi-cubic polynomial $f$ as
 \begin{align}
 	\label{eq:bicubic:long}
-	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  +\\
-	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3 \\
+	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  + \nonumber\\
+	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3.
 \end{align}
@@
 This gives us an equal number of equations to find 10 coefficients.
-As mentioned before this leads to an overdetermined system which can be solved by least squares.
+As mentioned before, this leads to an overdetermined system which can be solved by least squares.
 By substituting the 25 data points into the polynomial equation (\ref{eq:bicubic}), the following system of 
@@
 \end{equation*}
-
+%
 Using matrix notation, the system can be rewritten as
-
+%
 \begin{equation*}
@@
 \end{equation*}
-
+%
 Then, the least squares problem can be solved by using the normal equations,
-
+%
 \begin{equation*}
-	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} 
+	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k}.
 \end{equation*}
-
+%
 Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
-
+%
 $$ \mathbf{k} = \mathbf{B}\mathbf{f} $$
-
+%
 \begin{equation*}
@@
 	\label{eq:coefficients}
-	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{b_i}\mathbf{f} ,\\
-\end{equation}
-
+	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ k_i = \mathbf{b_i}\mathbf{f} ,\\
+\end{equation}
+%
 where $\mathbf{b_i}$ is the vector containing the $i-th$ row of $B$, i.e.
-
+%
 \begin{equation}
 	\label{eq:b_i_v}
-	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,25}
+	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \cdots	& b_{i,25}
 					\end{bmatrix}.\\
 \end{equation}
-
+%
 Let us rearrange $\mathbf{b_i}$ as a $5\times5$ matrix and denote it by $\mathbf{mask_i}$ as follows
@@
 	\label{eq:convolution_b}
-	\mathbf{K_i} = f \star \mathbf{mask_i} \: \mathtt{for} \: i \in \{1 \ldots 10\}\\
+	\mathbf{K_i} = f \star \mathbf{mask_i}.\\
 \end{equation}
@@
 
-The main idea of Haralick's algorithm, as stated in the original work, is to find "negative sloped zero crossings of the second derivative".
+The main idea of Haralick's algorithm, as stated in the original work, is to find ``negative sloped zero crossings of the second derivative".
 This means that it is actually looking for ascending step edges.
 According to Haralick's definition, a pixel will lie on such an edge if the following three conditions hold\footnote{Please note that there is a typo in Haralick's original paper.
-The statement of the idea is presented at the beggining of Section III, in the first column of page 42.
+The statement of the idea is presented at the begining of Section III, in the first column of page 42.
 The three conditions are presented at the end of the same section, in the second column.
@@
 \begin{enumerate}
-\item the first derivative is non zero.
-\item the second derivative is zero.
-\item the third derivative is negative.
+\item The first derivative is non zero.
+\item The second derivative is zero.
+\item The third derivative is negative.
 \end{enumerate}
@@
 
-Let $\theta$ be the gradient angle, defined clockwise with respect to the y-axis\footnote{Although it seems a non standard choice, we followed exactly from the original work}.
+Let $\theta$ be the gradient angle, defined clockwise with respect to the y-axis\footnote{Although it seems a non standard choice, we followed exactly the original work.}.
 We approximate it using the first order terms of the polynomial \ref{eq:bicubic} obtaining the following expression:
@@
 	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
-	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \\
+	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}.
 \end{align}
@@
 	x = \rho\sin{\theta} \\
-	y = \rho\cos{\theta} 
+	y = \rho\cos{\theta}.
 \end{align*}
@@
 
-By construction $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
+By construction, $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
 Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$.
@@
 \begin{equation}\label{eq:haralick:cond_b_orig}
-	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0
+	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0.	
 \end{equation}
@@
 \begin{algorithmic}[1]
-\REQUIRE $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
-\STATE Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
-\FORALL{pixel $i$ in image $im$}
-	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
-	\FOR{$j=1$ \TO $10$}
-		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
-	\ENDFOR
-	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
-	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
-	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
-		\STATE $im_{OUT}[i] \leftarrow 255$
-	\ELSE
-		\STATE $im_{OUT}[i] \leftarrow 0$
-	\ENDIF
-\ENDFOR
-\RETURN output image $\leftarrow im_{OUT}$
+\Require $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
+\State Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \Comment{See Table \ref{table:b_i}.}
+\ForAll{pixel $i$ in image $im$}
+	\State $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
+	\For{$j=1$ \bf{to} $10$}
+		\State $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
+	\EndFor
+	\State $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
+	\State $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
+	\If{{$|C_2/3C_3|\leq\rho_0$} \bf{and} {$C_3<0$}}
+		\State $edges[i] \leftarrow 1$
+	\Else
+		\State $edges[i] \leftarrow 0$
+	\EndIf
+\EndFor
+\Return $edges$
 \end{algorithmic}

### 2015-04-01 edge-detection-review-3rd-revision-v3.tex
finished correcting text

@@
 \newcommand{\demourl}{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}
+\algrenewcommand\Return{\State \algorithmicreturn{} }%
 
@@
 (such as Sobel masks), attempting to compute an approximation of the
-first derivative of the image \cite{Gonzalez2007Digital}. Section \ref{sec:first} describes such algorithms 
+first derivative of the image \cite{book:gonzalez:02:digital_image_processing}. Section \ref{sec:first} describes such algorithms 
 and serves as an introduction to a more sophisticated 
@@
 crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
-Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
+Haralick's algorithm \cite{segm:edge_region:haralick:84:digital_step} is an alternative approach based on the second derivative 
 which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
@@
 to compare their performance. Conclusions are detailed in Section \ref{sec:conclusions}.
-
-\nocite{IPOL}
 
@@
 the noise in the image.
-The algorithms that will presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
+The algorithms that will presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{segm:edge_region:haralick:84:digital_step}) 
 are based on the second derivative of the image, and take steps to reduce noise before 
@@
 	\State $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
-	\State $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
+	\State $im_{smooth} \leftarrow$ convolution($im$,$kernel$)
 	\State Define Laplacian operator $laplacian$ \Comment{See mask definition in Eq. \ref{eq:laplacian:mask}}
-	\State $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
+	\State $im_{lap} \leftarrow$ convolution($im_{smooth}$,$laplacian$)
 \Else
 \State $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
-\State $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
+\State $im_{lap} \leftarrow$ convolution($im$,$log\_kernel$)
 \EndIf
 \State $max_L \leftarrow 0$
-\ForAll{pixel $i$ in image $im_{LAPL}$}
-	\If{$im_{LAPL}[i]>max_L$}
-		\State $max_L \leftarrow im_{LAPL}[i]$
+\ForAll{pixel $i$ in image $im_{lap}$}
+	\If{$im_{lap}[i]>max_L$}
+		\State $max_L \leftarrow im_{lap}[i]$
 	\EndIf
 \EndFor
-\ForAll{pixel $i$ in image $im_{LAPL}$, except borders}
-	\ForAll{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
-		\If{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \bf{and} ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
+\ForAll{pixel $i$ in image $im_{lap}$, except borders}
+	\ForAll{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{lap}$}
+		\If{($\mbox{sign}(im_{lap}[p_1])\neq\mbox{sign}(im_{lap}[p_2])$) \bf{and} ($|im_{lap}[p_1]-im_{lap}[p_2]|>th_{ZC}$)}
 			\State $edges[i] \leftarrow 1$
@@
 
-In this section we present a detailed description of Haralick's original work \cite{bb20239}.
+In this section we present a detailed description of Haralick's original work \cite{segm:edge_region:haralick:84:digital_step}.
 The main idea of this algorithm is identical to that of the previous method: find zeros in 
@@
 
-Here, we present the interpolation method used in the original work of Haralick, although there are other options to do this \cite{getreuer}.
+Here, we present the interpolation method used in the original work of Haralick, although there are other options to do this \cite{im_proc:getreuer:2011:linear_methods_for_image_interpolation}.
 
@@
 %
-$$ \mathbf{k} = \mathbf{B}\mathbf{f} $$
+$$ \mathbf{k} = \mathbf{B}\mathbf{f}. $$
 %
@@
 
-The derivatives are obtained as follows.
+The derivatives are obtained as follows
 
@@
 \end{equation}
-
-By construction, $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
-Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$.
-This in turn implies that $C_1 \geq 0$.
-
+%
 The second condition that the second derivative is equal to zero becomes
-
+%
 \begin{equation}\label{eq:haralick:cond_b_orig}
@@
 With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes
-
-\begin{equation}
-\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
-\end{equation}
-
+%
+\begin{equation}
+\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 .
+\end{equation}
+%
 Thus, the solution is $\rho_*=-\frac{C_2}{3C_3} $ and imposing $|\rho_*|<\rho_0$ we obtain the second condition
-
-\begin{equation}
-\left| \frac{C_2}{3C_3} \right| < \rho_0,
-\end{equation}
-
+%
+\begin{equation}
+\left| \frac{C_2}{3C_3} \right| < \rho_0.
+\end{equation}
+%
 Finally,  the first condition is that the first derivative is non zero, 
-
+%
 \begin{equation}\label{eq:cond:first}
@@
 \end{equation}
-
+%
 If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain
-
+%
 \begin{equation}\label{eq:cond:first:optim}
-f'_{\theta}(\rho_*) = C_1 - \frac{C_2^2}{3C_3}  \\
-\end{equation}
-
-as $C_1>=0$ and $C_3<0$, $f'_{\theta}(\rho_*)$ will always be positive, thus the first condition always hold by construction and does not need to be verified.\\
+f'_{\theta}(\rho_*) = C_1 - \frac{C_2^2}{3C_3}. 
+\end{equation}
+%
+By construction, $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step.
+Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$.
+This in turn implies that $C_1 \geq 0$.
+As shown before, imposing the first condition implies that $C_3<0$, and we shown that $C_1>=0$, thus Equation \ref{eq:cond:first:optim} yields that $f'_{\theta}(\rho_*)$ will always be positive.
+This means that the first condition always hold by construction and does not need to be verified.\\
 
@@
 \Require $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
-\State Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \Comment{See Table \ref{table:b_i}.}
+\State Define the ten $5\times5$ masks ($mask_1, \dots, mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \Comment{See Table \ref{table:b_i}.}
+	\For{$j=1$ \bf{to} $10$} 
+		\State $K_j \leftarrow$ convolution($neighbors$,$mask_j$) \Comment{$K_j$ is the convolution image of Eq. \ref{eq:convolution_b}.}
+	\EndFor
 \ForAll{pixel $i$ in image $im$}
-	\State $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
 	\For{$j=1$ \bf{to} $10$}
-		\State $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
+		\State $k_j \leftarrow$ $K_j(i)$ \Comment{$K_j(i)$ is the value of image $K_j$ at location $i$.}
 	\EndFor
@@
 	\State $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
-	\If{{$|C_2/3C_3|\leq\rho_0$} \bf{and} {$C_3<0$}}
+	\If{{$|C_2/3C_3|<\rho_0$} \bf{and} {$C_3<0$}}
 		\State $edges[i] \leftarrow 1$
@@
 \begin{equation}
-	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
+	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y).
 \end{equation}
@@
 As mentioned before, the size $n$ of the kernel and the standard deviation $\sigma$ of the exponential function 
-are not independent. For the same $\sigma$ value , the function LoG has wider support than the
+are not independent. For the same $\sigma$ value, the function LoG has wider support than the
 Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to correcly sample the LoG kernel.\\
@@
 
-Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm (including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.
-Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}).
-
+Figure \ref{fig:result1} shows the output of each algorithm including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.
+Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}.
+In this simple case, the first derivative edge detection algorithms run better and faster.
+All algorithms show consistent results.
+Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.
+%
 \begin{figure}[h!]
@@
 \end{figure}
-
-In this simple case, the first derivative edge detection algorithms run better and faster.
-All algorithms show consistent results.
-Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.
-
+%
 \begin{figure}[t!]
@@
 Table \ref{exectime2} summarizes the execution times for each of the algorithms.
-
+%
 \begin{table}[t!]
@@
 \end{table}
-
+%
 In this example the difference in performance between the first derivative and the second derivative algorithms is more meaningful.
@@
 However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.
-
-Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes.
+Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones.
+This may be due the regularity that Haralick assumes.
 
@@
 They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to improve this, e.g.
-Canny edge detector \cite{Canny1986}).
+Canny edge detector \cite{segm:edge_detection:canny:86:computational_edge}).
 \\
@@
 This scale is drawn with a \emph{hot} colormap (black, red, yellow, white).
-Intuitively, this means that 'hotter' edges last longer across scales and thus are more meaningful.
+Intuitively, this means that ``hotter'' edges last longer across scales and thus are more meaningful.
 These results show two important properties.
@@
 If they were not included they will be seen as new contours in the saliency map.
-These results show empirical evidence that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini in \cite{book:morel:95:variational_pde}.
+These results show empirical evidence that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.
 
 In Figures \ref{fig:results:parameters:first:saliency:a} and \ref{fig:results:parameters:first:saliency:b} we compare the three first derivative algorithms by means of their saliency maps.
-As the results show, they all present a similiar behavior, however the are small differences in the results of Roberts vs the other two algorithms.
+As the results show, they all present a similiar behavior, however the are small differences in the results of Roberts versus the other two algorithms.
 For example in the latter, edges surrounding the segments of the orange are better defined, because they have more relative weight than other spurious borders.
@@
 In the case of the first derivative algorithms, they are very simple and the claims made by the original authors are straight forward to verify.
-The Roberts algorithm was presented in his Phd Thesis (\cite{im_proc:segmentation:roberts:1963:boundary_extraction}, Chapter IV ) as a step of a more complex 3D reconstruction system.
+The Roberts algorithm was presented in his Phd Thesis (\cite[Chapter IV]{im_proc:segmentation:roberts:1963:boundary_extraction}) as a step of a more complex 3D reconstruction system.
 For this reason, only qualitative results are shown on few images.
@@
 With respect to \emph{Prewitt} and \emph{Sobel} operators, they arise from an in-depth study of different ways to discretize the gradient.
-In \cite{im_proc:segmentation:prewitt:1970:object_enhancement} these discretizations are discussed and compared qualitatively.
+Prewitt  \cite{im_proc:segmentation:prewitt:1970:object_enhancement} discussed and compared these discretizations qualitatively.
 In this work we tested only two of the possible discretizations, and we obtained very similar results using both operators.
-On the other hand, they show better performance that the \emph{Roberts} mask;
+On the other hand, they show better performance that the \emph{Roberts} mask.
 
@@
 
-Regarding the Marr-Hildreth algorithm, the obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:84:digital_step}, where he made the following observations:
+Regarding the Marr-Hildreth algorithm, the obtained results validate the original claims \cite{segm:edge_region:marr:84:digital_step}:
 \begin{itemize}
-\item "A major difficulty with natural images is that changes can and do occur over a
-wide range of scales (Marr 1976a, b).
-No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales.
-"
-
-\item "The reason is that strict bandlimiting gives rise to sidelobes in the spatial spatial filter, and the consequence of these is that, in the zero-crossing image, strong intensity changes give rise to echoes as well as to the directly corresponding zero-crossings..."
-
-\item "Physical edges will produce roughly coincident zero-crossings in channels of nearby sizes.
+\item ``A major difficulty with natural images is that changes can and do occur over a wide range of scales (Marr 1976a, b).
+No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales."
+
+\item ``The reason is that strict bandlimiting gives rise to sidelobes in the spatial spatial filter, and the consequence of these is that, in the zero-crossing image, strong intensity changes give rise to echoes as well as to the directly corresponding zero-crossings[...]"
+
+\item ``Physical edges will produce roughly coincident zero-crossings in channels of nearby sizes.
 The spatial coincidence assumption asserts that the converse of this is true, that is the coincidence of zero crossings is sufficient evidence for the existence of a real physical edge.
@@
 The first claim is validated in our experiments by varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges.
-The same reasoning applies to the zero crossing threshold $t_{ZC}$ .
-
+The same reasoning applies to the zero crossing threshold $t_{ZC}$.
 The second observation is evidenced by the halos seen in the blades of the windmill, see inside left blade and left of the top blade in Figure \ref{fig:result3-e}.
-\footnote{Please look at these issues in the electronic version or the online demo, as they are not noticeable in the printed version.}
-
-The third claim is evidenced by the structure of the saliency maps of the threshold parameter (Figures \ref{fig:results:parameters:second:saliency:a} and \ref{fig:results:parameters:second:saliency:b}) and the smoothing parameter (Figure \ref{fig:results:parameters:marr:sigma:saliency}).
+\footnote{Please look at these issues in the electronic version or the online demo, as they are not noticeable in the printed version.} The third claim is evidenced by the structure of the saliency maps of the threshold parameter (Figures \ref{fig:results:parameters:second:saliency:a} and \ref{fig:results:parameters:second:saliency:b}) and the smoothing parameter (Figure \ref{fig:results:parameters:marr:sigma:saliency}).
 In both maps we observe that meaningful edges are coincident across scales and the spurious ones are not.
@@
 
-The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed.
+The second derivative algorithms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed.
 In real images, they have better behavior than first derivative algorithms.
@@
 Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significantly faster that the LoG one.
-The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
+The latter, while slower (since it needs a larger kernel to achieve similar results), is more accurate because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel).
 It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels.
@@
 Furthermore, we must remark that we focused on low level edge detectors, and thus none of the implemented methods ensure edge connectivity.
-This algorithms only serve as the input for a further edge integration step.
+These algorithms only serve as the input for further edge integration steps.
 There are many proposals in the literature aiming to integrate edgels into meaningful object boundaries.
-Examples of these approaches are the Canny edge detector \cite{Canny1986} or the meanignful level lines algorithm \cite{segm:cao:03:extracting_meaningful}.
+Examples of these approaches are the Canny edge detector \cite{segm:edge_detection:canny:86:computational_edge} or the meanignful level lines algorithm \cite{segm:cao:03:extracting_meaningful}.
 
@@
 In this sense, \emph{A contrario} methods \cite{segm:cao:03:extracting_meaningful} provide a better approach for setting parameters.
-For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.
-This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).
+For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori \cite{morphology:watershed:najman:1996:geodesic_saliency,im_proc:segmentation:arbelaez:2006:boundary_extraction}.
+This idea can be even extended to pick edges at different scales depending on the region of the image; this idea has been applied applied to regions \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions}.
 
@@
 We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
-A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
-In particular, quantitative challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation.
+Suitable frameworks for this kind of analysis are proposed in the literatura \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions,martin:2001:database_of_human_segmented, cardoso:2005:toward_a_generic_evaluation}, but this is clearly outside the scope of this work.
+In particular, quantitatively challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation.
 From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic image.
@@
 %with arbitrary images uploaded by themselves.
+
+\section*{Image Credits}
+
+\newcommand{\tw}{2cm}
+\begin{table}[ht] 
+  \begin{tabular}{|c|c|l|}
+  \hline
+  Image & Original & Credits \\
+  \hline
+  \includegraphics[width=\tw]{results/molino_bw.png} & \includegraphics[width=\tw]{results/molino.jpg} & \emph{Windmill}, H. Spont\'on, 2012.\\
+  \includegraphics[width=\tw]{results/square127.png} & \includegraphics[width=\tw]{results/square127.png} & \emph{Square127}, H. Spont\'on, 2012.\\
+   \includegraphics[width=\tw]{test/data/0006.png} & \includegraphics[width=\tw]{test/data/color/0006.png} & \emph{Cloth Sticks}, Haldo Spont\'on, 2012.\\
+\includegraphics[width=\tw]{test/data/0007.png} & \includegraphics[width=\tw]{test/data/color/0007.png} & \emph{Chairs}, R. Grompone, J. Jakubowicz, J.-M. Morel, G. Randall, 2010.\\
+   \includegraphics[width=\tw]{test/data/0005.png} & \includegraphics[width=\tw]{test/data/color/0005.png} & \emph{Piraeus}, R. Grompone, J. Jakubowicz, J.-M. Morel, G. Randall, 2010.\\
+   \includegraphics[width=\tw]{test/data/0008.png} & \includegraphics[width=\tw]{examples/barbara.jpg} & \emph{Barbara}, Standard test image.\\
+\includegraphics[width=\tw]{examples/lena_bw.png} & \includegraphics[width=\tw]{examples/lena.png} & \emph{Lena}, Standard test image.\\
+\includegraphics[width=\tw]{examples/oranges_bw.png} & \includegraphics[width=\tw]{examples/orange_full.jpg} & \emph{Oranges}, Unknown source, freely available on the internet.\\
+   \hline
+   \end{tabular}
+\end{table}
+
+\section*{Acknowledgements}
+We acknowledge support from the IPOL-LA project funded by Red Clara and ALICE-2 initiatives. We would like to thank the anonymous reviewers for the useful comments.
+We also thank Rafael Grompone for the great amount of work spent on improving the quality of the paper.
 

### 2015-04-02 edge-detection-review-3rd-revision-v4.tex
typo

--- 2015-04-01 edge-detection-review-3rd-revision-v3.tex
@@
 We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
-Suitable frameworks for this kind of analysis are proposed in the literatura \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions,martin:2001:database_of_human_segmented, cardoso:2005:toward_a_generic_evaluation}, but this is clearly outside the scope of this work.
+Suitable frameworks for this kind of analysis are proposed in the literature \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions,martin:2001:database_of_human_segmented, cardoso:2005:toward_a_generic_evaluation}, but this is clearly outside the scope of this work.
 In particular, quantitatively challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation.
